### 2017-08-03 Dtree.tex
First draft

Decision tree learning is a method for approximating a target function, in which learned function is represented by decision tree \citep{Mitchell_1997_Book}. \citet{Quinlan_1986} demonstrated that the technology for building decision trees from examples is fairly robust. He summarized an approach to synthesizing decision trees that has been used in a variety of systems, and he describes one such system, ID3, in detail. Although ID3 algorithm is successful in most classification problems, it does not handle the numeric attributes and only one attribute at a time is tested for making decisions. \citet{Quinlan_1993} extended the ID3 algorithm and presented the C4.5 algorithm. Later on they added boosting capability to C4.5 and called it C5.0. The improved algorithm accepts both continues and discrete features and solves over fitting problem by pruning. Although there are many other algorithm for implementing decision tree, in this study we use the C5.0 algorithm. Discussing the algorithm in details is out of the scope of this paper. There are fairly good amount of references to study those details \citep[e.g.,][]{Mitchell_1997_Book,Quinlan_1993,Hornik_2009}.\\
In general, decision tree measures the effectiveness of an attribute in classifying the training data through information gain. According to  \citet{Mitchell_1997_Book}, information gain of an attribute is the expected reduction in entropy caused by partitioning the examples according to this attribute. If the target can take on $c$ different values the entropy is defined as
\begin{equation}
Entropy(S) = \sum_{i=1}^{c} -p_{i}\ log_2\ p_i,
\end{equation}

where $p_i$ is proportion of $S$ belonging to class $i$. Entropy measures the homogeneity of examples or dataset. Higher entropy means data is  fairly uniformly distributed among different classes. On the other hand lower entropy means data unequally distributed among classes. The extreme case happens when all data belong to one class which results in entropy equal to zero. Given entropy as a measure of impurity in a collection of training example the measure of the effectiveness of attribute is defined through information gain which is expected reduction in entropy caused by partitioning the examples according to this attribute. Gain of the attribute A is defined as

\begin{equation}
Gain(S,A)\ =\ Entropy(S) - \sum_{v \in Values(A) } \frac{|S_v|}{|S|} Entropy(S_v)
\end{equation}

where $Values(A)$ is the set of all possible values of attribute A, and $S_v$ is the subset of $S$ for which attribute $A$ has value $v$. Information gain is used by the algorithm in order to determine which is the better attribute for classifying the training example. The attribute with higher information gain will be the first option to separate the training data based on and grow to the lower nodes ( For more detailed examples refer to chapter 3 of \citet{Mitchell_1997_Book}). \\
We discussed the basic idea behind the decision tree algorithm, however, there are more details specially in dealing with outliers and noises which we address them through pruning approaches. In this study we use C5.0 package \citep{C50_2015} in R to conduct the decision tree algorithm. We discuss the input parameters in the section.\ref{classification_model}.

### 2017-08-03 acknowledgments.tex
First draft

%
\begin{acknowledgments}
% 
acknowledgement here
% 
\end{acknowledgments}

### 2017-08-03 analysis_cluster.tex
First draft

\section{Clustering Analysis}

We start with clustering all dataset with 11 features. Visual presentation of results above 3 dimension is not possible. Therefor we present the clustering results based on a pair of features in Fig.~\ref{fig:CxCy_11}. 
\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_5.pdf}
    \caption{Results of clustering analysis with all 11 attributes (metrics). This results are subsample of 55 possible combination of metrics. Center of cluster is shown as crosses. At this plot ordinal, categorical clusters' numbers are not necessarily represent the poor, fair, good, and excellent. }
    \label{fig:CxCy_11}
\end{figure}
The figure represent a part of results in clustering with 11 features. There are 55 other possibility to present (e.g., C1 vs C7 or C2 vs C9). As one can see in some combination of features the clustering patterns are easily seen. For example in C3 vs C4 we can distinguish 4 region for clusters, although they are partly mixed together. This also correct for C6 vs C7 and C7 vs C8. In the latter one we observe less mixture of data than the former one. Also there are other types of figures where one feature does not add so much information to the clustering process. For example in C9 vs C10, C10 could be considered as irrelevant feature. Irrelevant feature does not add further information for clustering and redundant feature add the equivalent information as other feature \citep{Dy_2004_MLR}. The last combination of features are a good mixture of data and it is not possible to see the clustering boundaries. C1 vs C2 and C10 vs C11 are examples of such feature combination. It is worth to mention that at this point the cluster numbers are not consistent with ordinal categorical numbers of 1,2,3, and 4 representing poor, fair, good, and excellent, respectively.\\   
Although we don't expect to see the pattern that resulted from clustering using 11 features in presentation of only two, however, different studies show that the higher dimension reduce the effect of similarity based on distance. \citet{Parsons_2004_ACM} presented an illustrative example to show the effect of dimension in reducing the importance of the distance. Effect of higher dimensions in clustering has been well studied and many different methods are proposed to reduce this effect. Among them we can name  feature selection before, during, and after clustering, hybrid methods that use combination of methods to select the best subset of feature and subspace analysis. \citet{Dy_2004_MLR} addressed two issues involved in developing an automated feature subset selection algorithm for unlabeled data. They illustrated the irrelevant and redundant features and proposed methods for evaluating candidate features using two performance criteria. \\
Subspace analysis is another technique to address the challenges with higher dimensional data in clustering process. Subspace clustering is an extension of traditional clustering which looks for different pattern using subset of features.  \citet{Parsons_2004_ACM} provided a list of algorithm for conducting subspace clustering and also some potential applications for them. The most common factor among these algorithm is the process to find the a group of best subspaces through optimization process. An n-dimensional dataset has $2^n$ subspaces where it could be very costly and time consuming to evaluate all of them.\\
As we discussed earlier, there is not standard measures for evaluating clusters in the clustering literature also there is no single clustering assignment explains every application \citep{Dy_2004_MLR,Jain_1988_Book,Hartigan_1985_JOC}. The clustering process is strongly dependent on the application. A good example of the concept is clustering a whale, an elephant, and a tuna fish \citep{Jain_1999,Watanabe_1985_Book}.  Whales and elephants should be in the same cluster, because they both are mammals. However, regarding the scope of the clustering the user could put tuna fish and whales in the same cluster because both groups live in water. In this study we are interested in using those features who, in general, represent the simulation accuracy in 4 different categories. As a first step we used the \kmeans{} clustering approach for all features, however, because of mentioned reasons the results are not easy to discuss or even evaluate. Although we have four groups of data, the question is which one should be considered as poor, fair, good, or excellent groups. Therefore, we apply a modified method of subspace clustering approach to cluster the stations. As we discussed earlier and presented in figures the constrained \kmeans{} method effectively put the stations in a cluster with considering the fact that our constrain stations should not be in the same cluster. High number of iteration leads the clustering process to follow the clustering concept that we are looking for which is clustering stations as poor, fair, good, and excellent. In our case number of possible subspace is $2^{11}$ where each features have two options wether belong to subspace or not. However, because of preserving the distance based criteria effects we limited the number of features in the subspace to be 2,3, and 4 features. Therefor we have 330,165, and 55 unique subspaces for 4D,3D, and 2D, respectively. We conduct a \kmeans{} clustering analysis for each of these subspaces. After considering the constraint conditions, which is assigning cannot-link stations to different cluster and and repeat the \kmeans{} algorithm, in some cases, it is not possible to distinguish all four cannot-link stations in different clusters. In this study we ignore these cases. We only use those combination of features that gives 4 unique clusters for the constraint points (hypothetical stations), therefore, we know for sure that all data within same cluster let's say with metric value 3, should be considered as poor. We also control the clusters to be consistent (we reformat numbers to assign cluster 1 to all group of stations that our first constraint belongs to them and so on). Finally, using 550 unique subspace clustering results, we assign the most frequent class to the station.

Fig.~\ref{fig:C_cvm_comp} shows the results for different velocity models and components. A preliminary visually inspection shows that the behavior of scores with respect to the velocity model is similar. Therefor we merge data of different velocity models.  Fig.~\ref{fig:Figure_cluster_comp} shows the results.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_6.pdf}
    \caption{Box plot of data used in the study. Metrics shown for C1 to C11 for 3 components in separate plots for velocity models and clusters. Median of data are shown as a notch. Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data. Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
    \label{fig:C_cvm_comp}
\end{figure}

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_7.pdf}
    \caption{Box plot of data used in the study. Metrics shown for C1 to C11 for 3 components in separate plots for  clusters. Median of data are shown as a notch. Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data. Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
    \label{fig:Figure_cluster_comp}
\end{figure}

Although there are some differences in the scores based on different components, there are general similarity in all clusters for different metrics among their components. Fig.~\ref{fig:Figure_cluster_all_comp} shows the results with merging the components.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_8.pdf}
    \caption{Box plot of data used in the study. Metrics shown for C1 to C11  in separate plots for  clusters. Median of data are shown as a notch. Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data. Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
    \label{fig:Figure_cluster_all_comp}
\end{figure}

Fig.~\ref{fig:Figure_cluster_all_comp} represents the results from a broad perspective. It is necessary to do statistical significance tests to understand the effect of component, velocity model, earthquake, and frequency band in the results.  According to Fig.~\ref{fig:Figure_cluster_all_comp} scores C1, C2, and C11 has the least changes in different clusters. They suggest the idea that they may not helpful in defining a decision rule. On the other hand, other scores are fairly variable among different clusters. 
Scores C1, C2, and C10 are very similar in cluster 1 to 3. However, we can see a small increase in cluster 4 for these scores. Also the data variation becomes small and range of data moves higher. Scores C3-C9 have a considerable increment from cluster 1 to cluster 4. Although there are a broad variation of data in cluster 3 and cluster 4 for these scores, good amount of data (50\% based on the IQR) are in a small range and increasing with respect to the clusters. C11 has a constant behavior with a small reduction in median at cluster 4. Fig.~\ref{fig:Figure_cluster_barplot} shows number of stations in each cluster based on component.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=300px]
        {figures/pdf/Figure_9.pdf}
    \caption{Number of data in each cluster based on components after conducting constraint \kmeans{} clustering through subspace analysis.}
    \label{fig:Figure_cluster_barplot}
\end{figure}

### 2017-08-03 analysis_dtree.tex
First draft

\section{Classification model} \label{classification_model}

In the section.~\ref{methodology} we discussed the basics of decision tree method and we provided a brief history of available algorithm. In this section we develop a classification model. Although there are numerous methods to classify labeled data, however, in this study we are interested in a method to not only predict the class of simulation based on metrics, but also give an intuition about the decision process. Therefor we use tree based algorithm to develop the prediction model.\\

After preprocessing data the most important step in developing a prediction model is feature selection. The step becomes more relevant when data set has hundreds to tens of thousands of variable or features. Although our dataset does not have numerous features we do some analysis on features before developing the prediction model. Features of a dataset could have similar behavior. This characteristic is easy to understand by correlation criteria. The Pearson correlation coefficient between two features is defined as

\begin{equation}
R(i)=\frac{cov(X_i,X_j)}{\sqrt{var(X_i)var(X_j)}},
\end{equation}

where $cov$ is covariance function and $var$ is the variance function \citep{Guyon_2003}. Fig.~\ref{fig:Corplot} represent the Pearson Correlation factor among features.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=200px]
        {figures/pdf/Figure_10.pdf}
    \caption{Pearson correlation values between metrics. The plot is generated using Corrplot package in R \citep{Corrplot_2016}}
    \label{fig:Corplot}
\end{figure}

As we can see from the figure some metrics are highly correlated. Among them we can mention C1 vs C2 , C3 vs C4, C5 vs C6 and C8, and C6 vs C8 have correlation factor more than 90\%. Also we observe that C11 is mostly have a negative correlation with other features. Feature selection is a broad research field. \citet{Guyon_2003} presented a review of common methods to select features, construct new features or reduce the domain. They highlighted the fact that redundant variable or variables that are useless by themselves may be useful in combination with other variables. Clustering is a method to construct new features as an unsupervised learning or feature selection without having the labels in calculation. The idea is to replace a group of features mean with newly developed features. It helps to reduce the number of features and use all of the available information \citep{Duda_1973,Guyon_2003}. Since we don't have a long list of features and also we want to study all feature effects we leave all features to be used in the prediction model.

After analysis of each features the next step is subsetting data for training and testing process. A subset of data that is used for training and test should have about the equal distribution of target values. Equally distributed data helps training process to not biased to an special class also helps to have an accurate evaluation of the model during the test process. Imbalanced data may decrease the classifier performance. A dataset is imbalanced if the classification categories are not approximately equally represented. \citet{Branco_2015} provided a comprehensive review of different approaches toward imbalanced distribution of target variables. A mostly common approach is data preprocessing through resampling. According to \citet{Branco_2015}, resampling strategies based on diverse set of techniques such as: random under/over-sampling, distance methods, data cleaning approaches, clustering algorithm among them.\\
According to the clustering results, among all data with different components and velocity model there are 816,1253,879, and 76 data belong to cluster 1,2,3, and 4, respectively. \citet{Weiss_2003} discussed the fact that the perfectly balanced dataset does not always provide the optimal results. One option could be under sampling cluster 1 to 3 to be in the similar distribution of cluster 4. This may eliminate the useful examples leading to a worse performance. Therefor we 10 times over-sample the data with cluster 4 to have a distribution similar to other cluster. Oversampling may increase the likelihood of overfitting, however, in this case the oversampling rate is not high. On the other hand, using pruning method during training as well as evaluating algorithm with unseen test dataset we address the overfitting issue.  Choosing what fraction of data should be used for training and testing is an open question. We put 30 percent of data for test and 70 percent of data for training.

Different measures are defined for evaluation of performance of prediction models. Two main metrics are precision and recall. In a simple word, in a binary case, precision represent the fact that how many of predicted values as positive is actually positive and recall represent the idea that how many of positive cases are diagnosticated with the algorithm. Choosing between these two measures is dependent on the application. For example if there is a highly contagious disease and we want to put all contaminated persons in a quarantine, we need to focus on recall value, because even if we miss one case, it could distribute the virus, however, there is a trade off between recall and precision. In the mentioned example there should be some person that indicated as contaminated with virus where as they are not. Therefor, the algorithm precision is not high. Table~\ref{tab:confusion_def} which is called confusion matrix is a method to represent the overall functionality of the algorithm \citep{Branco_2015}.

\begin{table}
\centering
\caption{Confusion Matrix for binary problem}
\label{my-label}
\begin{tabular}{llll}
\hline
                                                 &                              & \multicolumn{2}{c}{Prediction}                              \\ \cline{3-4} 
                                                 &                              & \multicolumn{1}{c}{Positive} & \multicolumn{1}{c}{Negative} \\ \cline{2-4} 
\multicolumn{1}{c}{Reference} & \multicolumn{1}{c}{Positive} & TP                           & FN                           \\
\multicolumn{1}{c}{}                             & \multicolumn{1}{c}{Negative} & FP                           & TN                           \\ \hline
\end{tabular}
\label{tab:confusion_def}
\end{table}

According to Table~\ref{tab:confusion_def} and the definition, mathematically precision and recall are defined as
\begin{equation}
Precision = \frac{TP}{TP+FP}, ~ ~ Recall = \frac{TP}{TP+FN},
\end{equation}

Because of trade of between these two metrics other metrics are defined to generate a metric considering both values. Among them we use $F-measure$ (Rijsbergen1979) which is defined as

\begin{equation}
F_\beta = \frac{(1+\beta)^2.recall.precision}{\beta^2.recall+precision},
\end{equation}

In this study we assume $\beta = 1$ to give same importance to precision and recall. \\
Developing a prediction model has a broad discussion. In this study we are not presenting the best model, however, we present an intuition into the method and be able to present a simple model where a user could estimate the target class without using sophisticated methods. We divided data into training and test set. C5.0 package in R has two options to prune the tree. These parameters include: Confidence factor (CF) and the smallest number of samples that must be put in at least two of the splits (minCases). We tune the algorithm for these two parameters based on training with train data and test the results on test data based on F1 score. We also active the winnow option, which activate the internal feature selection algorithm of the C5.0 package. Fig.~\ref{fig:tune} represent the tuning parameters results. It's obvious that the minCases is dominant factor. In this study we choose $CF=0.2$. 
\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_11.pdf}
    \caption{Results of tuning C5.0 algorithm through grid search on confidence factor (CF) and the smallest number of samples that must be put in at least tow of the splits (minCases). Color variation represents F1-score.}
    \label{fig:tune}
\end{figure}

Fig.~\ref{fig:tuneCFp2} represent the variation of F1-score for different clusters (or classes) with respect to the minCases.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=200px]
        {figures/pdf/Figure_12.pdf}
    \caption{Variation of F1-score for $CF=0.2$ for different cluster. Mean-F1 score represent the geometrical mean of F1-score of other clusters.}
    \label{fig:tuneCFp2}
\end{figure}

A decision tree could grow to very low levels and predict all training data accurately unless there is duplicated data with two different target value. The decision tree algorithm in this study provides very good results even the for test data. However, the scope of this study is to generate a simple relationship to evaluate the simulation results. In the case of going to very sophisticated methods, other algorithms (random forest, conditional random forest, neural networks, ...) could be much more accurate. Consequently, In this study we do not activate the boosting capability of the C5.0 algorithm. According to  Fig.~\ref{fig:tuneCFp2} lower minCases and higher CF provides a better results. It is worth to not that the results are based on evaluating algorithm on test data, therefore, higher values are not representing overfitting. However, as we mentioned earlier, we are not presenting the best model, rather we are presenting a simplest model that in general represent the feature effects. 
First we start with a very simple model. A simple model needs a strong pruning process.  Therefore we use $minCases = 100, CF=0.2$. Fig.~\ref{fig:dtree1} shows the classification decision tree. 
\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_13.pdf}
    \caption{First prediction model (M1) for classifying ground motion simulation process based on goodness of fit scores of two metrics (i.e., C8 (Response Spectra) and C4(Total Energy)). Number above the box represent the number of training data that end up to that specific node. Bar plots represent the probability of each class under the related condition.  }
    \label{fig:dtree1}
\end{figure}
Table~\ref{tab:confusionmat_test_1} shows  the confusion matrix of applying the prediction model on test dataset and Table~\ref{tab:prec_recall_test_1} represent the precision, recall, and F1-score for each individual class.

\begin{table}
\centering
\caption{Confusion Matrix for first prediction model.}
%\label{my-label}
\begin{tabular}{cccccc}
\hline
                           & \multicolumn{5}{c}{Prediction} \\ \hline
{Reference} & C  & 1    & 2    & 3    & 4    \\
                           & 1     & 240  & 25   & 0    & 0    \\
                           & 2     & 13   & 299  & 21   & 0    \\
                           & 3     & 0    & 38   & 211  & 9   \\
                           & 4     & 0    & 0    & 31  & 226  \\ \hline
\end{tabular}
\label{tab:confusionmat_test_1}
\end{table}

\begin{table}
\centering
\caption{Precision, Recall, and F1-score of the first prediction model calculated from the confusion matrix.}
\begin{tabular}{llll}
\hline
        & Precision & Recall  & F1 Score \\ \cline{2-4} 
Class 1 & 0.9056604 & 0.9486166 & 0.9266409 \\
Class 2 & 0.8978979 & 0.8259669 & 0.8604317 \\
Class 3 & 0.8178295 & 0.8022814 & 0.8099808 \\
Class 4 & 0.8793774 & 0.9617021 & 0.9186992 \\ \hline
\end{tabular}
\label{tab:prec_recall_test_1}
\end{table}

Summary of attribute usage in decision tree are according to Table~\ref{tab:attribute_usage_1}

\begin{table}
\centering
\caption{Percentage of data that used the attribute to classify them.}
\begin{tabular}{ccc}
\hline
Data percentage (\%) & Attribute & Metric                \\ \hline
100.00               & C8        & Response Spectra      \\
46.78                 & C4        & Total Energy         \\ \hline
\end{tabular}
\label{tab:attribute_usage_1}
\end{table}

Now we decrease the $minClass$ to 20 to have a more accurate model we call this model M2. Fig.~\ref{fig:dtree2} represents the resultant model.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_14.pdf}
    \caption{Second classification model}
    \label{fig:dtree2}
\end{figure}

Confusion matrix, model scores, and attribute usage are presented in  Table~\ref{tab:confusionmat_test_2},  Table~\ref{tab:prec_recall_test_2}, and  Table~\ref{tab:attribute_usage_2}.

\begin{table}
\centering
\caption{Confusion matrix}
%\label{my-label}
\begin{tabular}{cccccc}
\hline
                           & \multicolumn{5}{c}{Prediction} \\ \hline
{Reference} & C  & 1    & 2    & 3    & 4    \\
                           & 1     & 242  & 20   & 0    & 0    \\
                           & 2     & 11   & 330  & 15   & 0    \\
                           & 3     & 0    & 12   & 238  & 4   \\
                           & 4     & 0    & 0    & 10  & 231  \\ \hline
\end{tabular}
\label{tab:confusionmat_test_2}
\end{table}

\begin{table}
\centering
\caption{Recall and precision}
\begin{tabular}{llll}
\hline
        & Precision & Recall  & F1 Score \\ \cline{2-4} 
Class 1 & 0.9236641 & 0.9565217 & 0.9398058 \\
Class 2 & 0.9269663 & 0.9116022 & 0.9192201 \\
Class 3 & 0.9370079 & 0.9049430 & 0.9206963 \\
Class 4 & 0.9585062 & 0.9829787 & 0.9705882 \\ \hline
\end{tabular}
\label{tab:prec_recall_test_2}
\end{table}

\begin{table}
\centering
\caption{Data usage percentile}
\begin{tabular}{ccc}
\hline
Data percentage (\%) & Attribute & Metric                \\ \hline
100.00                 & C8        & Response Spectra      \\
100.00                 & C4        & Total Energy                \\ 
31.87                 & C6        & Peak Velocity      \\
25.66                 & C5        & Peak Acceleration      \\
19.50                 & C10        & Cross Correlation       \\\hline
\end{tabular}
\label{tab:attribute_usage_2}
\end{table}

% error and accuracy also are other measures.
%Rijsbergen, C. V. (1979). Information retrieval. dept. of computer science, university of glasgow, 2nd edition.

### 2017-08-03 conclusions.tex
First draft

\section{Conclusions}

In this study we use two machine learning algorithms to address the long time question about choosing the appropriate metrics to evaluate the ground motion simulation process. Assuming hypothetical stations, we labeled the dataset through constraint \kmeans{} algorithm with subspace analysis. Using the labeled data we develop a decision tree algorithm to evaluate the ground motion simulation. We generate two models where the first model only uses two metrics to evaluate the stations and the second model which is more accurate and also more complicated than the first model uses 5 metrics. According to our analysis, Response Spectra is the most important metric in classifying the stations. After response spectra, Total Energy are mostly used. We present the results for separate velocity models and components, however, a comprehensive study is needed to address the significance of the results or variations to earthquake magnitude, velocity model, component, station distance from epicenter, earthquake depth, and so on. The proposed model could be used with high accuracy in determining the accuracy of simulation uniformly among physics based ground motion simulation developers.

### 2017-08-03 discussion.tex
First draft

\section{Discussion}

According to the results Response Spectra (C8) is the most important metric in classifying the stations. Data percentage in Table.~\ref{tab:attribute_usage_1} represent the amount of data that the attribute is used in classifying them.  After C8, Arias Intensity Value is the most effective metric. With $C8 \leq 4.07~ \& ~ 4.07 < C8 \leq 6.1$ and $6.1 < C8~ \&~ 6.87 \leq C4$ and $6.87 \leq C4 ~ \&  6.1 \leq C8$ one can classify the simulation as poor, fair, good, and excellent, respectively, with a high confidence.  In this case minimum F1-score occurs for Class 3 and maximum F1-score occurs for Class 1. Pruning process reduce the effect of oversampling in cluster 4. In the second model, which is more accurate from the first model for all metrics, other than Response Spectra and Total Energy, 3 other metrics including: Peak Velocity, Peak Acceleration, and Cross Correlation, also determines the classes.  Fig.~\ref{fig:C8dist} represents the variation of Response Spectra score in different distances for different velocity models and components. As one can see it is fairly well distributed in different score variation in respect to response. There is differences in median value in different components. The median value improves from CVMH+without GTL layer to CVMS.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_15.pdf}
    \caption{Variation of C8 (Response Spectra) score with distance. Velocity models are represented in different shapes and components in different colors.  }
    \label{fig:C8dist}
\end{figure}

Analysis of statistical significance for relevance of results to earthquake, velocity model, frequency band, magnitude, distance of stations, and components needs comprehensive study beyond clustering and classification. However, we illustrate the effects without making strong decision whether they affect the results or not. More comprehensive study is undergoing with \citet{Taborda_2016_GJI} dataset. \\
Finally, we represent the application of these models on classifying stations on a geographical representation. Fig.~\ref{fig:M1_dtree_gof} represent the application of the first decision tree algorithm on the simulation of Chino Hills earthquake. In order to keep consistency in color code, and be able to compare the results with other published results \citep[e.g.,][]{Taborda_2013_BSSA, Taborda_2014_BSSA, Taborda_2016_GJI} we assign 3,5,7, and 9 to cluster 1(poor),2(fair),3(good), and 4(excellent), respectively.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_16.pdf}
    \caption{M1 GOF-score for 3 velocity models and 3 components for Chino Hills earthquake simulation (max =4Hz) }
    \label{fig:M1_dtree_gof}
\end{figure}

Subsequently, Fig.~\ref{fig:M2_dtree_gof} represent of application of M2 metric on data.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_17.pdf}
    \caption{M2 GOF-score for 3 velocity models and 3 components for Chino Hills earthquake simulation (max =4Hz) }
    \label{fig:M2_dtree_gof}
\end{figure}

These figures are important from different point of view. As we mentioned before and illustrated in different figures, the GOF results for different component could be different. This can happen for many reason, for example not accurate orientation of station or simply location of station where makes difference for vertical and horizontal incidents. The fact that which component we should use as a final decision is beyond the scope of this paper. However, using different data in clustering and generating classification models should not affect the results. We consider the GOF of two seismograms (data and synthetic) as an observation to study the relationship between metrics not components. We only distinguished data based on components for presentation purposes. In all steps we use all data. This argument is also correct for velocity model. A velocity model could result in different accuracy and GOF scores. Not surprisingly, the algorithm will classify them in different classes. The model predicts different results for the same pair of data and synthetic for different components and velocity models simply because they are different.

### 2017-08-03 gof_database.tex
First draft

\section{Goodness-of-fit database}

State of California is an earthquake prone region with a long history of seismicity. The Los Aneles basin, due to its unique geologic sediments, has been studied from different perspective. Numerous high resolution velocity models are developed mainly to conduct forward ground motion simulation. Occurrence of frequent moderate magnitude earthquakes made the Los Angeles basin as a natural laboratory for seismic  studies specially testing the forward ground motion simulation modeling (give some references including Tabarda et al 2016). In a relatively recent study, \citet{Taborda_2014_BSSA} conducted a comprehensive study in order to assess the functionality of developed velocity models in the Los Angeles basin and surrounding areas. They simulated the 2008 Chino Hills earthquake using 3 velocity models including: CVM-S4, CVM-H, and CVM-H+GTL (for more details about these models see \citet{Taborda_2014_BSSA}.)
They computed the GOF scores based on \citet{Anderson_2004_Proc}, with minor modification introduced by \citet{Taborda_2013_BSSA} as discussed in the validation section.Fig.~\ref{fig:epicenters} shows the location of the epicenter and stations' locations.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=450px]
        {figures/pdf/Figure_1}
    \caption{(a) Region of interest and epicentral location of the 2008 Chino Hills earthquake (highlighted). Focal mechanisms are shown at the margins of the region of interest. Major quaternary faults in the area are shown in the back along with the main roads and county divisions. (b) Simulation area of interest. Dots indicate the location of the 336 stations considered in this study for validation. A subset of these stations include labels and highlights with larger symbols for later reference. The main local and interstate roads are also shown here. In both panels, the background shows the hillshade topography of the region. The color version of this figure is available only in the electronic edition. (figure and caption from 2014 paper. needs revision)}
    \label{fig:epicenters}
\end{figure}

We use \citet{Taborda_2014_BSSA} database where they computed 11 scores for 3 different velocity models. In this study we only analyze the results for broadband which in this case is \feq{0.1}{4}. Fig.~\ref{fig:data_box_plot} shows the box plot of all data that we used in this study which we separate it in velocity models and components.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_2.pdf}
    \caption{Box plot of data used in the study. Metrics are shown for C1 to C11 for 3 components and in separate plots for velocity models. Median of data are shown as a notch. Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data. Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
    \label{fig:data_box_plot}
\end{figure}

As we explained before, we use all data set regardless of velocity model and components. However, distinguishing data for velocity model and components could answer the question that if the goodness of fit scores dependent on them. Addressing this research question is beyond the scope of this paper. We distinguish data based on velocity models and components only in presenting data or results.

Fig.~\ref{fig:data_density}  shows the density distribution of each score for CVM-S4.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_3.pdf}
    \caption{Distribution of data (only CVMS-4) for different components in different scores based on different components. Each distribution has a unit area.}
    \label{fig:data_density}
\end{figure}

### 2017-08-03 introduction.tex
First draft

\section{Introduction}

The validation of ground motion synthetics has received increased attention over the last few years due to advances in physics based deterministic and hybrid simulation methods. Unless in very low frequencies (\fleq{0.5}), due to limitations and uncertainties in the ground motion simulation and velocity models, it is not possible to match the simulation and synthetics wiggle by wiggle. Therefor, in order to compare synthetic seismogram with recorded data as a complex time series, different metrics are defined. These metrics are means to characterize the how well the synthetic matches the statistical characteristic of observed data. In general, these metrics compute the goodness-of-fit (GOF) between data and synthetic in time and frequency domains. \citet{Anderson_2004_Proc} proposed 10 metrics to be used to characterized the GOF in different frequency bands. \citet{Kristekova_2006_BSSA,Kristekova_2009_GJI} developed envelope and phase misfit criteria as well as GOF metrics for realistic scenarios. \citet{Olsen_2010_SRL} presented an alternative GOF metrics for broadband synthetics with relatively high resolution of the small misfits, and \citet{Taborda_2013_BSSA} added modification to \citet{Anderson_2004_Proc} metrics.\\
Recently, due to advances in computational resources, high frequency 3D ground motion simulation on a regional scale were conducted in different regions using different methods (give some references.) The accuracy of simulation with observed data are measured using these GOF scores. However, researchers use these metrics in different perspectives which is highly dependent on their application. For example, for structural engineers who are interested in designing tall buildings, among other metrics, displacement could be an important factor. However, for rigid and semi-rigid structures like dams and nuclear power plants, acceleration could be more important. In general, for all purposes, the duration of strong ground morion could be an important factor. Analysis of the strong ground motion simulation for a specific application based on individual metrics is an acceptable practice, however, validating the simulation process only with individual score is not a valid approach. Our previous studies show that metrics could have completely different scores for a pair of synthetic and data \citep[e.g.,][]{Taborda_2016_GJI}. \citet{Taborda_2013_BSSA} defined a metric from a combination of  \citet{Anderson_2004_Proc} scores representing a general overview of GOF. However, the metric is a linear combination and somehow is biased by the authors preferences about the metrics. \\
Physics-based ground motion simulation community push the maximum frequency  of the simulation to higher values (\fmax{}$\geq 5 ~ Hz$) and there is a high demand for a metric to uniformly validate the simulations (refer to the high-f project). Although \citet{Anderson_2004_Proc}, defined different level for classifying the stations as poor, fair, good, and excellent based on individual metrics, the question becomes what if there is a station with completely different scores for different metrics. This research question motivated us to approach the problem from different perspective. In other words, if we assume we know the accuracy of the simulation process (let's say we classified it as good) what is the relationship between different GOF scores for that pair of stations. However, as far as the authors knowledge, there is no metric to be able to classify a pair of stations based on GOF scores as an estimation of the overall accuracy of the simulation. Therefore, as a first step, if we could label the pair of station for simulation accuracy we can discuss the relationships between GOF scores. To do this we need a good amount of dataset where calculated the GOF scores for a good amount of stations. \\  
 \citet{Taborda_2014_BSSA} conducted a comprehensive study to analyze the functionality of different velocity models in the Los Angeles basin. They used 2008 Chino Hills earthquakes based on 3 velocity models and conducted simulations with maximum frequency of \fmaxeq{4}. They compared the accuracy of the velocity models through using the GOF scores. The study generated an invaluable synthetic and data  GOF scores dataset.\\
The first and the most common method to categorizing multidimensional unlabeled data is clustering and specifically the distance based methods like \kmeans{} \citep{Macqueen_1967_Proc}. However, there is a drawback in this method. If the data is not explicitly distinguishable the final clusters are highly dependent on initial starting points of the algorithm. Therefor, for the use of stations with 11 dimensions, the ordinary \kmeans{} approach results in fairly different clusters after each analysis. The main reason for this issue is the algorithm is solely looking for data which are close together in terms of Euclidian distance. However, in our case we have one more extra information. Hypothetically, if we have a pair of synthetic and data with GOF score equal 3 for all metrics, according to \citet{Anderson_2004_Proc} we can classify the station as poor with a strong confidence. This is also true for GOF score of 5,7, and 9 for fair, good, and excellent classes, respectively. This background knowledge about the database could give a strong formation to the clustering process. \citet{Wagstaff_2001_Proc} defined a \kmeans{} clustering using background knowledge. Adding this feature to the algorithm improves the functionality and most importantly converge to the same clusters with any random initial cluster's centers. Using this approach, first we cluster stations then we statistically analyze each cluster data, based on these results we remove features with different trend from the analysis. Later on, in a supervised learning algorithm, we develop a classification algorithm based on their GOF scores.\\
In the following section,we present a summary of the validation metrics then we discuss the dataset. Later on we explain the methodology. We provide a brief summary of unsupervised learning which is ordinary and constrained clustering approaches. We also discuss the basics of decision tree algorithm as a supervised learning method. We discuss the clustering challenges and use subspace analysis. We develop two decision tree classifiers based on labeled data for future use in comparing data and synthetics. We conclude the study by applying the generated classification methods on \citet{Taborda_2014_BSSA} dataset.

### 2017-08-03 methodology.tex
First draft

\section{Methodology} \label{methodology}

As we discus in section.\ref{validation_metrics}, there is no consensus on a single metric who able to estimate the accuracy of the simulation or could classify the simulation based on given GOF scores. On the other hand we have a comprehensive dataset. Therefore we need to conduct a data mining and knowledge discovery process to get the classification pattern out of the data set. A common method for generating a decision rule having disjunctive characteristic is decision tree approach. The method is developing a rule for classifying data (here stations) based on different attributes (here GOF scores), this process also called supervised learning in machine learning community. However, decision tree needs labeled data. In our case labeled data would be a data with another attribute which shows the overall accuracy of the simulation (In other word the one metric that we are developing methods to generate it). Therefore we need to go one step back and generate label for the data. Generating labels for data is categorized as unsupervised learning or clustering. Different methods have been developed for clustering. The most commonly used method is \kmeans{}. However, \kmeans{} method gives different results based on initial values of cluster center especially where data is not clearly distinguishable. In result we use modified \kmeans{} approach. This method use background knowledge in the unsupervised learning process. Using background knowledge in unsupervised learning converts it into semi-supervised learning process.\\
In summary, first we label the data based on the constrained \kmeans{} approach, then we statistically analyze the each group results. We expect to see similarity among attributes within each classes. Later on, in a classification process, using decision tree we find the best hypothesis to classify a station based on GOF scores. In the rest of this section first we discuss the clustering method for both ordinary and constrained \kmeans{} approach then we discuss the decision tree algorithm and basics in separating data. We discuss the statistical analysis of the results in the results section. \\

\subsection{Clustring}
\input{clustering}
\subsection{Decision Tree}
\input{Dtree}

% More text for use
%========================= \\
 %The method uses background knowledge to direct the clustering in an appropriate direction so the results are close to real data. Obviously there will be some features that is not obey the general pattern among these features. We will put aside those features. Then we try to modify those features to be able to use again in the processing. In the other words we will scale those features to obey the general trend. Anderson(2004) divided the stations into four main groups based on those stations scores. In hypothetical case were two signals are identical these scores should be 10. However, if they are not the same, but the scoring method provides similar results we should get a fairly close score for all matrics. In real cases the scores are different from one another. However, if we could define a metrics that gives a similar level of goodness of fit for all of them then we can for sure categorized the station in an appropriate group. In this study we define 4 hypothetical stations with score of 3,5,7, and 9 which represent middle points for poor, fair, good and excellent goodness of fit categories. Any station close to these stations should reflect the idea that they are at the same group. However, they are slightly different and we want to know why and how we can make them similar. So having these hypothetical stations and forcing the constrained k-means algorithm to put these stations in different groups we are able to bring the background knowledge into the process. So in the study we start with one attribute and cluster stations. According to figures (2D and 3D) we can argue that the background knowledge assisted the k-means clustering to cluster the stations, realistically. Meanwhile it could not force the clusters to have a cluster centers as those points.  We assign the most frequent stations' cluster to that station. Then we study the score variation in each group and see if they have a uniform behavior if not we can further study the potential problems with that score and probable scaling factors to make it obey the pattern. We also analyze the effect of velocity model, earthquake, orientation, frequency band in the final results. 
%Based on having class for each station in the study we develop a decisition tree to provide the group of simulation for Los Angles Basin earthquakes.  \\

%To do list: \\

%Another approach to go is considering all data without separating for earthquake, velocity model, and so on. Then try to put them in one cluster. Those stations which are in the cluster but are in extreme case, remove from the data base. \\
%Analyze inside each cluster, provide normal distribution, if data looks ok keep it, if not remove those features. \\
%Normally majority of dimensions are irrelevant in high dimensional data. These irrelevant dimensions can hide clusters in noisy data and confuse the clustering algorithm. \\
%Try different method and cluster data, then find the points that clustered currently then use that points as a must link condition. \\
%We can not show data with 10 attributes in graph, however, we can generate a set of statistical figures of changing these values in the figure.\\ 
%If using a feature cause to not able to fit the station in any class we remove it or give a factor to put in the right class.\\
%in the end we want to have all stations as constrained. In other word, we start from 1D and if there is a station that is always in the same group we added it in constraint points. then we move to the 2D and we continue the same until we have all station as constraint.\\
%Constraint helps the clustering to get the same results after good amount of iterations.\\
%========================= \\

### 2017-08-03 buildpdf.sh

@@ -1,0 +1,16 @@
+
+MAIN=$1
+FINAL=main
+
+pdflatex ${MAIN}.tex
+bibtex   ${MAIN}.aux
+pdflatex ${MAIN}.tex
+pdflatex ${MAIN}.tex
+
+rm ${MAIN}.fff ${MAIN}.log ${MAIN}.aux ${MAIN}.ttt ${MAIN}.bbl ${MAIN}.lof ${MAIN}.blg ${MAIN}.spl ${MAIN}.out
+
+mv ${MAIN}.pdf ${FINAL}.pdf
+
+open -a Adobe\ Acrobat\ Pro.app ${FINAL}.pdf
+
+clear

### 2017-08-03 figures/codes/choose_constrained_cluster.m

@@ -1,0 +1,80 @@
+function my_index = choose_constrained_cluster(points,clusters)
+% function to distribute points between cluster through kmeans constraint.
+% Written by: Naeem Khoshnevis (nkhshnvs@memphis.edu)
+% Center for Earthquake Research and Information (CERI)
+% University of Memphis
+% July 6, 2017. Last update (July 6, 2017)
+
+
+if size(points,1) ~= size(clusters,1)
+   error('Points and clusters should be the same size.') 
+end
+
+distance_mat = zeros(size(clusters,1),size(clusters,1));
+my_index     = zeros(size(clusters,1),2);
+
+for i=1:size(clusters,1)
+    for j=1:size(clusters,1)
+        
+        distance_mat(i,j) = distanceM(points(j,1:2),clusters(i,1:2));
+        
+    end
+end
+
+
+link_mat = eye(size(clusters,1),size(points,1));
+
+cost = 1000000;
+
+for ff=1:2
+for ii=1:size(clusters,1)
+    for jj=1:size(clusters,1)
+        tmp_link_mat=link_mat;
+        tmp_i = tmp_link_mat(ii,:);
+        tmp_j = tmp_link_mat(jj,:);
+        tmp_link_mat(ii,:)=tmp_j;
+        tmp_link_mat(jj,:)=tmp_i;
+        cost1 = cluster_cost(distance_mat,tmp_link_mat);
+        
+        if cost1 < cost
+            link_mat = tmp_link_mat;
+            cost = cost1;
+        end
+        kk=1;
+        for ik=1:size(points,1)
+            for jk=1:size(clusters,1)
+                
+                if link_mat(jk,ik)==1
+                    my_index(kk,1)=jk;
+                    my_index(kk,2)=ik;
+                    kk=kk+1;
+                end
+                
+            end
+        end
+
+    end
+    
+end
+
+
+
+end
+end
+
+
+function cost = cluster_cost(distance_mat,link_mat)
+    new_dist = link_mat .* distance_mat;
+    cost = sum(new_dist(:));
+end
+
+% function d = distanceM(point,cluster)
+% d = sqrt((point(:,1)-cluster(:,1)).^2+(point(:,2)-cluster(:,2)).^2);
+% end
+
+function d = distanceM(point,cluster)
+d1 = (point - cluster).^2;
+d2 = sum(d1,2);
+d  = sqrt(d2);
+end
+

### 2017-08-03 figures/codes/figure1.m

@@ -1,0 +1,88 @@
+clc
+clear all 
+close all 
+
+% Figure  1 of the paper
+points = [
+1.9 3
+2 3.5
+2.5 3
+3 4
+];
+
+clusters = [
+0.5 0.5
+0.6 5
+2.4 3.5
+5.5 3.5
+];
+
+
+my_index = choose_constrained_cluster(points,clusters);
+
+
+figure
+subplot(2,1,2)
+        scatter(points(:,1),points(:,2))
+        hold on
+        scatter(clusters(:,1),clusters(:,2),'rs')
+        
+        for i=1:size(points,1)
+            ss = sprintf('%s%s','p',num2str(i));
+            text(points(i,1)+0.1,points(i,2)+0.1,ss)
+        end
+        
+        for j=1:size(clusters,1)
+            ss = sprintf('%s%s','c',num2str(j));
+            text(clusters(j,1)+0.1,clusters(j,2)+0.1,ss)
+        end
+        
+        
+        for i=1:size(my_index,1)
+            plot([clusters(my_index(i,1),1) points(my_index(i,2),1)],[clusters(my_index(i,1),2) points(my_index(i,2),2)])
+            hold on
+            
+        end
+        
+ xlim([0 6])
+ ylim([0 6])
+ xlabel('X')
+ ylabel('Y')
+ 
+my_index = [
+    3 1
+    3 2
+    3 3
+    3 4
+
+] ;
+ 
+subplot(2,1,1)
+
+
+        scatter(points(:,1),points(:,2))
+        hold on
+        scatter(clusters(:,1),clusters(:,2),'rs')
+        
+        for i=1:size(points,1)
+            ss = sprintf('%s%s','p',num2str(i));
+            text(points(i,1)+0.1,points(i,2)+0.1,ss)
+        end
+        
+        for j=1:size(clusters,1)
+            ss = sprintf('%s%s','c',num2str(j));
+            text(clusters(j,1)+0.1,clusters(j,2)+0.1,ss)
+        end
+        
+        
+        for i=1:size(my_index,1)
+            plot([clusters(my_index(i,1),1) points(my_index(i,2),1)],[clusters(my_index(i,1),2) points(my_index(i,2),2)])
+            hold on
+            
+        end
+        
+ xlim([0 6])
+ ylim([0 6])
+ xlabel('X')
+ ylabel('Y')
+

### 2017-08-03 gmt/comments.sh

@@ -1,0 +1,25 @@
+
+
+./plot-gof_copy.sh m1 cvms_ns cc 5 200 data4shell_NS_CVMS
+./plot-gof_copy.sh m1 cvms_ew cc 5 200 data4shell_EW_CVMS
+./plot-gof_copy.sh m1 cvms_ud cc 5 200 data4shell_UD_CVMS
+
+./plot-gof_copy.sh m1 cvmhy_ns cc 5 200 data4shell_NS_CVMHY
+./plot-gof_copy.sh m1 cvmhy_ew cc 5 200 data4shell_EW_CVMHY
+./plot-gof_copy.sh m1 cvmhy_ud cc 5 200 data4shell_UD_CVMHY
+
+./plot-gof_copy.sh m1 cvmhn_ns cc 5 200 data4shell_NS_CVMHN
+./plot-gof_copy.sh m1 cvmhn_ew cc 5 200 data4shell_EW_CVMHN
+./plot-gof_copy.sh m1 cvmhn_ud cc 5 200 data4shell_UD_CVMHN
+
+./plot-gof_copy.sh m2 cvms_ns cc 6 200 data4shell_NS_CVMS
+./plot-gof_copy.sh m2 cvms_ew cc 6 200 data4shell_EW_CVMS
+./plot-gof_copy.sh m2 cvms_ud cc 6 200 data4shell_UD_CVMS
+
+./plot-gof_copy.sh m2 cvmhy_ns cc 6 200 data4shell_NS_CVMHY
+./plot-gof_copy.sh m2 cvmhy_ew cc 6 200 data4shell_EW_CVMHY
+./plot-gof_copy.sh m2 cvmhy_ud cc 6 200 data4shell_UD_CVMHY
+
+./plot-gof_copy.sh m2 cvmhn_ns cc 6 200 data4shell_NS_CVMHN
+./plot-gof_copy.sh m2 cvmhn_ew cc 6 200 data4shell_EW_CVMHN
+./plot-gof_copy.sh m2 cvmhn_ud cc 6 200 data4shell_UD_CVMHN

### 2017-08-03 gmt/plot-gof-extract.sh

@@ -1,0 +1,94 @@
+
+clear
+
+# Set GMT constants
+# ------------------------------------------------------------------------------
+
+gmtset FONT_LABEL               = 12
+
+gmtset MAP_FRAME_PEN            = 0.75p,black
+gmtset MAP_FRAME_WIDTH          = 0.75p
+gmtset MAP_TICK_PEN_PRIMARY     = 0.75p
+gmtset MAP_GRID_PEN_PRIMARY     = 0.75p
+gmtset MAP_TICK_PEN_PRIMARY     = 0.75p
+gmtset MAP_TICK_LENGTH_PRIMARY  = 6p/3p
+gmtset MAP_LABEL_OFFSET         = 8p
+gmtset MAP_ANNOT_OFFSET_PRIMARY = 5p
+
+# Read arguments
+# ------------------------------------------------------------------------------
+
+CVM=$1
+TITLE=$2
+COLUMN=$3
+MAXCOUNT=$4
+
+XI=117000
+XF=173000
+YI=86000
+YF=133000
+
+  AZI=140
+  ELE=80
+PERSP=$AZI/$ELE
+
+# Assign static files data
+# ------------------------------------------------------------------------------
+
+DATADIR=../data-plain
+
+THESCORES=${DATADIR}/${CVM}-myscores.txt
+SHORE=${DATADIR}/xy-shoreline.txt
+EPICENTER=${DATADIR}/xy-epicenter.txt
+
+# Setup local constants
+# ------------------------------------------------------------------------------
+
+SCORES=scores.tmp
+DISTA=dista.tmp
+CPTFILE=gof.cpt
+OUTPUT=${CVM}-${TITLE}.ps
+STATIONS=stations.tmp
+MEDIA=media.tmp
+STDDEV=stddev.tmp
+
+# Captures the columns necessary for plotting and prints local files
+# ------------------------------------------------------------------------------
+
+awk -F' ' '{print $3, $4, $'$COLUMN'}' $THESCORES > $SCORES
+awk -F' ' '{print $3, $4}' $THESCORES > $STATIONS
+
+# Computes the grid for plotting the scores
+# ------------------------------------------------------------------------------
+
+surface $SCORES -Ggof.grd -I100 -R0/180000/0/135000 -C0.01 -T0.1
+
+# Surface map
[…]

### 2017-08-03 gmt/plot-gof-margin.sh

@@ -1,0 +1,229 @@
+
+clear
+
+set -x
+
+# Set GMT constants
+# ------------------------------------------------------------------------------
+
+gmt gmtset FONT_LABEL               = 12
+
+gmt gmtset MAP_FRAME_PEN            = 0.75p,black
+gmt gmtset MAP_FRAME_WIDTH          = 0.75p
+gmt gmtset MAP_TICK_PEN_PRIMARY     = 0.75p
+gmt gmtset MAP_GRID_PEN_PRIMARY     = 0.75p
+gmt gmtset MAP_TICK_PEN_PRIMARY     = 0.75p
+gmt gmtset MAP_TICK_LENGTH_PRIMARY  = 6p/3p
+gmt gmtset MAP_LABEL_OFFSET         = 8p
+gmt gmtset MAP_ANNOT_OFFSET_PRIMARY = 5p
+
+# Read arguments
+# ------------------------------------------------------------------------------
+
+     RUN=$1
+  METRIC=$2
+   TITLE=$3
+  COLUMN=$4
+MAXCOUNT=$5
+
+# Assign static files data
+# ------------------------------------------------------------------------------
+
+DATADIR=./
+
+# THESCORES=/Users/shima/Documents/Shima/Research/score/${METRIC}/${RUN}/scores.txt
+# SHORE=${DATADIR}/xy-shoreline.txt
+# EPICENTER=/Users/shima/Documents/Shima/RA-Work/Results/${METRIC}/xy-epicenter.txt
+
+THESCORES=sample-scores.txt
+SHORE=xy-shoreline.txt
+EPICENTER=xy-epicenter.txt
+
+# Setup local constants
+# ------------------------------------------------------------------------------
+
+SCORES=scores.tmp
+DISTA=dista.tmp
+CPTFILE=gof.cpt
+
+# OUTPUT=/Users/shima/Documents/Shima/Research/GOF-plot/${METRIC}/${RUN}/2-${RUN}-${METRIC}-${COLUMN}.ps
+OUTPUTNAME=goftest
+OUTPUTFILE=${OUTPUTNAME}.ps
+
+STATIONS=stations.tmp
+MEDIA=media.tmp
+STDDEV=stddev.tmp
+
+
+# ==============================================================================================================
+# PREPARING ALL THE DATA
+# ==============================================================================================================
+
+
+# Captures the columns necessary for plotting and prints local files
[…]

### 2017-08-03 gmt/plot-gof.sh

@@ -1,0 +1,159 @@
+
+clear
+
+# Set GMT constants
+# ------------------------------------------------------------------------------
+
+gmt gmtset FONT_LABEL               = 12
+
+gmt gmtset MAP_FRAME_PEN            = 0.75p,black
+gmt gmtset MAP_FRAME_WIDTH          = 0.75p
+gmt gmtset MAP_TICK_PEN_PRIMARY     = 0.75p
+gmt gmtset MAP_GRID_PEN_PRIMARY     = 0.75p
+gmt gmtset MAP_TICK_PEN_PRIMARY     = 0.75p
+gmt gmtset MAP_TICK_LENGTH_PRIMARY  = 6p/3p
+gmt gmtset MAP_LABEL_OFFSET         = 8p
+gmt gmtset MAP_ANNOT_OFFSET_PRIMARY = 5p
+
+# Read arguments
+# ------------------------------------------------------------------------------
+
+#CVM=$1
+TITLE=$1
+COLUMN=4
+MAXCOUNT=$2
+
+# Assign static files data
+# ------------------------------------------------------------------------------
+
+DATADIR= cluster_data
+
+THESCORES=${DATADIR}/data4shell_NS_CVMS.txt
+SHORE=xy-shoreline.txt
+EPICENTER=xy-epicenter.txt
+
+# Setup local constants
+# ------------------------------------------------------------------------------
+
+SCORES=scores.tmp
+DISTA=dista.tmp
+CPTFILE=gof.cpt
+OUTPUT=${TITLE}.ps
+STATIONS=stations.tmp
+MEDIA=media.tmp
+STDDEV=stddev.tmp
+
+# Captures the columns necessary for plotting and prints local files
+# ------------------------------------------------------------------------------
+
+awk -F' ' '{print $3, $3, $'$COLUMN'}' $THESCORES > $SCORES
+awk -F' ' '{print $4, $'$COLUMN'}' $THESCORES > $DISTA
+awk -F' ' '{print $2, $3}' $THESCORES > $STATIONS
+
+# Computes the media of all stations and prints to file
+# ------------------------------------------------------------------------------
+
+awk '{
+    s+=$'$COLUMN'
+} END {
+    media=s/NR
+} END {
[…]

### 2017-08-03 gmt/plot-gof_copy.sh

@@ -1,0 +1,162 @@
+
+clear
+
+# Set GMT constants
+# ------------------------------------------------------------------------------
+
+gmt gmtset FONT_LABEL               = 12
+
+gmt gmtset MAP_FRAME_PEN            = 0.75p,black
+gmt gmtset MAP_FRAME_WIDTH          = 0.75p
+gmt gmtset MAP_TICK_PEN_PRIMARY     = 0.75p
+gmt gmtset MAP_GRID_PEN_PRIMARY     = 0.75p
+gmt gmtset MAP_TICK_PEN_PRIMARY     = 0.75p
+gmt gmtset MAP_TICK_LENGTH_PRIMARY  = 6p/3p
+gmt gmtset MAP_LABEL_OFFSET         = 8p
+gmt gmtset MAP_ANNOT_OFFSET_PRIMARY = 5p
+
+# Read arguments
+# ------------------------------------------------------------------------------
+
+     RUN=$1
+  METRIC=$2
+   TITLE=$3
+  COLUMN=$4
+MAXCOUNT=$5
+FNAME=$6
+
+# Assign static files data
+# ------------------------------------------------------------------------------
+
+DATADIR=./cluster_data
+
+THESCORES=${DATADIR}/${FNAME}.txt
+SHORE=${DATADIR}/xy-shoreline.txt
+EPICENTER=${DATADIR}/xy-epicenter.txt
+
+# Setup local constants
+# ------------------------------------------------------------------------------
+
+SCORES=scores.tmp
+DISTA=dista.tmp
+CPTFILE=gof.cpt
+OUTPUT=${RUN}-${METRIC}.ps
+STATIONS=stations.tmp
+MEDIA=media.tmp
+STDDEV=stddev.tmp
+
+# Captures the columns necessary for plotting and prints local files
+# ------------------------------------------------------------------------------
+
+awk -F' ' '{print $2, $3, $'$COLUMN'}' $THESCORES > $SCORES
+awk -F' ' '{print $4, $'$COLUMN'}' $THESCORES > $DISTA
+awk -F' ' '{print $2, $3}' $THESCORES > $STATIONS
+
+# Computes the media of all stations and prints to file
+# ------------------------------------------------------------------------------
+
+awk '{
[…]

### 2017-08-20 abstract.tex
Abstract added.

% 
The validation of ground motion synthetics has received increased attention over the last few years due to the advances in physics-based deterministic and hybrid simulation methods. Unlike for low frequency simulations $(f \le 0.5 Hz)$, for which it has become reasonable to expect a good match between synthetics and data, in the case of high-frequency simulations $(1 Hz \le f)$ it is not possible to match results on a wiggle-by-wiggle basis. This is mostly due to the various complexities and uncertainties involved in earthquake ground motion modeling. Therefore, in order to compare synthetics with data we turn to different time series metrics, which are used as a means to characterize how the synthetics match the data on qualitative and statistical sense. In general, these metrics provide GOF scores that measure the level of similarity in the time and frequency domains. It is common for these scores to be scaled from 0 to 10, with 10 representing a perfect match. Although using individual metrics for particular applications is considered more adequate, there is no consensus or a unified method to classify the comparison between a set of synthetic and recorded seismograms when the various metrics offer different scores. We study the relationship among these metrics through 
semi-supervised and supervised learning approaches in order to labeling and classifying data, respectively. In particular we use constrained \kmeans{} clustering method, in which we define 4 hypothetical stations with scores 3, 5, 7, and 9 for all metrics. We put these stations in the category of cannot-link constraints. We generate the dataset through the validation of the results from a deterministic (physics-based) ground motion simulation for a moderate magnitude earthquake in the greater Los Angeles basin using three velocity models. The maximum frequency of the simulation is 4 Hz. The dataset involves over 300 stations and 11 metrics, or features, as they are understood in the machine learning lingo, where the metrics form a multi-dimensional space. We address the high-dimensional feature effects with a subspace-clustering analysis using 2,3, and 4 dimensional spaces. We label the dataset in each subspace as poor, fair, good, and excellent, where, hypothetical stations with score of 3,5,7, and 9 belongs them, respectively. We assign the final label based on the mode of subspaces for dataset.  Once the dataset has been labeled, we develop two decision trees using C5.0 algorithm to develop two prediction models to classify a simulation in four groups (poor, fair, good, and excellent) based on a reduced number of metrics. We address the unbalanced dataset issue through oversampling method and analyze the precision, recall and F1-score of the classifiers. The prediction models represent a simple, yet effective model to accurately classify the accuracy of the simulation process based on application-independent approach.

### 2017-09-01 Dtree.tex
Figures are pulished, two figures are added, and the manuscript is improved.

@@
+\subsection{DECISION TREE} \label{dtree_methodology}
+
 Decision tree learning is a method for approximating a target function, in which learned function is represented by decision tree \citep{Mitchell_1997_Book}.
@@
 Information gain is used by the algorithm in order to determine which is the better attribute for classifying the training example.
-The attribute with higher information gain will be the first option to separate the training data based on and grow to the lower nodes ( For more detailed examples refer to chapter 3 of \citet{Mitchell_1997_Book}).
-\\
+The attribute with higher information gain will be the first option to separate the training data based on that attribute and grow to the lower nodes (Refer to chapter 3 of \citet{Mitchell_1997_Book} for more detailed examples.) \\
 We discussed the basic idea behind the decision tree algorithm, however, there are more details specially in dealing with outliers and noises which we address them through pruning approaches.
-In this study we use C5.0 package \citep{C50_2015} in R to conduct the decision tree algorithm.
+In this study we use C5.0 package \citep{C50_2015} in R programming language \citep{R_2016_program} to develop the decision tree algorithms.
 We discuss the input parameters in the section.\ref{classification_model}.

### 2017-09-01 abstract.tex
Figures are pulished, two figures are added, and the manuscript is improved.

@@
 This is mostly due to the various complexities and uncertainties involved in earthquake ground motion modeling.
-Therefore, in order to compare synthetics with data we turn to different time series metrics, which are used as a means to characterize how the synthetics match the data on qualitative and statistical sense.
+Therefore, in order to compare synthetics with data we turn to different time series metrics, which are used as means to characterize how the synthetics match the data on qualitative and statistical sense.
 In general, these metrics provide GOF scores that measure the level of similarity in the time and frequency domains.
@@
 semi-supervised and supervised learning approaches in order to labeling and classifying data, respectively.
-In particular we use constrained \kmeans{} clustering method, in which we define 4 hypothetical stations with scores 3, 5, 7, and 9 for all metrics.
+In particular, we use constrained \kmeans{} clustering method, in which we define 4 hypothetical stations with scores 3, 5, 7, and 9 for all metrics.
 We put these stations in the category of cannot-link constraints.
@@
 We address the high-dimensional feature effects with a subspace-clustering analysis using 2,3, and 4 dimensional spaces.
-We label the dataset in each subspace as poor, fair, good, and excellent, where, hypothetical stations with score of 3,5,7, and 9 belongs them, respectively.
-We assign the final label based on the mode of subspaces for dataset.
-Once the dataset has been labeled, we develop two decision trees using C5.0 algorithm to develop two prediction models to classify a simulation in four groups (poor, fair, good, and excellent) based on a reduced number of metrics.
+We label the dataset in each subspace as poor, fair, good, and excellent, where, hypothetical stations with score of 3,5,7, and 9 belong them, respectively.
+We assign the final label based on the mode of subspaces for each station.
+Once the dataset has been labeled, we develop two decision trees using C5.0 algorithm as prediction models to classify a simulation in four groups (poor, fair, good, and excellent) based on a reduced number of metrics.
 We address the unbalanced dataset issue through oversampling method and analyze the precision, recall and F1-score of the classifiers.
 The prediction models represent a simple, yet effective model to accurately classify the accuracy of the simulation process based on application-independent approach.
+We analyze the different combinations of simulation accuracy classes for each components and present a guideline for making decision on accuracy level of simulation independent of components.
+We apply the proposed method on 2008 Chino Hills earthquake data and synthetic and present the results.

### 2017-09-01 analysis_cluster.tex
Figures are pulished, two figures are added, and the manuscript is improved.

@@
-\section{Clustering Analysis}
+\section{Dataset Clustering Analysis}
 
@@
 Center of cluster is shown as crosses.
-At this plot ordinal, categorical clusters' numbers are not necessarily represent the poor, fair, good, and excellent.
+At this plot ordinal, categorical clusters' numbers are not necessarily represent the poor, fair, good, and excellent clusters.
 }
@@
 \end{figure}
-The figure represent a part of results in clustering with 11 features.
+The figure presents a part of results in clustering with 11 features.
 There are 55 other possibility to present (e.g., C1 vs C7 or C2 vs C9).
@@
 In the latter one we observe less mixture of data than the former one.
-Also there are other types of figures where one feature does not add so much information to the clustering process.
+Also there are other types of combination where one feature does not add so much information to the clustering process.
 For example in C9 vs C10, C10 could be considered as irrelevant feature.
-Irrelevant feature does not add further information for clustering and redundant feature add the equivalent information as other feature \citep{Dy_2004_MLR}.
+Irrelevant feature does not add further information for clustering and redundant feature add the equivalent information as the other feature \citep{Dy_2004_MLR}.
 The last combination of features are a good mixture of data and it is not possible to see the clustering boundaries.
-C1 vs C2 and C10 vs C11 are examples of such feature combination.
-It is worth to mention that at this point the cluster numbers are not consistent with ordinal categorical numbers of 1,2,3, and 4 representing poor, fair, good, and excellent, respectively.\\
-Although we don't expect to see the pattern that resulted from clustering using 11 features in presentation of only two, however, different studies show that the higher dimension reduce the effect of similarity based on distance.
+C1 vs C2 and C10 vs C11 are examples of such features combinations.
+It is worth to mention that at this point the cluster numbers are not consistent with ordinal categorical numbers of 1, 2, 3, and 4 representing poor, fair, good, and excellent, respectively.\\
+Although we do not expect to see the pattern that resulted from clustering using 11 features in presentation of only two, however, different studies show that the higher dimension reduce the effect of similarity based on distance.
 \citet{Parsons_2004_ACM} presented an illustrative example to show the effect of dimension in reducing the importance of the distance.
-Effect of higher dimensions in clustering has been well studied and many different methods are proposed to reduce this effect.
-Among them we can name  feature selection before, during, and after clustering, hybrid methods that use combination of methods to select the best subset of feature and subspace analysis.
+Effect of higher dimensions in clustering has been well studied and many different methods are proposed to reduce this effect in the final results.
+Among them we can name  feature selection before, during, and after clustering, hybrid methods that use combination of methods to select the best subset of feature.
 \citet{Dy_2004_MLR} addressed two issues involved in developing an automated feature subset selection algorithm for unlabeled data.
@@
 \citet{Parsons_2004_ACM} provided a list of algorithm for conducting subspace clustering and also some potential applications for them.
-The most common factor among these algorithm is the process to find the a group of best subspaces through optimization process.
+The most common factor among these algorithm is the process to find a group of best subspaces through optimization process.
 An n-dimensional dataset has $2^n$ subspaces where it could be very costly and time consuming to evaluate all of them.\\
@@
 In this study we are interested in using those features who, in general, represent the simulation accuracy in 4 different categories.
-As a first step we used the \kmeans{} clustering approach for all features, however, because of mentioned reasons the results are not easy to discuss or even evaluate.
+As a first step we use the constrained \kmeans{} clustering approach for all features, however, because of mentioned reasons the results are not easy to discuss or even evaluate.
 Although we have four groups of data, the question is which one should be considered as poor, fair, good, or excellent groups.
@@
 Therefor we have 330,165, and 55 unique subspaces for 4D,3D, and 2D, respectively.
-We conduct a \kmeans{} clustering analysis for each of these subspaces.
-After considering the constraint conditions, which is assigning cannot-link stations to different cluster and and repeat the \kmeans{} algorithm, in some cases, it is not possible to distinguish all four cannot-link stations in different clusters.
+We conduct a constrained \kmeans{} clustering analysis for each of these subspaces and repeat the algorithm.
+In some cases, it is not possible to distinguish all four cannot-link stations in different clusters.
 In this study we ignore these cases.
@@
 
-Fig.~\ref{fig:C_cvm_comp} shows the results for different velocity models and components.
-A preliminary visually inspection shows that the behavior of scores with respect to the velocity model is similar.
+Fig.~\ref{fig:C_cvm_comp} shows the results for different velocity models and components after clustering the dataset.
+A preliminary visually inspection shows that the behavior of scores with respect to the velocity model is fairly similar.
 Therefor we merge data of different velocity models.
-Fig.~\ref{fig:Figure_cluster_comp} shows the results.
+and present them at Fig.~\ref{fig:Figure_cluster_comp}.
 
@@
         {figures/pdf/Figure_6.pdf}
-    \caption{Box plot of data used in the study.
-Metrics shown for C1 to C11 for 3 components in separate plots for velocity models and clusters.
-Median of data are shown as a notch.
+    \caption{Box plot of data used in the study after clustering.
+Metrics are shown for C1 to C11 for 3 components in separate plots for velocity models and clusters.
+Median of data are shown as notchs.
 Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data.
@@
         {figures/pdf/Figure_7.pdf}
-    \caption{Box plot of data used in the study.
+    \caption{Box plot of data used in the study after clustering and merging the results for different velocity models.
 Metrics shown for C1 to C11 for 3 components in separate plots for  clusters.
-Median of data are shown as a notch.
+Median of data are shown as notchs.
 Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data.
@@
 
-Although there are some differences in the scores based on different components, there are general similarity in all clusters for different metrics among their components.
+Although there are some differences in the scores based on different components, there are general similarities in all clusters for different metrics among their components.
 Fig.~\ref{fig:Figure_cluster_all_comp} shows the results with merging the components.
@@
         {figures/pdf/Figure_8.pdf}
-    \caption{Box plot of data used in the study.
+    \caption{Box plot of data used in the study after clustering and merging for different components and velocity models.
 Metrics shown for C1 to C11  in separate plots for  clusters.
-Median of data are shown as a notch.
+Median of data are shown as notchs.
 Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data.
@@
 
-Fig.~\ref{fig:Figure_cluster_all_comp} represents the results from a broad perspective.
-It is necessary to do statistical significance tests to understand the effect of component, velocity model, earthquake, and frequency band in the results.
+Fig.~\ref{fig:Figure_cluster_all_comp} presents the results from a broad perspective.
+Statistical tests can be helpful to understand the effect of component, velocity model, earthquake, and frequency band in the results.
 According to Fig.~\ref{fig:Figure_cluster_all_comp} scores C1, C2, and C11 has the least changes in different clusters.
-They suggest the idea that they may not helpful in defining a decision rule.
+It suggests the idea that they may not helpful in defining a decision rule.
 On the other hand, other scores are fairly variable among different clusters.

### 2017-09-01 analysis_dtree.tex
Figures are pulished, two figures are added, and the manuscript is improved.

@@
-\section{Classification model} \label{classification_model}
+\section{Classification model} \label{dtree}
 
-In the section.~\ref{methodology} we discussed the basics of decision tree method and we provided a brief history of available algorithm.
+In the section.~\ref{dtree_methodology} we discussed the basics of decision tree method and we provided a brief history of available algorithm.
 In this section we develop a classification model.
-Although there are numerous methods to classify labeled data, however, in this study we are interested in a method to not only predict the class of simulation based on metrics, but also give an intuition about the decision process.
+Although there are numerous methods to classify labeled data, however, in this study we are interested in a method that not only predict the class of simulation based on metrics, but also give an intuition about the decision process.
 Therefor we use tree based algorithm to develop the prediction model.\\
 
+Preprocessing data is an important part of data analysis.
+However, in this study we deal with a clean data without any missing values.
+All data are reported in the same scale (0-10), therefore, we do not normalize them.
 After preprocessing data the most important step in developing a prediction model is feature selection.
@@
         {figures/pdf/Figure_10.pdf}
-    \caption{Pearson correlation values between metrics. The plot is generated using Corrplot package in R \citep{Corrplot_2016}}
+    \caption{Pearson correlation values between metrics. The plot is generated using Corrplot package  \citep{Corrplot_2016}  in R programming language.}
     \label{fig:Corplot}
@@
 They highlighted the fact that redundant variable or variables that are useless by themselves may be useful in combination with other variables.
-Clustering is a method to construct new features as an unsupervised learning or feature selection without having the labels in calculation.
+Clustering can be used as a method to construct new features as an unsupervised learning or feature selection without having the labels in calculation.
 The idea is to replace a group of features mean with newly developed features.
 It helps to reduce the number of features and use all of the available information \citep{Duda_1973,Guyon_2003}.
-Since we don't have a long list of features and also we want to study all feature effects we leave all features to be used in the prediction model.
+Since we do not have a long list of features and also we want to study all features' effects we leave all features to be used in the prediction model.
 
@@
 Equally distributed data helps training process to not biased to an special class also helps to have an accurate evaluation of the model during the test process.
+Not equally distributed data is called imbalanced.
 Imbalanced data may decrease the classifier performance.
-A dataset is imbalanced if the classification categories are not approximately equally represented.
 \citet{Branco_2015} provided a comprehensive review of different approaches toward imbalanced distribution of target variables.
@@
 According to \citet{Branco_2015}, resampling strategies based on diverse set of techniques such as: random under/over-sampling, distance methods, data cleaning approaches, clustering algorithm among them.\\
-According to the clustering results, among all data with different components and velocity model there are 816,1253,879, and 76 data belong to cluster 1,2,3, and 4, respectively.
+According to the clustering results, among all data with different components and velocity model there are 816,1253,879, and 76 data belonging to cluster 1,2,3, and 4, respectively.
 \citet{Weiss_2003} discussed the fact that the perfectly balanced dataset does not always provide the optimal results.
@@
 Oversampling may increase the likelihood of overfitting, however, in this case the oversampling rate is not high.
-On the other hand, using pruning method during training as well as evaluating algorithm with unseen test dataset we address the overfitting issue.
+On the other hand, using pruning method during training as well as evaluating algorithm with unseen (test) dataset, we address the overfitting issues.
 Choosing what fraction of data should be used for training and testing is an open question.
-We put 30 percent of data for test and 70 percent of data for training.
+We put 30 percent of data for test and 70 percent of data for training process.
 
@@
 Two main metrics are precision and recall.
-In a simple word, in a binary case, precision represent the fact that how many of predicted values as positive is actually positive and recall represent the idea that how many of positive cases are diagnosticated with the algorithm.
+In a simple word, in a binary case, precision represent the fact that how many of predicted values as positive is actually positive and recall shows that how many of positive cases are diagnosticated with the algorithm.
 Choosing between these two measures is dependent on the application.
@@
 In the mentioned example there should be some person that indicated as contaminated with virus where as they are not.
-Therefor, the algorithm precision is not high.
-Table~\ref{tab:confusion_def} which is called confusion matrix is a method to represent the overall functionality of the algorithm \citep{Branco_2015}.
+Therefor, the algorithm precision is low.
+Table~\ref{tab:confusion_def} which is called confusion matrix is a method to present the overall functionality of the algorithm \citep{Branco_2015}.
 
@@
 
-According to Table~\ref{tab:confusion_def} and the definition, mathematically precision and recall are defined as
+According to Table~\ref{tab:confusion_def} and the definition, mathematically, precision and recall are defined as
 \begin{equation}
@@
 Developing a prediction model has a broad discussion.
-In this study we are not presenting the best model, however, we present an intuition into the method and be able to present a simple model where a user could estimate the target class without using sophisticated methods.
-We divided data into training and test set.
+In this study we are not presenting the best model, however, we present an intuition into the method and be able to present a simple model where a user could estimate the target class without using sophisticated algorithms or packages.
+We divid data into training and test sets.
 C5.0 package in R has two options to prune the tree.
 These parameters include: Confidence factor (CF) and the smallest number of samples that must be put in at least two of the splits (minCases).
-We tune the algorithm for these two parameters based on training with train data and test the results on test data based on F1 score.
-We also active the winnow option, which activate the internal feature selection algorithm of the C5.0 package.
-Fig.~\ref{fig:tune} represent the tuning parameters results.
+We tune the algorithm for these two parameters based on training with train data and test the results on test data based on F1-score.
+We also activate the winnow option, which use the internal feature selection algorithm of the C5.0 package.
+Fig.~\ref{fig:tune} presents the tuning parameters results.
 It's obvious that the minCases is dominant factor.
 In this study we choose $CF=0.2$.
+
 \begin{figure}
@@
         {figures/pdf/Figure_11.pdf}
-    \caption{Results of tuning C5.0 algorithm through grid search on confidence factor (CF) and the smallest number of samples that must be put in at least tow of the splits (minCases).
-Color variation represents F1-score.}
+    \caption{Results of tuning C5.0 algorithm through grid search on confidence factor (CF) and the smallest number of samples that must be put in at least two of the splits (minCases).
+Color variation represents F1-score.
+Gray area at C3 has F1-score less than 0.9.}
     \label{fig:tune}
@@
 
-Fig.~\ref{fig:tuneCFp2} represent the variation of F1-score for different clusters (or classes) with respect to the minCases.
+Fig.~\ref{fig:tuneCFp2} presents the variation of F1-score for different clusters (or classes in terms of supervised learning) with respect to the minCases.
 
@@
         {figures/pdf/Figure_12.pdf}
-    \caption{Variation of F1-score for $CF=0.2$ for different cluster. Mean-F1 score represent the geometrical mean of F1-score of other clusters.}
+    \caption{Variation of F1-score for $CF=0.2$ for different cluster. Mean-F1 score is  the geometrical mean of F1-score of all clusters.}
     \label{fig:tuneCFp2}
@@
 A decision tree could grow to very low levels and predict all training data accurately unless there is duplicated data with two different target value.
-The decision tree algorithm in this study provides very good results even the for test data.
+The decision tree algorithm in this study provides very good results for the test data.
 However, the scope of this study is to generate a simple relationship to evaluate the simulation results.
-In the case of going to very sophisticated methods, other algorithms (random forest, conditional random forest, neural networks, ...) could be much more accurate.
+In the case of going to very sophisticated methods, other algorithms (e.g., random forest, conditional random forest, neural networks, ...) can be much more accurate.
 Consequently, In this study we do not activate the boosting capability of the C5.0 algorithm.
 According to  Fig.~\ref{fig:tuneCFp2} lower minCases and higher CF provides a better results.
-It is worth to not that the results are based on evaluating algorithm on test data, therefore, higher values are not representing overfitting.
-However, as we mentioned earlier, we are not presenting the best model, rather we are presenting a simplest model that in general represent the feature effects.
-First we start with a very simple model. A simple model needs a strong pruning process.  Therefore we use $minCases = 100, CF=0.2$. Fig.~\ref{fig:dtree1} shows the classification decision tree. 
+It is worth to not that the results are based on evaluating algorithm on test data, therefore, higher F1-score are not suggesting overfitting.
+However, as we mentioned earlier, we do not present the best model, rather we present a simplest model that in general illustrate the feature effects.
+First we start with a very simple model. A simple model needs a strong pruning process.  Therefore we use $minCases = 120, CF=0.2$. Fig.~\ref{fig:dtree1} shows the classification decision tree. 
 \begin{figure}
@@
     \caption{First prediction model (M1) for classifying ground motion simulation process based on goodness of fit scores of two metrics (i.e., C8 (Response Spectra) and C4(Total Energy)).
-Number above the box represent the number of training data that end up to that specific node.
+Numbers above the box represent the number of training data that end up to that specific node.
 Bar plots represent the probability of each class under the related condition.
@@
 \end{figure}
-Table~\ref{tab:confusionmat_test_1} shows  the confusion matrix of applying the prediction model on test dataset and Table~\ref{tab:prec_recall_test_1} represent the precision, recall, and F1-score for each individual class.
+Table~\ref{tab:confusionmat_test_1} shows  the confusion matrix of applying the prediction model on test dataset and Table~\ref{tab:prec_recall_test_1} presents the precision, recall, and F1-score for each individual class.
 
@@
 \centering
-\caption{Confusion Matrix for first prediction model.}
+\caption{Confusion Matrix for first prediction model (M1) for the unseen (test) dataset.}
 %\label{my-label}
@@
 {Reference}         & C    & 1      & 2      & 3      & 4       \\
-                           & 1     & 240  & 25   & 0    & 0    \\
-                           & 2     & 13   & 299  & 21   & 0    \\
-                           & 3     & 0    & 38   & 211  & 9   \\
-                           & 4     & 0    & 0    & 31  & 226  \\ \hline
+                            & 1     & 242  & 11    & 0      & 0    \\
+                            & 2     & 13    & 313  & 16    & 0    \\
+                            & 3     & 0      & 42    & 207  & 15   \\
+                            & 4     & 0      & 0      & 23    & 231  \\ \hline
 \end{tabular}
@@
 \centering
-\caption{Precision, Recall, and F1-score of the first prediction model calculated from the confusion matrix.}
+\caption{Precision, Recall, and F1-score of the first prediction model (M1) calculated from the confusion matrix.}
 \begin{tabular}{llll}
@@
         & Precision & Recall  & F1 Score \\ \cline{2-4} 
-Class 1 & 0.9056604 & 0.9486166 & 0.9266409 \\
-Class 2 & 0.8978979 & 0.8259669 & 0.8604317 \\
-Class 3 & 0.8178295 & 0.8022814 & 0.8099808 \\
-Class 4 & 0.8793774 & 0.9617021 & 0.9186992 \\ \hline
+Class 1 & 0.9565217 & 0.9490196 & 0.9527559 \\
+Class 2 & 0.9152047 & 0.8551913 & 0.8841808 \\
+Class 3 & 0.7840909 & 0.8414634 & 0.8117647 \\
+Class 4 & 0.9094488 & 0.9390244 & 0.9240000 \\ \hline
 \end{tabular}
@@
 \centering
-\caption{Percentage of data that used the attribute to classify them.}
+\caption{Percentage of data that used the attribute to classify them in first prediction model (M1).}
 \begin{tabular}{ccc}
@@
 
-Now we decrease the $minClass$ to 20 to have a more accurate model we call this model M2. Fig.~\ref{fig:dtree2} represents the resultant model.
+Now we decrease the $minClass$ to 40 to have a more accurate model we call this model M2. Fig.~\ref{fig:dtree2} presents the resultant model.
 
@@
         {figures/pdf/Figure_14.pdf}
-    \caption{Second classification model}
+    \caption{First prediction model (M2) for classifying ground motion simulation process based on goodness of fit scores of two metrics (i.e., C8 (Response Spectra), C4(Total Energy), C6 (Peak Velocity), and C5 (Peak Acceleration)).
+Numbers above the box represent the number of training data that end up to that specific node.
+Bar plots represent the probability of each class under the related condition.
+}
     \label{fig:dtree2}
@@
 
-Confusion matrix, model scores, and attribute usage are presented in  Table~\ref{tab:confusionmat_test_2},  Table~\ref{tab:prec_recall_test_2}, and  Table~\ref{tab:attribute_usage_2}.
+Confusion matrix, model scores, and attribute usage for M2 are presented in  Table~\ref{tab:confusionmat_test_2},  Table~\ref{tab:prec_recall_test_2}, and  Table~\ref{tab:attribute_usage_2}.
 
@@
 \centering
-\caption{Confusion matrix}
+\caption{Confusion matrix of Model (M2) on the unseen (test) dataset.}
 %\label{my-label}
@@
 {Reference}        & C     & 1       & 2      & 3       & 4    \\
-                           & 1     & 242  & 20   & 0    & 0    \\
-                           & 2     & 11   & 330  & 15   & 0    \\
-                           & 3     & 0    & 12   & 238  & 4   \\
-                           & 4     & 0    & 0    & 10  & 231  \\ \hline
+                           & 1     & 248   & 14     & 0       & 0    \\
+                           & 2     & 7       & 343   & 18     & 0    \\
+                           & 3     & 0       & 9     & 221   & 33   \\
+                           & 4     & 0       & 0       & 7       & 213  \\ \hline
 \end{tabular}
@@
         & Precision & Recall  & F1 Score \\ \cline{2-4} 
-Class 1 & 0.9236641 & 0.9565217 & 0.9398058 \\
-Class 2 & 0.9269663 & 0.9116022 & 0.9192201 \\
-Class 3 & 0.9370079 & 0.9049430 & 0.9206963 \\
-Class 4 & 0.9585062 & 0.9829787 & 0.9705882 \\ \hline
+Class 1 & 0.9465649 &  0.9725490 &  0.9593810 \\
+Class 2 & 0.9320652 &  0.9371585 &  0.9346049 \\
+Class 3 & 0.8403042 &  0.8983740 &  0.8683694 \\
+Class 4 & 0.9681818 &  0.8658537 &  0.9141631 \\ \hline
 \end{tabular}
@@
 100.00                 & C8        & Response Spectra      \\
-100.00                 & C4        & Total Energy                \\ 
-31.87                 & C6        & Peak Velocity      \\
-25.66                 & C5        & Peak Acceleration      \\
-19.50                 & C10        & Cross Correlation       \\\hline
+60.23                   & C4        & Total Energy                \\ 
+36.99                   & C6        & Peak Velocity      \\
+21.54                   & C5        & Peak Acceleration      \\ \hline
 \end{tabular}
@@
 %Rijsbergen, C. V. (1979). Information retrieval. dept. of computer science, university of glasgow, 2nd edition.
+
+%Model M1:
+
+%Overall Statistics
+
+%               Accuracy : 0.8922          
+ %                95% CI : (0.8725, 0.9098)
+ %   No Information Rate : 0.3073          
+  %  P-Value [Acc > NIR] : < 2.2e-16
+
+  %                Kappa : 0.8551          
+% Mcnemar's Test P-Value : NA
+
+%Statistics by Class:
+
+%                     Class: 1 Class: 2 Class: 3 Class: 4
+%Sensitivity            0.9565   0.9152   0.7841   0.9094
+%Specificity            0.9849   0.9313   0.9541   0.9825
+%Pos Pred Value         0.9490   0.8552   0.8415   0.9390
+%Neg Pred Value         0.9872   0.9612   0.9343   0.9735
+%Prevalence             0.2273   0.3073   0.2372   0.2282
+%Detection Rate         0.2174   0.2812   0.1860   0.2075
+%Detection Prevalence   0.2291   0.3288   0.2210   0.2210
+%Balanced Accuracy      0.9707   0.9232   0.8691   0.9460
+
+%> f1S
+ % precision    recall        f1
+%1 0.9565217 0.9490196 0.9527559
+%2 0.9152047 0.8551913 0.8841808
+%3 0.7840909 0.8414634 0.8117647
+%4 0.9094488 0.9390244 0.9240000
+
+%Model M2:
+
+%Overall Statistics
+
+ %              Accuracy : 0.9209          
+  %               95% CI : (0.9035, 0.9361)
+  %  No Information Rate : 0.3306          
+   % P-Value [Acc > NIR] : < 2.2e-16
+
+    %              Kappa : 0.8934          
+% Mcnemar's Test P-Value : NA
+
+%Statistics by Class:
+
+ %                    Class: 1 Class: 2 Class: 3 Class: 4
+%Sensitivity            0.9466   0.9321   0.8403   0.9682
+%Specificity            0.9918   0.9691   0.9706   0.9630
+%Pos Pred Value         0.9725   0.9372   0.8984   0.8659
+%Neg Pred Value         0.9837   0.9665   0.9516   0.9919
+%Prevalence             0.2354   0.3306   0.2363   0.1977
+%Detection Rate         0.2228   0.3082   0.1986   0.1914
+%Detection Prevalence   0.2291   0.3288   0.2210   0.2210
+%Balanced Accuracy      0.9692   0.9506   0.9054   0.9656
+
+%> f1S
+ % precision    recall        f1
+%1 0.9465649 0.9725490 0.9593810
+%2 0.9320652 0.9371585 0.9346049
+%3 0.8403042 0.8983740 0.8683694
+%4 0.9681818 0.8658537 0.9141631

### 2017-09-01 conclusions.tex
Figures are pulished, two figures are added, and the manuscript is improved.

@@
 
-In this study we use two machine learning algorithms to address the long time question about choosing the appropriate metrics to evaluate the ground motion simulation process.
+In this study we use two machine learning algorithms to address the long-term question about choosing the appropriate metrics to evaluate the ground motion simulation process independent of application and components.
 Assuming hypothetical stations, we labeled the dataset through constraint \kmeans{} algorithm with subspace analysis.
-Using the labeled data we develop a decision tree algorithm to evaluate the ground motion simulation.
-We generate two models where the first model only uses two metrics to evaluate the stations and the second model which is more accurate and also more complicated than the first model uses 5 metrics.
+Using the labeled data we developed two decision trees algorithm to evaluate the ground motion simulation.
+We generate two models where the first model only uses two metrics to evaluate the stations and the second model which is more accurate and also more complicated than the first model uses 4 metrics.
 According to our analysis, Response Spectra is the most important metric in classifying the stations.
 After response spectra, Total Energy are mostly used.
-We present the results for separate velocity models and components, however, a comprehensive study is needed to address the significance of the results or variations to earthquake magnitude, velocity model, component, station distance from epicenter, earthquake depth, and so on.
-The proposed model could be used with high accuracy in determining the accuracy of simulation uniformly among physics based ground motion simulation developers.
+We studied different combination of components's classes and developed a model to predict a components independent class.
+We applied the models on 2008 Chino Hills earthquake simulation synthetics and data GOF scores and discussed the results.
+A comprehensive study is needed to address the significance of the results or variations to earthquake magnitude, velocity model, station distance from epicenter, earthquake depth, and so on.
+The proposed model could be used with high accuracy (about 90 \%) in determining the accuracy of simulation independent of application and components, uniformly among physics-based ground motion simulation developers.

### 2017-09-01 gof_database.tex
Figures are pulished, two figures are added, and the manuscript is improved.

@@
 Numerous high resolution velocity models are developed mainly to conduct forward ground motion simulation.
-Occurrence of frequent moderate magnitude earthquakes made the Los Angeles basin as a natural laboratory for seismic  studies specially testing the forward ground motion simulation modeling (give some references including Tabarda et al 2016).
+Occurrence of frequent moderate magnitude earthquakes made the Los Angeles basin as a natural laboratory for seismic  studies specially testing the forward ground motion simulation modeling.
 In a relatively recent study, \citet{Taborda_2014_BSSA} conducted a comprehensive study in order to assess the functionality of developed velocity models in the Los Angeles basin and surrounding areas.
 They simulated the 2008 Chino Hills earthquake using 3 velocity models including: CVM-S4, CVM-H, and CVM-H+GTL (for more details about these models see \citet{Taborda_2014_BSSA}.)
-They computed the GOF scores based on \citet{Anderson_2004_Proc}, with minor modification introduced by \citet{Taborda_2013_BSSA} as discussed in the validation section.Fig.~\ref{fig:epicenters} shows the location of the epicenter and stations' locations.
+They computed the GOF scores based on \citet{Anderson_2004_Proc}, with minor modification introduced by \citet{Taborda_2013_BSSA} as discussed in the validation section.
+Fig.~\ref{fig:epicenters} shows the location of the epicenter and seismic stations.
 
@@
         {figures/pdf/Figure_1}
-    \caption{(a) Region of interest and epicentral location of the 2008 Chino Hills earthquake (highlighted).
-Focal mechanisms are shown at the margins of the region of interest.
+    \caption{(a) Region of interest and epicentral location of the 2008 $M_W ~ 5.4$ Chino Hills earthquake (highlighted).
+Focal mechanism is shown at the margin of the region of interest.
 Major quaternary faults in the area are shown in the back along with the main roads and county divisions.
@@
 Dots indicate the location of the 336 stations considered in this study for validation.
-A subset of these stations include labels and highlights with larger symbols for later reference.
 The main local and interstate roads are also shown here.
 In both panels, the background shows the hillshade topography of the region.
-The color version of this figure is available only in the electronic edition.
-(figure and caption from 2014 paper.
-needs revision)}
+The color version of this figure is available only in the electronic edition.}
     \label{fig:epicenters}
@@
 In this study we only analyze the results for broadband which in this case is \feq{0.1}{4}.
-Fig.~\ref{fig:data_box_plot} shows the box plot of all data that we used in this study which we separate it in velocity models and components.
+Fig.~\ref{fig:data_box_plot} shows the box plot of all data that we used in this study which we separate them in velocity models and components.
+In total we use $336 \times 3 \times 3 = 3024$ individual pair of data and synthetic.
 
@@
 
-As we explained before, we use all data set regardless of velocity model and components.
-However, distinguishing data for velocity model and components could answer the question that if the goodness of fit scores dependent on them.
+As we explained before, we use all data set regardless of velocity models and components.
+However, distinguishing data for velocity model and components can answer the question that if the goodness of fit scores dependent on them.
 Addressing this research question is beyond the scope of this paper.
@@
 
-Fig.~\ref{fig:data_density}  shows the density distribution of each score for CVM-S4.
+Fig.~\ref{fig:data_density}  shows the density distribution of each score for CVM-S.
+The figure shows the variation of different metrics with components.
+Except in cross correlation (C10) in all metrics the up-down (UD) component scores are about the same or higher than horizontal components.
 
@@
         {figures/pdf/Figure_3.pdf}
-    \caption{Distribution of data (only CVMS-4) for different components in different scores based on different components. Each distribution has a unit area.}
+    \caption{Distribution of data (only CVM-S) for different components in different scores based on different components. Each distribution has a unit area.}
     \label{fig:data_density}

### 2017-09-01 introduction.tex
Figures are pulished, two figures are added, and the manuscript is improved.

@@
 
-The validation of ground motion synthetics has received increased attention over the last few years due to advances in physics based deterministic and hybrid simulation methods.
+The validation of ground motion synthetics has received increased attention over the last few years due to the advances in physics based deterministic and hybrid simulation methods.
 Unless in very low frequencies (\fleq{0.5}), due to limitations and uncertainties in the ground motion simulation and velocity models, it is not possible to match the simulation and synthetics wiggle by wiggle.
@@
 In general, these metrics compute the goodness-of-fit (GOF) between data and synthetic in time and frequency domains.
-\citet{Anderson_2004_Proc} proposed 10 metrics to be used to characterized the GOF in different frequency bands.
+\citet{Anderson_2004_Proc} proposed 10 metrics to be used to characterize the GOF in different frequency bands.
 \citet{Kristekova_2006_BSSA,Kristekova_2009_GJI} developed envelope and phase misfit criteria as well as GOF metrics for realistic scenarios.
 \citet{Olsen_2010_SRL} presented an alternative GOF metrics for broadband synthetics with relatively high resolution of the small misfits, and \citet{Taborda_2013_BSSA} added modification to \citet{Anderson_2004_Proc} metrics.\\
-Recently, due to advances in computational resources, high frequency 3D ground motion simulation on a regional scale were conducted in different regions using different methods (give some references.) The accuracy of simulation with observed data are measured using these GOF scores.
+These metrics have been widely used in different studies.
+Recently, due to advances in computational resources, 3D ground motion simulation on a regional scale were conducted in different regions using different methods \citep[e.g.,][]{Taborda_2016_GJI,Komatitsch_2004_BSSA,Bielak_2010_GJI,Roten_2016_GJI}.
+The accuracy of simulation with observed data are measured using these GOF scores.
 However, researchers use these metrics in different perspectives which is highly dependent on their application.
 For example, for structural engineers who are interested in designing tall buildings, among other metrics, displacement could be an important factor.
-However, for rigid and semi-rigid structures like dams and nuclear power plants, acceleration could be more important.
+However, for rigid and semi-rigid structures like dams and nuclear power plants, acceleration could be more important criteria.
 In general, for all purposes, the duration of strong ground morion could be an important factor.
 Analysis of the strong ground motion simulation for a specific application based on individual metrics is an acceptable practice, however, validating the simulation process only with individual score is not a valid approach.
-Our previous studies show that metrics could have completely different scores for a pair of synthetic and data \citep[e.g.,][]{Taborda_2016_GJI}.
-\citet{Taborda_2013_BSSA} defined a metric from a combination of  \citet{Anderson_2004_Proc} scores representing a general overview of GOF.
-However, the metric is a linear combination and somehow is biased by the authors preferences about the metrics.
+Previous studies show that metrics could have completely different scores for a pair of synthetic and data \citep[e.g.,][]{Taborda_2016_GJI}.
+\citet{Taborda_2013_BSSA} defined a metric from a combination of  \citet{Anderson_2004_Proc} scores presenting a general overview of GOF.
+However, the metric is a linear combination and somehow is biased by the authors' preferences about the metrics.
 \\
-Physics-based ground motion simulation community push the maximum frequency  of the simulation to higher values (\fmax{}$\geq 5 ~ Hz$) and there is a high demand for a metric to uniformly validate the simulations (refer to the high-f project).
-Although \citet{Anderson_2004_Proc}, defined different level for classifying the stations as poor, fair, good, and excellent based on individual metrics, the question becomes what if there is a station with completely different scores for different metrics.
+Physics-based ground motion simulation communities push the maximum frequency  of the simulation to higher values (\fmax{}$\geq 5 ~ Hz$) and there is a high demand for a metric to uniformly validate the simulations.
+Although \citet{Anderson_2004_Proc}, defined different level for classifying the stations as poor, fair, good, and excellent based on individual metrics, there is no consensus on the most important metrics, specially when they have significantly different GOF score.
+Therefore, the question "How one can estimate the accuracy of simulation based on all scores or which metrics can be good representatives of simulation accuracy on behalf of other metrics?" has not addressed.
 This research question motivated us to approach the problem from different perspective.
-In other words, if we assume we know the accuracy of the simulation process (let's say we classified it as good) what is the relationship between different GOF scores for that pair of stations.
+In other words, if we assume we know the accuracy of the simulation process (let's say we classified it as good) what is the relationship between different GOF scores for that pair of stations or which metrics are the best representer in classifying the stations.
 However, as far as the authors knowledge, there is no metric to be able to classify a pair of stations based on GOF scores as an estimation of the overall accuracy of the simulation.
-Therefore, as a first step, if we could label the pair of station for simulation accuracy we can discuss the relationships between GOF scores.
-To do this we need a good amount of dataset where calculated the GOF scores for a good amount of stations.
+Therefore, as a first step, if we label the pair of station for simulation accuracy we can discuss the relationships between GOF scores.
+To do this we need a dataset where involves GOF scores for a good amount of stations.
 \\
@@
 The study generated an invaluable synthetic and data  GOF scores dataset.\\
-The first and the most common method to categorizing multidimensional unlabeled data is clustering and specifically the distance based methods like \kmeans{} \citep{Macqueen_1967_Proc}.
+The first and the most common method in categorizing multidimensional unlabeled data is clustering and specifically the distance based methods like \kmeans{} \citep{Macqueen_1967_Proc}.
 However, there is a drawback in this method.
 If the data is not explicitly distinguishable the final clusters are highly dependent on initial starting points of the algorithm.
-Therefor, for the use of stations with 11 dimensions, the ordinary \kmeans{} approach results in fairly different clusters after each analysis.
-The main reason for this issue is the algorithm is solely looking for data which are close together in terms of Euclidian distance.
+Therefor, for stations with 11 dimensions, the ordinary \kmeans{} approach results in fairly different clusters after each analysis.
+The main reason for this issue is the algorithm solely looks for data which are close together in terms of Euclidean distance.
 However, in our case we have one more extra information.
 Hypothetically, if we have a pair of synthetic and data with GOF score equal 3 for all metrics, according to \citet{Anderson_2004_Proc} we can classify the station as poor with a strong confidence.
-This is also true for GOF score of 5,7, and 9 for fair, good, and excellent classes, respectively.
+This is also true for GOF score of 5, 7, and 9 for fair, good, and excellent classes, respectively.
 This background knowledge about the database could give a strong formation to the clustering process.
 \citet{Wagstaff_2001_Proc} defined a \kmeans{} clustering using background knowledge.
-Adding this feature to the algorithm improves the functionality and most importantly converge to the same clusters with any random initial cluster's centers.
-Using this approach, first we cluster stations then we statistically analyze each cluster data, based on these results we remove features with different trend from the analysis.
-Later on, in a supervised learning algorithm, we develop a classification algorithm based on their GOF scores.\\
+Adding this feature to the algorithm improves the functionality and most importantly converge to the same clusters with any random initial clusters' centers.
+Using this approach alongside the subspace analysis (more details are presented in the section.~\ref{clustering}) we cluster the stations, then in a supervised learning algorithm, we develop a classification algorithm based on their GOF scores.
+We tune the learning algorithm and present the functionality of each prediction model based on different metrics.
+It is possible to have different classes for different components of a pair if synthetic and data.
+We study the different combination of components' classes in different distances and propose a rough estimate on deciding about the final GOF class of a pair of synthetic and data which is application and component independent.\\
 In the following section,we present a summary of the validation metrics then we discuss the dataset.
 Later on we explain the methodology.
-We provide a brief summary of unsupervised learning which is ordinary and constrained clustering approaches.
+We provide a brief summary of unsupervised  and semi-supervised learning which are ordinary and constrained clustering approaches, respectively.
 We also discuss the basics of decision tree algorithm as a supervised learning method.
-We discuss the clustering challenges and use subspace analysis.
-We develop two decision tree classifiers based on labeled data for future use in comparing data and synthetics.
-We conclude the study by applying the generated classification methods on \citet{Taborda_2014_BSSA} dataset.
+We discuss the clustering challenges and use subspace analysis in the clustering analysis section.
+We also discuss the imbalanced data effects and solutions in the classification model section.
+We develop two decision tree classifiers (prediction model) based on labeled data for future use in comparing data and synthetics.
+We propose an algorithm to give a rough estimate to have component independent prediction.
+We conclude the study by applying the generated classification methods on \citet{Taborda_2014_BSSA} dataset and present the spatial variation of the simulation goodness of fit classes on the Los Angles basin for three velocity models and three components in component dependent and for three velocity models in component independent approaches.

### 2017-09-01 methodology.tex
Figures are pulished, two figures are added, and the manuscript is improved.

@@
 
-As we discus in section.\ref{validation_metrics}, there is no consensus on a single metric who able to estimate the accuracy of the simulation or could classify the simulation based on given GOF scores.
+As we discus in section.\ref{validation_metrics}, there is no consensus on a single metric who able to estimate the accuracy of the simulation or can classify the simulation based on given GOF scores.
 On the other hand we have a comprehensive dataset.
@@
 Generating labels for data is categorized as unsupervised learning or clustering.
-Different methods have been developed for clustering.
-The most commonly used method is \kmeans{}.
-However, \kmeans{} method gives different results based on initial values of cluster center especially where data is not clearly distinguishable.
-In result we use modified \kmeans{} approach.
-This method use background knowledge in the unsupervised learning process.
-Using background knowledge in unsupervised learning converts it into semi-supervised learning process.\\
-In summary, first we label the data based on the constrained \kmeans{} approach, then we statistically analyze the each group results.
-We expect to see similarity among attributes within each classes.
-Later on, in a classification process, using decision tree we find the best hypothesis to classify a station based on GOF scores.
-In the rest of this section first we discuss the clustering method for both ordinary and constrained \kmeans{} approach then we discuss the decision tree algorithm and basics in separating data.
-We discuss the statistical analysis of the results in the results section.
+Different methods are developed for clustering.
+The most commonly used method is \kmeans{}, however, it gives different results based on initial values of clusters' centers especially in the case that data is not clearly distinguishable.
+To overcome this drawback, we use modified \kmeans{} approach.
+In this method one can use background knowledge in the unsupervised learning process.
+Using background knowledge in unsupervised learning converts it into a semi-supervised learning process.\\
+In summary, first we label the data based on the constrained \kmeans{} approach, then, in a classification process, using decision tree we search for the best hypothesis to classify a station based on GOF scores.
+In the rest of this section first we discuss the clustering method for both ordinary and constrained \kmeans{} approach then we discuss the decision tree algorithm and basics in classifying data.
+We discuss the application results in the clustering analysis and classification model sections.
 \\
 
-\subsection{Clustring}
 \input{clustering}
-\subsection{Decision Tree}
 \input{Dtree}

### 2017-09-29 abstract.tex
completed abstract review, now working on introduction

--- 2017-08-20 abstract.tex
@@
 % 
-The validation of ground motion synthetics has received increased attention over the last few years due to the advances in physics-based deterministic and hybrid simulation methods.
+It has become common practice to validate ground motion simulation based on a variety of time- and frequency-domain metrics known in seismology and engineering and scaled to quantify the level of agreement between synthetics and data or other reference solutions.
+There is, however, no agreement about the importance or weight that it ought to be given to each individual metric.
+This leads to their selection often being subjective, either based on intended applications or personal preferences.
+In turn, this makes it difficult for simulators to identify modeling improvement needs, which could be easier if they could focused on a reduced number of alternative metrics.
+We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques used to label and classify  goodness-of-fit (GOF) results, with the objective prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation.
+In particular, we study the relationships that exist between eleven different metrics when understood as part of a multi-dimensional space using a constrained \kmeans{} method.
+We conduct a subspace clustering analysis to address the high-dimensional effects and this allows us to label the results from a large validation dataset into four categories (poor, fair, good, and excellent).
+We then develop two decision trees using a C5.0 algorithm.
+These decision trees help narrow the number of metrics leading to a validation prediction into the four categories.
+These decision trees can be understood as rapid predictors of the quality of a simulation, or as a data-informed classifiers that can help prioritize validation metrics.
+Our analysis indicates that among the eleven metrics considered here, the acceleration response spectra and total energy of velocity are the most dominant metrics ones, followed by the peak ground response in terms of velocity and acceleration.
+
+% OLD NAEEM VERSION
+% 
+% The validation of ground motion synthetics has received increased attention over the last few years due to the advances in physics-based deterministic and hybrid simulation methods.
 Unlike for low frequency simulations $(f \le 0.5 Hz)$, for which it has become reasonable to expect a good match between synthetics and data, in the case of high-frequency simulations $(1 Hz \le f)$ it is not possible to match results on a wiggle-by-wiggle basis.
 This is mostly due to the various complexities and uncertainties involved in earthquake ground motion modeling.
-Therefore, in order to compare synthetics with data we turn to different time series metrics, which are used as a means to characterize how the synthetics match the data on qualitative and statistical sense.
+Therefore, in order to compare synthetics with data we turn to different time series metrics, which are used as means to characterize how the synthetics match the data on qualitative and statistical sense.
 In general, these metrics provide GOF scores that measure the level of similarity in the time and frequency domains.
@@
 Although using individual metrics for particular applications is considered more adequate, there is no consensus or a unified method to classify the comparison between a set of synthetic and recorded seismograms when the various metrics offer different scores.
-We study the relationship among these metrics through
-semi-supervised and supervised learning approaches in order to labeling and classifying data, respectively.
-In particular we use constrained \kmeans{} clustering method, in which we define 4 hypothetical stations with scores 3, 5, 7, and 9 for all metrics.
+We study the relationship among these metrics through semi-supervised and supervised learning approaches in order to labeling and classifying data, respectively.
+In particular, we use constrained \kmeans{} clustering method, in which we define 4 hypothetical stations with scores 3, 5, 7, and 9 for all metrics.
 We put these stations in the category of cannot-link constraints.

### 2017-09-29 introduction.tex
completed abstract review, now working on introduction

--- 2017-08-03 introduction.tex
@@
-\section{Introduction}
+\introduction
 
-The validation of ground motion synthetics has received increased attention over the last few years due to advances in physics based deterministic and hybrid simulation methods.
+The verification and validation of ground motion synthetics has received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation.
+Various methods or schemes have been proposed to evaluate, quantitatively, the misfit or goodness-of-fit (GOF) between simulation synthetics and recorded data or reference alternative solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
+Among these, there are methods better suited for verification purposes, because they target the comparisons at a waveform level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others evaluate the comparisons through proxies at the structural response level \citep{Burks_BSSA_2014}.
+The remaining ones, on the other hand, target specific characteristics in the seismograms themselves in both time and frequency, and are thus better suited for high-frequency or broadband simulations, where the expectation of matching seismograms at the waveform level are less relevant, unless dealing with low-frequency (\fmaxleq{1}).
+From this group, \citep{Anderson_2004_Proc, Olsen_2010_SRL} are preferred because they offer a simple scoring scheme to normalize the GOF scores associated with different metrics on a common scale.
+
+Out of the various validation methods available, \citet{Anderson_2004_Proc} is perhaps the most widely used \citep[e.g.,]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
+We, in particular, have consistently relied on a modified version of this method for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake, and for a larger suite of moderate-magnitude events in southern California \citep{Taborda_2013_BSSA, Taborda_2014_BSSA, Taborda_2016_GJI}.
+In essence, Anderson's method evaluates the similarity in terms of 10 metrics based on the cumulative strength of the motion (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), the peak responses (in acceleration, velocity and displacement), the signals' frequency content (in the form of Fourier and response spectra), and their synchronization (through a cross correlation measure).
+While the strength
+
+However, the existence of various alternative methods, including those inspired by or modified after Anderson's criteria, suggest the need to better understand what sort of parameters ought to be taken into account to appropriately classify the validity of a simulation.
+
+Both \citet{Taborda_2013_BSSA} and \citet{Maufroy_2015_BSSA}, for instance, independently suggested the need to add
+
+, and have more recently used it
+
+% OLD NAEEM VERSION
+
+% The validation of ground motion synthetics has received increased attention over the last few years due to the advances in physics based deterministic and hybrid simulation methods.
 Unless in very low frequencies (\fleq{0.5}), due to limitations and uncertainties in the ground motion simulation and velocity models, it is not possible to match the simulation and synthetics wiggle by wiggle.
@@
 In general, these metrics compute the goodness-of-fit (GOF) between data and synthetic in time and frequency domains.
-\citet{Anderson_2004_Proc} proposed 10 metrics to be used to characterized the GOF in different frequency bands.
+\citet{Anderson_2004_Proc} proposed 10 metrics to be used to characterize the GOF in different frequency bands.
 \citet{Kristekova_2006_BSSA,Kristekova_2009_GJI} developed envelope and phase misfit criteria as well as GOF metrics for realistic scenarios.
-\citet{Olsen_2010_SRL} presented an alternative GOF metrics for broadband synthetics with relatively high resolution of the small misfits, and \citet{Taborda_2013_BSSA} added modification to \citet{Anderson_2004_Proc} metrics.\\
-Recently, due to advances in computational resources, high frequency 3D ground motion simulation on a regional scale were conducted in different regions using different methods (give some references.) The accuracy of simulation with observed data are measured using these GOF scores.
+\citet{Olsen_2010_SRL} presented an alternative GOF metrics for broadband synthetics with relatively high resolution of the small misfits, and \citet{Taborda_2013_BSSA} added modification to \citet{Anderson_2004_Proc} metrics.
+
+% These metrics have been widely used in different studies.
+Recently, due to advances in computational resources, 3D ground motion simulation on a regional scale were conducted in different regions using different methods \citep[e.g.,][]{Taborda_2016_GJI,Komatitsch_2004_BSSA,Bielak_2010_GJI,Roten_2016_GJI}.
+The accuracy of simulation with observed data are measured using these GOF scores.
 However, researchers use these metrics in different perspectives which is highly dependent on their application.
 For example, for structural engineers who are interested in designing tall buildings, among other metrics, displacement could be an important factor.
-However, for rigid and semi-rigid structures like dams and nuclear power plants, acceleration could be more important.
+However, for rigid and semi-rigid structures like dams and nuclear power plants, acceleration could be more important criteria.
 In general, for all purposes, the duration of strong ground morion could be an important factor.
 Analysis of the strong ground motion simulation for a specific application based on individual metrics is an acceptable practice, however, validating the simulation process only with individual score is not a valid approach.
-Our previous studies show that metrics could have completely different scores for a pair of synthetic and data \citep[e.g.,][]{Taborda_2016_GJI}.
-\citet{Taborda_2013_BSSA} defined a metric from a combination of  \citet{Anderson_2004_Proc} scores representing a general overview of GOF.
-However, the metric is a linear combination and somehow is biased by the authors preferences about the metrics.
-\\
-Physics-based ground motion simulation community push the maximum frequency  of the simulation to higher values (\fmax{}$\geq 5 ~ Hz$) and there is a high demand for a metric to uniformly validate the simulations (refer to the high-f project).
-Although \citet{Anderson_2004_Proc}, defined different level for classifying the stations as poor, fair, good, and excellent based on individual metrics, the question becomes what if there is a station with completely different scores for different metrics.
+Previous studies show that metrics could have completely different scores for a pair of synthetic and data \citep[e.g.,][]{Taborda_2016_GJI}.
+\citet{Taborda_2013_BSSA} defined a metric from a combination of  \citet{Anderson_2004_Proc} scores presenting a general overview of GOF.
+However, the metric is a linear combination and somehow is biased by the authors' preferences about the metrics.
+
+% Physics-based ground motion simulation communities push the maximum frequency  of the simulation to higher values (\fmax{}$\geq 5 ~ Hz$) and there is a high demand for a metric to uniformly validate the simulations.
+Although \citet{Anderson_2004_Proc}, defined different level for classifying the stations as poor, fair, good, and excellent based on individual metrics, there is no consensus on the most important metrics, specially when they have significantly different GOF score.
+Therefore, the question "How one can estimate the accuracy of simulation based on all scores or which metrics can be good representatives of simulation accuracy on behalf of other metrics?" has not addressed.
 This research question motivated us to approach the problem from different perspective.
-In other words, if we assume we know the accuracy of the simulation process (let's say we classified it as good) what is the relationship between different GOF scores for that pair of stations.
+In other words, if we assume we know the accuracy of the simulation process (let's say we classified it as good) what is the relationship between different GOF scores for that pair of stations or which metrics are the best representer in classifying the stations.
 However, as far as the authors knowledge, there is no metric to be able to classify a pair of stations based on GOF scores as an estimation of the overall accuracy of the simulation.
-Therefore, as a first step, if we could label the pair of station for simulation accuracy we can discuss the relationships between GOF scores.
-To do this we need a good amount of dataset where calculated the GOF scores for a good amount of stations.
-\\
- \citet{Taborda_2014_BSSA} conducted a comprehensive study to analyze the functionality of different velocity models in the Los Angeles basin.
+Therefore, as a first step, if we label the pair of station for simulation accuracy we can discuss the relationships between GOF scores.
+To do this we need a dataset where involves GOF scores for a good amount of stations.
+
+% \citet{Taborda_2014_BSSA} conducted a comprehensive study to analyze the functionality of different velocity models in the Los Angeles basin.
 They used 2008 Chino Hills earthquakes based on 3 velocity models and conducted simulations with maximum frequency of \fmaxeq{4}.
@@
 The study generated an invaluable synthetic and data  GOF scores dataset.\\
-The first and the most common method to categorizing multidimensional unlabeled data is clustering and specifically the distance based methods like \kmeans{} \citep{Macqueen_1967_Proc}.
+% The first and the most common method in categorizing multidimensional unlabeled data is clustering and specifically the distance based methods like \kmeans{} \citep{Macqueen_1967_Proc}.
 However, there is a drawback in this method.
 If the data is not explicitly distinguishable the final clusters are highly dependent on initial starting points of the algorithm.
-Therefor, for the use of stations with 11 dimensions, the ordinary \kmeans{} approach results in fairly different clusters after each analysis.
-The main reason for this issue is the algorithm is solely looking for data which are close together in terms of Euclidian distance.
+Therefor, for stations with 11 dimensions, the ordinary \kmeans{} approach results in fairly different clusters after each analysis.
+The main reason for this issue is the algorithm solely looks for data which are close together in terms of Euclidean distance.
 However, in our case we have one more extra information.
 Hypothetically, if we have a pair of synthetic and data with GOF score equal 3 for all metrics, according to \citet{Anderson_2004_Proc} we can classify the station as poor with a strong confidence.
-This is also true for GOF score of 5,7, and 9 for fair, good, and excellent classes, respectively.
+This is also true for GOF score of 5, 7, and 9 for fair, good, and excellent classes, respectively.
 This background knowledge about the database could give a strong formation to the clustering process.
 \citet{Wagstaff_2001_Proc} defined a \kmeans{} clustering using background knowledge.
-Adding this feature to the algorithm improves the functionality and most importantly converge to the same clusters with any random initial cluster's centers.
-Using this approach, first we cluster stations then we statistically analyze each cluster data, based on these results we remove features with different trend from the analysis.
-Later on, in a supervised learning algorithm, we develop a classification algorithm based on their GOF scores.\\
-In the following section,we present a summary of the validation metrics then we discuss the dataset.
+Adding this feature to the algorithm improves the functionality and most importantly converge to the same clusters with any random initial clusters' centers.
+Using this approach alongside the subspace analysis (more details are presented in the section.~\ref{clustering}) we cluster the stations, then in a supervised learning algorithm, we develop a classification algorithm based on their GOF scores.
+We tune the learning algorithm and present the functionality of each prediction model based on different metrics.
+It is possible to have different classes for different components of a pair if synthetic and data.
+We study the different combination of components' classes in different distances and propose a rough estimate on deciding about the final GOF class of a pair of synthetic and data which is application and component independent.\\
+% In the following section,we present a summary of the validation metrics then we discuss the dataset.
 Later on we explain the methodology.
-We provide a brief summary of unsupervised learning which is ordinary and constrained clustering approaches.
+We provide a brief summary of unsupervised  and semi-supervised learning which are ordinary and constrained clustering approaches, respectively.
 We also discuss the basics of decision tree algorithm as a supervised learning method.
-We discuss the clustering challenges and use subspace analysis.
-We develop two decision tree classifiers based on labeled data for future use in comparing data and synthetics.
-We conclude the study by applying the generated classification methods on \citet{Taborda_2014_BSSA} dataset.
+We discuss the clustering challenges and use subspace analysis in the clustering analysis section.
+We also discuss the imbalanced data effects and solutions in the classification model section.
+We develop two decision tree classifiers (prediction model) based on labeled data for future use in comparing data and synthetics.
+We propose an algorithm to give a rough estimate to have component independent prediction.
+We conclude the study by applying the generated classification methods on \citet{Taborda_2014_BSSA} dataset and present the spatial variation of the simulation goodness of fit classes on the Los Angles basin for three velocity models and three components in component dependent and for three velocity models in component independent approaches.

### 2017-10-02 introduction.tex
updates

@@
 
-The verification and validation of ground motion synthetics has received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation.
+Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation.
 Various methods or schemes have been proposed to evaluate, quantitatively, the misfit or goodness-of-fit (GOF) between simulation synthetics and recorded data or reference alternative solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
@@
 Out of the various validation methods available, \citet{Anderson_2004_Proc} is perhaps the most widely used \citep[e.g.,]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
-We, in particular, have consistently relied on a modified version of this method for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake, and for a larger suite of moderate-magnitude events in southern California \citep{Taborda_2013_BSSA, Taborda_2014_BSSA, Taborda_2016_GJI}.
-In essence, Anderson's method evaluates the similarity in terms of 10 metrics based on the cumulative strength of the motion (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), the peak responses (in acceleration, velocity and displacement), the signals' frequency content (in the form of Fourier and response spectra), and their synchronization (through a cross correlation measure).
-While the strength
+We, in particular, have consistently relied on a modified version of this method for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake, and for the evaluation of velocity models in southern California through the validation of a large suite of moderate-magnitude events \citep{Taborda_2013_BSSA, Taborda_2014_BSSA, Taborda_2016_GJI}.
+In essence, Anderson's method assess the similarity of two signals based on the average scores of 10 metrics.
+These metrics measure the cumulative strength of ground motion signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), the peak response (in acceleration, velocity and displacement), the signals' frequency content (in the form of Fourier and response spectra), and their synchronization (through cross-correlation).
+
+While the strength of method of \citet{Anderson_2004_Proc} is that its metrics are well understood in engineering and seismology, others have pointed out that some of these metrics are redundant or incomplete.
+Both \citet{Taborda_2013_BSSA} and \citet{Maufroy_2015_BSSA}, for instance, added a
+
+To remedy
 
@@
 
-Both \citet{Taborda_2013_BSSA} and \citet{Maufroy_2015_BSSA}, for instance, independently suggested the need to add
+Both , for instance, independently suggested the need to add
 

### 2017-10-04 abstract.tex
completed first revision of introduction

@@
 These decision trees can be understood as rapid predictors of the quality of a simulation, or as a data-informed classifiers that can help prioritize validation metrics.
-Our analysis indicates that among the eleven metrics considered here, the acceleration response spectra and total energy of velocity are the most dominant metrics ones, followed by the peak ground response in terms of velocity and acceleration.
+Our analysis indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.
 
+% ===========================================================================================
+%
 % OLD NAEEM VERSION

### 2017-10-04 introduction.tex
completed first revision of introduction

@@
 Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation.
-Various methods or schemes have been proposed to evaluate, quantitatively, the misfit or goodness-of-fit (GOF) between simulation synthetics and recorded data or reference alternative solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
-Among these, there are methods better suited for verification purposes, because they target the comparisons at a waveform level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others evaluate the comparisons through proxies at the structural response level \citep{Burks_BSSA_2014}.
-The remaining ones, on the other hand, target specific characteristics in the seismograms themselves in both time and frequency, and are thus better suited for high-frequency or broadband simulations, where the expectation of matching seismograms at the waveform level are less relevant, unless dealing with low-frequency (\fmaxleq{1}).
-From this group, \citep{Anderson_2004_Proc, Olsen_2010_SRL} are preferred because they offer a simple scoring scheme to normalize the GOF scores associated with different metrics on a common scale.
+Various methods or schemes have been proposed to evaluate, through direct quantitative comparisons or indirect statistical analysis, the similarity between simulation synthetics and recorded data or other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
+Among these methods, there are some that are better suited for verification purposes because they compare the signals at a waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for application validation purposes because they compare the behavior of systems (e.g., buildings or earth structures) when subjected to alternative input motions through proxies at the response level \citep{Burks_BSSA_2014}.
+The remaining methods, on the other hand, focus on the overall characteristics of seismograms, in both time and frequency, and are thus better suited for validating high-frequency or broadband simulations, where applications may be undefined and the expectation of matching seismograms at the waveform level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}).
+From this latter group, methods such as those of \citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL} are preferred because they offer simple schemes to quantify the goodness-of-fit (GOF) of synthetics with respect to data.
 
-Out of the various validation methods available, \citet{Anderson_2004_Proc} is perhaps the most widely used \citep[e.g.,]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
-We, in particular, have consistently relied on a modified version of this method for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake, and for the evaluation of velocity models in southern California through the validation of a large suite of moderate-magnitude events \citep{Taborda_2013_BSSA, Taborda_2014_BSSA, Taborda_2016_GJI}.
+Out of the various validation methods just mentioned, \citet{Anderson_2004_Proc} is perhaps the most widely used today \citep[e.g.,]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
+We, in particular, have consistently relied on a modified version of this method for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake, and for the evaluation of velocity models in southern California through the validation of a large suite of moderate-magnitude earthquake simulations \citep{Taborda_2013_BSSA, Taborda_2014_BSSA, Taborda_2016_GJI}.
 In essence, Anderson's method assess the similarity of two signals based on the average scores of 10 metrics.
-These metrics measure the cumulative strength of ground motion signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), the peak response (in acceleration, velocity and displacement), the signals' frequency content (in the form of Fourier and response spectra), and their synchronization (through cross-correlation).
+These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), the peak values (in acceleration, velocity and displacement), the signals' frequency content (in the form of Fourier and response spectra), and their synchronization (through cross-correlation).
 
-While the strength of method of \citet{Anderson_2004_Proc} is that its metrics are well understood in engineering and seismology, others have pointed out that some of these metrics are redundant or incomplete.
-Both \citet{Taborda_2013_BSSA} and \citet{Maufroy_2015_BSSA}, for instance, added a
+The strength of method proposed by \citet{Anderson_2004_Proc} is that its metrics are well understood in the engineering and seismology contexts.
+However, it has been pointed out that some of these metrics are redundant, or that other relevant metrics need to be added.
+\citet{Taborda_2013_BSSA}, for instance, included the strong motion duration \citep{Trifunac_1975_BSSA} as an additional metric, and averaged the Arias intensity- and energy integral-related scores to avoid duplication.
+In the same spirit, \citet{Maufroy_2015_BSSA} reduced the number of metrics limiting them to only those metrics with comparable units.
+Unfortunately, none of these alternatives address the underlying questions: what are the parameters that ought to be taken into account when validating ground motion simulations, and what level of priority should these be given.
 
-To remedy
+Lack of consensus about how to answer the above question, makes the choice of validation methods and the selection of the comparison metrics to be a subjective one, based mostly on personal preferences and---arguably---on the application intended for the simulation.
+In turn, the current multiplicity of metrics, if all are given the same weight in the analysis of validation results, makes it difficult for simulators to identify the sort of changes needed in their models that could lead to better ground motion predictions.
+We argue that this situation can be corrected if we can offer data-informed arguments they can help simulators focus on a reduced number of alternative metrics.
 
-However, the existence of various alternative methods, including those inspired by or modified after Anderson's criteria, suggest the need to better understand what sort of parameters ought to be taken into account to appropriately classify the validity of a simulation.
+To that end, we study the relationships that exist between eleven different metrics---those proposed by \citet{Anderson_2004_Proc}, plus strong motion duration---with the objective of offering a prioritized reduced number of metrics that can help predict validation results.
+We treat these metrics as independent variables defining a multi-dimensional space, and analyze them using machine learning techniques.
+We use a constrained \kmeans{} method and conduct a subspace clustering analysis to address high-dimensional effects in order to label the results from a large validation dataset into four categories (poor, fair, good, and excellent).
+The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}.
+Upon labeling the dataset, we develop two decision trees using a C5.0 algorithm to prioritize the metrics involved in the validation process.
+The decision trees narrow the number of metrics offering a prediction algorithm into the aforementioned validation categories.
+In summary, our results indicate that among the eleven metrics considered in the analysis, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.
+We test our prediction model and offer a discussion about its implications and potential use in future validation efforts.
 
-Both , for instance, independently suggested the need to add
-
-, and have more recently used it
-
+% ===========================================================================================
+%
 % OLD NAEEM VERSION
-
+% 
 % The validation of ground motion synthetics has received increased attention over the last few years due to the advances in physics based deterministic and hybrid simulation methods.

### 2017-10-04 introduction.tex
updates

@@
 Various methods or schemes have been proposed to evaluate, through direct quantitative comparisons or indirect statistical analysis, the similarity between simulation synthetics and recorded data or other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
-Among these methods, there are some that are better suited for verification purposes because they compare the signals at a waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for application validation purposes because they compare the behavior of systems (e.g., buildings or earth structures) when subjected to alternative input motions through proxies at the response level \citep{Burks_BSSA_2014}.
+Among these methods, there are some that are better suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for application validation purposes because they compare the behavior of systems (e.g., buildings or earth structures) when subjected to alternative input motions at the response level \citep{Burks_BSSA_2014}.
 The remaining methods, on the other hand, focus on the overall characteristics of seismograms, in both time and frequency, and are thus better suited for validating high-frequency or broadband simulations, where applications may be undefined and the expectation of matching seismograms at the waveform level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}).
@@
 
-Out of the various validation methods just mentioned, \citet{Anderson_2004_Proc} is perhaps the most widely used today \citep[e.g.,]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
+Out of the various validation methods available, \citet{Anderson_2004_Proc} is perhaps the most widely used today \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
 We, in particular, have consistently relied on a modified version of this method for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake, and for the evaluation of velocity models in southern California through the validation of a large suite of moderate-magnitude earthquake simulations \citep{Taborda_2013_BSSA, Taborda_2014_BSSA, Taborda_2016_GJI}.
-In essence, Anderson's method assess the similarity of two signals based on the average scores of 10 metrics.
-These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), the peak values (in acceleration, velocity and displacement), the signals' frequency content (in the form of Fourier and response spectra), and their synchronization (through cross-correlation).
 
-The strength of method proposed by \citet{Anderson_2004_Proc} is that its metrics are well understood in the engineering and seismology contexts.
+In essence, Anderson's method assess the similarity of two signals based on the average score of 10 metrics.
+These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), their peak values (in acceleration, velocity and displacement), frequency content (in terms of Fourier and response spectra), and their time synchronization (through cross-correlation).
+
+The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics are well understood in the engineering and seismology contexts.
 However, it has been pointed out that some of these metrics are redundant, or that other relevant metrics need to be added.
 \citet{Taborda_2013_BSSA}, for instance, included the strong motion duration \citep{Trifunac_1975_BSSA} as an additional metric, and averaged the Arias intensity- and energy integral-related scores to avoid duplication.
-In the same spirit, \citet{Maufroy_2015_BSSA} reduced the number of metrics limiting them to only those metrics with comparable units.
+In the same spirit, \citet{Maufroy_2015_BSSA} reduced the number of metrics limiting them to only those with comparable units.
 Unfortunately, none of these alternatives address the underlying questions: what are the parameters that ought to be taken into account when validating ground motion simulations, and what level of priority should these be given.
 
-Lack of consensus about how to answer the above question, makes the choice of validation methods and the selection of the comparison metrics to be a subjective one, based mostly on personal preferences and---arguably---on the application intended for the simulation.
+Lack of consensus about how to answer these questions makes the choice of validation methods and the selection of the comparison metrics to be a subjective one, mostly based on personal preferences and---arguably---on the application intended for the simulation.
 In turn, the current multiplicity of metrics, if all are given the same weight in the analysis of validation results, makes it difficult for simulators to identify the sort of changes needed in their models that could lead to better ground motion predictions.
-We argue that this situation can be corrected if we can offer data-informed arguments they can help simulators focus on a reduced number of alternative metrics.
+We think that this situation can be corrected if we can offer data-informed arguments that can help simulators justify focusing on a reduced number of alternative metrics.
 
@@
 The decision trees narrow the number of metrics offering a prediction algorithm into the aforementioned validation categories.
+
 In summary, our results indicate that among the eleven metrics considered in the analysis, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.

### 2018-01-20 abstract.tex
completed validation metrics section

@@
 % 
-It has become common practice to validate ground motion simulation based on a variety of time- and frequency-domain metrics known in seismology and engineering and scaled to quantify the level of agreement between synthetics and data or other reference solutions.
+It has become common practice to validate ground motion simulation based on a variety of time- and frequency-domain metrics scaled to quantify the level of agreement between synthetics and data, or with respect to other reference solutions.
 There is, however, no agreement about the importance or weight that it ought to be given to each individual metric.
 This leads to their selection often being subjective, either based on intended applications or personal preferences.
-In turn, this makes it difficult for simulators to identify modeling improvement needs, which could be easier if they could focused on a reduced number of alternative metrics.
-We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques used to label and classify  goodness-of-fit (GOF) results, with the objective prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation.
-In particular, we study the relationships that exist between eleven different metrics when understood as part of a multi-dimensional space using a constrained \kmeans{} method.
-We conduct a subspace clustering analysis to address the high-dimensional effects and this allows us to label the results from a large validation dataset into four categories (poor, fair, good, and excellent).
-We then develop two decision trees using a C5.0 algorithm.
+As a consequence, it is difficult for simulators to identify what modeling improvement are needed, which would be potentially easier if they could focused on a reduced number of alternative metrics.
+We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques used to label and classify goodness-of-fit results, with the objective prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation.
+In particular, we study the relationships that exist between eleven different metrics.
+In the process carried out, these metrics are understood as part of a multi-dimensional space.
+We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address high-dimensional effects.
+This allows us to label the results from a large validation dataset into four categories (poor, fair, good, and excellent) following common practice.
+We then develop two decision trees using a C4.5 algorithm.
 These decision trees help narrow the number of metrics leading to a validation prediction into the four categories.

### 2018-01-20 introduction.tex
completed validation metrics section

@@
 Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation.
-Various methods or schemes have been proposed to evaluate, through direct quantitative comparisons or indirect statistical analysis, the similarity between simulation synthetics and recorded data or other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
-Among these methods, there are some that are better suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for application validation purposes because they compare the behavior of systems (e.g., buildings or earth structures) when subjected to alternative input motions at the response level \citep{Burks_BSSA_2014}.
-The remaining methods, on the other hand, focus on the overall characteristics of seismograms, in both time and frequency, and are thus better suited for validating high-frequency or broadband simulations, where applications may be undefined and the expectation of matching seismograms at the waveform level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}).
-From this latter group, methods such as those of \citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL} are preferred because they offer simple schemes to quantify the goodness-of-fit (GOF) of synthetics with respect to data.
+Various methods or schemes have been proposed to evaluate, through direct quantitative comparisons or indirect statistical analysis, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
+Among these methods, there are some that are better suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for validation.
+Among the latter, some are better for direct application purposes because they compare the behavior of systems (e.g., buildings or earth structures) at the response level \citep{Burks_BSSA_2014}, while the remaining methods focus on the validation of the overall characteristics of seismograms.
+In this category, the comparisons are done both in time and frequency.
+This type of validation is better suited for high-frequency or broadband simulations, where applications may be undefined and for which the expectation of matching seismograms at the waveform level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}).
+Within this group, methods such as those proposed by \citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL} are preferred because they offer simple schemes to quantify the goodness-of-fit (GOF) of synthetics with respect to data in a broad sense.
 
-Out of the various validation methods available, \citet{Anderson_2004_Proc} is perhaps the most widely used today \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
+Out of the various validation methods just described, \citet{Anderson_2004_Proc} is perhaps the most widely used today \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
 We, in particular, have consistently relied on a modified version of this method for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake, and for the evaluation of velocity models in southern California through the validation of a large suite of moderate-magnitude earthquake simulations \citep{Taborda_2013_BSSA, Taborda_2014_BSSA, Taborda_2016_GJI}.
 
-In essence, Anderson's method assess the similarity of two signals based on the average score of 10 metrics.
+In essence, \citeauthor{Anderson_2004_Proc}'s method assesses the similarity of two signals based on the average score of 10 metrics.
 These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), their peak values (in acceleration, velocity and displacement), frequency content (in terms of Fourier and response spectra), and their time synchronization (through cross-correlation).
 
-The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics are well understood in the engineering and seismology contexts.
+The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics are well understood in both engineering and seismology contexts.
 However, it has been pointed out that some of these metrics are redundant, or that other relevant metrics need to be added.
-\citet{Taborda_2013_BSSA}, for instance, included the strong motion duration \citep{Trifunac_1975_BSSA} as an additional metric, and averaged the Arias intensity- and energy integral-related scores to avoid duplication.
-In the same spirit, \citet{Maufroy_2015_BSSA} reduced the number of metrics limiting them to only those with comparable units.
-Unfortunately, none of these alternatives address the underlying questions: what are the parameters that ought to be taken into account when validating ground motion simulations, and what level of priority should these be given.
+\citet{Taborda_2013_BSSA}, for instance, included---explicitly---the strong motion duration \citep{Trifunac_1975_BSSA} as an additional metric, and averaged the Arias intensity- and energy integral-related scores to avoid duplication.
+In the same spirit, \citet{Maufroy_2015_BSSA} reduced the number of metrics, limiting them to only those with comparable units.
+Unfortunately, none of these alternatives addresses the underlying questions: what are the parameters that ought to be taken into account when validating ground motion simulations, and what level of priority should these be given.
 
-Lack of consensus about how to answer these questions makes the choice of validation methods and the selection of the comparison metrics to be a subjective one, mostly based on personal preferences and---arguably---on the application intended for the simulation.
-In turn, the current multiplicity of metrics, if all are given the same weight in the analysis of validation results, makes it difficult for simulators to identify the sort of changes needed in their models that could lead to better ground motion predictions.
+Lack of consensus about how to answer these questions makes the choice of validation methods and the selection of the comparison metrics a subjective one, mostly based on personal preferences and---arguably---on the application intended for the simulation and/or the validation itself.
+At the same time, with the current multiplicity of metrics, to give to all metrics the same weight in a validation analysis, makes it difficult for simulators to identify the sort of changes needed in their models that could lead to better ground motion predictions.
 We think that this situation can be corrected if we can offer data-informed arguments that can help simulators justify focusing on a reduced number of alternative metrics.
 
-To that end, we study the relationships that exist between eleven different metrics---those proposed by \citet{Anderson_2004_Proc}, plus strong motion duration---with the objective of offering a prioritized reduced number of metrics that can help predict validation results.
+To that end, we study the relationships that exist between eleven different metrics---those proposed by \citet{Anderson_2004_Proc}, plus the strong motion duration---with the objective of offering a prioritized reduced number of metrics that can help predict validation results.
 We treat these metrics as independent variables defining a multi-dimensional space, and analyze them using machine learning techniques.
-We use a constrained \kmeans{} method and conduct a subspace clustering analysis to address high-dimensional effects in order to label the results from a large validation dataset into four categories (poor, fair, good, and excellent).
+We use a constrained \kmeans{} method \citep[e.g.,][]{Macqueen_1967_Proc, Wagstaff_2001_Proc} and conduct a subspace clustering analysis to address high-dimensional effects in order to label the results from a large validation dataset into four categories: poor, fair, good, and excellent.
+These categories adhere to the common practice in validation.
 The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}.
-Upon labeling the dataset, we develop two decision trees using a C5.0 algorithm to prioritize the metrics involved in the validation process.
+Upon labeling the dataset, we develop two decision trees using a C4.5 algorithm \citep[][]{Quinlan_1993_Book} to prioritize the metrics involved in the validation process.
 The decision trees narrow the number of metrics offering a prediction algorithm into the aforementioned validation categories.

### 2018-01-20 validation-metrics.tex
completed validation metrics section

\section{Validation Metrics} 
\label{sec:validation-metrics}

There are various methods and algorithms available to evaluate the similarity between two or more seismograms, both through direct quantitative comparisons or indirect statistical analysis \citep[e.g.,][]{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}. In earthquake ground motion simulation, they are primarily used for verification with respect to benchmark or analytical solutions, and for validation with respect to data (i.e., ground motion records). Here, we focus on the list of metrics proposed by \citet{Anderson_2004_Proc}, with an additional metric for duration, as suggested by others \citep{Olsen_2010_SRL, Maufroy_2015_BSSA} and as implemented in \citet{Taborda_2013_BSSA}.

\input{tab-metrics}

These metrics are listed in Table \ref{tab:metrics}. Following \citet{Anderson_2004_Proc}, when applied to a pair of signals, each one of these metrics yields a GOF score ranging from 0 to 10, where a value of 10 corresponds to two signals having identical characteristics. This scoring scale varies according to the following exponential function:
% 
\begin{equation}
\label{eq:gof-function}
	S \left( p_1, p_2 \right) = 10 \exp{ \left[ - \left( \frac{p_1 - p_2}{ \min\left( p_1, p_2 \right) } \right)^2 \right] }
	\, ,
\end{equation}
% 
\noindent
where $S$ is the GOF score that results from comparing values $p_1$ and $p_2$ from signals 1 and 2, respectively, for each one of the different metrics in Table \ref{tab:metrics}. \citeauthor{Anderson_2004_Proc} also provided guidelines for the process to be applied to the original signals as well as to those resulting from a sequential set of band-pass filters covering the whole frequency range of interest. In his method, this frequency-domain decomposition should have pass bands defined following a logarithmic distribution in order to give more weight to the low-frequencies, and the GOF scores of all sub-bands and the broadband be averaged into a single GOF final score.

\citet{Anderson_2004_Proc} calibrated the results of GOF scores comparing horizontal components of recorded seismograms and other simulations using the first ten metrics (C1 through C10) in Table \ref{tab:metrics}. He concluded that for a typical distribution of the scores, the GOF could be categorized into four groups: poor (for scores from 0 to 4), fair (4 to 6), good (6 to 8), and excellent (for scores from 8 to 10). While this categorization was arguably subjective, over the years it has become common practice and has been adopted or matched by alternative validation methods \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}, making the method itself a reference baseline for various validation studies \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.

Here, we focus our analysis on the eleven metrics included in Table \ref{tab:metrics}, and investigate the relationships that exist between them in order to prioritize a reduced number of metrics that can help predict the outcome category one would use to label the results of a given simulation.

% ===========================================================================================
%
% OLD NAEEM VERSION
% 
% Following the guidelines suggested by \citet{Anderson_2004_Proc}, \citet{Taborda_2014_BSSA} applied the scoring procedure to each pair of data and synthetic using compatible broadband sets, and series of bandpass filtered versions of the signals or sub-bands, SB1, SB2, SB3, SB4, and SB5. The signals bandpass filtered between 0.1 and 4 Hz for BB, and between 0.1 Hz and 0.25 Hz, and 0.25 and 0.5 Hz, and 0.5 and 1 Hz, and 1 and 2 Hz, and 2 and 4 Hz for SB1, SB2, SB3, SB4, and SB5, respectively. They also reported a scaled combination score of metrics and bands. However, since the combined scores are linear scaled combinations of metrics and frequency bands, in this specific study, they do not add new information to the data base.  \citet{Khoshnevis_2015_Proc} studied the effect of filtering on GOF score and presented that filtering approach can affect the GOF score and change the final classes. Therefore, in order to avoid extra source of uncertainty in the study, we only consider broadband GOF scores ($0.1-4~Hz$). Analysis of effect of bands in the results could be an appropriate subject for another study.

% Accuracy of ground motion simulation is evaluated based on a quantitative validation of the simulation ground motions. The validation process consist of comparisons between synthetics and data, at locations where records were available for the simulated events. For validating the simulation of strong motion complex time series, different metrics are defined. These metrics evaluate the similarity of two signals in time or frequency domain. \\

% \citet{Anderson_2004_Proc} proposed 10 individual quality measures to evaluate the credibility of synthetic seismograms.  He proposed to apply the metrics on 10 frequency bands with logarithmic spacing. The logarithmic spacing puts more emphasis on lower frequencies where these frequencies are more amenable to waveform fitting and are particularly important for response of large structures. A comprehensive score (S1) could be achieved by the average of the scores of seismograms filtered in each of the valid frequency bands individually, as well as the broadband seismogram. An alternative score (S2) is obtained by averaging the scores on the ten individual criteria for only the accelerogram filtered to allow all frequencies to pass. \citet{Anderson_2004_Proc} scaled the scores into range of 0 to 10 with 10 giving perfect agreement. He suggested to consider a score below 4 as a poor fit, a score of 4-6 as a fair fit, a score of 6 to 8 as a good fit, and a score over 8 is an excellent fit. However, some of these metrics like integral measures are potentially very easy to fit and some of them like cross correlation is by far the most difficult of all parameters to achieve high score. Therefore, for a pair of synthetic and data some of metrics could belong to good and some of them could belong to fair fit classes.\\

% In other study, \citet{Kristekova_2006_BSSA} developed quantitative misfit criteria for comparison of seismograms. The misfit criteria are defined based on the time-frequency (TF) representation of the seismograms obtained at the continues wavelet transform. Based on local TF envelop and phase differences, they defined envelop and phase misfit dependent on both time and frequency. With projection of TF misfit onto frequency or time domain, they computed frequency- or time-dependent misfits, respectively. The method proposed  by \citet{Kristekova_2006_BSSA} needs to consider one of signals as a reference signal. Therefore, \citet{Kristekova_2009_GJI} presented an extension of the theory of the TF misfit criteria. They defined locally and globally normalized TF criteria where locally normalized misfits can be used if it is important to investigate relatively small part of the signal. The globally normalized misfits can be used for quantifying an overall level of disagreement. In the extended version of the metrics, they defined a misfit criteria for 3 components of a signal and also they developed TF GOF criteria for more realistic scenario where there is a high level of disagreement and it is not possible to consider one of signals as a reference signal. They classified the GOF scores as poor, fair, good, and excellent based on GOF numerical values.\\

% \citet{Olsen_2010_SRL} argued that the \citet{Kristekova_2006_BSSA,Kristekova_2009_GJI}  metrics are suitable for low frequency signals. Therefor, they proposed a new GOF method for validation of Broadband synthetics. They used some of metrics that are used by \citet{Anderson_2004_Proc}, however, they used different scoring approach which generates a relatively high-resolution representation of small misfits. They also used inelastic and elastic response spectra ration (IE ratio) as a structural engineering specific metric. They proposed to combine the metrics for final decisions based on user defined application-based weights.\\

% Recently, \citet{Taborda_2013_BSSA} conducted a simulation for 2008 Chino Hills earthquake and proposed a minor modification in the \citet{Anderson_2004_Proc} metrics. They added an eleventh parameter to the Anderson's scores as strong motion duration.\\

% A comprehensive validation approach should include metrics in both time and frequency domain. All mentioned metrics \citep[i.e.,][]{Anderson_2004_Proc,Kristekova_2006_BSSA,Kristekova_2009_GJI,Olsen_2010_SRL} have this characteristic. However, in this study we are mostly dependent on available data based on \citet{Taborda_2014_BSSA} where they used the GOF method proposed by \citet{Anderson_2004_Proc}, with minor modification introduced by \citet{Taborda_2013_BSSA}. The method compares synthetics against data using 11 individual parameters, namely: Arias Intensity Integral (C1), Energy Integral (C2), Arias Intensity Value (C3), Total Energy (C4), Peak Acceleration (C5), Peak Velocity (C6), Peak Displacement (C7), Response Spectra (C8), Fourier Amplitude Spectrum (C9), Cross Correlation (C10), and Strong Phase Duration (C11). Each parameter is mapped on to a numerical scale ranging from 0 to 10, where a score of 10 corresponds to a perfect  match between two signals.\\
% Following the guidelines suggested by \citet{Anderson_2004_Proc}, \citet{Taborda_2014_BSSA} applied the scoring procedure to each pair of data and synthetic using compatible broadband sets, and series of bandpass filtered versions of the signals or sub-bands, SB1, SB2, SB3, SB4, and SB5. The signals bandpass filtered between 0.1 and 4 Hz for BB, and between 0.1 Hz and 0.25 Hz, and 0.25 and 0.5 Hz, and 0.5 and 1 Hz, and 1 and 2 Hz, and 2 and 4 Hz for SB1, SB2, SB3, SB4, and SB5, respectively. They also reported a scaled combination score of metrics and bands. However, since the combined scores are linear scaled combinations of metrics and frequency bands, in this specific study, they do not add new information to the data base.  \citet{Khoshnevis_2015_Proc} studied the effect of filtering on GOF score and presented that filtering approach can affect the GOF score and change the final classes. Therefore, in order to avoid extra source of uncertainty in the study, we only consider broadband GOF scores ($0.1-4~Hz$). Analysis of effect of bands in the results could be an appropriate subject for another study.

### 2018-01-21 study-dataset.tex
up to the dataset section

\section{Study Dataset}

We select the simulation results and validation analysis done by \citet{Taborda_2014_BSSA} as our reference dataset. In that study, the authors carried out deterministic simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake using a finite-element approach. The simulations, corresponding to a kinematic finite-fault model of the earthquakes, were obtained for a maximum frequency \fmaxeq{4} and a minimum shear wave velocity of \vsmineq{200}. In total, \citet{Taborda_2014_BSSA} did three simulations, each for a different velocity model \citep[CVM-S4, CVM-H, CVM-H+GTL,][]{Small_2017_SRL}. The modeling domain covered an area of \adomain{180}{135}{km} that included all the major sedimentary basins and other relevant geologic structures in the greater Los Angeles region. Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three component of motion (NS, EW, and UD). The simulation domain and the stations used for the analysis are shown in Figure \ref{fig:chino-hills}, and the validation results obtained by \citet{Taborda_2014_BSSA} for the GOF analysis of the broadband signals is shown using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps} for the three simulation sets.

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-02}
    \caption{Foo}
    \label{fig:ref-gof-maps}
\end{figure*}

Here, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the component of motions. Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be discussed next was done using the whole dataset of GOF scores. The motivation behind this choice was that the dataset, as a sample of GOF scores, was independent of the simulation and serves here as a generic set for the purpose of identifying the correlations that exist between the different metrics in Table \ref{tab:metrics}. As such, given the simulations for each velocity model (3), the components of motion (3), and the number of stations used in the validation (336) gives us a large enough sample. Figure \ref{fig:data-box-plot} reinforces precisely this idea by comparing the statistical distribution of the broadband GOF scores of the simulations classified by velocity models and components. It is clear that although there are differences between them, these are minor and negligible; in other words, the statistical distribution of results for each metric is about the same independently of model or component.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/pdf/figure-03}
    \caption{Statistical distribution of the GOF scores used in this study based on the validation results obtained by \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component of motion (NS, EW, UD). In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range (Q3-Q1), with outliers shown as scattered dots.}
    \label{fig:data-box-plot}
\end{figure}

% ===========================================================================================
%
% OLD NAEEM VERSION
% 
% State of California is an earthquake prone region with a long history of seismicity. The Los Aneles basin, due to its unique geologic sediments, has been studied from different perspective. Numerous high resolution velocity models are developed mainly to conduct forward ground motion simulation. Occurrence of frequent moderate magnitude earthquakes made the Los Angeles basin as a natural laboratory for seismic  studies specially testing the forward ground motion simulation modeling. In a relatively recent study, \citet{Taborda_2014_BSSA} conducted a comprehensive study in order to assess the functionality of developed velocity models in the Los Angeles basin and surrounding areas. They simulated the 2008 Chino Hills earthquake using 3 velocity models including: CVM-S4, CVM-H, and CVM-H+GTL (for more details about these models see \citet{Taborda_2014_BSSA}.)
% They computed the GOF scores based on \citet{Anderson_2004_Proc}, with minor modification introduced by \citet{Taborda_2013_BSSA} as discussed in the validation section. Fig.~\ref{fig:epicenters} shows the location of the epicenter and seismic stations.

% \begin{figure}
%     \centering
%     \includegraphics
%        % [width=\columnwidth]
%         [width=450px]
%         {figures/pdf/Figure_1}
%     \caption{(a) Region of interest and epicentral location of the 2008 $M_W ~ 5.4$ Chino Hills earthquake (highlighted). Focal mechanism is shown at the margin of the region of interest. Major quaternary faults in the area are shown in the back along with the main roads and county divisions. (b) Simulation area of interest. Dots indicate the location of the 336 stations considered in this study for validation. The main local and interstate roads are also shown here. In both panels, the background shows the hillshade topography of the region. The color version of this figure is available only in the electronic edition.}
%     \label{fig:epicenters}
% \end{figure}

% We use \citet{Taborda_2014_BSSA} database where they computed 11 scores for 3 different velocity models. In this study we only analyze the results for broadband which in this case is \feq{0.1}{4}. Fig.~\ref{fig:data_box_plot} shows the box plot of all data that we used in this study which we separate them in velocity models and components. In total we use $336 \times 3 \times 3 = 3024$ individual pair of data and synthetic.

% \begin{figure}
%     \centering
%     \includegraphics
%        % [width=\columnwidth]
%         [width=\textwidth]
%         {figures/pdf/Figure_2.pdf}
%     \caption{Box plot of data used in the study. Metrics are shown for C1 to C11 for 3 components and in separate plots for velocity models. Median of data are shown as a notch. Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data. Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
%     \label{fig:data_box_plot}
% \end{figure}

% As we explained before, we use all data set regardless of velocity models and components. However, distinguishing data for velocity model and components can answer the question that if the goodness of fit scores dependent on them. Addressing this research question is beyond the scope of this paper. We distinguish data based on velocity models and components only in presenting data or results.

% Fig.~\ref{fig:data_density}  shows the density distribution of each score for CVM-S. The figure shows the variation of different metrics with components. Except in cross correlation (C10) in all metrics the up-down (UD) component scores are about the same or higher than horizontal components.

% \begin{figure}
%     \centering
%     \includegraphics
%        % [width=\columnwidth]
%         [width=\textwidth]
%         {figures/pdf/Figure_3.pdf}
%     \caption{Distribution of data (only CVM-S) for different components in different scores based on different components. Each distribution has a unit area.}
%     \label{fig:data_density}
% \end{figure}

### 2018-01-21 study-dataset.tex
minor fixes

@@
 The motivation behind this choice was that the dataset, as a sample of GOF scores, was independent of the simulation and serves here as a generic set for the purpose of identifying the correlations that exist between the different metrics in Table \ref{tab:metrics}.
-As such, given the simulations for each velocity model (3), the components of motion (3), and the number of stations used in the validation (336) gives us a large enough sample.
+As such, given the simulations for each velocity model (3), the components of motion (3), and the number of stations used in the validation (336) gives us a large enough sample of 3,024 GOF scores.
 Figure \ref{fig:data-box-plot} reinforces precisely this idea by comparing the statistical distribution of the broadband GOF scores of the simulations classified by velocity models and components.

### 2018-01-21 study-dataset.tex
done with intro to the methodology section

--- 2018-01-21 study-dataset.tex
@@
 In that study, the authors carried out deterministic simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake using a finite-element approach.
-The simulations, corresponding to a kinematic finite-fault model of the earthquakes, were obtained for a maximum frequency \fmaxeq{4} and a minimum shear wave velocity of \vsmineq{200}.
-In total, \citet{Taborda_2014_BSSA} did three simulations, each for a different velocity model \citep[CVM-S4, CVM-H, CVM-H+GTL,][]{Small_2017_SRL}.
+The simulations, done for a kinematic finite-fault model of the earthquake, were computed for a maximum frequency \fmaxeq{4} and a minimum shear wave velocity \vsmineq{200}.
+In total, \citet{Taborda_2014_BSSA} did three simulations, each for a different velocity model \citep[CVM-S4, CVM-H, CVM-H+GTL, see][]{Small_2017_SRL}.
 The modeling domain covered an area of \adomain{180}{135}{km} that included all the major sedimentary basins and other relevant geologic structures in the greater Los Angeles region.
-Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three component of motion (NS, EW, and UD).
-The simulation domain and the stations used for the analysis are shown in Figure \ref{fig:chino-hills}, and the validation results obtained by \citet{Taborda_2014_BSSA} for the GOF analysis of the broadband signals is shown using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps} for the three simulation sets.
+Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three component of motion (EW, NS, and UD).
+The simulation domain and the stations used for the analysis are shown in Figure \ref{fig:chino-hills}, and the validation results obtained by \citet{Taborda_2014_BSSA} for the GOF analysis of the broadband signals using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps} for the three simulation sets.
 
@@
     \includegraphics[width=\textwidth]{figures/pdf/figure-02}
-    \caption{Foo}
+    \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between simulations for three different velocity models and data recorded at ground motion monitoring stations.}
     \label{fig:ref-gof-maps}
@@
 Here, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the component of motions.
-Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be discussed next was done using the whole dataset of GOF scores.
-The motivation behind this choice was that the dataset, as a sample of GOF scores, was independent of the simulation and serves here as a generic set for the purpose of identifying the correlations that exist between the different metrics in Table \ref{tab:metrics}.
-As such, given the simulations for each velocity model (3), the components of motion (3), and the number of stations used in the validation (336) gives us a large enough sample.
-Figure \ref{fig:data-box-plot} reinforces precisely this idea by comparing the statistical distribution of the broadband GOF scores of the simulations classified by velocity models and components.
-It is clear that although there are differences between them, these are minor and negligible; in other words, the statistical distribution of results for each metric is about the same independently of model or component.
+Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be described in the following section was done using the whole dataset of scores.
+The motivation behind this choice was that the dataset, as a sample of GOF values, was independent of the simulation and serves here as a generic set for the purpose of identifying the correlations that exist between the different metrics in Table \ref{tab:metrics}.
+As such, given the simulations for each velocity model (3), the motion components (3), and the number of stations used in the validation (336) gave us a large enough sample of 3,024 GOF scores.
+Figure \ref{fig:data-box-plot} illustrates this idea by comparing the statistical distribution of the broadband GOF scores of the simulations classified by velocity models and components.
+It is clear that although there are differences between them, these are negligible; in other words, the statistical distribution of results for each metric is about the same independently of model or component.
 

### 2018-01-21 introduction.tex
minor fixes

@@
 We treat these metrics as independent variables defining a multi-dimensional space, and analyze them using machine learning techniques.
-We use a constrained \kmeans{} method \citep[e.g.,][]{Macqueen_1967_Proc, Wagstaff_2001_Proc} and conduct a subspace clustering analysis to address high-dimensional effects in order to label the results from a large validation dataset into four categories: poor, fair, good, and excellent.
+We use a constrained \kmeans{} method \citep[e.g.,][]{Macqueen_1967_Proc, Wagstaff_2001_Proc} and conduct a subspace clustering analysis to address high-dimensional effects in order to add labels to the dataset in terms of four categories: poor, fair, good, and excellent.
 These categories adhere to the common practice in validation.

### 2018-01-22 study-dataset.tex
minor fixes
working on the clustering section

@@
 
-We select the simulation results and validation analysis done by \citet{Taborda_2014_BSSA} as our reference dataset.
+We select the simulation results and validation analysis done by \citet{Taborda_2014_BSSA} as our reference dataset, hereafter the dataset.
 In that study, the authors carried out deterministic simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake using a finite-element approach.
@@
 Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three component of motion (EW, NS, and UD).
-The simulation domain and the stations used for the analysis are shown in Figure \ref{fig:chino-hills}, and the validation results obtained by \citet{Taborda_2014_BSSA} for the GOF analysis of the broadband signals using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps} for the three simulation sets.
+The simulation domain and the stations used for the analysis are shown in Figure \ref{fig:chino-hills}, and a sample of the validation results obtained by \citet{Taborda_2014_BSSA} using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps}.
+This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, averaged for all metrics, and the three component of motion.
+This type of so-called GOF maps are commonly used in validation studies.
 
-\begin{figure*}
+\begin{figure*}[t]
     \centering
@@
 \end{figure*}
+% 
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/pdf/figure-03}
+    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component of motion (NS, EW, UD).
+In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range, with outliers shown as scattered dots.}
+    \label{fig:data-box-plot}
+\end{figure}
 
@@
 It is clear that although there are differences between them, these are negligible; in other words, the statistical distribution of results for each metric is about the same independently of model or component.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\columnwidth]{figures/pdf/figure-03}
-    \caption{Statistical distribution of the GOF scores used in this study based on the validation results obtained by \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component of motion (NS, EW, UD).
-In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range (Q3-Q1), with outliers shown as scattered dots.}
-    \label{fig:data-box-plot}
-\end{figure}
 

### 2018-01-24 decision-tree.tex
minor progress

\subsection{Decision Tree Construction}
\label{sec:decision-tree}

Working...

% ===========================================================================================
%
% OLD NAEEM VERSION
% 
% Decision tree learning is a method for approximating a target function, in which learned function is represented by decision tree \citep{Mitchell_1997_Book}. \citet{Quinlan_1986} demonstrated that the technology for building decision trees from examples is fairly robust. He summarized an approach to synthesizing decision trees that has been used in a variety of systems, and he describes one such system, ID3, in detail. Although ID3 algorithm is successful in most classification problems, it does not handle the numeric attributes and only one attribute at a time is tested for making decisions. \citet{Quinlan_1993} extended the ID3 algorithm and presented the C4.5 algorithm. Later on they added boosting capability to C4.5 and called it C5.0. The improved algorithm accepts both continues and discrete features and solves over fitting problem by pruning. Although there are many other algorithm for implementing decision tree, in this study we use the C5.0 algorithm. Discussing the algorithm in details is out of the scope of this paper. There are fairly good amount of references to study those details \citep[e.g.,][]{Mitchell_1997_Book,Quinlan_1993,Hornik_2009}.\\
% In general, decision tree measures the effectiveness of an attribute in classifying the training data through information gain. According to  \citet{Mitchell_1997_Book}, information gain of an attribute is the expected reduction in entropy caused by partitioning the examples according to this attribute. If the target can take on $c$ different values the entropy is defined as
% \begin{equation}
% Entropy(S) = \sum_{i=1}^{c} -p_{i}\ log_2\ p_i,
% \end{equation}

% where $p_i$ is proportion of $S$ belonging to class $i$. Entropy measures the homogeneity of examples or dataset. Higher entropy means data is  fairly uniformly distributed among different classes. On the other hand lower entropy means data unequally distributed among classes. The extreme case happens when all data belong to one class which results in entropy equal to zero. Given entropy as a measure of impurity in a collection of training example the measure of the effectiveness of attribute is defined through information gain which is expected reduction in entropy caused by partitioning the examples according to this attribute. Gain of the attribute A is defined as

% \begin{equation}
% Gain(S,A)\ =\ Entropy(S) - \sum_{v \in Values(A) } \frac{|S_v|}{|S|} Entropy(S_v)
% \end{equation}

% where $Values(A)$ is the set of all possible values of attribute A, and $S_v$ is the subset of $S$ for which attribute $A$ has value $v$. Information gain is used by the algorithm in order to determine which is the better attribute for classifying the training example. The attribute with higher information gain will be the first option to separate the training data based on that attribute and grow to the lower nodes (Refer to chapter 3 of \citet{Mitchell_1997_Book} for more detailed examples.) \\
% We discussed the basic idea behind the decision tree algorithm, however, there are more details specially in dealing with outliers and noises which we address them through pruning approaches. In this study we use C5.0 package \citep{C50_2015} in R programming language \citep{R_2016_program} to develop the decision tree algorithms. We discuss the input parameters in the section.\ref{classification_model}.

### 2018-01-26 analysis_cluster.tex
a first complete pass on clustering method, still needing to be reduced.

@@
 Therefor we present the clustering results based on a pair of features in Fig.~\ref{fig:CxCy_11}.
-\begin{figure}
+
+\begin{figure*}
     \centering
-    \includegraphics
-       % [width=\columnwidth]
-        [width=\textwidth]
-        {figures/pdf/Figure_5.pdf}
+    \includegraphics[width=\textwidth]{figures/pdf/Figure_5.pdf}
     \caption{Results of clustering analysis with all 11 attributes (metrics).
@@
     \label{fig:CxCy_11}
-\end{figure}
+\end{figure*}
+
 The figure presents a part of results in clustering with 11 features.
@@
 C1 vs C2 and C10 vs C11 are examples of such features combinations.
-It is worth to mention that at this point the cluster numbers are not consistent with ordinal categorical numbers of 1, 2, 3, and 4 representing poor, fair, good, and excellent, respectively.\\
+It is worth to mention that at this point the cluster numbers are not consistent with ordinal categorical numbers of 1, 2, 3, and 4 representing poor, fair, good, and excellent, respectively.
+
 Although we do not expect to see the pattern that resulted from clustering using 11 features in presentation of only two, however, different studies show that the higher dimension reduce the effect of similarity based on distance.
@@
 They illustrated the irrelevant and redundant features and proposed methods for evaluating candidate features using two performance criteria.
-\\
+
 Subspace analysis is another technique to address the challenges with higher dimensional data in clustering process.
@@
 The most common factor among these algorithm is the process to find a group of best subspaces through optimization process.
-An n-dimensional dataset has $2^n$ subspaces where it could be very costly and time consuming to evaluate all of them.\\
+An n-dimensional dataset has $2^n$ subspaces where it could be very costly and time consuming to evaluate all of them.
+
 As we discussed earlier, there is not standard measures for evaluating clusters in the clustering literature also there is no single clustering assignment explains every application \citep{Dy_2004_MLR,Jain_1988_Book,Hartigan_1985_JOC}.
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
-    \includegraphics
-       % [width=\columnwidth]
-        [width=\textwidth]
-        {figures/pdf/Figure_6.pdf}
+    \includegraphics[width=\textwidth]{figures/pdf/Figure_6.pdf}
     \caption{Box plot of data used in the study after clustering.
@@
     \label{fig:C_cvm_comp}
-\end{figure}
+\end{figure*}
 
-\begin{figure}
+\begin{figure*}
     \centering
-    \includegraphics
-       % [width=\columnwidth]
-        [width=\textwidth]
-        {figures/pdf/Figure_7.pdf}
+    \includegraphics[width=\textwidth]{figures/pdf/Figure_7.pdf}
     \caption{Box plot of data used in the study after clustering and merging the results for different velocity models.
@@
     \label{fig:Figure_cluster_comp}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
-    \includegraphics
-       % [width=\columnwidth]
-        [width=\textwidth]
-        {figures/pdf/Figure_8.pdf}
+    \includegraphics[width=\textwidth]{figures/pdf/Figure_8.pdf}
     \caption{Box plot of data used in the study after clustering and merging for different components and velocity models.
@@
     \label{fig:Figure_cluster_all_comp}
-\end{figure}
+\end{figure*}
 
@@
     \centering
-    \includegraphics
-       % [width=\columnwidth]
-        [width=300px]
-        {figures/pdf/Figure_9.pdf}
+    \includegraphics[width=\columnwidth]{figures/pdf/Figure_9.pdf}
     \caption{Number of data in each cluster based on components after conducting constraint \kmeans{} clustering through subspace analysis.}

### 2018-01-26 study-dataset.tex
clustring first pass revised

@@
     \includegraphics[width=\textwidth]{figures/pdf/figure-02}
-    \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between simulations for three different velocity models and data recorded at ground motion monitoring stations.}
+    \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between simulations for three different velocity models and data recorded at ground motion monitoring stations.
+The color version of this figure is available only in the electronic edition.}
     \label{fig:ref-gof-maps}
@@
     \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component of motion (NS, EW, UD).
-In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range, with outliers shown as scattered dots.}
+In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range, with outliers shown as scattered dots.
+The color version of this figure is available only in the electronic edition.}
     \label{fig:data-box-plot}

### 2018-01-30 analysis_dtree.tex
preparing for dtree method section

@@
 Although there are numerous methods to classify labeled data, however, in this study we are interested in a method that not only predict the class of simulation based on metrics, but also give an intuition about the decision process.
-Therefor we use tree based algorithm to develop the prediction model.\\
+Therefor we use tree based algorithm to develop the prediction model.
 
@@
 The Pearson correlation coefficient between two features is defined as
-
+% 
 \begin{equation}
@@
 \end{equation} 
-
+% 
 where $cov$ is covariance function and $var$ is the variance function \citep{Guyon_2003}. Fig.~\ref{fig:Corplot} represent the Pearson Correlation factor among features.
@@
     \centering
-    \includegraphics
-       % [width=\columnwidth]
-        [width=200px]
-        {figures/pdf/Figure_10.pdf}
+    \includegraphics[width=\columnwidth]{figures/pdf/Figure_10.pdf}
     \caption{Pearson correlation values between metrics. The plot is generated using Corrplot package  \citep{Corrplot_2016}  in R programming language.}
@@
 A mostly common approach is data preprocessing through resampling.
-According to \citet{Branco_2015}, resampling strategies based on diverse set of techniques such as: random under/over-sampling, distance methods, data cleaning approaches, clustering algorithm among them.\\
+According to \citet{Branco_2015}, resampling strategies based on diverse set of techniques such as: random under/over-sampling, distance methods, data cleaning approaches, clustering algorithm among them.
+
 According to the clustering results, among all data with different components and velocity model there are 816,1253,879, and 76 data belonging to cluster 1,2,3, and 4, respectively.
@@
 
-In this study we assume $\beta = 1$ to give same importance to precision and recall. \\
+In this study we assume $\beta = 1$ to give same importance to precision and recall.
+
 Developing a prediction model has a broad discussion.
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
@@
     \label{fig:tune}
-\end{figure}
+\end{figure*}
 
@@
 However, as we mentioned earlier, we do not present the best model, rather we present a simplest model that in general illustrate the feature effects.
+
 First we start with a very simple model. A simple model needs a strong pruning process.  Therefore we use $minCases = 120, CF=0.2$. Fig.~\ref{fig:dtree1} shows the classification decision tree.
-\begin{figure}
+
+\begin{figure*}
     \centering
@@
     \label{fig:dtree1}
-\end{figure}
+\end{figure*}
 Table~\ref{tab:confusionmat_test_1} shows  the confusion matrix of applying the prediction model on test dataset and Table~\ref{tab:prec_recall_test_1} presents the precision, recall, and F1-score for each individual class.
@@
     \label{fig:dtree2}
-\end{figure}
+\end{figure*}
 

### 2018-01-30 decision-tree.tex
clustring first pass revised
reduced clustering section
preparing for dtree method section

@@
-\subsection{Decision Tree Construction}
+\subsection{Decision Tree Design}
 \label{sec:decision-tree}
 
-Working...
+The construction of a decision tree is a supervised learning process to obtain a target function used to measure the effectiveness of an attribute to classify data.
+There are a good number of approaches to develop decision trees \citep[e.g.,][]{Quinlan_1986_ML, Mitchell_1997_Book, Hornik_2009_}
+
+We use the C4.5 algorithm developed by \citet{Quinlan_1993_Book}...
+
+% In machine learning, decision trees are classified as a supervised method,
 
@@
 % 
-% Decision tree learning is a method for approximating a target function, in which learned function is represented by decision tree \citep{Mitchell_1997_Book}.
+Decision tree learning is a method for approximating a target function, in which learned function is represented by decision tree \citep{Mitchell_1997_Book}.
 \citet{Quinlan_1986} demonstrated that the technology for building decision trees from examples is fairly robust.
@@
 Although ID3 algorithm is successful in most classification problems, it does not handle the numeric attributes and only one attribute at a time is tested for making decisions.
-\citet{Quinlan_1993} extended the ID3 algorithm and presented the C4.5 algorithm.
+\citet{Quinlan_1993_Book} extended the ID3 algorithm and presented the C4.5 algorithm.
 Later on they added boosting capability to C4.5 and called it C5.0.
@@
 Discussing the algorithm in details is out of the scope of this paper.
-There are fairly good amount of references to study those details \citep[e.g.,][]{Mitchell_1997_Book,Quinlan_1993,Hornik_2009}.\\
-% In general, decision tree measures the effectiveness of an attribute in classifying the training data through information gain.
+There are fairly good amount of references to study those details \citep[e.g.,][]{Mitchell_1997_Book, Quinlan_1993_Book, Hornik_2009_}.
+
+In general, decision tree measures the effectiveness of an attribute in classifying the training data through information gain.
 According to  \citet{Mitchell_1997_Book}, information gain of an attribute is the expected reduction in entropy caused by partitioning the examples according to this attribute.
 If the target can take on $c$ different values the entropy is defined as
-% \begin{equation}
-% Entropy(S) = \sum_{i=1}^{c} -p_{i}\ log_2\ p_i,
-% \end{equation}
-
-% where $p_i$ is proportion of $S$ belonging to class $i$.
+% 
+\begin{equation}
+Entropy(S) = \sum_{i=1}^{c} -p_{i}\ log_2\ p_i,
+\end{equation}
+% 
+where $p_i$ is proportion of $S$ belonging to class $i$.
 Entropy measures the homogeneity of examples or dataset.
@@
 Gain of the attribute A is defined as
+% 
+\begin{equation}
+Gain(S,A)\ =\ Entropy(S) - \sum_{v \in Values(A) } \frac{|S_v|}{|S|} Entropy(S_v)
+\end{equation} 
+% 
+where $Values(A)$ is the set of all possible values of attribute A, and $S_v$ is the subset of $S$ for which attribute $A$ has value $v$.
+Information gain is used by the algorithm in order to determine which is the better attribute for classifying the training example.
+The attribute with higher information gain will be the first option to separate the training data based on that attribute and grow to the lower nodes (Refer to chapter 3 of \citet{Mitchell_1997_Book} for more detailed examples.)
 
-% \begin{equation}
-% Gain(S,A)\ =\ Entropy(S) - \sum_{v \in Values(A) } \frac{|S_v|}{|S|} Entropy(S_v)
-% \end{equation}
-
-% where $Values(A)$ is the set of all possible values of attribute A, and $S_v$ is the subset of $S$ for which attribute $A$ has value $v$.
-Information gain is used by the algorithm in order to determine which is the better attribute for classifying the training example.
-The attribute with higher information gain will be the first option to separate the training data based on that attribute and grow to the lower nodes (Refer to chapter 3 of \citet{Mitchell_1997_Book} for more detailed examples.) \\
-% We discussed the basic idea behind the decision tree algorithm, however, there are more details specially in dealing with outliers and noises which we address them through pruning approaches.
+We discussed the basic idea behind the decision tree algorithm, however, there are more details specially in dealing with outliers and noises which we address them through pruning approaches.
 In this study we use C5.0 package \citep{C50_2015} in R programming language \citep{R_2016_program} to develop the decision tree algorithms.
-We discuss the input parameters in the section.\ref{classification_model}.
+We discuss the input parameters in the section \ref{classification_model}.

### 2018-02-02 abstract.tex
minor fixes
done with methodology

@@
 We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address high-dimensional effects.
-This allows us to label the results from a large validation dataset into four categories (poor, fair, good, and excellent) following common practice.
-We then develop two decision trees using a C4.5 algorithm.
+This allows us to label a given validation dataset into four categories (poor, fair, good, and excellent) following common practice.
+We then develop two decision trees using a C5.0 algorithm.
 These decision trees help narrow the number of metrics leading to a validation prediction into the four categories.

### 2018-02-02 decision-tree.tex
done with methodology

--- 2018-01-24 decision-tree.tex
@@
-\subsection{Decision Tree Construction}
+\subsection{Decision Trees}
 \label{sec:decision-tree}
 
-Working...
+Building decision trees is a supervised learning process used to approximate a target function as a sequence of disjunctive conditions designed to measure the effectiveness of a set of attributes to classify data and predict outcomes.
+The theory behind decision trees is well established \citep[e.g.,][]{Quinlan_1986_ML, Mitchell_1997_Book}.
+Decision trees are, however, non-unique.
+They depend on the dataset, the user parameters, and the algorithm employed to build them \citep{Murthy_1998_DMKD}.
+Therefore, before describing our choice of the decision tree algorithm and parameters, we revise three aspects about the dataset: data classes, attributes equivalence, and dataset balance.
+
+First, regarding data classes we refer to whether the classes of data are properly distinguishable or not, and to whether the data attributes contribute to those classes or not.
+In our case, data classes are handled by means of the clustering and labeling process.
+As we described in the previous section, at the end of the clustering process, all the data-points (or samples) in our dataset get labeled with one out of four outcome classes (or clusters) defined in terms of the validation categories (poor, fair, good, excellent).
+
+Second, attributes equivalence refers to whether the data attributes are comparable with each other on equivalent terms or not.
+There are cases, either because of the method or because of the data, that the data attributes need to be standardized.
+In artificial neural networks, for instance, the data needs to be normalized on a $[-1,1]$ or $[0,1]$ scale.
+Normalization is also necessary when data attributes come in different unit scales or value ranges \citep[e.g.,][]{Wu_2010_JH}.
+In our case, this is not necessary because (a) all the data attributes---defined by the 11 GOF metrics---are on the same dimensionless numerical scale (varying from 0 to 10), and (b) because the algorithm we use is not susceptible to the attributes scale.
+
+Last, there is the issue of balance.
+A balanced dataset is one that has about the same number of samples per class.
+Algorithms used for building decision trees tend to perform better with balanced datasets \citep[e.g.,][]{Branco_2016_ACMCS, Weiss_2003_JAIR}.
+Imbalanced datasets, on the other hand, are those with a significant disparity in the number of samples in each class.
+Imbalanced datasets can be improved by a resampling process.
+There are two basic resampling approaches: undersampling and oversampling \citep{Branco_2016_ACMCS}.
+Undersampling discards data from the subsets with larger number of samples.
+Oversampling replicates data until reaching appropriate number of samples.
+Both processes are done by randomly picking data to either be discarded or replicated.
+In our case, as we will see later, we apply oversampling to one of our dataset classes.
+This process has consequences, mainly that it increases the likelihood of overfitting.
+A matter we discuss later.
+
+Upon completing the clustering and resampling processes, the next step is to subdivide the dataset in two, a training dataset (with 70 percent of the samples picked randomly), and a testing dataset (with the remaining 30 percent).
+We first use the training dataset to build a large suite of potential decision trees, and then use the testing dataset to evaluate and pick the best possible decision tree(s) among them.
+There are several algorithms for building decision trees \citep[e.g., ID3, C4.5, C5.0, CART; see][]{Quinlan_1986_ML, Quinlan_1993_Book, Quinlan_1996_JAIR, Breiman_1984_Book}.
+Here we use the C5.0 implementation \citep{Kuhn_2017_Manual}, which is an extension of the C4.5 algorithm developed by \citet{Quinlan_1993_Book, Quinlan_1996_JAIR}.
+
+In essence, given a dataset, the process of building the tree consists of recursively identifying the attributes in the training data that are most likely to predict an outcome.
+The process is recursive because new branches and decision nodes are created based on the remaining data at every branch and level, until the tree reaches a certain depth or when other conditions are met.
+The effectiveness of an attribute $A$ in classifying any subset $S$ from the training data is measured through the information gain $G$ as
+% 
+\begin{equation}
+	\label{eq:gain}
+	G(S,A) = E(S) - \sum_{v \, \in \, A[a]} \frac{ \left| S_v \right| }{ \left| S \right| } E \left( S_v \right)
+	\, ,
+\end{equation} 
+% 
+where $A[a]$ is the set of all possible values of attribute $A$, $S_v$ is the subset of $S$ for which attribute $A$ has value $v$, and $E(\cdot)$ is the entropy function given by
+% 
+\begin{equation}
+	E(S) = \sum_{i=1}^{c} \Big( -p_i \log_2 \left(p_i\right) \Big)
+	\, ,
+\end{equation}
+% 
+where $p_i$ is the proportion of $S$ belonging to class $i$, and $c$ are the different values that a given target attribute can take.
+
+Entropy measures the homogeneity of a set of data.
+The higher the entropy, the more even the distribution of the data across classes.
+Conversely, an extremely low value of entropy would mean most of the data falls within a particular class.
+With this in mind, the information gain $G$ of an attribute $A$ for the dataset $S$, or $G(S,A)$ in equation (\ref{eq:gain}), is the expected reduction in entropy caused by partitioning the dataset according to attribute $A$ \citep{Mitchell_1997_Book}.
+For selection of a threshold for a continuous attribute refer to \citet{Quinlan_1996_JAIR}.
+
+In the particular case of the C5.0 algorithm the result of this process depends on several parameters, logical and numerical.
+We set the options to do feature selection and evaluate possible advanced splits of data, but do not use the boosted models option \citep[see][]{Kuhn_2017_Manual}; and use the confidence factor (CF) and the smallest number of samples that must be placed in at least two of the splits ($S_{\min}$, also known as \textit{minCases}) to fine tune the prediction models.
+In particular, we build a suite of trees for varying values of CF and $S_{\min}$.
+
+Each tree resulting from this process has different qualities.
+We are interested in selecting a tree that is highly effective, but with a low level of complexity.
+In other words, we seek a tree with a good ratio of accuracy for predicting the classification outcome of the data, while using a reduced number of attributes (i.e., GOF metrics) in only a few number of steps (i.e., tree nodes).
+The number of metrics and the number of nodes are directly seen from the topology of the trees, and are often proportional (shallow trees tend to use less nodes).
+In general, smaller trees are preferable because they are easy to understand and often more accurate predictors \citep{Quinlan_1996_JAIR}.
+
+The effectiveness of the tree, on the other hand, needs to be measured. To that end we use the factor $F_{\beta}$ proposed by \citet{Rijsbergen_1979_Book} as
+% 
+\begin{equation}
+	\label{eq:f}
+	F_{\beta} = \frac{ (1 + \beta^2) P R}{\beta^2 P + R}
+	\, ,
+\end{equation}
+% 
+where $P$ and $R$ are the precision and recall factors, respectively, and $\beta$ is a weighting factor between the two.
+$P$ measures the fraction of data samples that are correctly classified within a category, whereas $R$ measures the fraction of the data classified within that category regardless of whether it was correctly classified or not.
+Mathematically, they are defined as
+% 
+\begin{equation}
+	P = \frac{ \mathrm{TP} }{ \mathrm{TP} + \mathrm{FP} } 
+	\quad \mathrm{and} \quad
+	R = \frac{ \mathrm{TP} }{ \mathrm{TP} + \mathrm{FN} }
+	\, .
+\end{equation}
+% 
+Here, TP, FP, and FN are the number of samples in the testing dataset considered to be true-positives, false-positives, and false-negatives, respectively.
+These values are typically ordered in what is known as a confusion matrix.
+For the simplest case of two categories, a confusion matrix takes the form
+% 
+\begin{equation}
+\nonumber
+\begin{array}{rr@{}rl}
+	&					&	\multicolumn{2}{c}{\mathrm{Prediction}}		\\
+	&					&	\mathrm{Postive}	&	\mathrm{Negative}	\\[0.75ex]
+	\mathrm{Actual~Value}		
+	&	\begin{array}{r}
+			\mathrm{Positive} \\[1ex]
+			\mathrm{Negative} 
+		\end{array}
+	&	\left[~
+		\begin{array}{c}
+			\mathrm{TP} \\[1ex]
+			\mathrm{FP} 
+		\end{array}\right.
+	&
+		\left.
+		\begin{array}{cc}
+			~\mathrm{FN} \\[1ex]
+			~\mathrm{TN} \\[1ex]
+		\end{array}~\right]
+\end{array}
+\end{equation}
+% 
+where TN is the number of true-negative samples.
+
+We compute $F_{\beta}$ in equation (\ref{eq:f}) using $\beta = 1$ to give equal weight to $P$ and $R$ \citep{McCarthy_1995_Proc}.
+This is done for all the trees obtained using all possible combinations of CF and $S_{\min}$ values.
+Then, we select the tree with the highest level of effectiveness as measure by $F_1$, commensurate to such tree having a low complexity, i.e., a reduced number of metrics and a shallow depth, which is precisely the goal set at the start.
+Note that $F_1$ is obtained based on the testing dataset as opposed to the training dataset.
+The results as applied to our dataset are presented next.
 
@@
 % Decision tree learning is a method for approximating a target function, in which learned function is represented by decision tree \citep{Mitchell_1997_Book}.
-\citet{Quinlan_1986} demonstrated that the technology for building decision trees from examples is fairly robust.
+\citet{Quinlan_1986_ML} demonstrated that the technology for building decision trees from examples is fairly robust.
 He summarized an approach to synthesizing decision trees that has been used in a variety of systems, and he describes one such system, ID3, in detail.
 Although ID3 algorithm is successful in most classification problems, it does not handle the numeric attributes and only one attribute at a time is tested for making decisions.
-\citet{Quinlan_1993} extended the ID3 algorithm and presented the C4.5 algorithm.
+\citet{Quinlan_1993_Book} extended the ID3 algorithm and presented the C4.5 algorithm.
 Later on they added boosting capability to C4.5 and called it C5.0.
@@
 Discussing the algorithm in details is out of the scope of this paper.
-There are fairly good amount of references to study those details \citep[e.g.,][]{Mitchell_1997_Book,Quinlan_1993,Hornik_2009}.\\
+There are fairly good amount of references to study those details \citep[e.g.,][]{Mitchell_1997_Book, Quinlan_1993_Book, Hornik_2009_CS}.
+
 % In general, decision tree measures the effectiveness of an attribute in classifying the training data through information gain.
@@
 If the target can take on $c$ different values the entropy is defined as
+% % 
 % \begin{equation}
-% Entropy(S) = \sum_{i=1}^{c} -p_{i}\ log_2\ p_i,
+% 	Entropy(S) = \sum_{i=1}^{c} \left[ -p_i \log_2 \left(p_i\right) \right]
+% 	\, ,
 % \end{equation}
-
+% % 
 % where $p_i$ is proportion of $S$ belonging to class $i$.
@@
 Gain of the attribute A is defined as
-
+% % 
 % \begin{equation}
@@
 % \end{equation} 
-
+% % 
 % where $Values(A)$ is the set of all possible values of attribute A, and $S_v$ is the subset of $S$ for which attribute $A$ has value $v$.
 Information gain is used by the algorithm in order to determine which is the better attribute for classifying the training example.
-The attribute with higher information gain will be the first option to separate the training data based on that attribute and grow to the lower nodes (Refer to chapter 3 of \citet{Mitchell_1997_Book} for more detailed examples.) \\
+The attribute with higher information gain will be the first option to separate the training data based on that attribute and grow to the lower nodes (Refer to chapter 3 of \citet{Mitchell_1997_Book} for more detailed examples.)
+
 % We discussed the basic idea behind the decision tree algorithm, however, there are more details specially in dealing with outliers and noises which we address them through pruning approaches.
 In this study we use C5.0 package \citep{C50_2015} in R programming language \citep{R_2016_program} to develop the decision tree algorithms.
-We discuss the input parameters in the section.\ref{classification_model}.
+We discuss the input parameters in the section \ref{classification_model}.

### 2018-02-02 introduction.tex
done with methodology

@@
 The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}.
-Upon labeling the dataset, we develop two decision trees using a C4.5 algorithm \citep[][]{Quinlan_1993_Book} to prioritize the metrics involved in the validation process.
+Upon labeling the dataset, we develop two decision trees using a C5.0 algorithm \citep[][]{Quinlan_1993_Book, Quinlan_1996_JAIR} to prioritize the metrics involved in the validation process.
 The decision trees narrow the number of metrics offering a prediction algorithm into the aforementioned validation categories.

### 2018-02-02 results.tex
done with methodology

\section{Results}
\label{sec:results}

\subsection{Preprocessing Results}

Working...

\subsection{Decision Trees}

Working...

In total, we built \textcolor{red}{XXX} trees corresponding to \textcolor{red}{XXX} CF values (varying between 0 and 1 at intervals of \textcolor{red}{0.1}) and \textcolor{red}{XXX} $S_{\min}$ values (varying between 1 and \textcolor{red}{XXX} at intervals of \textcolor{red}{XXX}).

% \input{analysis_cluster}

% \input{analysis_dtree}

### 2018-02-03 metrics.tex
did some cleaning

--- 2018-01-20 validation-metrics.tex
@@
 In his method, this frequency-domain decomposition should have pass bands defined following a logarithmic distribution in order to give more weight to the low-frequencies, and the GOF scores of all sub-bands and the broadband be averaged into a single GOF final score.
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{figures/pdf/figure-01}
+    \caption{\textbf{(a)} Region of interest used by \citet{Taborda_2014_BSSA} for the simulations of the 2008 $M_W ~ 5.4$ Chino Hills earthquake, including the epicenter, focal mechanism, and major quaternary faults.
+In the background, the main roads and topography are shown for visual reference.
+\textbf{(b)} Distribution of the 336 stations (gray dots) used by \citet{Taborda_2014_BSSA} for the validation analysis of their simulations within the modeling domain shown in part (a).
+Roads, city names, and the hillshade topography are also shown here in the background for reference.
+The color version of this figure is available only in the electronic edition.}
+    \label{fig:chino-hills}
+\end{figure*}
 

### 2018-02-03 results.tex
did some cleaning

@@
 Working...
+
+\begin{figure*}[t]
+	\centering
+	\includegraphics[width=\textwidth]{figures/pdf/figure-05}
+	\caption{A sample of the results from the multi-dimensional clustering analysis for pairs of consecutive metrics C1 through C11.
+These results are part of the 55 possible sub-spaces with 2 features (metrics).
+The cluster centers are shown with a cross symbol.
+Although for some pairs it may be obvious, the clusters 1 through 4 do not necessarily correspond in that order to the validation classes poor, fair, good, and excellent.
+The color version of this figure is available only in the electronic edition.}
+	\label{fig:clusters-sample}
+\end{figure*}
 

### 2018-02-03 buildpdf.sh

@@ -11,6 +11,6 @@ rm ${MAIN}.fff ${MAIN}.log ${MAIN}.aux ${MAIN}.ttt ${MAIN}.bbl ${MAIN}.lof ${MAI
 
 mv ${MAIN}.pdf ${FINAL}.pdf
 
-open -a Adobe\ Acrobat\ Pro.app ${FINAL}.pdf
+open -a Adobe\ Acrobat.app ${FINAL}.pdf
 
 clear

### 2018-02-04 results.tex
working on results

@@
 \subsection{Preprocessing Results}
-
-Working...
 
@@
 The color version of this figure is available only in the electronic edition.}
-	\label{fig:clusters-sample}
+	\label{fig:clusters}
 \end{figure*}
 
-\subsection{Decision Trees}
+As explained in the methodology, the first step is to subject the dataset through a clustering process to properly label the data according to the classes to be used by the decision tree algorithm.
+Figure \ref{fig:clusters} presents a sample of the results obtained after the clustering process is completed.
+Because visual representation above 3 dimensions is not possible, in this figure we limit ourselves to present two-dimensional subspaces, i.e., for two features (or metrics).
+In total, as we mentioned before, there are 55 subspaces with 2 features, and 550 total in our analysis (done for subspaces with 2, 3, and 4).
+
+Figure \ref{fig:clusters} shows 6 subspaces for an ordered sequence of metric pairs from C1 to C11 (see Table \ref{tab:metrics}).
+As it can be seen, not all feature combinations lead to clearly distinguishable clusters.
+In this sample we include well-partitioned sets (e.g., C3-C4, C5-C6, C7-C8), and some others where the data is all mixed (e.g., C1-C2, C10-C11).
+Note also that although some data distribution may hint the existence of clusters, it is considered irrelevant because knowledge about one feature does not add value to the other.
+That is the case of the C9-C10 subspace.
+Conversely, there are subspaces with a clear correlation between the data, but with indistinguishable clusters.
+That is the case of the C1-C2 subspace.
+Such subspaces do not contribute information to the clustering process \citep{Dy_2004_MLR}.
+
+\textcolor{red}{I will discard Figure 6, and will promote Figure 7 to be the new Figure 6 as the one showing the final distribution of the data after clustering is all done.}
 
 Working...
+
+\subsection{Decision Trees}
 

### 2018-02-04 results.tex
improved new figure 6

@@
 \label{sec:results}
-
-\subsection{Preprocessing Results}
-
-\begin{figure*}[t]
-	\centering
-	\includegraphics[width=\textwidth]{figures/pdf/figure-05}
-	\caption{A sample of the results from the multi-dimensional clustering analysis for pairs of consecutive metrics C1 through C11.
-These results are part of the 55 possible sub-spaces with 2 features (metrics).
-The cluster centers are shown with a cross symbol.
-Although for some pairs it may be obvious, the clusters 1 through 4 do not necessarily correspond in that order to the validation classes poor, fair, good, and excellent.
-The color version of this figure is available only in the electronic edition.}
-	\label{fig:clusters}
-\end{figure*}
 
@@
 
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{figures/pdf/figure-05}
+	\caption{A sample of the results from the multi-dimensional clustering analysis for pairs of consecutive metrics C1 through C11.
+These results are part of the 55 possible sub-spaces with 2 features (metrics).
+The cluster centers are shown with a cross symbol.
+Although for some pairs it may be obvious, the clusters 1 through 4 do not necessarily correspond in that order to the validation classes poor, fair, good, and excellent.
+The color version of this figure is available only in the electronic edition.}
+	\label{fig:clusters}
+\end{figure*}
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{figures/pdf/figure-06}
+	\caption{Statistical distribution of the dataset as partitioned into the four validation categories (poor, fair, good, excellent) after the clustering analysis.
+The distributions are shown in the form of box-plots for each metric (C1 through C11), and discretized in terms of the component of motion (NS, EW, UD) for visualization purposes.
+In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range, with outliers shown as scattered dots.
+The color version of this figure is available only in the electronic edition.}
+	\label{fig:boxed-clusters}
+\end{figure*}
+
+Figure 
 \textcolor{red}{I will discard Figure 6, and will promote Figure 7 to be the new Figure 6 as the one showing the final distribution of the data after clustering is all done.}
@@
 Working...
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/pdf/Figure_9.pdf}
+	\caption{Number of data in each cluster based on components after conducting constraint \kmeans{} clustering through subspace analysis.}
+	\label{fig:Figure_cluster_barplot}
+\end{figure}
+
+% Fig.~\ref{fig:C_cvm_comp} shows the results for different velocity models and components after clustering the dataset.
+A preliminary visually inspection shows that the behavior of scores with respect to the velocity model is fairly similar.
+Therefor we merge data of different velocity models.
+and present them at Fig.~\ref{fig:Figure_cluster_comp}.
+
+% \begin{figure*}
+%     \centering
+%     \includegraphics[width=\textwidth]{figures/pdf/Figure_6.pdf}
+%     \caption{Box plot of data used in the study after clustering.
+Metrics are shown for C1 to C11 for 3 components in separate plots for velocity models and clusters.
+Median of data are shown as notchs.
+Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data.
+Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
+%     \label{fig:C_cvm_comp}
+% \end{figure*}
+
+% \begin{figure*}
+%     \centering
+%     \includegraphics[width=\textwidth]{figures/pdf/Figure_7.pdf}
+%     \caption{Box plot of data used in the study after clustering and merging the results for different velocity models.
+Metrics shown for C1 to C11 for 3 components in separate plots for  clusters.
+Median of data are shown as notchs.
+Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data.
+Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
+%     \label{fig:Figure_cluster_comp}
+% \end{figure*}
+
+% Although there are some differences in the scores based on different components, there are general similarities in all clusters for different metrics among their components.
+Fig.~\ref{fig:Figure_cluster_all_comp} shows the results with merging the components.
+
+% Fig.~\ref{fig:Figure_cluster_all_comp} presents the results from a broad perspective.
+Statistical tests can be helpful to understand the effect of component, velocity model, earthquake, and frequency band in the results.
+According to Fig.~\ref{fig:Figure_cluster_all_comp} scores C1, C2, and C11 has the least changes in different clusters.
+It suggests the idea that they may not helpful in defining a decision rule.
+On the other hand, other scores are fairly variable among different clusters.
+% Scores C1, C2, and C10 are very similar in cluster 1 to 3.
+However, we can see a small increase in cluster 4 for these scores.
+Also the data variation becomes small and range of data moves higher.
+Scores C3-C9 have a considerable increment from cluster 1 to cluster 4.
+Although there are a broad variation of data in cluster 3 and cluster 4 for these scores, good amount of data (50\% based on the IQR) are in a small range and increasing with respect to the clusters.
+C11 has a constant behavior with a small reduction in median at cluster 4.
+Fig.~\ref{fig:Figure_cluster_barplot} shows number of stations in each cluster based on component.
 

### 2018-02-04 decision-tree.tex
making progress with results section

@@
 We are interested in selecting a tree that is highly effective, but with a low level of complexity.
-In other words, we seek a tree with a good ratio of accuracy for predicting the classification outcome of the data, while using a reduced number of attributes (i.e., GOF metrics) in only a few number of steps (i.e., tree nodes).
-The number of metrics and the number of nodes are directly seen from the topology of the trees, and are often proportional (shallow trees tend to use less nodes).
+In other words, we seek a tree with a good ratio of accuracy for predicting the classification outcome of the data, while using a reduced number of attributes (GOF metrics) in only a few number of steps (as given by a small number of decision nodes or a tree with shallow depth).
+The number of metrics, the number of nodes, and the depth of a tree are directly seen from the topology of the tree, and are often proportional (shallow trees tend to use less nodes, and thus less metrics).
 In general, smaller trees are preferable because they are easy to understand and often more accurate predictors \citep{Quinlan_1996_JAIR}.

### 2018-02-04 results.tex
making progress with results section

--- 2018-02-03 results.tex
@@
 
-\subsection{Preprocessing Results}
+As explained in the methodology, the first step is to subject the dataset through a clustering process to properly label the data according to the classes to be used by the decision tree algorithm.
+Figure \ref{fig:clusters} presents a sample of the results obtained after the clustering process is completed.
+Because visual representation above 3 dimensions is not possible, in this figure we limit ourselves to present two-dimensional subspaces, i.e., subspaces for two features (or metrics).
+Recall that there are 55 subspaces with 2 features, and a total of 550 for the sum of the subspaces with 2, 3, and 4 features.
 
-Working...
+In particular, Fig.~\ref{fig:clusters} shows six subspaces for an ordered sequence of metric pairs from C1 to C11 (see Table \ref{tab:metrics}).
+Although it would be impractical to discuss all the possible combinations, some points can be made from those shown in this figure.
+For instance, not all feature combinations lead to clearly distinguishable clusters.
+In this sample we include well-partitioned sets such as C3-C4, C5-C6, C7-C8, and some others where the data-points are mixed across classes, such as subspaces C1-C2 and C10-C11.
+Note also that although some data distribution may hint the existence of clusters, they may nonetheless be irrelevant because knowledge about one feature does not add insight about the other.
+That is the case of the C9-C10 subspace.
+Conversely, there are subspaces with a clear correlation between the data, but with indistinguishable clusters.
+That is the case of the C1-C2 subspace.
+These latter two types of subspaces do not contribute information to the clustering process \citep{Dy_2004_MLR}.
 
-\begin{figure*}[t]
+\begin{figure*}
 	\centering
@@
 The color version of this figure is available only in the electronic edition.}
-	\label{fig:clusters-sample}
+	\label{fig:clusters}
 \end{figure*}
 
-\subsection{Decision Trees}
+\begin{figure*}
+	\centering
+	\includegraphics[width=\textwidth]{figures/pdf/figure-06}
+	\caption{Statistical distribution of the dataset as partitioned into the four validation categories (poor, fair, good, excellent) after the clustering analysis.
+The distributions are shown in the form of box-plots for each metric (C1 through C11), and discretized in terms of the component of motion (NS, EW, UD) for visualization purposes.
+In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range, with outliers shown as scattered dots.
+The color version of this figure is available only in the electronic edition.}
+	\label{fig:boxed-clusters}
+\end{figure*}
 
-Working...
+Figure \ref{fig:boxed-clusters} shows the statistical distribution of the data in the form of box-plots for each metric and component of motion once partitioned into the four validation classes.
+Separately, we have prepared similar plots for the different velocity models used in the simulations from where we obtained our dataset.
+As in Fig.~\ref{fig:data-box-plot} before the clustering, and in Fig.~\ref{fig:data-box-plot} after the clustering process, we observe that the velocity models and the components of motion do not have a strong influence on the statistical distribution of the data, that is why we merge all the data samples into a single dataset at the before processing the data and we continue forward with that approach after clustering and labeling.
 
-In total, we built \textcolor{red}{XXX} trees corresponding to \textcolor{red}{XXX} CF values (varying between 0 and 1 at intervals of \textcolor{red}{0.1}) and \textcolor{red}{XXX} $S_{\min}$ values (varying between 1 and \textcolor{red}{XXX} at intervals of \textcolor{red}{XXX}).
+In total, the initial dataset is partitioned into groups of 816, 1253, 879 and 76 data samples for the poor, fair, good and excellent classes, respectively.
+These groups are shown in Fig.~\ref{fig:count-classes}, which also show the contribution of the component of motion to the final count of samples in each class for illustrative purposes.
+The main aspect to notice in this figure and the quantities just provided is the fact that the number of samples in the excellent class is significantly less than those of the poor, fair, and good classes.
 
-% \input{analysis_cluster}
+As a consequence, before moving on with the decision tree analysis, it is necessary to do resampling on the subset of the excellent class.
+In order to maintain a good number of samples for the other three classes, we used the oversampling approach described in the previous section.
+To that end, we replicated data randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes}.
+By common standards, an oversample rate of $\times$10 is considered acceptable \textcolor{red}{(reference)}.
+(Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary.) Once this process was completed we went on with the decision tree analysis.
 
-% \input{analysis_dtree}
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{figures/pdf/figure-07}
+	\caption{Number of data samples in each class (poor, far, good, excellent) after conducting a multi-dimensional constrained \kmeans{} clustering process using subspace analysis for 2, 3, and 4 features (GOF metrics).
+Here, for illustrative purposes, the filling of the bars distinguishes the component of motion of the compared signals as classified.
+The dashed-line bar indicates the number of samples in the excellent class after oversampling.}
+	\label{fig:count-classes}
+\end{figure}
+
+In total we generated 20,000 trees using the C5.0 algorithm for combinations of the parameters CF and $S_{\min}$.
+CF was chosen to vary between 0 and 1 at intervals of size 0.01, and $S_{\min}$ between 1 and 200 at unitary intervals.
+Not surprisingly, for such a small size intervals, the algorithm often reached the same tree for different CF and $S_{\min}$ values.
+Therefore, in reality, from the 20,000 we ran the algorithm, we only found 66 unique trees.
+For each one of these unique threes, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}) and extracted the total number of nodes in the trees and their depth.
+
+Figure... (working)

### 2018-02-05 method.tex
did some cleanng heare and there

\section{Data Analysis Method} 
\label{sec:method}

We are interested in proposing an algorithm with a reduced and prioritized number of validation metrics based on previously acquired validation data (i.e., our dataset). A common method to do this is to identify rules with disjunctive characteristics in the form of a decision tree that leads to an outcome representative of the overall analysis. In our case, we define such outcome in terms of four categories representative of the quality of the validation: poor, fair, good, excellent. The development of such a decision tree requires a proper classification of the data, which can be done through a clustering process. The following sections explain the data processing analysis we put in place to cluster the dataset and obtain two alternative decision trees.

\input{clustering}

\input{decision-tree}

### 2018-02-07 results.tex
modified figures 5, 6, 7 for results section. pending to modify text
work in progress update

@@
 
-As explained in the methodology, the first step is to subject the dataset through a clustering process to properly label the data according to the classes to be used by the decision tree algorithm.
-Figure \ref{fig:clusters} presents a sample of the results obtained after the clustering process is completed.
-Because visual representation above 3 dimensions is not possible, in this figure we limit ourselves to present two-dimensional subspaces, i.e., subspaces for two features (or metrics).
-Recall that there are 55 subspaces with 2 features, and a total of 550 for the sum of the subspaces with 2, 3, and 4 features.
+As explained in the methodology, the first step is to perform clustering process on the dataset in order to properly label the samples according to the classes to be used in the decision tree algorithm.
+Figure \ref{fig:clusters} presents a sample of the results obtained after the constrained \kmeans{} clustering process is completed, including subspacing.
+Recall that there are a total of 550 subspaces.
+Fig.~\ref{fig:clusters} shows three examples for each one of the types of subspaces, that is, three for each one of the 2-, 3-, and 4-feature analyses.
+However, since it is not practical or possible to present three- or higher-dimensional plots, we display the results from a two-dimensional perspective into the the subspaces by picking 2 metrics at a time.
+(Note that in the only cases in which this is a direct view into the clustering results is in the cases of the 2-feature analysis.)
 
-In particular, Fig.~\ref{fig:clusters} shows six subspaces for an ordered sequence of metric pairs from C1 to C11 (see Table \ref{tab:metrics}).
+% There are some characteristics of the results shown in Fig.~\ref{fig:clusters} worth
+
+subspaces for an ordered sequence of metric pairs from C1 to C11 (see Table \ref{tab:metrics}).
 Although it would be impractical to discuss all the possible combinations, some points can be made from those shown in this figure.
@@
 
-\begin{figure*}
+\begin{figure*}[ht!]
 	\centering
@@
 \end{figure*}
+
+Figure \ref{fig:boxed-clusters} shows the statistical distribution of the data in the form of box-plots for each metric and component of motion once partitioned into the four validation classes.
+Separately, we have prepared similar plots for the different velocity models used in the simulations from where we obtained our dataset.
+As in Fig.~\ref{fig:data-box-plot} before the clustering, and in Fig.~\ref{fig:data-box-plot} after the clustering process, we observe that the velocity models and the components of motion do not have a strong influence on the statistical distribution of the data, that is why we merge all the data samples into a single dataset at the before processing the data and we continue forward with that approach after clustering and labeling.
+
+In total, the initial dataset is partitioned into groups of 816, 1253, 879 and 76 data samples for the poor, fair, good and excellent classes, respectively.
+These groups are shown in Fig.~\ref{fig:count-classes}, which also show the contribution of the component of motion to the final count of samples in each class for illustrative purposes.
+The main aspect to notice in this figure and the quantities just provided is the fact that the number of samples in the excellent class is significantly less than those of the poor, fair, and good classes.
 
@@
 
-Figure \ref{fig:boxed-clusters} shows the statistical distribution of the data in the form of box-plots for each metric and component of motion once partitioned into the four validation classes.
-Separately, we have prepared similar plots for the different velocity models used in the simulations from where we obtained our dataset.
-As in Fig.~\ref{fig:data-box-plot} before the clustering, and in Fig.~\ref{fig:data-box-plot} after the clustering process, we observe that the velocity models and the components of motion do not have a strong influence on the statistical distribution of the data, that is why we merge all the data samples into a single dataset at the before processing the data and we continue forward with that approach after clustering and labeling.
-
-In total, the initial dataset is partitioned into groups of 816, 1253, 879 and 76 data samples for the poor, fair, good and excellent classes, respectively.
-These groups are shown in Fig.~\ref{fig:count-classes}, which also show the contribution of the component of motion to the final count of samples in each class for illustrative purposes.
-The main aspect to notice in this figure and the quantities just provided is the fact that the number of samples in the excellent class is significantly less than those of the poor, fair, and good classes.
-
 As a consequence, before moving on with the decision tree analysis, it is necessary to do resampling on the subset of the excellent class.
@@
 To that end, we replicated data randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes}.
-By common standards, an oversample rate of $\times$10 is considered acceptable \textcolor{red}{(reference)}.
+By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}.
 (Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary.) Once this process was completed we went on with the decision tree analysis.

### 2018-02-08 results.tex
added nodes-depth-attribute figures and worked on results section

@@
 
-As explained in the methodology, the first step is to perform clustering process on the dataset in order to properly label the samples according to the classes to be used in the decision tree algorithm.
-Figure \ref{fig:clusters} presents a sample of the results obtained after the constrained \kmeans{} clustering process is completed, including subspacing.
+As explained in the methodology, the first step is to carry out a clustering process on the dataset in order to properly label the samples according to the classes to be used in the decision tree algorithm.
+Figure \ref{fig:clusters} presents a sample of the results obtained after applying the constrained \kmeans{} method, including subspacing.
 Recall that there are a total of 550 subspaces.
-Fig.~\ref{fig:clusters} shows three examples for each one of the types of subspaces, that is, three for each one of the 2-, 3-, and 4-feature analyses.
+Figure \ref{fig:clusters} shows three examples for each one of the subspaces considered, that is, three for each one of the 2-, 3-, and 4-feature subspace analyses.
 However, since it is not practical or possible to present three- or higher-dimensional plots, we display the results from a two-dimensional perspective into the the subspaces by picking 2 metrics at a time.
-(Note that in the only cases in which this is a direct view into the clustering results is in the cases of the 2-feature analysis.)
+(Note that the only cases in which this is a direct view into the clustering results are the 2-feature analysis cases.)
 
-% There are some characteristics of the results shown in Fig.~\ref{fig:clusters} worth
-
-subspaces for an ordered sequence of metric pairs from C1 to C11 (see Table \ref{tab:metrics}).
-Although it would be impractical to discuss all the possible combinations, some points can be made from those shown in this figure.
-For instance, not all feature combinations lead to clearly distinguishable clusters.
-In this sample we include well-partitioned sets such as C3-C4, C5-C6, C7-C8, and some others where the data-points are mixed across classes, such as subspaces C1-C2 and C10-C11.
-Note also that although some data distribution may hint the existence of clusters, they may nonetheless be irrelevant because knowledge about one feature does not add insight about the other.
-That is the case of the C9-C10 subspace.
-Conversely, there are subspaces with a clear correlation between the data, but with indistinguishable clusters.
-That is the case of the C1-C2 subspace.
-These latter two types of subspaces do not contribute information to the clustering process \citep{Dy_2004_MLR}.
+Some aspects of Fig.~\ref{fig:clusters} are worth discussing.
+Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case.
+In the C5-C9 and C7-C9 combinations in the 3-feature analysis, for instance, the excellent cluster has a limited presence.
+This is due to the influence of the lower C9 values (see Fig.~\ref{fig:data-box-plot}) on the correlations with C5 and C7.
+This is not necessarily a bad thing, as it helps to understand the different relationships between the various GOF metrics.
+In the best of cases, the combinations C1-C2 and C5-C8 in the 2- and 4-feature analysis, for instance, exhibit an almost perfect proportionality between the metrics.
+These are called redundant relationships, where knowledge about a feature provides a direct view into another feature's results.
+On the other hand, the combination C1-C7 in the 2-feature analysis presents an example of an irrelevant relationship, where knowledge about a feature provides no insight about the other.
+Subspaces with redundant relationships are useful for reducing the number of features.
+Irrelevant relationships, however, do not contribute much to the whole process \citep{Dy_2004_MLR}.
 
@@
 	\includegraphics[width=\textwidth]{figures/pdf/figure-05}
-	\caption{A sample of the results from the multi-dimensional clustering analysis for pairs of consecutive metrics C1 through C11.
-These results are part of the 55 possible sub-spaces with 2 features (metrics).
-The cluster centers are shown with a cross symbol.
-Although for some pairs it may be obvious, the clusters 1 through 4 do not necessarily correspond in that order to the validation classes poor, fair, good, and excellent.
+	\caption{A sample of the results from the multi-dimensional clustering analysis showing the clusters from a two-dimensional perspective into the relationships between different GOF metrics.
+In each case, the labels near the upper-left corner indicate the features considered in the subspace analysis, and the features associated to the horizontal ($x$) and vertical ($y$) axes in the plot.
+The top, middle, and bottom rows correspond to 2-, 3-, and 4-feature subspace analyses.
+In each case, the poor, fair, good, and excellent clusters are indicated with circle, triangle, square, and cross symbols, respectively.
 The color version of this figure is available only in the electronic edition.}
@@
 
-Figure \ref{fig:boxed-clusters} shows the statistical distribution of the data in the form of box-plots for each metric and component of motion once partitioned into the four validation classes.
-Separately, we have prepared similar plots for the different velocity models used in the simulations from where we obtained our dataset.
-As in Fig.~\ref{fig:data-box-plot} before the clustering, and in Fig.~\ref{fig:data-box-plot} after the clustering process, we observe that the velocity models and the components of motion do not have a strong influence on the statistical distribution of the data, that is why we merge all the data samples into a single dataset at the before processing the data and we continue forward with that approach after clustering and labeling.
-
-In total, the initial dataset is partitioned into groups of 816, 1253, 879 and 76 data samples for the poor, fair, good and excellent classes, respectively.
-These groups are shown in Fig.~\ref{fig:count-classes}, which also show the contribution of the component of motion to the final count of samples in each class for illustrative purposes.
-The main aspect to notice in this figure and the quantities just provided is the fact that the number of samples in the excellent class is significantly less than those of the poor, fair, and good classes.
+Figure \ref{fig:boxed-clusters} shows the statistical distribution (box-plots) of the samples once the dataset is partitioned into the four validation classes.
+This is similar to Fig.~\ref{fig:data-box-plot}, but after the clustering process is completed.
+Separately, we prepared similar plots to look at the influence of the velocity models and components of motions on the results of the clustering process and, as observed before, they had no significant differences with the aggregate of all the samples in the dataset.
 
@@
 
-As a consequence, before moving on with the decision tree analysis, it is necessary to do resampling on the subset of the excellent class.
-In order to maintain a good number of samples for the other three classes, we used the oversampling approach described in the previous section.
-To that end, we replicated data randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes}.
+In total, the clustering process results in 816, 1253, 879 and 76 data samples for the poor, fair, good and excellent classes, respectively.
+These groups are shown in Fig.~\ref{fig:count-classes}.
+As it can be seen in this figure, the number of samples in the excellent class is significantly less than those in the poor, fair, and good classes.
+Therefore, before moving on with the decision tree analysis, it is necessary to resample the subset of the excellent class.
+We used the oversampling approach described in the previous section.
+We replicated data from the excellent class randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.
 By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}.
@@
 Here, for illustrative purposes, the filling of the bars distinguishes the component of motion of the compared signals as classified.
-The dashed-line bar indicates the number of samples in the excellent class after oversampling.}
+The dashed-line bar indicates the number of samples in the excellent class after oversampling.
+The color version of this figure is available only in the electronic edition.}
 	\label{fig:count-classes}
@@
 
-In total we generated 20,000 trees using the C5.0 algorithm for combinations of the parameters CF and $S_{\min}$.
-CF was chosen to vary between 0 and 1 at intervals of size 0.01, and $S_{\min}$ between 1 and 200 at unitary intervals.
-Not surprisingly, for such a small size intervals, the algorithm often reached the same tree for different CF and $S_{\min}$ values.
-Therefore, in reality, from the 20,000 we ran the algorithm, we only found 66 unique trees.
-For each one of these unique threes, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}) and extracted the total number of nodes in the trees and their depth.
+In total we generated 20,000 trees using the C5.0 algorithm for all possible combinations of the parameters CF and $S_{\min}$, where CF was chosen to vary between 0 and 1 at intervals of size 0.01, and $S_{\min}$ was chosen to vary between 1 and 200 at unitary intervals.
+Despite our choice for small intervals, the algorithm often reached recurrent tree structures for different CF and $S_{\min}$ values.
+Therefore, in reality, from the 20,000 combinations for which we ran the algorithm, only 66 unique trees were found.
 
-Figure... (working)
+For each one of these unique trees, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}), and extracted the total number of nodes in the trees and their depth.
+Figure \ref{fig:nodes-depths} shows the results of $F_1$ for all the trees and its distribution in terms of the number of attributes used in the trees as a function of the number of nodes and the depth of each tree.
+Recall that we are interested in finding a sequence of decisions (represented by disjunctive decision nodes in a tree) that can lead to good GOF predictions (i.e, high values of $F_1$) using a reduced number of attributes (GOF metrics).
+In general, all the trees obtained with the C5.0 algorithm are good in terms of the effectiveness factor ($F_1$ values close to 1).
+Then our choice comes down to using a reduced number of metrics.
+Having several trees with 2, 3, and 4 metrics (as opposed to 11), the following factor in the decision is choosing trees with algorithms using a small number of steps to reach the prediction.
+This is given by a combination between the depth of the tree and the number of nodes in the tree (i.e., trees with low complexity).
+
+\begin{figure*}[t]
+	\centering
+	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
+	\caption{Trees effectiveness results in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed by the number of attributes (GOF metrics) as a function of \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees.
+The rings around the left-most bottom dots indicate selected trees shown in Fig.~\ref{fig:trees}.
+The color version of this figure is available only in the electronic edition.}
+	\label{fig:nodes-depth}
+\end{figure*}
+
+Based on these criteria, we selected three candidate trees: T1, T2, and T3. These trees are shown in Fig.~\ref{fig:trees}.
+
+\begin{figure*}
+	\centering
+	% \includegraphics[width=\textwidth]{figures/pdf/figure-09}
+	\caption{}
+	\label{fig:trees}
+\end{figure*}

### 2018-02-08 results.tex
finished figure 9

@@
 	\centering
-	% \includegraphics[width=\textwidth]{figures/pdf/figure-09}
-	\caption{}
+	\includegraphics[width=0.86\textwidth]{figures/pdf/figure-09}
+	\caption{Selected trees \textbf{(a)} T1, \textbf{(b)} T2, and \textbf{(c)} T3.
+In each case the decision nodes of the trees contain the code corresponding to the metric (see Table \ref{tab:metrics}) used and the branches beneath each node show the limit value to be used to select the next level in the tree.
+At the bottom, normalized histograms are shown to indicate the distribution of the samples at each leaf-node according to the validation classes with codes P, F, G, and E for poor, fair, good, and excellent, respectively.
+At the top of each histogram is the total count of samples at the corresponding leaf node.
+The histograms highlight the dominant validation class.
+The color version of this figure is available only in the electronic edition.}
 	\label{fig:trees}

### 2018-02-08 results.tex
working on results section, final aspects

@@
 
-Based on these criteria, we selected three candidate trees: T1, T2, and T3. These trees are shown in Fig.~\ref{fig:trees}.
+Based on these criteria, we selected three candidate trees: T1, T2, and T3.
+These trees are shown in Fig.~\ref{fig:trees}.
+T1 is the simplest of the three, and T2 and T3 share some of the logarithmic structure on the branches to the right-hand side.
+T3 is the more complex of them.
+More complex trees tend to be deeper, use more nodes, and thus employ more attributes.
+In total T1, T2, and T3 use 2, 3 and 4 attributes respectively.
+All coincide in the use of the total energy (C4) and response spectra (C8) as key metrics.
+T2 adds peak acceleration (C5), and T3 adds peak velocity (C5) to the previous metrics.
 
@@
 \end{figure*}
+
+\begin{table}%[t]
+	\centering
+	\caption{Confusion matrix results for decision tree T1}
+	\label{tab:t1:confusion}
+	\small
+	\begin{tabular}{cccccc}
+		& 		& \multicolumn{4}{c}{Prediction}\\ 
+		\cline{3-6}
+		&   	& P  	& F     & G  	& E 	\\
+		\cline{2-6}
+		\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{Actual}}}  
+		& P 	& 242 	& 11 	& 0     & 0    	\\
+		& F 	& 13 	& 313   & 16    & 0    	\\
+		& G 	& 0 	& 42    & 207   & 15   	\\
+		& E 	& 0 	& 0     & 23    & 231	\\ 
+		\cline{2-6}
+	\end{tabular}
+\end{table}
+
+\begin{table}%[t]
+	\centering
+	\caption{Confusion matrix results for decision tree T3}
+	\label{tab:t3:confusion}
+	\small
+	\begin{tabular}{cccccc}
+		& 		& \multicolumn{4}{c}{Prediction}\\ 
+		\cline{3-6}
+		&   	& P  	& F     & G  	& E 	\\
+		\cline{2-6}
+		\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{Actual}}}  
+		& P 	& 248 	& 14 	& 0     & 0    	\\
+		& F 	& 7 	& 343   & 18    & 0    	\\
+		& G 	& 0 	& 9 	& 221   & 33   	\\
+		& E 	& 0 	& 0     & 7    	& 213	\\ 
+		\cline{2-6}
+	\end{tabular}
+\end{table}
+
+At the bottom of each tree in Fig.~\ref{fig:trees} are the histograms of the data that land on each leaf node.
+(Note that the count of samples here is done based on the training dataset.) The decision tree assigns the final validation category based on the dominant class in each leaf.
+As such, the samples in the second leaf-node from the right in T3 are categorized as good despite a significant portion of them being excellent.
+This is a natural trade-off embedded in the use of decision trees which actually results in more effective trees.
+To illustrate this, Tables \ref{tab:t1:confusion} and \ref{tab:t3:confusion} show the confusion matrix results for T1 and T3, respectively; and Tables \ref{tab:t1:f1} and \ref{tab:t3:f1} show the corresponding results for the $F_1$ scores.
+
+working...
+
+Values are obtained using the testing dataset. P, F, G, and E stand for the validation classes poor, fair, good, and excellent, respectively.
+
+% As we mentioned in reference to the results shown in Fig.~\ref{nodes-depth}, all trees led to relatively high values of the effectiveness score (i.e., $F_1 > 0.9$).
+As an example, Table \ref{tab:t1:confusion} shows the count of samples in the confusion matrix for tree T1 and Table \ref{tab:t1:f1} show the results of the precision and recall factors, along with the individual-leaves and leaves average $F_1$ scores.
+
+% All trees having high $F_1$ factors, we highlight (with darker color) the bars in those histogram where a dominance of a class is strong to help us understand their differences.

### 2018-02-09 results.tex
finished results section

@@
 
+\begin{table}
+	\centering
+	\caption{T1 precision, recall, and $F_1$ values per class}
+	\label{tab:t1:f1}
+	\small
+	\begin{tabular}{lccc}
+	Class   & $P$	& $R$  	& $F_1$ \\ 
+	\hline
+	P 		& 0.96 	& 0.95 	& 0.95 	\\
+	F 		& 0.92 	& 0.86 	& 0.88 	\\
+	G 		& 0.78 	& 0.84 	& 0.81 	\\
+	E 		& 0.91 	& 0.94 	& 0.92 	\\
+	\hline
+	Mean 	&		&		& 0.90	\\
+	\end{tabular}
+\end{table}
+
+\begin{table}
+	\centering
+	\caption{T3 precision, recall, and $F_1$ values per class}
+	\label{tab:t3:f1}
+	\small
+	\begin{tabular}{lccc}
+	Class   & $P$	& $R$  	& $F_1$ \\ 
+	\hline
+	P 		& 0.95 	& 0.97 	& 0.95 	\\
+	F 		& 0.93 	& 0.93 	& 0.93 	\\
+	G 		& 0.84 	& 0.89 	& 0.86 	\\
+	E 		& 0.96 	& 0.86 	& 0.91 	\\
+	\hline
+	Mean 	&		&		& 0.91	\\
+	\end{tabular}
+\end{table}
+
 At the bottom of each tree in Fig.~\ref{fig:trees} are the histograms of the data that land on each leaf node.
@@
 This is a natural trade-off embedded in the use of decision trees which actually results in more effective trees.
-To illustrate this, Tables \ref{tab:t1:confusion} and \ref{tab:t3:confusion} show the confusion matrix results for T1 and T3, respectively; and Tables \ref{tab:t1:f1} and \ref{tab:t3:f1} show the corresponding results for the $F_1$ scores.
 
-working...
+The actual effectiveness (i.e., the average $F_1$ values shown in Fig.~\ref{fig:nodes-depth}), however, is measured based on the testing dataset.
+Recall that $F_1$ depends on the number of true predictions and false predictions measured by the confusion matrix and the precision and recall factors.
+Tables \ref{tab:t1:confusion} and \ref{tab:t3:confusion} show the confusion matrices for T1 and T3, respectively; and Tables \ref{tab:t1:f1} and \ref{tab:t3:f1} show the corresponding results for $P$, $R$, and $F_1$.
+As it can be seen in the Tables \ref{tab:t1:confusion} and \ref{tab:t3:confusion}, both trees lead to strong diagonal confusion matrices, meaning that the classification into poor, fair, good and excellent is well defined.
+The differences between both matrices are actually minor and small with respect to their diagonal values, and the $F_1$ values in Tables \ref{tab:t1:f1} and \ref{tab:t3:f1} are all near to or above 0.9.
+The values obtained for tree T2 are consistent with this.
 
-Values are obtained using the testing dataset. P, F, G, and E stand for the validation classes poor, fair, good, and excellent, respectively.
-
-% As we mentioned in reference to the results shown in Fig.~\ref{nodes-depth}, all trees led to relatively high values of the effectiveness score (i.e., $F_1 > 0.9$).
-As an example, Table \ref{tab:t1:confusion} shows the count of samples in the confusion matrix for tree T1 and Table \ref{tab:t1:f1} show the results of the precision and recall factors, along with the individual-leaves and leaves average $F_1$ scores.
-
-% All trees having high $F_1$ factors, we highlight (with darker color) the bars in those histogram where a dominance of a class is strong to help us understand their differences.
+The final selection of a preferred tree, then, comes down to reducing the complexity.
+We favor tree T1 because it is based only in three decision steps, and two GOF metrics, C4: energy, and C8: response spectra.

### 2018-02-09 testing.tex
started verification test section

\section{Verification Test}

We now use the selected tree T1 on the original simulation results from \citet{Taborda_2014_BSSA}, for different velocity models and components of motion. The results obtained with the T1 validation algorithm are shown in Fig... \textcolor{red}{[work in progress]}.

### 2018-02-10 results.tex
prepared verification test figures

@@
 In the best of cases, the combinations C1-C2 and C5-C8 in the 2- and 4-feature analysis, for instance, exhibit an almost perfect proportionality between the metrics.
-These are called redundant relationships, where knowledge about a feature provides a direct view into another feature's results.
-On the other hand, the combination C1-C7 in the 2-feature analysis presents an example of an irrelevant relationship, where knowledge about a feature provides no insight about the other.
-Subspaces with redundant relationships are useful for reducing the number of features.
-Irrelevant relationships, however, do not contribute much to the whole process \citep{Dy_2004_MLR}.
+That means either one of the metrics in those combinations is redundant.
+Relationships with redundant features are those where knowledge about one of the features provides a direct view into the other.
+On the other hand, the combination C1-C7 in the 2-feature analysis presents an example of an irrelevant feature.
+We say C1 is irrelevant because it provides no insight about the outcome of C7.
+Identifying subspaces with redundant and irrelevant features is important because, on the one hand, the former help reduce the number of necessary features, while on the other hand, the latter can essentially be discarded because of their weak contribution to the decision making process \citep{Dy_2004_MLR}.
 
@@
 In each case, the poor, fair, good, and excellent clusters are indicated with circle, triangle, square, and cross symbols, respectively.
+The empty circles indicate the location of the artificial cannot-link stations we introduced as background knowledge to the clustering process.
 The color version of this figure is available only in the electronic edition.}
@@
 
-Figure \ref{fig:boxed-clusters} shows the statistical distribution (box-plots) of the samples once the dataset is partitioned into the four validation classes.
+Figure \ref{fig:boxed-clusters} shows the statistical distribution (box-plots) of the samples once the dataset is partitioned into the four GOF validation classes.
 This is similar to Fig.~\ref{fig:data-box-plot}, but after the clustering process is completed.
@@
 	\caption{Statistical distribution of the dataset as partitioned into the four validation categories (poor, fair, good, excellent) after the clustering analysis.
-The distributions are shown in the form of box-plots for each metric (C1 through C11), and discretized in terms of the component of motion (NS, EW, UD) for visualization purposes.
+The distributions are shown in the form of box-plots for each metric (C1 through C11).
 In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range, with outliers shown as scattered dots.
@@
 By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}.
-(Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary.) Once this process was completed we went on with the decision tree analysis.
+(Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary.
+Instead, we used a strong pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.
 
@@
 	\includegraphics[width=\columnwidth]{figures/pdf/figure-07}
-	\caption{Number of data samples in each class (poor, far, good, excellent) after conducting a multi-dimensional constrained \kmeans{} clustering process using subspace analysis for 2, 3, and 4 features (GOF metrics).
-Here, for illustrative purposes, the filling of the bars distinguishes the component of motion of the compared signals as classified.
+	\caption{Number of data samples in each class (poor, fair, good, excellent) after conducting a multi-dimensional constrained \kmeans{} clustering process using subspace analysis for 2, 3, and 4 features (GOF metrics).
 The dashed-line bar indicates the number of samples in the excellent class after oversampling.
@@
 For each one of these unique trees, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}), and extracted the total number of nodes in the trees and their depth.
-Figure \ref{fig:nodes-depths} shows the results of $F_1$ for all the trees and its distribution in terms of the number of attributes used in the trees as a function of the number of nodes and the depth of each tree.
+Figure \ref{fig:nodes-depth} shows the results of $F_1$ for all the trees and its distribution in terms of the number of attributes used in the trees as a function of the number of nodes and the depth of each tree.
 Recall that we are interested in finding a sequence of decisions (represented by disjunctive decision nodes in a tree) that can lead to good GOF predictions (i.e, high values of $F_1$) using a reduced number of attributes (GOF metrics).
@@
 	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
-	\caption{Trees effectiveness results in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed by the number of attributes (GOF metrics) as a function of \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees.
+	\caption{Accuracy of tree predictions in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed with respect to the number of attributes (GOF metrics) as a function of: \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees.
 The rings around the left-most bottom dots indicate selected trees shown in Fig.~\ref{fig:trees}.
@@
 These trees are shown in Fig.~\ref{fig:trees}.
-T1 is the simplest of the three, and T2 and T3 share some of the logarithmic structure on the branches to the right-hand side.
-T3 is the more complex of them.
-More complex trees tend to be deeper, use more nodes, and thus employ more attributes.
+T1 is the simplest of the three, and T2 and T3 share some of their topology on the right-hand side.
+T3 is the most complex of them.
+More complex trees tend to be deeper, have more nodes, and employ more attributes.
+This leads to higher $F_1$ values, but they may not necessarily be more practical.
 In total T1, T2, and T3 use 2, 3 and 4 attributes respectively.

### 2018-02-10 archived/discussion.tex
finished resutls section

\section{Discussion}

According to the results Response Spectra (C8) is the most important metric in classifying the stations. Data percentage in Table.~\ref{tab:attribute_usage_1} presents the amount of data that the attribute is used in classifying them.  After Response Spectra (C8), Total Energy (C4) is the most effective metric. With $C8 \leq 3.87~$ and $~ 3.87 < C8 \leq 6.1$ and $6.1 < C8~ \&~ C4 \leq 6.86$ and $6.86 \leq C4 ~ \&  6.1 \leq C8$ one can classify the simulation as poor, fair, good, and excellent, respectively, with a high confidence.  In this case minimum F1-score occurs for Class 3 and maximum F1-score occurs for Class 1. Pruning process reduce the effect of oversampling in cluster 4. In the second model, which is more accurate than the first model, other than Response Spectra and Total Energy, 2 other metrics including: Peak Velocity and Peak Acceleration are important in determining the final classes. With changing the tuning parameters (i.e., minClass and CF), which is increasing the functionality of the algorithm, other than the discussed metrics, some other metrics also get involved in the prediction model. However, in that case the model is prone to overfitting. We prefer a model with simple procedure as well as a fairly good F1-score.  Obviously, there are some stations that we cannot classify them correctly. It happens because of the prediction model and initial assumption about the class of stations. All prediction models have some uncertainty in predicting the final results.\\
  It is also relevant to study the variation of scores (specifically C8) with distance. Fig.~\ref{fig:C8dist} presents the variation of Response Spectra score in different distances for different velocity models and components. As one can see it is fairly well distributed for different components and velocity models with respect to distance. There is differences in median value in different components. The median value improves from CVM-H to CVM-S. In general UD components has better GOF score than horizontal components.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_15.pdf}
    \caption{Variation of C8 (Response Spectra) score with distance. Velocity models are represented in different shapes and components in different colors.  }
    \label{fig:C8dist}
\end{figure}

Analysis of statistical significance for relevance of results to earthquake, velocity model, frequency band, magnitude, distance of stations, and components needs comprehensive study beyond clustering and classification. If after a comprehensive study, one finds any significant factors, the factors should be used as an input in the prediction models. In this study, we illustrate the effects without making strong decision whether they affect the results or not. More comprehensive study is undergoing with \citet{Taborda_2016_GJI} dataset. \\

So far we developed two prediction models to get a pair of signals GOF scores and return their accuracy classes. However, in the case of real simulation, there are 3 components which they could have different GOF class. This situation is very similar to this study's original research question. In that case we had different metrics with different scores which make it hard to define the final score/class of a pair of signals. We addressed that issue using constrained \kmeans{} algorithm and decision tree model. However, in this case there is no background knowledge. To the authors' knowledge there is no study about the final class of a pair of stations based on different classes of each components. In this study we focus on developing a model to predict the accuracy of the GOF of a pair of stations based on individual GOF scores. However, we report the number of stations with different possible combination in Table~\ref{tab:station_comp_cvm}.

\begin{table}
\centering
\caption{Number of stations in different score combination}
\begin{tabular}{clccc}
\hline
No & GOF class combination                                                                                   & CVM-H+GTL    & CVM-H     &  CVM-S              \\ \hline
1 & All components are in the same class                                                               &   91               & 96               &  91                    \\
2 & Horizontal components are the same and Vertical has higher class                 &   66               & 80               &  70                     \\
3 & Horizontal components are the same and Vertical has lower class                 &    26              &  27               &   23                 \\
4 & One H and V components are the same and the other H has lower class       &   65               &  45               &  57                   \\
5 & One H and V components are the same and the other H has higher class      &   50                & 47              &  52                    \\
6 & All components classes are different and vertical has the highest class           &    27             & 27                &  31                    \\
7 & All components classes are different and vertical has the middle class            &      6             & 4                   &  3                     \\
8 & All components classes are different and vertical has the lowest class            &      5              & 10                &   9                      \\ \hline

\end{tabular}
\label{tab:station_comp_cvm}
\end{table}

According to Table~\ref{tab:station_comp_cvm}, majority of stations have all components at the same class. This makes it easier to classify them according to the common class. In the cases that horizontal components are in the same class, it is common to have a better match at the vertical component. In those cases where vertical component and one of horizontal components are in the same class, there is no any specific trend on the class of the other horizontal component, and it can be a better or worse match.  In those cases that all components are different, it is more common for vertical component to have a better class. It is very rare to have two horizontal component with two not successive classes. Fig.~\ref{fig:distance_comp_score_combination} presents the  Table~\ref{tab:station_comp_cvm} in a histogram plot for different velocity models. The histograms defined based on the number of stations in each categories and in radius of 10 km increments form the epicenter. We merge No.2 $\&$ 3 and 4 $\&$ 5, and 6 $\&$ 7 $\&$ 8 to have a better presentation in the histogram plots.\\

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_16.pdf}
    \caption{Number of stations with different combination of class for each components in 10 km incremental distances from epicenter.}
    \label{fig:distance_comp_score_combination}
\end{figure}

Number of stations with all components in the same class is about the same in all velocity models. CVM-H provides more stations with the same class. It is worth to mention that the common class can be any class out of the classes categories. Therefore, in this section we do not discuss the accuracy of the velocity model. In the stations closer to the epicenter  (less than 10 km) each velocity model has different behavior. In all distances (other than less than 10 km for CVM-S) all components with the same class or horizontal components with the same class is higher than all components are in different classes. Rotation of stations during or before an earthquake can affect the scores of horizontal components. In those cases the vertical components are trustable.  If two horizontal components are different, there is a chance of a problem in the orientation of the seismic station. Therefore, for stations with all components in the same group we consider the final class as the same, for stations where two horizontal components are different we consider the class of vertical component. According to the Table~\ref{tab:station_comp_cvm} at some cases one of the horizontal components has even better match, however, the mismatch between two horizontal components suggest the idea that the better class is not authentic. In those stations with two horizontal components are in the same class we consider that class as a final class. This option is conservative because the data show that in the case of two horizontal components in the same class it is common to the vertical component to have a better class.

Finally, we present the application of these models on classifying stations on a spatial geographical plots. Fig.~\ref{fig:M1_dtree_gof} shows the application of the first decision tree algorithm (M1) on the simulation of Chino Hills earthquake. In order to keep consistency in color code, and be able to comempare the results with other published results \citep[e.g.,][]{Taborda_2013_BSSA, Taborda_2014_BSSA, Taborda_2016_GJI} we assign 3,5,7, and 9 to cluster 1(poor),2(fair),3(good), and 4(excellent), respectively.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_17.pdf}
    \caption{M1 GOF-score for 3 velocity models and 3 components for 2008 $Mw 5.4$ Chino Hills earthquake simulation (max =4Hz) }
    \label{fig:M1_dtree_gof}
\end{figure}

Subsequently, Fig.~\ref{fig:M2_dtree_gof} represents the application of M2 metric on the data.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_18.pdf}
    \caption{M2 GOF-score for 3 velocity models and 3 components for 2008 $Mw 5.4$ Chino Hills earthquake simulation ($f_{max} =4Hz$) }
    \label{fig:M2_dtree_gof}
\end{figure}

As we discussed before, the class for different components could be different. We discussed the reasons for these differences and we can visually observe them in Fig.~\ref{fig:M1_dtree_gof,M2_dtree_gof}. We expect different results for 3 different velocity models. Although these models have many details in common (see \citet{Taborda_2014_BSSA} for more details), however, they have major differences and can cause different GOF scores. There are are more stations with "Excellent" class in model $M1$ than $M2$. That is because M1 is a simplified method and using only two metrics, predicts most of the stations as excellent class. Adding $C5$ to the decision process, $M2$ model tries to be more accurate in splitting data among "Good" and "Excellent" classes.\\
Although, the decision tree can be grown to very lower levels and be able to predict almost all the results accurately, however, due to more complicated algorithm for the user and also overfitting problem, we do not develop that complicated prediction model. However, if we take a look at $M2$  model, which we consider it as more accurate model than $M1$, out of 9 end nodes, 7 nodes (i.e., 3, 6, 8, 12, 13, 14, and 17) have very high probability of predicting the accurate model. There are two nodes (i.e., 7 and 16) which have fairly close probability of final classes. Adding more metrics with help to reduce this kind of uncertainties and achieve better classification.
In developing these models we did not leave out the possible noises or outliers. One approach can be removing all data that are out of Interquartile Range (IQR). However, not all data are considered as outliers for all metrics. Therefore, leaving out these data, we can loose many stations information. It could be a good independent study to go one by on for each of the stations with wrong classification and study the possible reasons for misclassification. Since we labeled the data by another algorithm , there is always a chance that the labeling system has some error. 
Looking at the confusion matrices (Table.~\ref{tab:confusionmat_test_1}, Table.~\ref{tab:confusionmat_test_2}) the algorithm in both models can predict the correct class with overall accuracy of 89\% and 92\% for M1,M2 models respectively. The accuracy is calculated based on dividing the sum of the number of true values (on diagonal) over sum of all numbers. In case of misclassification, there is only one class to misclassify. In other words there is no station with class 1 that the model classifies it as 3 or 4.  There is not station as class 2 that is classified as class 4. There is not stations with class 3 that is classified as class 1 and there is no model with class 4 that is classified as class 1 or 2.\\

These results suggest the idea that the prediction model is robust and the slight error is understandable and we tolerate that amount of error because of the trade of between more complicated formulation and the final error values. In the end, we generate the component independent class for the dataset for each velocity models based on earlier discussion.  Fig.~\ref{fig:CVM_final_class} present the results. CVM-S has more stations in "Excellent", "good", and "Fair" classes than CVM-H and CVM-H+GTL models and provide better results in general than the other two models.

\begin{figure}
    \centering
    \includegraphics
       % [width=\columnwidth]
        [width=\textwidth]
        {figures/pdf/Figure_19.pdf}
    \caption{Classification of GOF class of synthetic and data independent of components. Synthetic data are generated for 2008 Chino Hills Mw 5.4 earthquake in \citet{Taborda_2014_BSSA} with $f_{max} =4Hz$}
    \label{fig:CVM_final_class}
\end{figure}

### 2018-02-10 testing.tex
prepared verification test figures
finished resutls section

@@
 
-We now use the selected tree T1 on the original simulation results from \citet{Taborda_2014_BSSA}, for different velocity models and components of motion.
-The results obtained with the T1 validation algorithm are shown in Fig...
-\textcolor{red}{[work in progress]}.
+We now use tree T1 on the original simulation results from \citet{Taborda_2014_BSSA}, for different velocity models and components of motion.
+In this case we no longer aggregate all data from the simulation but now look at individual simulations done with different velocity models, and the three components of motion separately, as it would normally be done during any ground motion simulation validation.
+The results obtained with the T1 validation algorithm are shown in Fig.~\ref{fig:res-gof-maps}.
+These results are in essence very similar to those obtained by \citet{Taborda_2014_BSSA}, though using the T1 algorithm, we now rely only on the two metrics C4 and C8.
+
+A relative drawback from the T1 algorithm is that because its outcome are GOF validation classes (poor, fair, good, excellent) as opposed to GOF validation scores (with values from 0 to 10), then once one obtains the results for the validation process for different components of motion (as shown in the rows of Fig.~\ref{fig:res-gof-maps}) then there is no natural way of taking averages as one would do for scores in a numerical scale.
+Therefore, in order to combine results from the three components into a single validation classification for each station, we define the following rules:
+
+\begin{itemize}
+	\setlength\itemsep{0ex}
+	\item If the three components share the same validation class, then that class is assigned as the combination result.
+	\item If two components share the same validation class, but one differs, then the class shared by the former two is assigned as the combination result, thus favoring the majority.
+	\item If all three components have different validation classes, then the final combination result is set as the lower class of the three, penalizing the lack of conclusive results.
+\end{itemize}
+
+We call the result of applying these rules a tree T1 combination.
+Applying them to the validation results shown in Fig.~\ref{fig:res-gof-maps} for the case of the simulation done using the model CVM-S is shown in Fig.~\ref{fig:avg-gof-maps}, which compares the outcome with the result of the average scores obtained using the traditional \citet{Anderson_2004_Proc}-type GOF method.
+This figure exemplifies the difference between the numeric GOF validation-scores scale and the proposed T1 GOF validation classes.
+At the same time, it highlights the similarities between the outcomes.
+
+\begin{figure*}%[t]
+	\centering
+	\includegraphics[width=\textwidth]{figures/pdf/figure-10}
+	\caption{Results obtained using the validation algorithm of the selected decision tree T1 on the 2008 Chinno Hills earthquake simulation results reported by \citet{Taborda_2014_BSSA}.
+Here, the validation process is carried out separately for each component of motion (EW, NS, UD), for three different simulations corresponding to simulations done using the souther California velocity models: CVM-S, CVM-H, CVM-H+GTL.
+The contours maps are drawn for illustration purposes only based on the spatial distribution of the GOF validation class assigned to each station.
+The bottom histograms show the count of stations in each validation class.
+Stations are indicated with dots, and the epicenter with a Star.
+The color version of this figure is available only in the electronic edition.}
+	\label{fig:res-gof-maps}
+\end{figure*}
+
+In our view, the T1 results are simpler to assimilate and equally informative.
+Looking at both plots in Fig.~\ref{fig:avg-gof-maps} it is possible to argue they lead to the same conclusions and present a similar picture of the validation of the simulation, not only globally but at specific locations, with only minor differences---especially if one is to consider the assumed equivalence between scores and classes (0--4: poor; 4--6: fair; 6--8:good; 8--10: excellent).
+
+\begin{figure*}[th!]
+	\centering
+	\includegraphics[width=\textwidth]{figures/pdf/figure-11}
+	\caption{Comparison of GOF validation scores obtained using an 11-metric \citet{Anderson_2004_Proc}-type GOF scoring (left) and the T1 combination GOF validation classification (right).
+The top plots show maps with a distribution of the scores or classes outcome at each station.
+The contours are drawn for illustration purposes only.
+The bottom plots show histograms with the count of stations for each GOF score interval or validation class.}
+	\label{fig:avg-gof-maps}
+\end{figure*}

### 2018-02-10 acknowledgments.tex
completed abstract review, now working on introduction
finished first complete version

@@
-%
-\begin{acknowledgments}
-% 
-acknowledgement here
-% 
-\end{acknowledgments}
+\section{Acknowledgments}
+\small
+
+This work was supported by National Science Foundation (NSF) awards: ``SI2-SSI: Community Software for Extreme-Scale Computing in Earthquake System Science'' (ACI-1450451), and ``Improving Earthquake Forecasting and Seismic Hazard Analysis Through Extreme-Scale Simulations'' (OAC-1713792); and Southern California Earthquake Center (SCEC) awards: ``An Examination of the Correlations Between Different Goodness-of-Fit Metrics Based on a Large Dataset of Ground Motion Simulation Validations'' (16-067), and ``Application of Machine Learning in Deterministic Ground Motion Simulation'' (17-239).
+SCEC is funded by NSF Cooperative Agreement EAR-1600087 and U.S.~Geological Survey Cooperative Agreement G17AC00047.
+The SCEC contribution number for this paper is 8015.

### 2018-02-10 conclusions.tex
finished first complete version

--- 2017-08-03 conclusions.tex
@@
 
-In this study we use two machine learning algorithms to address the long time question about choosing the appropriate metrics to evaluate the ground motion simulation process.
+We present the result of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of prioritize and reduce the number of metrics.
+As a result of the data processing analysis, we propose a simplified algorithm based on a decision tree which uses only two metrics as opposed to the initial eleven available in the dataset.
+In particular, the proposed algorithm uses a three disjunctive decisions based on comparisons between data and synthetics in terms of the total energy and acceleration response spectra.
+We also propose rules to allow for the new class-based validation criteria to combine results obtained from different components of motion, and demonstrate that the results obtained with the proposed algorithm using two metrics are comparable to those obtained with the score-based validation results used in other recent validation studies.
+Similarly, one could implement similar rules to combine results from a frequency-band analysis, using, for example, the mode or majority validation class.
+
+We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulation done for a particular earthquake in a particular region.
+In a future follow-up study it would be ideal to refine the tree adding additional simulations from different earthquakes, regions, models, etcetera, in order to arrive to a sound and robust decision tree.
+The procedural steps laid out here, nonetheless, remains valid---and it is likely that the metrics of energy and response spectra will prevail as those among the most decisive ones.
+
+This latter point is perhaps the most important conclusion of our study.
+Identifying the total energy and response spectra metrics as the decisive ones is a key contribution to validation of ground motion simulations.
+On the one hand, this contributes to clarifying a standing question in the area of validation, and on the other hand, it signals to simulators where the focus should be in the search for improvements in their models.
+
+% OLD NAEEM VERSION
+% -----------------
+
+% In this study we use two machine learning algorithms to address the long-term question about choosing the appropriate metrics to evaluate the ground motion simulation process independent of application and components.
 Assuming hypothetical stations, we labeled the dataset through constraint \kmeans{} algorithm with subspace analysis.
-Using the labeled data we develop a decision tree algorithm to evaluate the ground motion simulation.
-We generate two models where the first model only uses two metrics to evaluate the stations and the second model which is more accurate and also more complicated than the first model uses 5 metrics.
+Using the labeled data we developed two decision trees algorithm to evaluate the ground motion simulation.
+We generate two models where the first model only uses two metrics to evaluate the stations and the second model which is more accurate and also more complicated than the first model uses 4 metrics.
 According to our analysis, Response Spectra is the most important metric in classifying the stations.
 After response spectra, Total Energy are mostly used.
-We present the results for separate velocity models and components, however, a comprehensive study is needed to address the significance of the results or variations to earthquake magnitude, velocity model, component, station distance from epicenter, earthquake depth, and so on.
-The proposed model could be used with high accuracy in determining the accuracy of simulation uniformly among physics based ground motion simulation developers.
+We studied different combination of components's classes and developed a model to predict a components independent class.
+We applied the models on 2008 Chino Hills earthquake simulation synthetics and data GOF scores and discussed the results.
+A comprehensive study is needed to address the significance of the results or variations to earthquake magnitude, velocity model, station distance from epicenter, earthquake depth, and so on.
+The proposed model could be used with high accuracy (about 90 \%) in determining the accuracy of simulation independent of application and components, uniformly among physics-based ground motion simulation developers.

### 2018-02-10 data-resources.tex
finished first complete version

\section{Data and Resources}

Calculations, data processing, and some initial figures were done using R, the language and environment for statistical computing and graphics (\url{https://www.r-project.org}, last accessed February 2018). Additional calculations and figures were done using MATLAB B (\url{http://www.mathworks.com}, last accessed February 2018). Map figures were prepared using the Generic Mapping Tools, GMT (\url{http://gmt.soest.hawaii.edu/}, last accessed February 2018). Additional editing of figures was done using Adobe Illustrator (\url{http://www.adobe.com/Illustrator‎}, last accessed February 2018). The validation dataset used here was readily available to the authors and can be provided without restrictions upon request.

### 2018-02-11 abstract.tex
did some cleanng heare and there
som edits before submission

@@
 This allows us to label a given validation dataset into four categories (poor, fair, good, and excellent) following common practice.
-We then develop two decision trees using a C5.0 algorithm.
-These decision trees help narrow the number of metrics leading to a validation prediction into the four categories.
+We then develop a family of decision trees using a C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four categories.
 These decision trees can be understood as rapid predictors of the quality of a simulation, or as a data-informed classifiers that can help prioritize validation metrics.
 Our analysis indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.
-
-% ===========================================================================================
-%
-% OLD NAEEM VERSION
-% 
-% The validation of ground motion synthetics has received increased attention over the last few years due to the advances in physics-based deterministic and hybrid simulation methods.
-Unlike for low frequency simulations $(f \le 0.5 Hz)$, for which it has become reasonable to expect a good match between synthetics and data, in the case of high-frequency simulations $(1 Hz \le f)$ it is not possible to match results on a wiggle-by-wiggle basis.
-This is mostly due to the various complexities and uncertainties involved in earthquake ground motion modeling.
-Therefore, in order to compare synthetics with data we turn to different time series metrics, which are used as means to characterize how the synthetics match the data on qualitative and statistical sense.
-In general, these metrics provide GOF scores that measure the level of similarity in the time and frequency domains.
-It is common for these scores to be scaled from 0 to 10, with 10 representing a perfect match.
-Although using individual metrics for particular applications is considered more adequate, there is no consensus or a unified method to classify the comparison between a set of synthetic and recorded seismograms when the various metrics offer different scores.
-We study the relationship among these metrics through semi-supervised and supervised learning approaches in order to labeling and classifying data, respectively.
-In particular, we use constrained \kmeans{} clustering method, in which we define 4 hypothetical stations with scores 3, 5, 7, and 9 for all metrics.
-We put these stations in the category of cannot-link constraints.
-We generate the dataset through the validation of the results from a deterministic (physics-based) ground motion simulation for a moderate magnitude earthquake in the greater Los Angeles basin using three velocity models.
-The maximum frequency of the simulation is 4 Hz.
-The dataset involves over 300 stations and 11 metrics, or features, as they are understood in the machine learning lingo, where the metrics form a multi-dimensional space.
-We address the high-dimensional feature effects with a subspace-clustering analysis using 2,3, and 4 dimensional spaces.
-We label the dataset in each subspace as poor, fair, good, and excellent, where, hypothetical stations with score of 3,5,7, and 9 belong them, respectively.
-We assign the final label based on the mode of subspaces for each station.
-Once the dataset has been labeled, we develop two decision trees using C5.0 algorithm as prediction models to classify a simulation in four groups (poor, fair, good, and excellent) based on a reduced number of metrics.
-We address the unbalanced dataset issue through oversampling method and analyze the precision, recall and F1-score of the classifiers.
-The prediction models represent a simple, yet effective model to accurately classify the accuracy of the simulation process based on application-independent approach.
-We analyze the different combinations of simulation accuracy classes for each components and present a guideline for making decision on accuracy level of simulation independent of components.
-We apply the proposed method on 2008 Chino Hills earthquake data and synthetic and present the results.

### 2018-02-11 conclusions.tex
som edits before submission

@@
 
-We present the result of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of prioritize and reduce the number of metrics.
+We present the result of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of prioritize and reduce the number of metrics, and develop an application independent decision algorithm.
 As a result of the data processing analysis, we propose a simplified algorithm based on a decision tree which uses only two metrics as opposed to the initial eleven available in the dataset.
-In particular, the proposed algorithm uses a three disjunctive decisions based on comparisons between data and synthetics in terms of the total energy and acceleration response spectra.
-We also propose rules to allow for the new class-based validation criteria to combine results obtained from different components of motion, and demonstrate that the results obtained with the proposed algorithm using two metrics are comparable to those obtained with the score-based validation results used in other recent validation studies.
+In particular, the proposed algorithm uses three (disjunctive decision) steps based on the values obtained for the metrics of the total energy and acceleration response spectra.
+we also propose rules to allow for the new class-based validation criteria to combine results obtained from different components of motion, and demonstrate that the results obtained with the proposed algorithm using two metrics are comparable to those obtained with the score-based validation results used in other recent validation studies.
 Similarly, one could implement similar rules to combine results from a frequency-band analysis, using, for example, the mode or majority validation class.
 
-We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulation done for a particular earthquake in a particular region.
-In a future follow-up study it would be ideal to refine the tree adding additional simulations from different earthquakes, regions, models, etcetera, in order to arrive to a sound and robust decision tree.
-The procedural steps laid out here, nonetheless, remains valid---and it is likely that the metrics of energy and response spectra will prevail as those among the most decisive ones.
+We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulations done for a particular earthquake in a particular region.
+In a future follow-up study it would be ideal to refine the tree adding other simulation datasets (i.e., from different earthquakes, regions, models, etcetera) in order to arrive to a sufficiently sound and robust decision tree.
+The procedural steps laid out here, nonetheless, remains valid.
+We showed that the procedure and background information used for clustering and decision making is stable, and it is likely that the metrics of energy and response spectra (along with peak acceleration and velocity as suggested by the additional trees) will prevail as those among the most decisive ones.
 

### 2018-02-11 data-resources.tex
som edits before submission

@@
 
-Calculations, data processing, and some initial figures were done using R, the language and environment for statistical computing and graphics (\url{https://www.r-project.org}, last accessed February 2018).
-Additional calculations and figures were done using MATLAB B (\url{http://www.mathworks.com}, last accessed February 2018).
+Calculations, data processing, and some initial figures were done using R, the language and environment for statistical computing and graphics (\url{https://www.r-project.org}, last accessed February 2018), and the C5.0 package developed by \citet{Kuhn_2017_Manual}.
+Additional calculations and figures were done using MATLAB (\url{http://www.mathworks.com}, last accessed February 2018).
 Map figures were prepared using the Generic Mapping Tools, GMT (\url{http://gmt.soest.hawaii.edu/}, last accessed February 2018).

### 2018-02-11 dataset.tex
did some cleaning
did some cleanng heare and there
som edits before submission

--- 2018-01-26 study-dataset.tex
@@
 % 
-\begin{figure}
+\begin{figure}[ht!]
     \centering
@@
     \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component of motion (NS, EW, UD).
-In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range, with outliers shown as scattered dots.
+In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots.
 The color version of this figure is available only in the electronic edition.}
@@
 It is clear that although there are differences between them, these are negligible; in other words, the statistical distribution of results for each metric is about the same independently of model or component.
-
-% ===========================================================================================
-%
-% OLD NAEEM VERSION
-% 
-% State of California is an earthquake prone region with a long history of seismicity.
-The Los Aneles basin, due to its unique geologic sediments, has been studied from different perspective.
-Numerous high resolution velocity models are developed mainly to conduct forward ground motion simulation.
-Occurrence of frequent moderate magnitude earthquakes made the Los Angeles basin as a natural laboratory for seismic  studies specially testing the forward ground motion simulation modeling.
-In a relatively recent study, \citet{Taborda_2014_BSSA} conducted a comprehensive study in order to assess the functionality of developed velocity models in the Los Angeles basin and surrounding areas.
-They simulated the 2008 Chino Hills earthquake using 3 velocity models including: CVM-S4, CVM-H, and CVM-H+GTL (for more details about these models see \citet{Taborda_2014_BSSA}.)
-% They computed the GOF scores based on \citet{Anderson_2004_Proc}, with minor modification introduced by \citet{Taborda_2013_BSSA} as discussed in the validation section.
-Fig.~\ref{fig:epicenters} shows the location of the epicenter and seismic stations.
-
-% \begin{figure}
-%     \centering
-%     \includegraphics
-%        % [width=\columnwidth]
-%         [width=450px]
-%         {figures/pdf/Figure_1}
-%     \caption{(a) Region of interest and epicentral location of the 2008 $M_W ~ 5.4$ Chino Hills earthquake (highlighted).
-Focal mechanism is shown at the margin of the region of interest.
-Major quaternary faults in the area are shown in the back along with the main roads and county divisions.
-(b) Simulation area of interest.
-Dots indicate the location of the 336 stations considered in this study for validation.
-The main local and interstate roads are also shown here.
-In both panels, the background shows the hillshade topography of the region.
-The color version of this figure is available only in the electronic edition.}
-%     \label{fig:epicenters}
-% \end{figure}
-
-% We use \citet{Taborda_2014_BSSA} database where they computed 11 scores for 3 different velocity models.
-In this study we only analyze the results for broadband which in this case is \feq{0.1}{4}.
-Fig.~\ref{fig:data_box_plot} shows the box plot of all data that we used in this study which we separate them in velocity models and components.
-In total we use $336 \times 3 \times 3 = 3024$ individual pair of data and synthetic.
-
-% \begin{figure}
-%     \centering
-%     \includegraphics
-%        % [width=\columnwidth]
-%         [width=\textwidth]
-%         {figures/pdf/Figure_2.pdf}
-%     \caption{Box plot of data used in the study.
-Metrics are shown for C1 to C11 for 3 components and in separate plots for velocity models.
-Median of data are shown as a notch.
-Thick lines represent the IQR (Interquartile Range, Q3-Q1) of data.
-Outliers (data less than Q1-1.5*IQR and greater than Q3+1.5*IQR) are shown as scatter dots above and below plots if applicable.}
-%     \label{fig:data_box_plot}
-% \end{figure}
-
-% As we explained before, we use all data set regardless of velocity models and components.
-However, distinguishing data for velocity model and components can answer the question that if the goodness of fit scores dependent on them.
-Addressing this research question is beyond the scope of this paper.
-We distinguish data based on velocity models and components only in presenting data or results.
-
-% Fig.~\ref{fig:data_density}  shows the density distribution of each score for CVM-S.
-The figure shows the variation of different metrics with components.
-Except in cross correlation (C10) in all metrics the up-down (UD) component scores are about the same or higher than horizontal components.
-
-% \begin{figure}
-%     \centering
-%     \includegraphics
-%        % [width=\columnwidth]
-%         [width=\textwidth]
-%         {figures/pdf/Figure_3.pdf}
-%     \caption{Distribution of data (only CVM-S) for different components in different scores based on different components. Each distribution has a unit area.}
-%     \label{fig:data_density}
-% \end{figure}
+Consequently, we combine all the data into a single dataset.

### 2018-02-11 introduction.tex
did some cleanng heare and there
som edits before submission

--- 2018-01-21 introduction.tex
@@
 The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}.
-Upon labeling the dataset, we develop two decision trees using a C4.5 algorithm \citep[][]{Quinlan_1993_Book} to prioritize the metrics involved in the validation process.
+Upon labeling the dataset, we develop a family of decision trees using a C5.0 algorithm \citep[][]{Quinlan_1993_Book, Quinlan_1996_JAIR}, and select a few to prioritize the metrics involved in the validation process.
 The decision trees narrow the number of metrics offering a prediction algorithm into the aforementioned validation categories.
@@
 We test our prediction model and offer a discussion about its implications and potential use in future validation efforts.
-
-% ===========================================================================================
-%
-% OLD NAEEM VERSION
-% 
-% The validation of ground motion synthetics has received increased attention over the last few years due to the advances in physics based deterministic and hybrid simulation methods.
-Unless in very low frequencies (\fleq{0.5}), due to limitations and uncertainties in the ground motion simulation and velocity models, it is not possible to match the simulation and synthetics wiggle by wiggle.
-Therefor, in order to compare synthetic seismogram with recorded data as a complex time series, different metrics are defined.
-These metrics are means to characterize the how well the synthetic matches the statistical characteristic of observed data.
-In general, these metrics compute the goodness-of-fit (GOF) between data and synthetic in time and frequency domains.
-\citet{Anderson_2004_Proc} proposed 10 metrics to be used to characterize the GOF in different frequency bands.
-\citet{Kristekova_2006_BSSA,Kristekova_2009_GJI} developed envelope and phase misfit criteria as well as GOF metrics for realistic scenarios.
-\citet{Olsen_2010_SRL} presented an alternative GOF metrics for broadband synthetics with relatively high resolution of the small misfits, and \citet{Taborda_2013_BSSA} added modification to \citet{Anderson_2004_Proc} metrics.
-
-% These metrics have been widely used in different studies.
-Recently, due to advances in computational resources, 3D ground motion simulation on a regional scale were conducted in different regions using different methods \citep[e.g.,][]{Taborda_2016_GJI,Komatitsch_2004_BSSA,Bielak_2010_GJI,Roten_2016_GJI}.
-The accuracy of simulation with observed data are measured using these GOF scores.
-However, researchers use these metrics in different perspectives which is highly dependent on their application.
-For example, for structural engineers who are interested in designing tall buildings, among other metrics, displacement could be an important factor.
-However, for rigid and semi-rigid structures like dams and nuclear power plants, acceleration could be more important criteria.
-In general, for all purposes, the duration of strong ground morion could be an important factor.
-Analysis of the strong ground motion simulation for a specific application based on individual metrics is an acceptable practice, however, validating the simulation process only with individual score is not a valid approach.
-Previous studies show that metrics could have completely different scores for a pair of synthetic and data \citep[e.g.,][]{Taborda_2016_GJI}.
-\citet{Taborda_2013_BSSA} defined a metric from a combination of  \citet{Anderson_2004_Proc} scores presenting a general overview of GOF.
-However, the metric is a linear combination and somehow is biased by the authors' preferences about the metrics.
-
-% Physics-based ground motion simulation communities push the maximum frequency  of the simulation to higher values (\fmax{}$\geq 5 ~ Hz$) and there is a high demand for a metric to uniformly validate the simulations.
-Although \citet{Anderson_2004_Proc}, defined different level for classifying the stations as poor, fair, good, and excellent based on individual metrics, there is no consensus on the most important metrics, specially when they have significantly different GOF score.
-Therefore, the question "How one can estimate the accuracy of simulation based on all scores or which metrics can be good representatives of simulation accuracy on behalf of other metrics?" has not addressed.
-This research question motivated us to approach the problem from different perspective.
-In other words, if we assume we know the accuracy of the simulation process (let's say we classified it as good) what is the relationship between different GOF scores for that pair of stations or which metrics are the best representer in classifying the stations.
-However, as far as the authors knowledge, there is no metric to be able to classify a pair of stations based on GOF scores as an estimation of the overall accuracy of the simulation.
-Therefore, as a first step, if we label the pair of station for simulation accuracy we can discuss the relationships between GOF scores.
-To do this we need a dataset where involves GOF scores for a good amount of stations.
-
-% \citet{Taborda_2014_BSSA} conducted a comprehensive study to analyze the functionality of different velocity models in the Los Angeles basin.
-They used 2008 Chino Hills earthquakes based on 3 velocity models and conducted simulations with maximum frequency of \fmaxeq{4}.
-They compared the accuracy of the velocity models through using the GOF scores.
-The study generated an invaluable synthetic and data  GOF scores dataset.\\
-% The first and the most common method in categorizing multidimensional unlabeled data is clustering and specifically the distance based methods like \kmeans{} \citep{Macqueen_1967_Proc}.
-However, there is a drawback in this method.
-If the data is not explicitly distinguishable the final clusters are highly dependent on initial starting points of the algorithm.
-Therefor, for stations with 11 dimensions, the ordinary \kmeans{} approach results in fairly different clusters after each analysis.
-The main reason for this issue is the algorithm solely looks for data which are close together in terms of Euclidean distance.
-However, in our case we have one more extra information.
-Hypothetically, if we have a pair of synthetic and data with GOF score equal 3 for all metrics, according to \citet{Anderson_2004_Proc} we can classify the station as poor with a strong confidence.
-This is also true for GOF score of 5, 7, and 9 for fair, good, and excellent classes, respectively.
-This background knowledge about the database could give a strong formation to the clustering process.
-\citet{Wagstaff_2001_Proc} defined a \kmeans{} clustering using background knowledge.
-Adding this feature to the algorithm improves the functionality and most importantly converge to the same clusters with any random initial clusters' centers.
-Using this approach alongside the subspace analysis (more details are presented in the section.~\ref{clustering}) we cluster the stations, then in a supervised learning algorithm, we develop a classification algorithm based on their GOF scores.
-We tune the learning algorithm and present the functionality of each prediction model based on different metrics.
-It is possible to have different classes for different components of a pair if synthetic and data.
-We study the different combination of components' classes in different distances and propose a rough estimate on deciding about the final GOF class of a pair of synthetic and data which is application and component independent.\\
-% In the following section,we present a summary of the validation metrics then we discuss the dataset.
-Later on we explain the methodology.
-We provide a brief summary of unsupervised  and semi-supervised learning which are ordinary and constrained clustering approaches, respectively.
-We also discuss the basics of decision tree algorithm as a supervised learning method.
-We discuss the clustering challenges and use subspace analysis in the clustering analysis section.
-We also discuss the imbalanced data effects and solutions in the classification model section.
-We develop two decision tree classifiers (prediction model) based on labeled data for future use in comparing data and synthetics.
-We propose an algorithm to give a rough estimate to have component independent prediction.
-We conclude the study by applying the generated classification methods on \citet{Taborda_2014_BSSA} dataset and present the spatial variation of the simulation goodness of fit classes on the Los Angles basin for three velocity models and three components in component dependent and for three velocity models in component independent approaches.

### 2018-02-11 dataset.tex
final round of edits--hopefully--before submission

@@
     \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component of motion (NS, EW, UD).
-In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots.
+In each case, the median is indicated by a notch in the box, the boxes represents the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), and the line shows the range of the data, with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots.
 The color version of this figure is available only in the electronic edition.}

### 2018-02-11 results.tex
finished first complete version
som edits before submission
final round of edits--hopefully--before submission

@@
 On the other hand, the combination C1-C7 in the 2-feature analysis presents an example of an irrelevant feature.
-We say C1 is irrelevant because it provides no insight about the outcome of C7.
+We say C1 is irrelevant because it provides no insight about the outcome of C7 or that of the clusters.
 Identifying subspaces with redundant and irrelevant features is important because, on the one hand, the former help reduce the number of necessary features, while on the other hand, the latter can essentially be discarded because of their weak contribution to the decision making process \citep{Dy_2004_MLR}.
@@
 
+\begin{figure*}[t]
+	\centering
+	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
+	\caption{Accuracy of tree predictions in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed with respect to the number of attributes (GOF metrics) as a function of: \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees.
+The rings around some of the dots indicate select trees used as reference in the text, other figures and tables.
+The color version of this figure is available only in the electronic edition.}
+	\label{fig:nodes-depth}
+\end{figure*}
+
+\begin{table}[t]
+	\centering
+	\caption{Confusion matrix results for decision tree T1}
+	\label{tab:t1:confusion}
+	\small
+	\begin{tabular}{cccccc}
+		& 		& \multicolumn{4}{c}{Prediction}\\ 
+		\cline{3-6}
+		&   	& P  	& F     & G  	& E 	\\
+		\cline{2-6}
+		\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{Actual}}}  
+		& P 	& 242 	& 11 	& 0     & 0    	\\
+		& F 	& 13 	& 313   & 16    & 0    	\\
+		& G 	& 0 	& 42    & 207   & 15   	\\
+		& E 	& 0 	& 0     & 23    & 231	\\ 
+		\cline{2-6}
+	\end{tabular}
+\end{table}
+
+\begin{table}[t]
+	\centering
+	\caption{Confusion matrix results for decision tree T3}
+	\label{tab:t3:confusion}
+	\small
+	\begin{tabular}{cccccc}
+		& 		& \multicolumn{4}{c}{Prediction}\\ 
+		\cline{3-6}
+		&   	& P  	& F     & G  	& E 	\\
+		\cline{2-6}
+		\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{Actual}}}  
+		& P 	& 248 	& 14 	& 0     & 0    	\\
+		& F 	& 7 	& 343   & 18    & 0    	\\
+		& G 	& 0 	& 9 	& 221   & 33   	\\
+		& E 	& 0 	& 0     & 7    	& 213	\\ 
+		\cline{2-6}
+	\end{tabular}
+\end{table}
+
 For each one of these unique trees, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}), and extracted the total number of nodes in the trees and their depth.
@@
 This is given by a combination between the depth of the tree and the number of nodes in the tree (i.e., trees with low complexity).
-
-\begin{figure*}[t]
-	\centering
-	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
-	\caption{Accuracy of tree predictions in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed with respect to the number of attributes (GOF metrics) as a function of: \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees.
-The rings around the left-most bottom dots indicate selected trees shown in Fig.~\ref{fig:trees}.
-The color version of this figure is available only in the electronic edition.}
-	\label{fig:nodes-depth}
-\end{figure*}
 
@@
 
-\begin{table}%[t]
-	\centering
-	\caption{Confusion matrix results for decision tree T1}
-	\label{tab:t1:confusion}
-	\small
-	\begin{tabular}{cccccc}
-		& 		& \multicolumn{4}{c}{Prediction}\\ 
-		\cline{3-6}
-		&   	& P  	& F     & G  	& E 	\\
-		\cline{2-6}
-		\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{Actual}}}  
-		& P 	& 242 	& 11 	& 0     & 0    	\\
-		& F 	& 13 	& 313   & 16    & 0    	\\
-		& G 	& 0 	& 42    & 207   & 15   	\\
-		& E 	& 0 	& 0     & 23    & 231	\\ 
-		\cline{2-6}
-	\end{tabular}
-\end{table}
-
-\begin{table}%[t]
-	\centering
-	\caption{Confusion matrix results for decision tree T3}
-	\label{tab:t3:confusion}
-	\small
-	\begin{tabular}{cccccc}
-		& 		& \multicolumn{4}{c}{Prediction}\\ 
-		\cline{3-6}
-		&   	& P  	& F     & G  	& E 	\\
-		\cline{2-6}
-		\parbox[t]{2mm}{\multirow{4}{*}{\rotatebox[origin=c]{90}{Actual}}}  
-		& P 	& 248 	& 14 	& 0     & 0    	\\
-		& F 	& 7 	& 343   & 18    & 0    	\\
-		& G 	& 0 	& 9 	& 221   & 33   	\\
-		& E 	& 0 	& 0     & 7    	& 213	\\ 
-		\cline{2-6}
-	\end{tabular}
-\end{table}
-
-\begin{table}
+\begin{table}[t]
 	\centering
@@
 
-\begin{table}
+\begin{table}[t]
 	\centering
@@
 
-The final selection of a preferred tree, then, comes down to reducing the complexity.
-We favor tree T1 because it is based only in three decision steps, and two GOF metrics, C4: energy, and C8: response spectra.
+Another aspect of interest is the level of participation of each metric in the analysis of data as GOF values are run through the decision nodes in the trees.
+The results of such level of participation are listed in Table \ref{tab:participation} for trees T1, T2, T3, and T66.
+(The tree T66 is the most complex of them all as inferred from the number of metrics involved and the number of nodes and levels as seen in Fig.~\ref{fig:nodes-depth}.) The percentages in this table represent the amount of data that is seen by a decision node associated by any particular metric.
+The nodes are not unique and metrics are often used in different nodes in a given tree, thus these percentage are cumulative for each individual metric.
+A percentage of 100 means a given metric has the opportunity to see all data samples through a single nodes or multiple ones.
+A low percentage means only a few data samples are evaluated by that metric, and a null percentage means the metric plays no role in the decision tree (it is not present in any node).
+
+Notice that almost independently of the tree (T3 being the exception), the metrics C4 (total energy) and C8 (response spectra) are consistently the most relevant ones in the decision algorithms.
+They are followed by C5 (peak acceleration) and C6 (peak velocity).
+On the other end, notice that C11 (strong phase duration) plays no role whatsoever in any of the trees, and that C1 (Arias integral), C2 (energy integral), and C3 (Arias intensity) have only small to marginal participations.
+The remaining metrics---C7 (peak displacement), C9 (Fourier spectrum), and C10 (Cross correlation)---have a somewhat significant participation, but only in the more complex trees.
+
+A low participation does not necessarily mean that a given metric is not important.
+Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
+What the results we present here mean, in the context of GOF validation, is that in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more decisive in reaching an accurate conclusion about validation quality than the duration (C11).
+
+In the end, the final selection of a preferred tree comes down to reducing complexity in the analysis without compromising the interpretation of results.
+We favor tree T1 because it is based only in three decision steps, and two GOF metrics: C4 (total energy), and C8 (response spectra).
+
+\begin{table}[t]
+	\centering
+	\caption{Metrics data analysis participation in select trees (in percent)}
+	\label{tab:participation}
+	\small
+	\begin{tabular}{lrrrr}
+	Metric 	& T1 	& T2	& T3 	& T66 	\\ 
+	\hline
+	C1 		& --	& --	& -- 	& 1.5	\\
+	C2 		& -- 	& --	& -- 	& 0.5	\\
+	C3 		& -- 	& -- 	& -- 	& 9.6	\\
+	\rowcolor{light-light-gray}
+	C4 		& 46.8	& 52.4	& 60.2	& 100.0	\\
+	\rowcolor{light-light-gray}
+	C5 		& -- 	& 21.5	& 21.5	& 72.2	\\
+	\rowcolor{light-light-gray}
+	C6 		& -- 	& -- 	& 37.0	& 56.3	\\
+	C7 		& -- 	& -- 	& -- 	& 19.4	\\
+	\rowcolor{light-light-gray}
+	C8 		& 100.0	& 100.0	& 100.0	& 100.0	\\
+	C9 		& -- 	& -- 	& --	& 23.8	\\
+	C10		& -- 	& -- 	& -- 	& 22.2	\\
+	C11		& -- 	& -- 	& -- 	& -- 	\\
+	\hline
+	\end{tabular}
+\end{table}

### 2018-02-11 results.tex
fixing size of figures for submission, added style package for submission
last changes before submission

@@
 Figure \ref{fig:clusters} shows three examples for each one of the subspaces considered, that is, three for each one of the 2-, 3-, and 4-feature subspace analyses.
-However, since it is not practical or possible to present three- or higher-dimensional plots, we display the results from a two-dimensional perspective into the the subspaces by picking 2 metrics at a time.
+However, since it is not practical or possible to present three- or higher-dimensional plots, we display the results from a two-dimensional perspective into the subspaces by picking 2 metrics at a time.
 (Note that the only cases in which this is a direct view into the clustering results are the 2-feature analysis cases.)
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/pdf/figure-07}
+	\includegraphics[width=0.45\textwidth]{figures/pdf/figure-07}
 	\caption{Number of data samples in each class (poor, fair, good, excellent) after conducting a multi-dimensional constrained \kmeans{} clustering process using subspace analysis for 2, 3, and 4 features (GOF metrics).
@@
 
-Notice that almost independently of the tree (T3 being the exception), the metrics C4 (total energy) and C8 (response spectra) are consistently the most relevant ones in the decision algorithms.
+Notice that in all trees the metrics C4 (total energy) and C8 (response spectra) are consistently the most relevant ones in the decision algorithms.
 They are followed by C5 (peak acceleration) and C6 (peak velocity).

### 2018-02-11 testing.tex
finished first complete version
last changes before submission

@@
 
-\begin{figure*}%[t]
+\begin{figure*}%[th!]
 	\centering
 	\includegraphics[width=\textwidth]{figures/pdf/figure-10}
-	\caption{Results obtained using the validation algorithm of the selected decision tree T1 on the 2008 Chinno Hills earthquake simulation results reported by \citet{Taborda_2014_BSSA}.
-Here, the validation process is carried out separately for each component of motion (EW, NS, UD), for three different simulations corresponding to simulations done using the souther California velocity models: CVM-S, CVM-H, CVM-H+GTL.
+	\caption{Results obtained using the validation algorithm of the selected decision tree T1 on the 2008 Chino Hills earthquake simulation results reported by \citet{Taborda_2014_BSSA}.
+Here, the validation process is carried out separately for each component of motion (EW, NS, UD), for three different simulations corresponding to simulations done using the southern California velocity models: CVM-S, CVM-H, CVM-H+GTL.
 The contours maps are drawn for illustration purposes only based on the spatial distribution of the GOF validation class assigned to each station.

### 2018-02-11 abstract.tex
final changes before submission

@@
 % 
-It has become common practice to validate ground motion simulation based on a variety of time- and frequency-domain metrics scaled to quantify the level of agreement between synthetics and data, or with respect to other reference solutions.
-There is, however, no agreement about the importance or weight that it ought to be given to each individual metric.
+It has become common practice to validate ground motion simulations based on a variety of time and frequency metrics scaled to quantify the level of agreement between synthetics and data, or with respect to reference solutions.
+There is, however, no agreement about the importance or weight that it ought to be given to each metric.
 This leads to their selection often being subjective, either based on intended applications or personal preferences.
-As a consequence, it is difficult for simulators to identify what modeling improvement are needed, which would be potentially easier if they could focused on a reduced number of alternative metrics.
-We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques used to label and classify goodness-of-fit results, with the objective prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation.
+As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could focused on a reduced number of metrics.
+We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques that help label and classify goodness-of-fit results, with the objective prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation.
 In particular, we study the relationships that exist between eleven different metrics.
@@
 We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address high-dimensional effects.
-This allows us to label a given validation dataset into four categories (poor, fair, good, and excellent) following common practice.
+This allows us to label a given validation dataset into four categories (poor, fair, good, excellent) following common practice.
 We then develop a family of decision trees using a C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four categories.

### 2018-03-28 response.tex
response draft added.

\setlength{\paperwidth}{210mm}
\setlength{\paperheight}{297mm}% fixed.

\documentclass{article}

\usepackage[review]{myreviewpckg}

\usepackage{textcomp}
\usepackage{simplemargins}

\clubpenalty=10000  % Orphan - First of paragraph left behind
\widowpenalty=10000 % Widow  - Last of paragraph sent ahead

\setallmargins{1in}

\begin{document}

% ******************************************************************
% *************************** REVIEWER 1 ***************************
% ******************************************************************

\begin{center}
	\bf
	\large
	Authors' Response to Reviewer 1
\end{center}

\noindent
We thank the reviewer for his/her comments and suggestions, which helped us improve the manuscript. To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font. In the following, we provide a response to each comment. The original comments are in italic black font, followed by our responses in regular blue font.
\vspace{2ex}
\newline

\introcomment{~}{%
This is an interesting article for the readers of the BSSA journal. The authors propose a new approach based on machine learning techniques to determine the importance of validation metrics for ground motion simulation. The field of ground motion simulation validation (GMSV) is relatively new and of interest to both seismologists who develop simulation models and engineers who would like to use these simulations in engineering applications. One of the most challenging questions to address is "which validation metric(s) is best to use?" as there are so many different metrics in this field and so many different opinions. It has been difficult to address this question because the answer varies depending on the application of the simulations (e.g., GMPE development, PSHA, structural analysis response of interest, etc). This paper is interesting because it provides a statistical approach to reduce the number of validation metrics and select a few representatives from a bigger set without compromising the final validation results. The authors have done a great job providing background on the topic. However, their application is very narrow, they only consider the 11 Anderson2004+ parameters (as they've noted there are other parameters) for one simulation method for one earthquake, and as a result their conclusions are only valid for this particular case. They have mentioned this limitation in the conclusions, but the abstract and introduction do not explain the limitations and the application-dependency of the proposed approach. Both Abstract and Introduction need modification to specify the application for which the results and conclusions are true (i.e., the
application here is to narrow down the 11 Anderson2004+ parameters to select fewer parameters that give similar results as the original 11 for the simulations they chose in this paper). Currently, the Abstract and Introduction give the impression that the conclusions are true for any engineering application or simulation method. Furthermore, there should be a short discussion in the paper on how the users can apply the same method for different applications, for example for structural responses such as drift ratio or responses that are sensitive to high-frequency content of simulations (theoretically, this method should be applicable to any response of interest if they replace the GOF scores used in this study? If true this should be mentioned as it will be of interest to others).
}

\response{%
Will be added. 
}

\comment{1}{%
The terminologies used in the paper are not clear in many cases. For example, what are "features" "attributes" "dimensions" "trees, nodes and leafs" "class"? It was very difficult to
read the section on Data Analysis Method and Introduction since these terminologies were not clearly defined or described in terms of the variables used in this paper. I was confused what
"data" and "results" were referring to: GOF score data? GOF metric data? GOF category? Something other than GOF? I suggest that the authors streamline the paper and the terminologies to be consistent and clear in the final version.
}

\response{%
response
}

\comment{2}{%
The terms "common practice" or "common standards" have been mentioned several times. Because GMSV is such a new field (and I don't think there is consensus, I for one don't agree
with some statements but I agree that others may), I think these terms should be removed. Just say this is your opinion, or it is based on the references that you have listed.
}

\response{%
Response 
}

\comment{3}{%
In abstract, "semi-supervised" and "supervised" learning techniques are mentioned, but in the paper these terms are not explained, what part of the learning techniques is semi-supervised
and what does that mean?
}

\response{%
Response
}

\comment{4}{%
In introduction, the difference between "verification" and "validation" should be discussed.
}

\response{%
Response
}

\comment{5}{%
In abstract and introduction, what does a "direct application" mean? it's not clear. Similarly, in Validation Metrics section: "direct quantitative comparisons" and "indirect statistical analysis"
}

\response{%
Response
}

\comment{6}{%
Page 4, line 53: add "relatively" before "well understood"
}

\response{%
Response
}

\comment{7}{%
The two questions posed in the introduction lines 59-61: I don't think the first one is really the question, all parameters should be considered if anyone has thought of them as important (i.e.,
has proposed them). The second question is the main question and the one being addressed in this paper.
}

\response{%
Response
}

\comment{8}{%
You are treating these metrics as "independent variables", but as you have also noted many of them are not. For example, Arias intensity and energy are correlate. Does this compromise your conclusions? If not explain why in the Results section.
}

\response{%
Response
}

\comment{9}{%
Page 7, line 133: delete the last sentence
}

\response{%
Response
}

\comment{10}{%
Page 14, line 310: you say "... the latter can essentially be discarded..." I don't understand this statement, wouldn't this mean that both C1 and C7 are important to consider since they are
independent of each other (one gives information that the other one doesn't)?
}

\response{%
Response
}

\comment{11}{%
Page 14 line 323: It's not clear how you "replicate data", provide more detail.
}

\response{%
Response
}

\comment{12}{%
I don't think "confusion matrix" given in tables is explained in the paper!
}

\response{%
Response
}

\comment{13}{%
Page 17, lines 391-394: this sentence should be re-worded, what is an "accurate conclusion"? You should be very specific that your conclusion is that instead of using all 11 validation metrics you can select a few in order to get the same categorization as all 11 metrics would have given you.}

\response{%
Response
}

\comment{14}{%
Page 18, line 405: this is implying that fig 10 is similar to fig2! I personally cannot see this in the figures! They look very different. It might be better to regenerate fig 11a but with different color scales that only vary by four choices, then the comparison would be much easier.
}

\response{%
Response
}

\comment{15}{%
Figure 8: I don't know what you mean by y-label "decision attributes". Do you mean "the number of GOF metrics"? why there is no data point for 1 or 11?
}

\response{%
Response
}

\comment{16}{%
The conclusion basically says that response spectrum is the most important validation metric. This is not surprising to any engineer or modeler. The big question is what other parameters
describe the differences between a simulation and a recorded motion if their response spectra matched? Or in other words, what are the other important validation parameters aside from
response spectrum (many have argued that's duration, but no statistical proof). Can you use your method to answer this question? Can you condition your machine learning process on Sa
being the same (for example use a subset of data with similar Sa or artificially make them equal), then decide what's the next important parameter? Would the result be the same as in your
conclusions? Or would it be influenced by removal of Sa as a parameter (this is possible since there are correlations between Sa and other metrics).
}

\response{%
Response
}

Typos:\\
\\

\comment{1}{%
Page 2, line 8: change focused to focus
}

\response{%
Response
}

\comment{2}{%
Page 2, line 20: change classifiers to classifier
}

\response{%
Response
}

\comment{3}{%
Page 13, line 185: change measure to measured
}

\response{%
Response
}

\comment{4}{%
Page 16, line 354: the last C5 should be C6?
}

\response{%
Response
}

% ******************************************************************
% *************************** REVIEWER 2 ***************************
% ******************************************************************

\newpage

\begin{center}
	\bf
	\large
	Authors' Response to Reviewer 2
	\end{center}

\noindent
We thank the reviewer for his/her comments and suggestions, which helped us improve the manuscript. To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font. In the following, we provide a response to each comment. The original comments are in italic black font, followed by our responses in regular blue font.
\vspace{2ex}
\newline

\introcomment{~}{%
Review of Khoshnevis and Taborda for BSSA
}

\introcomment{~}{%
Earthquake ground motion simulations are usually validated against observational data or empirically-derived solutions. However, exactly which parameters to compare (e.g., peak
ground velocity, duration, spectral content) in order to assess the goodness-of-fit, is not always clear. This paper uses machine learning techniques and eleven possible metrics for comparing (recorded and simulated) ground motions to determine the parameters that are most predictive of goodness-of-fit. This is an interesting paper, and will help future researchers
prioritize which metrics to focus on when validating ground motions simulations. The paper is generally well written. As a strong ground motion seismologist (i.e., I assume I'm part of the audience you hope to reach), there is a lot of machine learning-related jargon that makes certain sections difficult to follow. It would also be useful to have some discussion regarding which parameters were found to be repetitive, incase future authors wish to look at more than the four recommended parameters.
}

\comment{1}{%
L. 8 "... could \textbf{focus} ..."
}

\response{
Response
}

\comment{2}{%
L. 11 "... the objective \textbf{of} ..."
}

\response{
Response
}

\comment{3}{%
L. 25-41 The frequent use of phrases such as "Among these methods,"; "Among the latter,"; "In this category,"; and "Within this group," make the text difficult to follow.
}

\response{
Response
}

\comment{4}{%
L. 42-43 I would suggest moving the description of the Anderson (2004) method (i.e., L. 48+) to directly follow this sentence.
}

\response{
Response
}

\comment{5}{%
L. 59-61/L. 391-394 It should be stated somewhere that this is really intended for \textbf{engineering purposes}.
}

\response{
Response
}

\comment{6}{%
L. 61 This sentence should end with a "?"
}

\response{
Response
}

\comment{7}{%
L. 98-99 It is not clear how the equation for S is applied to values that are a function of period (e.g., the response or Fourier spectrum). Do you follow the guidelines provided by Anderson? If
so, what frequency range/bands are you examining?
}

\response{
Response
}

\comment{8}{%
L. 128/132/137/etc. "... the three component\textbf{s} ..."
}

\response{
Response
}

\comment{9}{%
L. 236 Where is the discussion on the consequences of overfitting? The only other place I see overfitting mentioned is L. 328.
}

\response{
Response
}

\comment{10}{%
L. 236 Remove "A matter we discuss later."
}

\response{
Response
}

\comment{11}{%
L. 241-243 It is not helpful to list different algorithm names for making decision trees, without any information about them or how they differ.
}

\response{
Response
}

\comment{12}{%
L. 257-258 I don't understand this sentence. What is a continuous attribute?
}

\response{
Response
}

\comment{13}{%
L. 259-264 This paragraph contains a lot of jargon.
}

\response{
Response
}

\comment{14}{%
L. 300 Isn't this also true for the C5-C7 combination?
}

\response{
Response
}

\comment{15}{%
L. 328 What is "a strong pruning process"?
}

\response{
Response
}

\comment{16}{%
L. 433 "We present the result\textbf{s} of..."
}

\response{
Response
}

\comment{17}{%
L. 435 "... goal of prioritiz\textbf{ing} and reduc\textbf{ing}..."
}

\response{
Response
}

\comment{18}{%
L. 443 Remove "Similarly" or "similar" (redundant)
}

\response{
Response
}

\comment{19}{%
L. 450 "... remain valid."
}

\response{
Response
}

\comment{20}{%
L. 455 "... as the decisive \textbf{parameters} ..." ? (This short concluding paragraph would be much stronger without the use of unnecessary phrases, i.e., "This latter point", "One the one hand",
"on the other hand."
}

\response{
Response
}

\end{document}

### 2018-03-30 response.tex
First draft of revision.

@@
 
+% *****************************************************************
+% *************************Associate Editor *********************
+% *****************************************************************
+
+\begin{center}
+	\bf
+	\large
+	Authors' Response to Associate Editor (Dr. Thomas Brocher)
+\end{center}
+
+\noindent
+We thank Dr.
+Thomas Brocher for his comments and suggestions, which helped us improve the manuscript.
+To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font and the removed text in red font.
+In the following, we provide a response to each comment.
+The original comments are in italic black font, followed by our responses in regular blue font.
+\vspace{2ex}
+\newline
+
+\introcomment{~}{%
+I agree with the peer reviewers that this paper addresses a topical subject of great importance to BSSA readers.
+The science is in the process of transitioning from purely observational based analysis of ground motions to a mixture of observations and simulations of ground motions.
+The paper is a useful contribution to this discussion, is generally well written, and carefully
+describes the methodology used and its limitations. The peer reviewers have made some very helpful suggestions for improving the paper that I hope the authors will strongly consider.\\
+In particular, I agree with their comment that the limitations of the study, i.e., the choice of 11 metrics and earthquake simulations for only one earthquake, needs to be included in the abstract and the conclusions.
+In addition, I also agree that all the terms and methodology should be well defined for the BSSA readership, so that the paper can be understood by them without resort to frequent online searches.
+I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit.
+It clearly shows that only these metrics are change significantly with the goodness of fit (P,F, G, E).
+Plotting these values as a function of P, F, G, and E together on the same figure might allow the reader to see this
+relationship more clearly and might help provide a heuristic understanding of how C4, C5, C6, and C8 are diagnostic whereas C11 is constant with P, F, G, and E, and thus not diagnostic.
+In fact, in my opinion, Figure 6 is a simple and easy way of determining which metrics are diagnostic that might even eliminate the need for the decision trees.
+In any case, more
+discussion of this figure might help the authors make the case for the main conclusions.
+}
+
+\response{%
+
+Thanks for the constructive suggestions.
+We agree with you and peer reviewers that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
+The confusion happened because of integrating three different topics into a method to carry out this research.
+One technical term in one field is different in another field and that cause a confusion.
+For example cluster, class, and label all are referring to the same concept, however, it depends on the application.
+We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.\\
+We also modified Figure 6 in order to easily track the variation of different metrics for final classes.
+We do agree with you that Figure 6 clearly gives a good understanding about the metrics, however, we add decision trees, because they can convert the qualitative judgment to quantitative conditional form to be able to exactly decide on the final class based on their numerical GOF scores.
+\\
+We added text to abstract and conclusion to give a heads up to readers about 11 metrics and one earthquake with 3 velocity model  simulations.\\
+}
+
 % ******************************************************************
@@
 % ******************************************************************
-
+\newpage
 \begin{center}
@@
 We thank the reviewer for his/her comments and suggestions, which helped us improve the manuscript.
-To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font.
+To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font and the removed text in red font.
 In the following, we provide a response to each comment.
@@
 \response{%
-Will be added. 
+The reviewer pointed out a very important topic that we believe we did not discuss it well in the manuscript.
+The manuscript presents steps of moving from considerable amount of data with specialties knowledge (background knowledge) to a final product that gives a better understanding about the data.
+According the reviewer suggestion we added a paragraph  to discuss this important point.
+Also we modified abstract an introduction for acknowledging the limitations with the study.
 }
@@
 \response{%
-response
+We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
+The confusion happened because of integrating three different topics into a method to carry out this research.
+One technical term in one field is different in another field and that cause a confusion.
+For example cluster, class, and label all are referring to the same concept, however, it depends on the application.
+We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.\\
 }
@@
 \response{%
-Response 
+Thanks for the comment. We fixed it. 
 }
@@
 \response{%
-Response
+We briefly explained these terms \textbf{Data Analysis Method} section. We also added clarification about semi-supervised and supervised learning into  \textbf{Introduction} section.
 }
@@
 \response{%
-Response
+A discussion about verification and validation is added. 
 }
@@
 \response{%
-Response
+\color{red}{Ricardo, please address this comment.}
 }
@@
 \response{%
-Response
+Added. Thanks.
 }
@@
 \response{%
-Response
+We modified the text for clarification and also to acknowledge each individual proposed metrics' importance. 
 }
@@
 You are treating these metrics as "independent variables", but as you have also noted many of them are not.
-For example, Arias intensity and energy are correlate.
+For example, Arias intensity and energy are correlated.
 Does this compromise your conclusions?
@@
 \response{%
-Response
+This is a very interesting comment.
+You are totally right.
+There are fairly high correlation between Arias intensity and energy integral.
+However, decision tree algorithm is robust enough to realize this similarity and pick the best one in each decision making process.
+As we remove the peak ground acceleration and response spectra from the prediction model, the algorithm looks for other best parameters.
+In case of hundreds of metrics it would be a reasonable choice to do feature selection analysis and choose one metric out of several correlated metrics.
+However, in this study, we only deal with 11 metrics and the results show that it works fairly well.
 }
@@
 \response{%
-Response
+Deleted. Thanks.
 }
@@
 \response{%
-Response
+In that specific subspace (C1,C7), C1 does not help in clustering process.
+That is why we call it irrelevant feature, because one can define the clusters based on only defining boundaries on C7.
+C1 can be important, however, it does not add any positive or negative values in decision making process in this subspace analysis.
 }
@@
 \response{%
-Response
+Thanks. More details provided. 
 }
@@
 \response{%
-Response
+In \textbf{Decision Trees} section we provided a definition of confusion matrix for the simplest case of two categories. The idea can be extended for more than two (in our case four) categories. 
 }
@@
 \response{%
-Response
+You are right. It could be misleading. We re-worded the sentence. 
 }
@@
 \response{%
-Response
+That is correct.
+The statement convey part of the message.
+Although we expect to see similarities between the current results and previous studies, we keep in mind that this study is based on lack of consensus on using GOF metrics for validation purposes.
+Since the previous study is representing a uniform average of all metrics, we do not expect to match all stations.
+Therefore, we prefer to keep it as numerical.
+We modified the text accordingly.
 }
@@
 \response{%
-Response
+That is right. It should be the number of decision attributes. We modified the figure accordingly. 
 }
@@
 \response{%
-Response
+That is a very good suggestion.
+We excluded C8(response spectra) and C5(peak acceleration) from analysis and reproduce similar trees with same complexities in terms of pruning process with previous four trees.
+As we expected, most important, metrics are still C4 and C6.
+C11 also become somewhat effective in the most complicated tree.
 }
@@
 \response{%
-Response
+Done. Thanks.
 }
@@
 \response{%
-Response
+Fixed it. Thanks.
 }
@@
 \response{%
-Response
+Changed. Thanks.
 }
@@
 \response{%
-Response
+Fixed. Thanks.
 }
@@
 
+\response{
+We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
+The confusion happened because of integrating three different topics into a method to carry out this research.
+One technical term in one field is different in another field and that cause a confusion.
+For example cluster, class, and label all are referring to the same concept, however, it depends on the application.
+We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.\\
+We also studied the models without response spectra (C8) and without response spectra (C8) and peak ground acceleration (C5).\\ 
+}
+
 \comment{1}{%
@@
 \response{
-Response
+Done. Thanks.
 }
@@
 \response{
-Response
+Added. Thanks.
 }
@@
 \response{
-Response
+\color{red}{Ricardo, please address this comment.}
 }
@@
 \comment{4}{%
-L. 42-43 I would suggest moving the description of the Anderson (2004) method (i.e., L. 48+) to directly follow this sentence.
-}
+L. 42-43 I would suggest moving the description of the Anderson (2004) method (i.e., L. 48+) to directly follow this sentence.}
 
 \response{
-Response
+Thanks for the suggestion. We moved the description of the Anderson (2004) method (i.e., L. 48+) to the mentioned location. It reads better now. 
 }
@@
 \response{
-Response
+Thanks for pointing this out.
+Since Anderson developed metrics for engineering application, we assumed it is implicitly mentioned.
+However, we agree that it should be explicitly clarified.
+We modified the text accordingly.
 }
@@
 \response{
-Response
+Fixed. Thanks.
 }
@@
 \response{
-Response
+In this study we only use broadband results. We added a clarification into the text. Thanks for the comment.  
 }
@@
 \response{
-Response
+Removed. Thanks.
 }
@@
 \response{
-Response
+We added a brief description about them. Thanks.
 }
@@
 \response{
-Response
+We meant to say numerical continues attribute. Thanks for the comment. Fixed. 
 }
@@
 \response{
-Response
+Thanks for the comment. Additional clarification added. 
 }
@@
 \response{
-Response
+That is right. Added.
 }
@@
 \response{
-Response
+We modified it to heavy pruning process and also added its explanation. 
 }
@@
 \response{
-Response
+Modified accordingly. Thanks.
 }
@@
 \response{
-Response
+Modified. Thanks.
 }
@@
 \response{
-Response
+Extra "s" deleted. Thanks.
 }

### 2018-03-30 abstract.tex
First draft of revision.

@@
 This leads to their selection often being subjective, either based on intended applications or personal preferences.
-As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could focused on a reduced number of metrics.
-We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques that help label and classify goodness-of-fit results, with the objective prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation.
-In particular, we study the relationships that exist between eleven different metrics.
+As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could \myremove{focused} \myrevision{focus} on a reduced number of metrics.
+We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques that help label and classify goodness-of-fit results, with the objective \myrevision{of} prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation.
+\myrevision{We use a dataset of three physic-based regional scale ground motion simulations based on one earthquake and three velocity models with eleven goodness of fit metrics.} In particular, we study the relationships that exist between \myremove{eleven} \myrevision{these} different metrics.
 In the process carried out, these metrics are understood as part of a multi-dimensional space.
 We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address high-dimensional effects.
-This allows us to label a given validation dataset into four categories (poor, fair, good, excellent) following common practice.
+This allows us to label a given validation dataset into four categories (poor, fair, good, excellent) following \myrevision{previous studies} \myremove{common practice}.
 We then develop a family of decision trees using a C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four categories.
-These decision trees can be understood as rapid predictors of the quality of a simulation, or as a data-informed classifiers that can help prioritize validation metrics.
+These decision trees can be understood as rapid predictors of the quality of a simulation, or as  \myremove{a} data-informed classifiers that can help prioritize validation metrics.
 Our analysis indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.
+\myrevision{Adding other simulation datasets (i.e., from different earthquakes, regions, models, metrics, etcetera) will improve the results in terms of developing a sufficiently sound and robust decision tree, however, the procedural steps laid out in this study, nonetheless, remain valid.}

### 2018-03-30 conclusions.tex
First draft of revision.

@@
 
-We present the result of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of prioritize and reduce the number of metrics, and develop an application independent decision algorithm.
+We present the result\myrevision{s} of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of \myremove{prioritize} \myrevision{prioritizing} and \myremove{reduce} \myrevision{reducing} the number of metrics, and develop an application independent decision algorithm.
 As a result of the data processing analysis, we propose a simplified algorithm based on a decision tree which uses only two metrics as opposed to the initial eleven available in the dataset.
 In particular, the proposed algorithm uses three (disjunctive decision) steps based on the values obtained for the metrics of the total energy and acceleration response spectra.
-we also propose rules to allow for the new class-based validation criteria to combine results obtained from different components of motion, and demonstrate that the results obtained with the proposed algorithm using two metrics are comparable to those obtained with the score-based validation results used in other recent validation studies.
-Similarly, one could implement similar rules to combine results from a frequency-band analysis, using, for example, the mode or majority validation class.
+\myremove{w}\myrevision{W}e also propose rules to allow for the new class-based validation criteria to combine results obtained from different components of motion, and demonstrate that the results obtained with the proposed algorithm using two metrics are comparable to those obtained with the score-based validation results used in other recent validation studies.
+\myremove{Similarly} \myrevision{Also}, one could implement similar rules to combine results from a frequency-band analysis, using, for example, the mode or majority validation class.
 
-We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulations done for a particular earthquake in a particular region.
-In a future follow-up study it would be ideal to refine the tree adding other simulation datasets (i.e., from different earthquakes, regions, models, etcetera) in order to arrive to a sufficiently sound and robust decision tree.
-The procedural steps laid out here, nonetheless, remains valid.
+We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulations done for a particular earthquake \myrevision{(one earthquake and three velocity models)} in a particular region.
+In a future follow-up study it would be ideal to refine the tree adding other simulation datasets (i.e., from different earthquakes, regions, models, \myrevision{metrics,} etcetera) in order to arrive to a sufficiently sound and robust decision tree.
+The procedural steps laid out here, nonetheless, remain\myremove{s} valid.
 We showed that the procedure and background information used for clustering and decision making is stable, and it is likely that the metrics of energy and response spectra (along with peak acceleration and velocity as suggested by the additional trees) will prevail as those among the most decisive ones.

### 2018-03-30 dataset.tex
fixing size of figures for submission, added style package for submission
First draft of revision.

@@
 The modeling domain covered an area of \adomain{180}{135}{km} that included all the major sedimentary basins and other relevant geologic structures in the greater Los Angeles region.
-Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three component of motion (EW, NS, and UD).
+Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three component\myrevision{s} of motion (EW, NS, and UD).
 The simulation domain and the stations used for the analysis are shown in Figure \ref{fig:chino-hills}, and a sample of the validation results obtained by \citet{Taborda_2014_BSSA} using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps}.
-This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, averaged for all metrics, and the three component of motion.
-This type of so-called GOF maps are commonly used in validation studies.
+This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, averaged for all metrics, and the three component\myrevision{s} of motion.
+\myremove{This type of so-called GOF maps are commonly used in validation studies.}
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/pdf/figure-03}
-    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component of motion (NS, EW, UD).
+    \includegraphics[width=0.48\textwidth]{figures/pdf/figure-03}
+    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component\myrevision{s} of motion (NS, EW, UD).
 In each case, the median is indicated by a notch in the box, the boxes represents the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), and the line shows the range of the data, with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots.

### 2018-03-30 decision-tree.tex
did some cleanng heare and there
last changes before submission
First draft of revision.

@@
 The theory behind decision trees is well established \citep[e.g.,][]{Quinlan_1986_ML, Mitchell_1997_Book}.
-Decision trees are, however, non-unique.
+\myrevision{A decision tree can be used to classify a case by starting at the root of the tree and moving through it until the leaf is encountered.
+Each tree is a structure including a root, decision nodes and leaf nodes.
+Decision nodes specifies some test to carried out on attributes and leaf nodes indicate the final class \citep{Quinlan_1993_Book}.} Decision trees are, however, non-unique.
 They depend on the dataset, the user parameters, and the algorithm employed to build them \citep{Murthy_1998_DMKD}.
@@
 This process has consequences, mainly that it increases the likelihood of overfitting.
-A matter we discuss later.
+\myremove{A matter we discuss later.} \myrevision{Overfitting happens when the prediction algorithm predicts the final results too closely or exactly to the training data, however, it provides non-accurate results for unseen or test data.}
 
@@
 There are several algorithms for building decision trees \citep[e.g., ID3, C4.5, C5.0, CART; see][]{Quinlan_1986_ML, Quinlan_1993_Book, Quinlan_1996_JAIR, Breiman_1984_Book}.
-Here we use the C5.0 implementation \citep{Kuhn_2017_Manual}, which is an extension of the C4.5 algorithm developed by \citet{Quinlan_1993_Book, Quinlan_1996_JAIR}.
+\myrevision{C5.0 is considered as the developed version of ID3 and C4.5, where it does not have any limitations in handling missing data, numerical attributes, and developing boosted models.
+Internal algorithm of CART is different than other methods, however, discussing technical details of these algorithms are beyond the scope of this study.} Here we use the C5.0 implementation \citep{Kuhn_2017_Manual}, which is an extension of the C4.5 algorithm developed by \citet{Quinlan_1993_Book, Quinlan_1996_JAIR}.
 
@@
 With this in mind, the information gain $G$ of an attribute $A$ for the dataset $S$, or $G(S,A)$ in equation (\ref{eq:gain}), is the expected reduction in entropy caused by partitioning the dataset according to attribute $A$ \citep{Mitchell_1997_Book}.
-For selection of a threshold for a continuous attribute refer to \citet{Quinlan_1996_JAIR}.
+For selection of a threshold for a \myrevision{numerical} continuous attribute refer to \citet{Quinlan_1996_JAIR}.
 
 In the particular case of the C5.0 algorithm the result of this process depends on several parameters, logical and numerical.
-We set the options to do feature selection and evaluate possible advanced splits of data, but do not use the boosted models option \citep[see][]{Kuhn_2017_Manual}; and use the confidence factor (CF) and the smallest number of samples that must be placed in at least two of the splits ($S_{\min}$, also known as \textit{minCases}) to fine tune the prediction models.
-In particular, we build a suite of trees for varying values of CF and $S_{\min}$.
+We set the options to do feature selection \myrevision{or winnowing.
+Therefore, the algorithm choose to use the most important attributes and ignore others.
+We also activate the evaluation of possible advanced splits of data.
+Sometimes data is very close to the splitting threshold where assigning a hard threshold may incorrectly change the classification, whereas, that small difference could happen because of numerical errors or error in measurement.
+Therefore, advanced split can define a soft threshold to consider different probabilities in assigning data to different classes.
+A decision tree can grow so deep to finally classify all training data correctly.
+However, the tree will be very complex and will suffer from overfitting.
+In other words it will have very poor functionality for test data.
+Therefore, decision tree algorithms (here C5.0) first grow the tree up to the most complex tree then based on user defined factors start to prune the grown tree.The pruning process helps to reduce the complexity of the tree and also increase it's generalizability.
+Consequently the pruned tree's leaf nodes will have training data which belongs to more than one class which the final class will be the mode of that leaf node.
+In this study we use two parameters to control the pruning process including confidence or certainty factor (CF) and the smallest number of samples that must be put in at least two of the splits ($S_{\min}$, also known as \textit{minCases}).
+C5.0 algorithm uses error-based pruning method (see \citet{Quinlan_1993_Book} for more details).
+The objective is to minimize the observed error rate.
+Therefore, error estimates for leaf nodes and subtrees are computed assuming that they were used to classify a set of unseen cases with a predicted error rate which is computed based on CF and binomial distribution of number of misclassified events (E) and all training data at that leaf (N).
+Decreasing CF and increasing $S_{\min}$ cause more heavy pruning process \citep[see][]{Kuhn_2017_Manual}.} \myremove{and evaluate possible advanced splits of data, but do not use the boosted models option \citep[see][]{Kuhn_2017_Manual}; and use the confidence factor (CF) and the smallest number of samples that must be placed in at least two of the splits ($S_{\min}$, also known as \textit{minCases}) to fine tune the prediction models.} In particular, we build a suite of trees for varying values of CF and $S_{\min}$.
 
@@
 	&					&	\multicolumn{2}{c}{\mathrm{Prediction}}		\\
-	&					&	\mathrm{Postive}	&	\mathrm{Negative}	\\[0.75ex]
+	&					&	\mathrm{Positive}	&	\mathrm{Negative}	\\[0.75ex]
 	\mathrm{Actual~Value}		
@@
 This is done for all the trees obtained using all possible combinations of CF and $S_{\min}$ values.
-Then, we select the tree with the highest level of effectiveness as measure by $F_1$, commensurate to such tree having a low complexity, i.e., a reduced number of metrics and a shallow depth, which is precisely the goal set at the start.
+Then, we select the tree with the highest level of effectiveness as \myremove{measure} \myrevision{measured} by $F_1$, commensurate to such tree having a low complexity, i.e., a reduced number of metrics and a shallow depth, which is precisely the goal set at the start.
 Note that $F_1$ is obtained based on the testing dataset as opposed to the training dataset.
 The results as applied to our dataset are presented next.
-
-% ===========================================================================================
-%
-% OLD NAEEM VERSION
-% 
-% Decision tree learning is a method for approximating a target function, in which learned function is represented by decision tree \citep{Mitchell_1997_Book}.
-\citet{Quinlan_1986_ML} demonstrated that the technology for building decision trees from examples is fairly robust.
-He summarized an approach to synthesizing decision trees that has been used in a variety of systems, and he describes one such system, ID3, in detail.
-Although ID3 algorithm is successful in most classification problems, it does not handle the numeric attributes and only one attribute at a time is tested for making decisions.
-\citet{Quinlan_1993_Book} extended the ID3 algorithm and presented the C4.5 algorithm.
-Later on they added boosting capability to C4.5 and called it C5.0.
-The improved algorithm accepts both continues and discrete features and solves over fitting problem by pruning.
-Although there are many other algorithm for implementing decision tree, in this study we use the C5.0 algorithm.
-Discussing the algorithm in details is out of the scope of this paper.
-There are fairly good amount of references to study those details \citep[e.g.,][]{Mitchell_1997_Book, Quinlan_1993_Book, Hornik_2009_CS}.
-
-% In general, decision tree measures the effectiveness of an attribute in classifying the training data through information gain.
-According to  \citet{Mitchell_1997_Book}, information gain of an attribute is the expected reduction in entropy caused by partitioning the examples according to this attribute.
-If the target can take on $c$ different values the entropy is defined as
-% % 
-% \begin{equation}
-% 	Entropy(S) = \sum_{i=1}^{c} \left[ -p_i \log_2 \left(p_i\right) \right]
-% 	\, ,
-% \end{equation}
-% % 
-% where $p_i$ is proportion of $S$ belonging to class $i$.
-Entropy measures the homogeneity of examples or dataset.
-Higher entropy means data is  fairly uniformly distributed among different classes.
-On the other hand lower entropy means data unequally distributed among classes.
-The extreme case happens when all data belong to one class which results in entropy equal to zero.
-Given entropy as a measure of impurity in a collection of training example the measure of the effectiveness of attribute is defined through information gain which is expected reduction in entropy caused by partitioning the examples according to this attribute.
-Gain of the attribute A is defined as
-% % 
-% \begin{equation}
-% Gain(S,A)\ =\ Entropy(S) - \sum_{v \in Values(A) } \frac{|S_v|}{|S|} Entropy(S_v)
-% \end{equation} 
-% % 
-% where $Values(A)$ is the set of all possible values of attribute A, and $S_v$ is the subset of $S$ for which attribute $A$ has value $v$.
-Information gain is used by the algorithm in order to determine which is the better attribute for classifying the training example.
-The attribute with higher information gain will be the first option to separate the training data based on that attribute and grow to the lower nodes (Refer to chapter 3 of \citet{Mitchell_1997_Book} for more detailed examples.)
-
-% We discussed the basic idea behind the decision tree algorithm, however, there are more details specially in dealing with outliers and noises which we address them through pruning approaches.
-In this study we use C5.0 package \citep{C50_2015} in R programming language \citep{R_2016_program} to develop the decision tree algorithms.
-We discuss the input parameters in the section \ref{classification_model}.

### 2018-03-30 introduction.tex
First draft of revision.

@@
 Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation.
+\myrevision{Validation is the process concerned with determining whether a conceptual model and the results it leads to are accurate representations of the physical system under study.
+This entails a comparison of simulation results with experimental data and observations.
+Verification, on the other hand, is concerned with the correctness and accuracy of a solution approach and its proper implementation with respect to the intended conceptual model, which involves a comparison of simulation results with exact results, if available, or with other alternative or approximate results \citep{Taborda_2013_BSSA}}.
 Various methods or schemes have been proposed to evaluate, through direct quantitative comparisons or indirect statistical analysis, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
@@
 Out of the various validation methods just described, \citet{Anderson_2004_Proc} is perhaps the most widely used today \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
+\myrevision{ In essence, \citeauthor{Anderson_2004_Proc}'s method assesses the similarity of two signals based on the average score of 10 metrics.
+These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), their peak values (in acceleration, velocity and displacement), frequency content (in terms of Fourier and response spectra), and their time synchronization (through cross-correlation).}
+
 We, in particular, have consistently relied on a modified version of this method for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake, and for the evaluation of velocity models in southern California through the validation of a large suite of moderate-magnitude earthquake simulations \citep{Taborda_2013_BSSA, Taborda_2014_BSSA, Taborda_2016_GJI}.
 
-In essence, \citeauthor{Anderson_2004_Proc}'s method assesses the similarity of two signals based on the average score of 10 metrics.
+\myremove{In essence, \citeauthor{Anderson_2004_Proc}'s method assesses the similarity of two signals based on the average score of 10 metrics.
 These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), their peak values (in acceleration, velocity and displacement), frequency content (in terms of Fourier and response spectra), and their time synchronization (through cross-correlation).
+}
 
-The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics are well understood in both engineering and seismology contexts.
+The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics are \myrevision{relatively} well understood in both engineering and seismology contexts.
 However, it has been pointed out that some of these metrics are redundant, or that other relevant metrics need to be added.
@@
 In the same spirit, \citet{Maufroy_2015_BSSA} reduced the number of metrics, limiting them to only those with comparable units.
-Unfortunately, none of these alternatives addresses the underlying questions: what are the parameters that ought to be taken into account when validating ground motion simulations, and what level of priority should these be given.
+Unfortunately, none of these alternatives addresses the underlying questions: what are the \myrevision{most important} parameters that ought to be taken into account when validating ground motion simulations \myrevision{for engineering applications without specific target application (e.g., designing tall buildings or power plants)}, and what level of priority should these be given\myremove{.}\myrevision{?}
 
@@
 We use a constrained \kmeans{} method \citep[e.g.,][]{Macqueen_1967_Proc, Wagstaff_2001_Proc} and conduct a subspace clustering analysis to address high-dimensional effects in order to add labels to the dataset in terms of four categories: poor, fair, good, and excellent.
-These categories adhere to the common practice in validation.
-The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}.
+These categories \myrevision{are commonly used in previous validation studies} \myremove{adhere to the common practice in validation}.
+\myrevision{In the machine learning linguistic, \kmeans{} clustering approach is considered as an unsupervised learning method.
+However, when we add background knowledge into the process it becomes constrained \kmeans{} method and is considered as a semi-supervised learning method.} The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}.
 Upon labeling the dataset, we develop a family of decision trees using a C5.0 algorithm \citep[][]{Quinlan_1993_Book, Quinlan_1996_JAIR}, and select a few to prioritize the metrics involved in the validation process.
 The decision trees narrow the number of metrics offering a prediction algorithm into the aforementioned validation categories.
+\myrevision{Decision trees are considered as supervised learning algorithms.}
 
-In summary, our results indicate that among the eleven metrics considered in the analysis, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.
+In summary, our results indicate that \myrevision{for \citeauthor{Anderson_2004_Proc}'s metrics based the results of 2008 $M_w$ 5.4 Chino Hills earthquake for three velocity models,} among the eleven metrics considered in the analysis, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.
 We test our prediction model and offer a discussion about its implications and potential use in future validation efforts.

### 2018-03-30 method.tex
First draft of revision.

@@
 The development of such a decision tree requires a proper classification of the data, which can be done through a clustering process.
-The following sections explain the data processing analysis we put in place to cluster the dataset and obtain two alternative decision trees.
+\myrevision{In this study we integrate three different topics (i.e., ground motion simulation, semi-supervised, and supervised learning), therefore, the same concept may have different technical terms in different part of the paper based on their application.
+In machine learning linguistic, attributes and features, mostly are used interchangeably.
+In this study, we use features for clustering and attributes for classification purposes.
+The reason for choosing these terms is keeping consistency with the references that we use to develop the methods.
+Therefore, in this study feature, attribute, and metric all refer to the eleven GOF scores.
+Consequently, dimension refers to the number of features in each constrained \kmeans{} clustering analysis.
+For example, 3-Dimensional subspace analysis is application of constrained \kmeans{} on three out of eleven features.
+Each observation with eleven metrics is categorized in one of Poor(p), Fair(F), Good(g), and Excellent (E) categories.
+The process of assigning these categories to each observation observation is called clustering or labeling.
+The resulted dataset is called the labeled dataset.
+When we use these labeled dataset in classification, the labels or clusters are called classes.
+Therefore, label, class, and cluster all refer to four categories of GOF level.
+}
+
+The following sections explain the data processing analysis we put in place to cluster the dataset and obtain \myremove{two} \myrevision{three} alternative decision trees.
 

### 2018-03-30 metrics.tex
did some cleanng heare and there
som edits before submission
First draft of revision.

@@
 Following \citet{Anderson_2004_Proc}, when applied to a pair of signals, each one of these metrics yields a GOF score ranging from 0 to 10, where a value of 10 corresponds to two signals having identical characteristics.
-This scoring scale varies according to the following exponential function:
+This scoring scale varies according to the following exponential function
 % 
@@
 In his method, this frequency-domain decomposition should have pass bands defined following a logarithmic distribution in order to give more weight to the low-frequencies, and the GOF scores of all sub-bands and the broadband be averaged into a single GOF final score.
+\myrevision{However, in this study, in order to reduce the uncertainties that can come from different filtering process \citep[e.g., see ][]{Khoshnevis_2015_Proc} we use broadband results (\feq{0}{4}).
+For metrics which are functions of frequency (or period) like Fourier spectra and response spectra, the final result is the mean value of application of equation(\ref{eq:gof-function}) on all frequencies (or periods).}
 
@@
 He concluded that for a typical distribution of the scores, the GOF could be categorized into four groups: poor (for scores from 0 to 4), fair (4 to 6), good (6 to 8), and excellent (for scores from 8 to 10).
-While this categorization was arguably subjective, over the years it has become common practice and has been adopted or matched by alternative validation methods \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}, making the method itself a reference baseline for various validation studies \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
+While this categorization was arguably subjective, over the years it has  \myrevision{ been frequently used and} \myremove{become common practice and has} been adopted or matched by alternative validation methods \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}, making the method itself a reference baseline for various validation studies \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
 
 Here, we focus our analysis on the eleven metrics included in Table \ref{tab:metrics}, and investigate the relationships that exist between them in order to prioritize a reduced number of metrics that can help predict the outcome category one would use to label the results of a given simulation.
-
-% ===========================================================================================
-%
-% OLD NAEEM VERSION
-% 
-% Following the guidelines suggested by \citet{Anderson_2004_Proc}, \citet{Taborda_2014_BSSA} applied the scoring procedure to each pair of data and synthetic using compatible broadband sets, and series of bandpass filtered versions of the signals or sub-bands, SB1, SB2, SB3, SB4, and SB5.
-The signals bandpass filtered between 0.1 and 4 Hz for BB, and between 0.1 Hz and 0.25 Hz, and 0.25 and 0.5 Hz, and 0.5 and 1 Hz, and 1 and 2 Hz, and 2 and 4 Hz for SB1, SB2, SB3, SB4, and SB5, respectively.
-They also reported a scaled combination score of metrics and bands.
-However, since the combined scores are linear scaled combinations of metrics and frequency bands, in this specific study, they do not add new information to the data base.
-\citet{Khoshnevis_2015_Proc} studied the effect of filtering on GOF score and presented that filtering approach can affect the GOF score and change the final classes.
-Therefore, in order to avoid extra source of uncertainty in the study, we only consider broadband GOF scores ($0.1-4~Hz$).
-Analysis of effect of bands in the results could be an appropriate subject for another study.
-
-% Accuracy of ground motion simulation is evaluated based on a quantitative validation of the simulation ground motions.
-The validation process consist of comparisons between synthetics and data, at locations where records were available for the simulated events.
-For validating the simulation of strong motion complex time series, different metrics are defined.
-These metrics evaluate the similarity of two signals in time or frequency domain.
-\\
-
-% \citet{Anderson_2004_Proc} proposed 10 individual quality measures to evaluate the credibility of synthetic seismograms.
-He proposed to apply the metrics on 10 frequency bands with logarithmic spacing.
-The logarithmic spacing puts more emphasis on lower frequencies where these frequencies are more amenable to waveform fitting and are particularly important for response of large structures.
-A comprehensive score (S1) could be achieved by the average of the scores of seismograms filtered in each of the valid frequency bands individually, as well as the broadband seismogram.
-An alternative score (S2) is obtained by averaging the scores on the ten individual criteria for only the accelerogram filtered to allow all frequencies to pass.
-\citet{Anderson_2004_Proc} scaled the scores into range of 0 to 10 with 10 giving perfect agreement.
-He suggested to consider a score below 4 as a poor fit, a score of 4-6 as a fair fit, a score of 6 to 8 as a good fit, and a score over 8 is an excellent fit.
-However, some of these metrics like integral measures are potentially very easy to fit and some of them like cross correlation is by far the most difficult of all parameters to achieve high score.
-Therefore, for a pair of synthetic and data some of metrics could belong to good and some of them could belong to fair fit classes.\\
-
-% In other study, \citet{Kristekova_2006_BSSA} developed quantitative misfit criteria for comparison of seismograms.
-The misfit criteria are defined based on the time-frequency (TF) representation of the seismograms obtained at the continues wavelet transform.
-Based on local TF envelop and phase differences, they defined envelop and phase misfit dependent on both time and frequency.
-With projection of TF misfit onto frequency or time domain, they computed frequency- or time-dependent misfits, respectively.
-The method proposed  by \citet{Kristekova_2006_BSSA} needs to consider one of signals as a reference signal.
-Therefore, \citet{Kristekova_2009_GJI} presented an extension of the theory of the TF misfit criteria.
-They defined locally and globally normalized TF criteria where locally normalized misfits can be used if it is important to investigate relatively small part of the signal.
-The globally normalized misfits can be used for quantifying an overall level of disagreement.
-In the extended version of the metrics, they defined a misfit criteria for 3 components of a signal and also they developed TF GOF criteria for more realistic scenario where there is a high level of disagreement and it is not possible to consider one of signals as a reference signal.
-They classified the GOF scores as poor, fair, good, and excellent based on GOF numerical values.\\
-
-% \citet{Olsen_2010_SRL} argued that the \citet{Kristekova_2006_BSSA,Kristekova_2009_GJI}  metrics are suitable for low frequency signals.
-Therefor, they proposed a new GOF method for validation of Broadband synthetics.
-They used some of metrics that are used by \citet{Anderson_2004_Proc}, however, they used different scoring approach which generates a relatively high-resolution representation of small misfits.
-They also used inelastic and elastic response spectra ration (IE ratio) as a structural engineering specific metric.
-They proposed to combine the metrics for final decisions based on user defined application-based weights.\\
-
-% Recently, \citet{Taborda_2013_BSSA} conducted a simulation for 2008 Chino Hills earthquake and proposed a minor modification in the \citet{Anderson_2004_Proc} metrics.
-They added an eleventh parameter to the Anderson's scores as strong motion duration.\\
-
-% A comprehensive validation approach should include metrics in both time and frequency domain.
-All mentioned metrics \citep[i.e.,][]{Anderson_2004_Proc,Kristekova_2006_BSSA,Kristekova_2009_GJI,Olsen_2010_SRL} have this characteristic.
-However, in this study we are mostly dependent on available data based on \citet{Taborda_2014_BSSA} where they used the GOF method proposed by \citet{Anderson_2004_Proc}, with minor modification introduced by \citet{Taborda_2013_BSSA}.
-The method compares synthetics against data using 11 individual parameters, namely: Arias Intensity Integral (C1), Energy Integral (C2), Arias Intensity Value (C3), Total Energy (C4), Peak Acceleration (C5), Peak Velocity (C6), Peak Displacement (C7), Response Spectra (C8), Fourier Amplitude Spectrum (C9), Cross Correlation (C10), and Strong Phase Duration (C11).
-Each parameter is mapped on to a numerical scale ranging from 0 to 10, where a score of 10 corresponds to a perfect  match between two signals.\\
-% Following the guidelines suggested by \citet{Anderson_2004_Proc}, \citet{Taborda_2014_BSSA} applied the scoring procedure to each pair of data and synthetic using compatible broadband sets, and series of bandpass filtered versions of the signals or sub-bands, SB1, SB2, SB3, SB4, and SB5.
-The signals bandpass filtered between 0.1 and 4 Hz for BB, and between 0.1 Hz and 0.25 Hz, and 0.25 and 0.5 Hz, and 0.5 and 1 Hz, and 1 and 2 Hz, and 2 and 4 Hz for SB1, SB2, SB3, SB4, and SB5, respectively.
-They also reported a scaled combination score of metrics and bands.
-However, since the combined scores are linear scaled combinations of metrics and frequency bands, in this specific study, they do not add new information to the data base.
-\citet{Khoshnevis_2015_Proc} studied the effect of filtering on GOF score and presented that filtering approach can affect the GOF score and change the final classes.
-Therefore, in order to avoid extra source of uncertainty in the study, we only consider broadband GOF scores ($0.1-4~Hz$).
-Analysis of effect of bands in the results could be an appropriate subject for another study.

### 2018-03-30 results.tex
First draft of revision.

@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/pdf/figure-06}
+	\includegraphics[width=\textwidth]{figures/pdf/figure-06_mod}
 	\caption{Statistical distribution of the dataset as partitioned into the four validation categories (poor, fair, good, excellent) after the clustering analysis.
@@
 In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range, with outliers shown as scattered dots.
-The color version of this figure is available only in the electronic edition.}
+The color version of this figure is available only in the electronic edition.
+\myrevision{Dashed lines represent the boundaries of poor, fair, good, and excellent categories according to  \citeauthor{Anderson_2004_Proc}'s method.}}
 	\label{fig:boxed-clusters}
@@
 We used the oversampling approach described in the previous section.
-We replicated data from the excellent class randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.
-By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}.
+We replicated data from the excellent class \myrevision{through random selection of data from that class until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.
+This oversampling is considered as a $\times$10 factor.} \myremove{randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.} By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}.
 (Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary.
-Instead, we used a strong pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.
+Instead, we used a \myremove{strong} \myrevision{heavy} pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.
 
@@
 
+%\begin{figure*}[t]
+%	\centering
+%	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
+%	\caption{Accuracy of tree predictions in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed with respect to the number of attributes (GOF metrics) as a function of: \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees.
+The rings around some of the dots indicate select trees used as reference in the text, other figures and tables.
+The color version of this figure is available only in the electronic edition.}
+%	\label{fig:nodes-depth}
+%\end{figure*}
+
 \begin{figure*}[t]
 	\centering
-	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
+	\includegraphics[width=\textwidth]{figures/pdf/figure-08_mod}
 	\caption{Accuracy of tree predictions in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed with respect to the number of attributes (GOF metrics) as a function of: \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees.
@@
 More complex trees tend to be deeper, have more nodes, and employ more attributes.
-This leads to higher $F_1$ values, but they may not necessarily be more practical.
+\myrevision{complexity of trees is controlled by pruning process.} This leads to higher $F_1$ values, but they may not necessarily be more practical.
 In total T1, T2, and T3 use 2, 3 and 4 attributes respectively.
 All coincide in the use of the total energy (C4) and response spectra (C8) as key metrics.
-T2 adds peak acceleration (C5), and T3 adds peak velocity (C5) to the previous metrics.
+T2 adds peak acceleration (C5), and T3 adds peak velocity \myremove{(C5)} \myrevision{(C6)} to the previous metrics.
 
@@
 As such, the samples in the second leaf-node from the right in T3 are categorized as good despite a significant portion of them being excellent.
-This is a natural trade-off embedded in the use of decision trees which actually results in more effective trees.
+This is a natural trade-off embedded in the use of \myrevision{pruned} decision trees which actually results in more effective trees.
 
@@
 
-Notice that in all trees the metrics C4 (total energy) and C8 (response spectra) are consistently the most relevant ones in the decision algorithms.
+Notice that in all trees \myrevision{,using all metrics,} the metrics C4 (total energy) and C8 (response spectra) are consistently the most relevant ones in the decision algorithms.
 They are followed by C5 (peak acceleration) and C6 (peak velocity).
@@
 
+\myrevision{It is obvious that C8(response spectra) and C5(peak acceleration) are the most important metrics in developing the prediction models.
+Therefore, in order to understand other parameters importance we developed two sets of trees according to complexity of the original trees (the structure of generated trees are not presented here), while once dropping C8 (response spectra) and once dropping C8 (response spectra) and C5(peak acceleration) from the prediction attributes.
+According to Table \ref{tab:participation}, these trees also emphasize the importance of C4 and C6.
+However, in the absence of C5(peak acceleration) and C8 (response spectra),  C3 (Arias intensity) and C9 (Fourier spectrum) become more effective in developing prediction models.
+Also C11(strong phase duration) becomes somewhat important in the most complicated tree.}
+
 A low participation does not necessarily mean that a given metric is not important.
 Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
-What the results we present here mean, in the context of GOF validation, is that in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more decisive in reaching an accurate conclusion about validation quality than the duration (C11).
+What the results we present here mean, in the context of GOF validation \myrevision{for engineering applications} , is that in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more decisive in \myrevision{narrowing eleven validation metrics into fewer numbers in order to get the same categorization as all eleven metrics}  \myremove{reaching an accurate conclusion about validation quality} than the duration (C11).
 
@@
 
+%\begin{table}[t]
+%	\centering
+%	\caption{Metrics data analysis participation in select trees (in percent)}
+%	\label{tab:participation}
+%	\small
+%	\begin{tabular}{lrrrr}
+%	Metric 	& T1 	& T2	& T3 	& T66 	\\ 
+%	\hline
+%	C1 		& --	& --	& -- 	& 1.5	\\
+%	C2 		& -- 	& --	& -- 	& 0.5	\\
+%	C3 		& -- 	& -- 	& -- 	& 9.6	\\
+%	\rowcolor{light-light-gray}
+%	C4 		& 46.8	& 52.4	& 60.2	& 100.0	\\
+%	\rowcolor{light-light-gray}
+%	C5 		& -- 	& 21.5	& 21.5	& 72.2	\\
+%	\rowcolor{light-light-gray}
+%	C6 		& -- 	& -- 	& 37.0	& 56.3	\\
+%	C7 		& -- 	& -- 	& -- 	& 19.4	\\
+%	\rowcolor{light-light-gray}
+%	C8 		& 100.0	& 100.0	& 100.0	& 100.0	\\
+%	C9 		& -- 	& -- 	& --	& 23.8	\\
+%	C10		& -- 	& -- 	& -- 	& 22.2	\\
+%	C11		& -- 	& -- 	& -- 	& -- 	\\
+%	\hline
+%	\end{tabular}
+%\end{table}
+
 \begin{table}[t]
 	\centering
-	\caption{Metrics data analysis participation in select trees (in percent)}
+	\caption{Metrics data analysis participation in select trees (in percent). \myrevision{Stars represent the trees with same complexity in pruning process, however, not exact tree structures.}}
 	\label{tab:participation}
 	\small
-	\begin{tabular}{lrrrr}
-	Metric 	& T1 	& T2	& T3 	& T66 	\\ 
+	\myrevision{
+	\begin{tabular}{lrrrrrrrrrrrr}
+	Metric 	& T1 	& T2	& T3 	& T66  & T1* 	& T2*	& T3* 	& T66*  & T1* 	& T2*	& T3* 	& T66* 	\\ 
 	\hline
-	C1 		& --	& --	& -- 	& 1.5	\\
-	C2 		& -- 	& --	& -- 	& 0.5	\\
-	C3 		& -- 	& -- 	& -- 	& 9.6	\\
+	&\multicolumn{4}{c}{Including C8} & \multicolumn{4}{c}{Without C8}&  \multicolumn{4}{c}{Without C5 and C8}\\
+	\hline
+	C1 		& --	& --	& -- 	& 1.5 & --	& --	& -- 	& --- & --	& --	& -- 	& --	\\
+	C2 		& -- 	& --	& -- 	& 0.5 & --	& --	& -- 	& 2.27 & --	& --	& -- 	& --	\\
+	C3 		& -- 	& -- 	& -- 	& 9.6 & --	& --	& -- 	& 53.22 & --	& 60.35	& 61.19 	& 99.11	\\
 	\rowcolor{light-light-gray}
-	C4 		& 46.8	& 52.4	& 60.2	& 100.0	\\
+	C4 		& 46.8	& 52.4	& 60.2	& 100.0  & 44.35	& 81.77	& 81.77 	& 83.20 & 47.44	& 48.67	& 80.54 	& 54.64	\\
 	\rowcolor{light-light-gray}
-	C5 		& -- 	& 21.5	& 21.5	& 72.2	\\
+	C5 		& -- 	& 21.5	& 21.5	& 72.2 & 100.0	& 100.0	& 100.0 	& 100.0 & --	& --	& -- 	& --	\\
 	\rowcolor{light-light-gray}
-	C6 		& -- 	& -- 	& 37.0	& 56.3	\\
-	C7 		& -- 	& -- 	& -- 	& 19.4	\\
+	C6 		& -- 	& -- 	& 37.0	& 56.3 & 58.65	& 63.51	& 62.77 	& 82.39 & 100.0	& 100	& 100 	& 100\\
+	C7 		& -- 	& -- 	& -- 	& 19.4 & --	& --	& -- 	& 32.37 & --	& --	& -- 	& 43.24	\\
 	\rowcolor{light-light-gray}
-	C8 		& 100.0	& 100.0	& 100.0	& 100.0	\\
-	C9 		& -- 	& -- 	& --	& 23.8	\\
-	C10		& -- 	& -- 	& -- 	& 22.2	\\
-	C11		& -- 	& -- 	& -- 	& -- 	\\
+	C8 		& 100.0	& 100.0	& 100.0	& 100.0 & --	& --	& -- 	& --	 & --	& --	& -- 	& --\\
+	C9 		& -- 	& -- 	& --	& 23.8 & --	& --	& -- 	& 48.63 & --	& --	& -- 	& 72.22	\\
+	C10		& -- 	& -- 	& -- 	& 22.2 & --	& --	& -- 	& 24.93 & --	& --	& -- 	& 20.46	\\
+	C11		& -- 	& -- 	& -- 	& -- 	    & --	& --	& -- 	& --- & --	& --	& -- 	& 8.09\\
 	\hline
 	\end{tabular}
+	}
 \end{table}

### 2018-03-30 testing.tex
First draft of revision.

@@
 In our view, the T1 results are simpler to assimilate and equally informative.
-Looking at both plots in Fig.~\ref{fig:avg-gof-maps} it is possible to argue they lead to the same conclusions and present a similar picture of the validation of the simulation, not only globally but at specific locations, with only minor differences---especially if one is to consider the assumed equivalence between scores and classes (0--4: poor; 4--6: fair; 6--8:good; 8--10: excellent).
+Looking at both plots in Fig.~\ref{fig:avg-gof-maps} it is possible to argue they lead to \myremove{the same} \myrevision{similar} conclusions and present a similar picture of the validation of the simulation, not only globally but at specific locations, with only minor differences---especially if one is to consider the assumed equivalence between scores and classes (0--4: poor; 4--6: fair; 6--8:good; 8--10: excellent).
+\myrevision{However, we do not expect to see the same results in both figures.
+\citet{Taborda_2014_BSSA} results is developed based on assigning uniform weight on all metrics, however, this study research question is addressing the importance of different metrics and implying that assigning uniform weights might not provide accurate results.}
 
@@
 \end{figure*}
+
+\myrevision{It is worth mentioning that, in this study, we presented an integration of scientists knowledge and field data using machine learning algorithms to develop an acceptable and robust prediction model.
+This method can be used in any field where there are some field data as well as expertise knowledge about the data.
+Validation metrics are note limited to \citeauthor{Anderson_2004_Proc}'s metrics.
+Therefore, adding more metrics from different field can be used through the proposed method to develop prediction models.
+}

### 2018-04-05 response.tex
minor things starting to review

--- 2018-03-28 response.tex
@@
-\setlength{\paperwidth}{210mm}
-\setlength{\paperheight}{297mm}% fixed.
+% \setlength{\paperwidth}{210mm}
+% \setlength{\paperheight}{297mm}% fixed.
 
@@
 
+% *****************************************************************
+% *********************** Associate Editor ************************
+% *****************************************************************
+
+\begin{center}
+	\bf
+	\large
+	Authors' Response to Associate Editor (Dr. Thomas Brocher)
+\end{center}
+
+\noindent
+We thank Dr.
+Thomas Brocher for his comments and suggestions, which helped us improve the manuscript.
+To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font and the removed text in red font.
+In the following, we provide a response to each comment.
+The original comments are in italic black font, followed by our responses in regular blue font.
+\vspace{2ex}
+\newline
+
+\introcomment{~}{%
+I agree with the peer reviewers that this paper addresses a topical subject of great importance to BSSA readers.
+The science is in the process of transitioning from purely observational based analysis of ground motions to a mixture of observations and simulations of ground motions.
+The paper is a useful contribution to this discussion, is generally well written, and carefully describes the methodology used and its limitations.
+The peer reviewers have made some very helpful suggestions for improving the paper that I hope the authors will strongly consider.
+}
+
+\introcomment{~}{%
+In particular, I agree with their comment that the limitations of the study, i.e., the choice of 11 metrics and earthquake simulations for only one earthquake, needs to be included in the abstract and the conclusions.
+In addition, I also agree that all the terms and methodology should be well defined for the BSSA readership, so that the paper can be understood by them without resort to frequent online searches.
+}
+
+\introcomment{~}{%
+I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit.
+It clearly shows that only these metrics are change significantly with the goodness of fit (P,F, G, E).
+Plotting these values as a function of P, F, G, and E together on the same figure might allow the reader to see this relationship more clearly and might help provide a heuristic understanding of how C4, C5, C6, and C8 are diagnostic whereas C11 is constant with P, F, G, and E, and thus not diagnostic.
+In fact, in my opinion, Figure 6 is a simple and easy way of determining which metrics are diagnostic that might even eliminate the need for the decision trees.
+In any case, more
+discussion of this figure might help the authors make the case for the main conclusions.
+}
+
+\response{%
+Thanks for the constructive suggestions.
+We agree with you and peer reviewers that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
+The confusion happened because of integrating three different topics into a method to carry out this research.
+One technical term in one field is different in another field and that cause a confusion.
+For example cluster, class, and label all are referring to the same concept, however, it depends on the application.
+We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.
+\vspace{2ex}
+\newline 
+We also modified Figure 6 in order to easily track the variation of different metrics for final classes.
+We do agree with you that Figure 6 clearly gives a good understanding about the metrics, however, we add decision trees, because they can convert the qualitative judgment to quantitative conditional form to be able to exactly decide on the final class based on their numerical GOF scores.
+\vspace{2ex}
+\newline 
+We added text to abstract and conclusion to give a heads up to readers about 11 metrics and one earthquake with 3 velocity model  simulations.
+}
+
 % ******************************************************************
@@
 
+\newpage
 \begin{center}
@@
 They have mentioned this limitation in the conclusions, but the abstract and introduction do not explain the limitations and the application-dependency of the proposed approach.
-Both Abstract and Introduction need modification to specify the application for which the results and conclusions are true (i.e., the
-application here is to narrow down the 11 Anderson2004+ parameters to select fewer parameters that give similar results as the original 11 for the simulations they chose in this paper).
+Both Abstract and Introduction need modification to specify the application for which the results and conclusions are true (i.e., the application here is to narrow down the 11 Anderson2004+ parameters to select fewer parameters that give similar results as the original 11 for the simulations they chose in this paper).
 Currently, the Abstract and Introduction give the impression that the conclusions are true for any engineering application or simulation method.
@@
 \response{%
-Will be added. 
+The reviewer pointed out a very important topic that we believe we did not discuss it well in the manuscript.
+The manuscript presents steps of moving from considerable amount of data with specialties knowledge (background knowledge) to a final product that gives a better understanding about the data.
+According the reviewer suggestion we added a paragraph  to discuss this important point.
+Also we modified abstract an introduction for acknowledging the limitations with the study.
 }
 
-\comment{1}{%
-The terminologies used in the paper are not clear in many cases. For example, what are "features" "attributes" "dimensions" "trees, nodes and leafs" "class"? It was very difficult to
-read the section on Data Analysis Method and Introduction since these terminologies were not clearly defined or described in terms of the variables used in this paper. I was confused what
-"data" and "results" were referring to: GOF score data?
+\end{document}
+
+% \comment{1}{
+% The terminologies used in the paper are not clear in many cases.
+For example, what are "features" "attributes" "dimensions" "trees, nodes and leafs" ``class''?
+It was very difficult to read the section on Data Analysis Method and Introduction since these terminologies were not clearly defined or described in terms of the variables used in this paper.
+I was confused what
+% "data" and "results" were referring to: GOF score data?
 GOF metric data?
@@
 I suggest that the authors streamline the paper and the terminologies to be consistent and clear in the final version.
-}
+% }
 
-\response{%
-response
-}
+% \response{%
+% We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
+The confusion happened because of integrating three different topics into a method to carry out this research.
+One technical term in one field is different in another field and that cause a confusion.
+For example cluster, class, and label all are referring to the same concept, however, it depends on the application.
+We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.\\
+% }
 
-\comment{2}{%
-The terms "common practice" or "common standards" have been mentioned several times. Because GMSV is such a new field (and I don't think there is consensus, I for one don't agree
-with some statements but I agree that others may), I think these terms should be removed. Just say this is your opinion, or it is based on the references that you have listed.
-}
+% \comment{2}{%
+% The terms "common practice" or "common standards" have been mentioned several times. Because GMSV is such a new field (and I don't think there is consensus, I for one don't agree
+% with some statements but I agree that others may), I think these terms should be removed. Just say this is your opinion, or it is based on the references that you have listed.
+% }
 
-\response{%
-Response 
-}
+% \response{%
+% Thanks for the comment. We fixed it. 
+% }
 
-\comment{3}{%
-In abstract, "semi-supervised" and "supervised" learning techniques are mentioned, but in the paper these terms are not explained, what part of the learning techniques is semi-supervised
-and what does that mean?
-}
+% \comment{3}{%
+% In abstract, "semi-supervised" and "supervised" learning techniques are mentioned, but in the paper these terms are not explained, what part of the learning techniques is semi-supervised
+% and what does that mean?
+% }
 
-\response{%
-Response
-}
+% \response{%
+% We briefly explained these terms \textbf{Data Analysis Method} section. We also added clarification about semi-supervised and supervised learning into  \textbf{Introduction} section.
+% }
 
-\comment{4}{%
-In introduction, the difference between "verification" and "validation" should be discussed.
-}
+% \comment{4}{%
+% In introduction, the difference between "verification" and "validation" should be discussed.
+% }
 
-\response{%
-Response
-}
+% \response{%
+% A discussion about verification and validation is added. 
+% }
 
-\comment{5}{%
-In abstract and introduction, what does a "direct application" mean? it's not clear. Similarly, in Validation Metrics section: "direct quantitative comparisons" and "indirect statistical analysis"
-}
+% \comment{5}{%
+% In abstract and introduction, what does a "direct application" mean? it's not clear. Similarly, in Validation Metrics section: "direct quantitative comparisons" and "indirect statistical analysis"
+% }
 
-\response{%
-Response
-}
+% \response{%
+% \color{red}{Ricardo, please address this comment.}
+% }
 
-\comment{6}{%
-Page 4, line 53: add "relatively" before "well understood"
-}
+% \comment{6}{%
+% Page 4, line 53: add "relatively" before "well understood"
+% }
 
-\response{%
-Response
-}
+% \response{%
+% Added. Thanks.
+% }
 
-\comment{7}{%
-The two questions posed in the introduction lines 59-61: I don't think the first one is really the question, all parameters should be considered if anyone has thought of them as important (i.e.,
-has proposed them). The second question is the main question and the one being addressed in this paper.
-}
+% \comment{7}{%
+% The two questions posed in the introduction lines 59-61: I don't think the first one is really the question, all parameters should be considered if anyone has thought of them as important (i.e.,
+% has proposed them). The second question is the main question and the one being addressed in this paper.
+% }
 
-\response{%
-Response
-}
+% \response{%
+% We modified the text for clarification and also to acknowledge each individual proposed metrics' importance. 
+% }
 
-\comment{8}{%
-You are treating these metrics as "independent variables", but as you have also noted many of them are not.
-For example, Arias intensity and energy are correlate.
+% \comment{8}{%
+% You are treating these metrics as "independent variables", but as you have also noted many of them are not.
+For example, Arias intensity and energy are correlated.
 Does this compromise your conclusions?
 If not explain why in the Results section.
-}
+% }
 
-\response{%
-Response
-}
+% \response{%
+% This is a very interesting comment.
+You are totally right.
+There are fairly high correlation between Arias intensity and energy integral.
+However, decision tree algorithm is robust enough to realize this similarity and pick the best one in each decision making process.
+As we remove the peak ground acceleration and response spectra from the prediction model, the algorithm looks for other best parameters.
+In case of hundreds of metrics it would be a reasonable choice to do feature selection analysis and choose one metric out of several correlated metrics.
+However, in this study, we only deal with 11 metrics and the results show that it works fairly well.
+% }
 
-\comment{9}{%
-Page 7, line 133: delete the last sentence
-}
+% \comment{9}{%
+% Page 7, line 133: delete the last sentence
+% }
 
-\response{%
-Response
-}
+% \response{%
+% Deleted. Thanks.
+% }
 
-\comment{10}{%
-Page 14, line 310: you say "... the latter can essentially be discarded..." I don't understand this statement, wouldn't this mean that both C1 and C7 are important to consider since they are
-independent of each other (one gives information that the other one doesn't)?
-}
+% \comment{10}{%
+% Page 14, line 310: you say "... the latter can essentially be discarded..." I don't understand this statement, wouldn't this mean that both C1 and C7 are important to consider since they are
+% independent of each other (one gives information that the other one doesn't)?
+% }
 
-\response{%
-Response
-}
+% \response{%
+% In that specific subspace (C1,C7), C1 does not help in clustering process.
+That is why we call it irrelevant feature, because one can define the clusters based on only defining boundaries on C7.
+C1 can be important, however, it does not add any positive or negative values in decision making process in this subspace analysis.
+% }
 
-\comment{11}{%
-Page 14 line 323: It's not clear how you "replicate data", provide more detail.
-}
+% \comment{11}{%
+% Page 14 line 323: It's not clear how you "replicate data", provide more detail.
+% }
 
-\response{%
-Response
-}
+% \response{%
+% Thanks. More details provided. 
+% }
 
-\comment{12}{%
-I don't think "confusion matrix" given in tables is explained in the paper!
-}
+% \comment{12}{%
+% I don't think "confusion matrix" given in tables is explained in the paper!
+% }
 
-\response{%
-Response
-}
+% \response{%
+% In \textbf{Decision Trees} section we provided a definition of confusion matrix for the simplest case of two categories. The idea can be extended for more than two (in our case four) categories. 
+% }
 
-\comment{13}{%
-Page 17, lines 391-394: this sentence should be re-worded, what is an "accurate conclusion"?
+% \comment{13}{%
+% Page 17, lines 391-394: this sentence should be re-worded, what is an "accurate conclusion"?
 You should be very specific that your conclusion is that instead of using all 11 validation metrics you can select a few in order to get the same categorization as all 11 metrics would have given you.}
 
-\response{%
-Response
-}
+% \response{%
+% You are right. It could be misleading. We re-worded the sentence. 
+% }
 
-\comment{14}{%
-Page 18, line 405: this is implying that fig 10 is similar to fig2!
+% \comment{14}{%
+% Page 18, line 405: this is implying that fig 10 is similar to fig2!
 I personally cannot see this in the figures!
@@
 It might be better to regenerate fig 11a but with different color scales that only vary by four choices, then the comparison would be much easier.
-}
+% }
 
-\response{%
-Response
-}
+% \response{%
+% That is correct.
+The statement convey part of the message.
+Although we expect to see similarities between the current results and previous studies, we keep in mind that this study is based on lack of consensus on using GOF metrics for validation purposes.
+Since the previous study is representing a uniform average of all metrics, we do not expect to match all stations.
+Therefore, we prefer to keep it as numerical.
+We modified the text accordingly.
+% }
 
-\comment{15}{%
-Figure 8: I don't know what you mean by y-label "decision attributes". Do you mean "the number of GOF metrics"? why there is no data point for 1 or 11?
-}
+% \comment{15}{%
+% Figure 8: I don't know what you mean by y-label "decision attributes". Do you mean "the number of GOF metrics"? why there is no data point for 1 or 11?
+% }
 
-\response{%
-Response
-}
+% \response{%
+% That is right. It should be the number of decision attributes. We modified the figure accordingly. 
+% }
 
-\comment{16}{%
-The conclusion basically says that response spectrum is the most important validation metric. This is not surprising to any engineer or modeler. The big question is what other parameters
-describe the differences between a simulation and a recorded motion if their response spectra matched? Or in other words, what are the other important validation parameters aside from
-response spectrum (many have argued that's duration, but no statistical proof). Can you use your method to answer this question? Can you condition your machine learning process on Sa
-being the same (for example use a subset of data with similar Sa or artificially make them equal), then decide what's the next important parameter? Would the result be the same as in your
-conclusions? Or would it be influenced by removal of Sa as a parameter (this is possible since there are correlations between Sa and other metrics).
-}
+% \comment{16}{%
+% The conclusion basically says that response spectrum is the most important validation metric. This is not surprising to any engineer or modeler. The big question is what other parameters
+% describe the differences between a simulation and a recorded motion if their response spectra matched? Or in other words, what are the other important validation parameters aside from
+% response spectrum (many have argued that's duration, but no statistical proof). Can you use your method to answer this question? Can you condition your machine learning process on Sa
+% being the same (for example use a subset of data with similar Sa or artificially make them equal), then decide what's the next important parameter? Would the result be the same as in your
+% conclusions? Or would it be influenced by removal of Sa as a parameter (this is possible since there are correlations between Sa and other metrics).
+% }
 
-\response{%
-Response
-}
+% \response{%
+% That is a very good suggestion.
+We excluded C8(response spectra) and C5(peak acceleration) from analysis and reproduce similar trees with same complexities in terms of pruning process with previous four trees.
+As we expected, most important, metrics are still C4 and C6.
+C11 also become somewhat effective in the most complicated tree.
+% }
 
-Typos:\\
-\\
+% Typos:\\
+% \\
 
-\comment{1}{%
-Page 2, line 8: change focused to focus
-}
+% \comment{1}{%
+% Page 2, line 8: change focused to focus
+% }
 
-\response{%
-Response
-}
+% \response{%
+% Done. Thanks.
+% }
 
-\comment{2}{%
-Page 2, line 20: change classifiers to classifier
-}
+% \comment{2}{%
+% Page 2, line 20: change classifiers to classifier
+% }
 
-\response{%
-Response
-}
+% \response{%
+% Fixed it. Thanks.
+% }
 
-\comment{3}{%
-Page 13, line 185: change measure to measured
-}
+% \comment{3}{%
+% Page 13, line 185: change measure to measured
+% }
 
-\response{%
-Response
-}
+% \response{%
+% Changed. Thanks.
+% }
 
-\comment{4}{%
-Page 16, line 354: the last C5 should be C6?
-}
+% \comment{4}{%
+% Page 16, line 354: the last C5 should be C6?
+% }
 
-\response{%
-Response
-}
+% \response{%
+% Fixed. Thanks.
+% }
 
-% ******************************************************************
-% *************************** REVIEWER 2 ***************************
-% ******************************************************************
+% % ******************************************************************
+% % *************************** REVIEWER 2 ***************************
+% % ******************************************************************
 
-\newpage
+% \newpage
 
-\begin{center}
-	\bf
-	\large
-	Authors' Response to Reviewer 2
-	\end{center}
+% \begin{center}
+% 	\bf
+% 	\large
+% 	Authors' Response to Reviewer 2
+% 	\end{center}
 
-\noindent
-We thank the reviewer for his/her comments and suggestions, which helped us improve the manuscript.
-To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font.
+% \noindent
+% We thank the reviewer for his/her comments and suggestions, which helped us improve the manuscript.
+To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font and the removed text in red font.
 In the following, we provide a response to each comment.
 The original comments are in italic black font, followed by our responses in regular blue font.
-\vspace{2ex}
-\newline
+% \vspace{2ex}
+% \newline
 
-\introcomment{~}{%
-Review of Khoshnevis and Taborda for BSSA
-}
+% \introcomment{~}{%
+% Review of Khoshnevis and Taborda for BSSA
+% }
 
-\introcomment{~}{%
-Earthquake ground motion simulations are usually validated against observational data or empirically-derived solutions. However, exactly which parameters to compare (e.g., peak
-ground velocity, duration, spectral content) in order to assess the goodness-of-fit, is not always clear.
+% \introcomment{~}{%
+% Earthquake ground motion simulations are usually validated against observational data or empirically-derived solutions. However, exactly which parameters to compare (e.g., peak
+% ground velocity, duration, spectral content) in order to assess the goodness-of-fit, is not always clear.
 This paper uses machine learning techniques and eleven possible metrics for comparing (recorded and simulated) ground motions to determine the parameters that are most predictive of goodness-of-fit.
 This is an interesting paper, and will help future researchers
-prioritize which metrics to focus on when validating ground motions simulations.
+% prioritize which metrics to focus on when validating ground motions simulations.
 The paper is generally well written.
@@
 It would also be useful to have some discussion regarding which parameters were found to be repetitive, incase future authors wish to look at more than the four recommended parameters.
-}
+% }
 
-\comment{1}{%
-L. 8 "... could \textbf{focus} ..."
-}
+% \response{
+% We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
+The confusion happened because of integrating three different topics into a method to carry out this research.
+One technical term in one field is different in another field and that cause a confusion.
+For example cluster, class, and label all are referring to the same concept, however, it depends on the application.
+We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.\\
+% We also studied the models without response spectra (C8) and without response spectra (C8) and peak ground acceleration (C5).\\ 
+% }
 
-\response{
-Response
-}
+% \comment{1}{%
+% L. 8 "... could \textbf{focus} ..."
+% }
 
-\comment{2}{%
-L. 11 "... the objective \textbf{of} ..."
-}
+% \response{
+% Done. Thanks.
+% }
 
-\response{
-Response
-}
+% \comment{2}{%
+% L. 11 "... the objective \textbf{of} ..."
+% }
 
-\comment{3}{%
-L. 25-41 The frequent use of phrases such as "Among these methods,"; "Among the latter,"; "In this category,"; and "Within this group," make the text difficult to follow.
-}
+% \response{
+% Added. Thanks.
+% }
 
-\response{
-Response
-}
+% \comment{3}{%
+% L. 25-41 The frequent use of phrases such as "Among these methods,"; "Among the latter,"; "In this category,"; and "Within this group," make the text difficult to follow.
+% }
 
-\comment{4}{%
-L. 42-43 I would suggest moving the description of the Anderson (2004) method (i.e., L. 48+) to directly follow this sentence.
-}
+% \response{
+% \color{red}{Ricardo, please address this comment.}
+% }
 
-\response{
-Response
-}
+% \comment{4}{%
+% L. 42-43 I would suggest moving the description of the Anderson (2004) method (i.e., L. 48+) to directly follow this sentence.}
 
-\comment{5}{%
-L. 59-61/L. 391-394 It should be stated somewhere that this is really intended for \textbf{engineering purposes}.
-}
+% \response{
+% Thanks for the suggestion. We moved the description of the Anderson (2004) method (i.e., L. 48+) to the mentioned location. It reads better now. 
+% }
 
-\response{
-Response
-}
+% \comment{5}{%
+% L. 59-61/L. 391-394 It should be stated somewhere that this is really intended for \textbf{engineering purposes}.
+% }
 
-\comment{6}{%
-L. 61 This sentence should end with a "?"
-}
+% \response{
+% Thanks for pointing this out.
+Since Anderson developed metrics for engineering application, we assumed it is implicitly mentioned.
+However, we agree that it should be explicitly clarified.
+We modified the text accordingly.
+% }
 
-\response{
-Response
-}
+% \comment{6}{%
+% L. 61 This sentence should end with a "?"
+% }
 
-\comment{7}{%
-L. 98-99 It is not clear how the equation for S is applied to values that are a function of period (e.g., the response or Fourier spectrum). Do you follow the guidelines provided by Anderson? If
-so, what frequency range/bands are you examining?
-}
+% \response{
+% Fixed. Thanks.
+% }
 
-\response{
-Response
-}
+% \comment{7}{%
+% L. 98-99 It is not clear how the equation for S is applied to values that are a function of period (e.g., the response or Fourier spectrum). Do you follow the guidelines provided by Anderson? If
+% so, what frequency range/bands are you examining?
+% }
 
-\comment{8}{%
-L. 128/132/137/etc. "... the three component\textbf{s} ..."
-}
+% \response{
+% In this study we only use broadband results. We added a clarification into the text. Thanks for the comment.  
+% }
 
-\response{
-Response
-}
+% \comment{8}{%
+% L. 128/132/137/etc. "... the three component\textbf{s} ..."
+% }
 
-\comment{9}{%
-L. 236 Where is the discussion on the consequences of overfitting? The only other place I see overfitting mentioned is L. 328.
-}
+% \response{
+% Fixed. Thanks.
+% }
 
-\response{
-Response
-}
+% \comment{9}{%
+% L. 236 Where is the discussion on the consequences of overfitting? The only other place I see overfitting mentioned is L. 328.
+% }
 
-\comment{10}{%
-L. 236 Remove "A matter we discuss later."
-}
+% \response{
+% Added. Thanks.
+% }
 
-\response{
-Response
-}
+% \comment{10}{%
+% L. 236 Remove "A matter we discuss later."
+% }
 
-\comment{11}{%
-L. 241-243 It is not helpful to list different algorithm names for making decision trees, without any information about them or how they differ.
-}
+% \response{
+% Removed. Thanks.
+% }
 
-\response{
-Response
-}
+% \comment{11}{%
+% L. 241-243 It is not helpful to list different algorithm names for making decision trees, without any information about them or how they differ.
+% }
 
-\comment{12}{%
-L. 257-258 I don't understand this sentence. What is a continuous attribute?
-}
+% \response{
+% We added a brief description about them. Thanks.
+% }
 
-\response{
-Response
-}
+% \comment{12}{%
+% L. 257-258 I don't understand this sentence. What is a continuous attribute?
+% }
 
-\comment{13}{%
-L. 259-264 This paragraph contains a lot of jargon.
-}
+% \response{
+% We meant to say numerical continues attribute. Thanks for the comment. Fixed. 
+% }
 
-\response{
-Response
-}
+% \comment{13}{%
+% L. 259-264 This paragraph contains a lot of jargon.
+% }
 
-\comment{14}{%
-L. 300 Isn't this also true for the C5-C7 combination?
-}
+% \response{
+% Thanks for the comment. Additional clarification added. 
+% }
 
-\response{
-Response
-}
+% \comment{14}{%
+% L. 300 Isn't this also true for the C5-C7 combination?
+% }
 
-\comment{15}{%
-L. 328 What is "a strong pruning process"?
-}
+% \response{
+% That is right. Added.
+% }
 
-\response{
-Response
-}
+% \comment{15}{%
+% L. 328 What is "a strong pruning process"?
+% }
 
-\comment{16}{%
-L. 433 "We present the result\textbf{s} of..."
-}
+% \response{
+% We modified it to heavy pruning process and also added its explanation. 
+% }
 
-\response{
-Response
-}
+% \comment{16}{%
+% L. 433 "We present the result\textbf{s} of..."
+% }
 
-\comment{17}{%
-L. 435 "... goal of prioritiz\textbf{ing} and reduc\textbf{ing}..."
-}
+% \response{
+% Added. Thanks.
+% }
 
-\response{
-Response
-}
+% \comment{17}{%
+% L. 435 "... goal of prioritiz\textbf{ing} and reduc\textbf{ing}..."
+% }
 
-\comment{18}{%
-L. 443 Remove "Similarly" or "similar" (redundant)
-}
+% \response{
+% Modified accordingly. Thanks.
+% }
 
-\response{
-Response
-}
+% \comment{18}{%
+% L. 443 Remove "Similarly" or "similar" (redundant)
+% }
 
-\comment{19}{%
-L. 450 "... remain valid."
-}
+% \response{
+% Modified. Thanks.
+% }
 
-\response{
-Response
-}
+% \comment{19}{%
+% L. 450 "... remain valid."
+% }
 
-\comment{20}{%
-L. 455 "... as the decisive \textbf{parameters} ..." ? (This short concluding paragraph would be much stronger without the use of unnecessary phrases, i.e., "This latter point", "One the one hand",
-"on the other hand."
-}
+% \response{
+% Extra "s" deleted. Thanks.
+% }
 
-\response{
-Response
-}
+% \comment{20}{%
+% L. 455 "... as the decisive \textbf{parameters} ..." ? (This short concluding paragraph would be much stronger without the use of unnecessary phrases, i.e., "This latter point", "One the one hand",
+% "on the other hand."
+% }
 
-\end{document}
+% \response{
+% \color{red}{Ricardo, please address this comment.}
+% }
+
+% \end{document}

### 2018-04-05 buildpdf.sh

@@ -1,6 +1,5 @@
 
 MAIN=$1
-FINAL=main
 
 pdflatex ${MAIN}.tex
 bibtex   ${MAIN}.aux
@@ -9,8 +8,8 @@ pdflatex ${MAIN}.tex
 
 rm ${MAIN}.fff ${MAIN}.log ${MAIN}.aux ${MAIN}.ttt ${MAIN}.bbl ${MAIN}.lof ${MAIN}.blg ${MAIN}.spl ${MAIN}.out
 
-mv ${MAIN}.pdf ${FINAL}.pdf
+mv ${MAIN}.pdf ${MAIN}.pdf
 
-open -a Adobe\ Acrobat.app ${FINAL}.pdf
+open -a Adobe\ Acrobat.app ${MAIN}.pdf
 
 clear

### 2018-04-06 response.tex
went over the simple changes, more attention needed for the remaining ones.

--- 2018-03-28 response.tex
@@
 
+% *****************************************************************
+% *********************** Associate Editor ************************
+% *****************************************************************
+
+\begin{center}
+	\bf
+	\large
+	Authors' Response to Associate Editor (Dr. Thomas Brocher)
+\end{center}
+
+% \noindent
+% We thank Dr.
+Thomas Brocher for his comments and suggestions, which helped us improve the manuscript.
+To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font and the removed text in red font.
+In the following, we provide a response to each comment.
+The original comments are in italic black font, followed by our responses in regular blue font.
+% \vspace{2ex}
+% \newline
+%
+% \introcomment{~}{%
+% I agree with the peer reviewers that this paper addresses a topical subject of great importance to BSSA readers.
+The science is in the process of transitioning from purely observational based analysis of ground motions to a mixture of observations and simulations of ground motions.
+The paper is a useful contribution to this discussion, is generally well written, and carefully describes the methodology used and its limitations.
+The peer reviewers have made some very helpful suggestions for improving the paper that I hope the authors will strongly consider.
+% }
+%
+% \introcomment{~}{%
+% In particular, I agree with their comment that the limitations of the study, i.e., the choice of 11 metrics and earthquake simulations for only one earthquake, needs to be included in the abstract and the conclusions.
+In addition, I also agree that all the terms and methodology should be well defined for the BSSA readership, so that the paper can be understood by them without resort to frequent online searches.
+% }
+%
+% \introcomment{~}{%
+% I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit.
+It clearly shows that only these metrics are change significantly with the goodness of fit (P,F, G, E).
+Plotting these values as a function of P, F, G, and E together on the same figure might allow the reader to see this relationship more clearly and might help provide a heuristic understanding of how C4, C5, C6, and C8 are diagnostic whereas C11 is constant with P, F, G, and E, and thus not diagnostic.
+In fact, in my opinion, Figure 6 is a simple and easy way of determining which metrics are diagnostic that might even eliminate the need for the decision trees.
+In any case, more
+% discussion of this figure might help the authors make the case for the main conclusions.
+% }
+%
+% \response{%
+% Thanks for the constructive suggestions.
+We agree with you and peer reviewers that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
+The confusion happened because of integrating three different topics into a method to carry out this research.
+One technical term in one field is different in another field and that cause a confusion.
+For example cluster, class, and label all are referring to the same concept, however, it depends on the application.
+We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.
+% \vspace{2ex}
+% \newline
+% We also modified Figure 6 in order to easily track the variation of different metrics for final classes.
+We do agree with you that Figure 6 clearly gives a good understanding about the metrics, however, we add decision trees, because they can convert the qualitative judgment to quantitative conditional form to be able to exactly decide on the final class based on their numerical GOF scores.
+% \vspace{2ex}
+% \newline
+% We added text to abstract and conclusion to give a heads up to readers about 11 metrics and one earthquake with 3 velocity model  simulations.
+% }
+
 % ******************************************************************
@@
 We thank the reviewer for his/her comments and suggestions, which helped us improve the manuscript.
-To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font.
-In the following, we provide a response to each comment.
+To aid the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font.
+Here, we provide a response to each comment.
 The original comments are in italic black font, followed by our responses in regular blue font.
@@
 
-\introcomment{~}{%
-This is an interesting article for the readers of the BSSA journal.
+% \introcomment{~}{%
+% This is an interesting article for the readers of the BSSA journal.
 The authors propose a new approach based on machine learning techniques to determine the importance of validation metrics for ground motion simulation.
 The field of ground motion simulation validation (GMSV) is relatively new and of interest to both seismologists who develop simulation models and engineers who would like to use these simulations in engineering applications.
-One of the most challenging questions to address is "which validation metric(s) is best to use?" as there are so many different metrics in this field and so many different opinions.
+One of the most challenging questions to address is ``which validation metric(s) is best to use?'' as there are so many different metrics in this field and so many different opinions.
 It has been difficult to address this question because the answer varies depending on the application of the simulations (e.g., GMPE development, PSHA, structural analysis response of interest, etc).
@@
 If true this should be mentioned as it will be of interest to others).
-}
-
-\response{%
-Will be added. 
-}
-
-\comment{1}{%
-The terminologies used in the paper are not clear in many cases. For example, what are "features" "attributes" "dimensions" "trees, nodes and leafs" "class"? It was very difficult to
-read the section on Data Analysis Method and Introduction since these terminologies were not clearly defined or described in terms of the variables used in this paper. I was confused what
-"data" and "results" were referring to: GOF score data?
+% }
+%
+% \response{%
+% The reviewer pointed out a very important topic that we believe we did not discuss it well in the manuscript.
+The manuscript presents steps of moving from considerable amount of data with specialties knowledge (background knowledge) to a final product that gives a better understanding about the data.
+According the reviewer suggestion we added a paragraph  to discuss this important point.
+Also we modified abstract an introduction for acknowledging the limitations with the study.
+% }
+%
+% \comment{1}{
+% The terminologies used in the paper are not clear in many cases.
+For example, what are ``features'' ``attributes'' ``dimensions'' ``trees, nodes and leafs'' ``class''?
+It was very difficult to read the section on Data Analysis Method and Introduction since these terminologies were not clearly defined or described in terms of the variables used in this paper.
+I was confused what ``data'' and ``results'' were referring to: GOF score data?
 GOF metric data?
@@
 I suggest that the authors streamline the paper and the terminologies to be consistent and clear in the final version.
+% }
+%
+% \response{%
+% We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
+The confusion happened because of integrating three different topics into a method to carry out this research.
+One technical term in one field is different in another field and that cause a confusion.
+For example cluster, class, and label all are referring to the same concept, however, it depends on the application.
+We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.
+% }
+%
+% \comment{2}{%
+% The terms ``common practice'' or ``common standards'' have been mentioned several times.
+Because GMSV is such a new field (and I don't think there is consensus, I for one don't agree with some statements but I agree that others may), I think these terms should be removed.
+Just say this is your opinion, or it is based on the references that you have listed.
+% }
+%
+% \response{%
+% Thanks for the comment. We fixed it.
+% }
+%
+% \comment{3}{%
+% In abstract, ``semi-supervised'' and ``supervised'' learning techniques are mentioned, but in the paper these terms are not explained, what part of the learning techniques is semi-supervised
+% and what does that mean?
+% }
+%
+% \response{%
+% We briefly explained these terms \textbf{Data Analysis Method} section. We also added clarification about semi-supervised and supervised learning into  \textbf{Introduction} section.
+% }
+%
+% \comment{4}{%
+% In introduction, the difference between ``verification'' and ``validation'' should be discussed.
+% }
+%
+% \response{%
+% A discussion about verification and validation is added.
+% }
+%
+% \comment{5}{%
+% In abstract and introduction, what does a ``direct application'' mean?
+it's not clear.
+Similarly, in Validation Metrics section: ``direct quantitative comparisons'' and ``indirect statistical analysis''.
+% }
+%
+% \response{%
+% \color{red}{Ricardo, please address this comment.}
+% }
+
+\comment{6}{%
+Page 4, line 53: add ``relatively'' before ``well understood''.
 }
@@
 \response{%
-response
+We appreciate the suggestion made by the reviewer, but we disagree.
+There should be nothing relative about the fact that these metrics are well understood by seismologists and engineers, and the strength of Anderson's method relying on this characteristic has been pointed out before by others.
+That being said, we modified the sentence with a choice of words that we have used in two previous papers published in BSSA, which convey the same idea we wanted to transmit.
 }
 
-\comment{2}{%
-The terms "common practice" or "common standards" have been mentioned several times. Because GMSV is such a new field (and I don't think there is consensus, I for one don't agree
-with some statements but I agree that others may), I think these terms should be removed. Just say this is your opinion, or it is based on the references that you have listed.
-}
-
-\response{%
-Response 
-}
-
-\comment{3}{%
-In abstract, "semi-supervised" and "supervised" learning techniques are mentioned, but in the paper these terms are not explained, what part of the learning techniques is semi-supervised
-and what does that mean?
-}
-
-\response{%
-Response
-}
-
-\comment{4}{%
-In introduction, the difference between "verification" and "validation" should be discussed.
-}
-
-\response{%
-Response
-}
-
-\comment{5}{%
-In abstract and introduction, what does a "direct application" mean? it's not clear. Similarly, in Validation Metrics section: "direct quantitative comparisons" and "indirect statistical analysis"
-}
-
-\response{%
-Response
-}
-
-\comment{6}{%
-Page 4, line 53: add "relatively" before "well understood"
-}
-
-\response{%
-Response
-}
-
-\comment{7}{%
-The two questions posed in the introduction lines 59-61: I don't think the first one is really the question, all parameters should be considered if anyone has thought of them as important (i.e.,
-has proposed them). The second question is the main question and the one being addressed in this paper.
-}
-
-\response{%
-Response
-}
-
-\comment{8}{%
-You are treating these metrics as "independent variables", but as you have also noted many of them are not.
-For example, Arias intensity and energy are correlate.
+% \comment{7}{%
+% The two questions posed in the introduction lines 59--61: I don't think the first one is really the question, all parameters should be considered if anyone has thought of them as important (i.e., has proposed them).
+The second question is the main question and the one being addressed in this paper.
+% }
+%
+% \response{%
+% We modified the text for clarification and also to acknowledge each individual proposed metrics' importance.
+% }
+%
+% \comment{8}{%
+% You are treating these metrics as ``independent variables'', but as you have also noted many of them are not.
+For example, Arias intensity and energy are correlated.
 Does this compromise your conclusions?
 If not explain why in the Results section.
-}
-
-\response{%
-Response
-}
+% }
+%
+% \response{%
+% This is a very interesting comment.
+You are totally right.
+There are fairly high correlation between Arias intensity and energy integral.
+However, decision tree algorithm is robust enough to realize this similarity and pick the best one in each decision making process.
+As we remove the peak ground acceleration and response spectra from the prediction model, the algorithm looks for other best parameters.
+In case of hundreds of metrics it would be a reasonable choice to do feature selection analysis and choose one metric out of several correlated metrics.
+However, in this study, we only deal with 11 metrics and the results show that it works fairly well.
+% }
 
@@
 \response{%
-Response
+We followed this suggestion and deleted the sentence.
 }
 
-\comment{10}{%
-Page 14, line 310: you say "... the latter can essentially be discarded..." I don't understand this statement, wouldn't this mean that both C1 and C7 are important to consider since they are
-independent of each other (one gives information that the other one doesn't)?
+% \comment{10}{%
+% Page 14, line 310: you say ``...the latter can essentially be discarded...'' I don't understand this statement, wouldn't this mean that both C1 and C7 are important to consider since they are independent of each other (one gives information that the other one doesn't)?
+% }
+%
+% \response{%
+% In that specific subspace (C1,C7), C1 does not help in clustering process.
+That is why we call it irrelevant feature, because one can define the clusters based on only defining boundaries on C7.
+C1 can be important, however, it does not add any positive or negative values in decision making process in this subspace analysis.
+% }
+%
+% \comment{11}{%
+% Page 14, line 323: It's not clear how you ``replicate data'', provide more detail.
+% }
+%
+% \response{%
+% Thanks. More details provided.
+% }
+%
+% \comment{12}{%
+% I don't think ``confusion matrix'' given in tables is explained in the paper!
+% }
+%
+% \response{%
+% In \textbf{Decision Trees} section we provided a definition of confusion matrix for the simplest case of two categories. The idea can be extended for more than two (in our case four) categories.
+% }
+%
+% \comment{13}{%
+% Page 17, lines 391--394: this sentence should be re-worded, what is an ``accurate conclusion''?
+You should be very specific that your conclusion is that instead of using all 11 validation metrics you can select a few in order to get the same categorization as all 11 metrics would have given you.}
+%
+% \response{%
+% You are right. It could be misleading. We re-worded the sentence.
+% }
+%
+% \comment{14}{%
+% Page 18, line 405: this is implying that fig 10 is similar to fig2!
+I personally cannot see this in the figures!
+They look very different.
+It might be better to regenerate fig 11a but with different color scales that only vary by four choices, then the comparison would be much easier.
+% }
+%
+% \response{%
+% That is correct.
+The statement convey part of the message.
+Although we expect to see similarities between the current results and previous studies, we keep in mind that this study is based on lack of consensus on using GOF metrics for validation purposes.
+Since the previous study is representing a uniform average of all metrics, we do not expect to match all stations.
+Therefore, we prefer to keep it as numerical.
+We modified the text accordingly.
+% }
+
+% \comment{15}{%
+% Figure 8: I don't know what you mean by y-label ``decision attributes''. Do you mean ``the number of GOF metrics''? why there is no data point for 1 or 11?
+% }
+
+% \response{% *** needs attention and modification in text too ***
+% That is right.
+It should be the number of decision attributes.
+We modified the figure accordingly.
+The reason why there is no data for 1 is because there are no trees that would classify the data using a single metric into the four classes because each node is a binary decision, so one needs at least 2 metrics with 3 or more nodes with a minimum depth of 2.
+And the reason why there is no data for 11 metrics is because when using all metrics, C11 (duration) never had a participation---as found by the algorithm.
+The new Table 6 shows that if we forcefully eliminate
+%
+% ...
+%
+% }
+
+% \comment{16}{%
+% The conclusion basically says that response spectrum is the most important validation metric.
+This is not surprising to any engineer or modeler.
+The big question is what other parameters describe the differences between a simulation and a recorded motion if their response spectra matched?
+Or in other words, what are the other important validation parameters aside from response spectrum (many have argued that's duration, but no statistical proof).
+Can you use your method to answer this question?
+Can you condition your machine learning process on Sa being the same (for example use a subset of data with similar Sa or artificially make them equal), then decide what's the next important parameter?
+Would the result be the same as in your conclusions?
+Or would it be influenced by removal of Sa as a parameter (this is possible since there are correlations between Sa and other metrics).
+% }
+%
+% \response{%
+% That is a very good suggestion.
+We excluded C8(response spectra) and C5(peak acceleration) from analysis and reproduce similar trees with same complexities in terms of pruning process with previous four trees.
+As we expected, most important, metrics are still C4 and C6.
+C11 also become somewhat effective in the most complicated tree.
+% }
+
+\textbf{Typos:}
+
+\comment{1}{%
+Page 2, line 8: change ``focused'' to ``focus''.
 }
@@
 \response{%
-Response
+Done. Thanks.
 }
 
-\comment{11}{%
-Page 14 line 323: It's not clear how you "replicate data", provide more detail.
+\comment{2}{%
+Page 2, line 20: change ``classifiers'' to ``classifier''
 }
@@
 \response{%
-Response
+In fact, we were referring to ``these decision trees'', so we removed the ``a'' before ``data-informed qualifiers'' instead. The plural form is now consistent. Thank you.
 }
 
-\comment{12}{%
-I don't think "confusion matrix" given in tables is explained in the paper!
+\comment{3}{%
+Page 13, line 285: change ``measure'' to ``measured''.
 }
@@
 \response{%
-Response
-}
-
-\comment{13}{%
-Page 17, lines 391-394: this sentence should be re-worded, what is an "accurate conclusion"?
-You should be very specific that your conclusion is that instead of using all 11 validation metrics you can select a few in order to get the same categorization as all 11 metrics would have given you.}
-
-\response{%
-Response
-}
-
-\comment{14}{%
-Page 18, line 405: this is implying that fig 10 is similar to fig2!
-I personally cannot see this in the figures!
-They look very different.
-It might be better to regenerate fig 11a but with different color scales that only vary by four choices, then the comparison would be much easier.
-}
-
-\response{%
-Response
-}
-
-\comment{15}{%
-Figure 8: I don't know what you mean by y-label "decision attributes". Do you mean "the number of GOF metrics"? why there is no data point for 1 or 11?
-}
-
-\response{%
-Response
-}
-
-\comment{16}{%
-The conclusion basically says that response spectrum is the most important validation metric. This is not surprising to any engineer or modeler. The big question is what other parameters
-describe the differences between a simulation and a recorded motion if their response spectra matched? Or in other words, what are the other important validation parameters aside from
-response spectrum (many have argued that's duration, but no statistical proof). Can you use your method to answer this question? Can you condition your machine learning process on Sa
-being the same (for example use a subset of data with similar Sa or artificially make them equal), then decide what's the next important parameter? Would the result be the same as in your
-conclusions? Or would it be influenced by removal of Sa as a parameter (this is possible since there are correlations between Sa and other metrics).
-}
-
-\response{%
-Response
-}
-
-Typos:\\
-\\
-
-\comment{1}{%
-Page 2, line 8: change focused to focus
-}
-
-\response{%
-Response
-}
-
-\comment{2}{%
-Page 2, line 20: change classifiers to classifier
-}
-
-\response{%
-Response
-}
-
-\comment{3}{%
-Page 13, line 185: change measure to measured
-}
-
-\response{%
-Response
+We changed it. Thanks.
 }
@@
 \response{%
-Response
+Yes, we fixed it.
 }
@@
 
-\introcomment{~}{%
-Review of Khoshnevis and Taborda for BSSA
-}
-
-\introcomment{~}{%
-Earthquake ground motion simulations are usually validated against observational data or empirically-derived solutions. However, exactly which parameters to compare (e.g., peak
-ground velocity, duration, spectral content) in order to assess the goodness-of-fit, is not always clear.
+% \introcomment{~}{%
+% Review of Khoshnevis and Taborda for BSSA
+% }
+%
+% \introcomment{~}{%
+% Earthquake ground motion simulations are usually validated against observational data or empirically-derived solutions.
+However, exactly which parameters to compare (e.g., peak ground velocity, duration, spectral content) in order to assess the goodness-of-fit, is not always clear.
 This paper uses machine learning techniques and eleven possible metrics for comparing (recorded and simulated) ground motions to determine the parameters that are most predictive of goodness-of-fit.
-This is an interesting paper, and will help future researchers
-prioritize which metrics to focus on when validating ground motions simulations.
+This is an interesting paper, and will help future researchers prioritize which metrics to focus on when validating ground motions simulations.
 The paper is generally well written.
@@
 It would also be useful to have some discussion regarding which parameters were found to be repetitive, incase future authors wish to look at more than the four recommended parameters.
-}
+% }
+%
+% \response{
+% We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
+The confusion happened because of integrating three different topics into a method to carry out this research.
+One technical term in one field is different in another field and that cause a confusion.
+For example cluster, class, and label all are referring to the same concept, however, it depends on the application.
+We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.
+% \vspace{2ex}
+% \newline
+% We also studied the models without response spectra (C8) and without response spectra (C8) and peak ground acceleration (C5).
+% }
 
 \comment{1}{%
-L. 8 "... could \textbf{focus} ..."
+Line 8: ``... could \textbf{focus} ...''
 }
@@
 \comment{2}{%
-L. 11 "... the objective \textbf{of} ..."
+Line 11: ``...the objective \textbf{of}...''
 }
@@
 \response{
-Response
+Added. Thanks.
 }
 
-\comment{3}{%
-L. 25-41 The frequent use of phrases such as "Among these methods,"; "Among the latter,"; "In this category,"; and "Within this group," make the text difficult to follow.
+% \comment{3}{%
+% Lines 25--41: The frequent use of phrases such as ``Among these methods,''; ``Among the latter,''; ``In this category,''; and ``Within this group,'' make the text difficult to follow.
+% }
+%
+% \response{
+% \color{red}{Ricardo, please address this comment.}
+% }
+%
+% \comment{4}{%
+% Lines 42--43: I would suggest moving the description of the Anderson (2004) method (i.e., Line 48+) to directly follow this sentence.}
+%
+% \response{
+% Thanks for the suggestion. We moved the description of the Anderson (2004) method (i.e., Line 48+) to the mentioned location. It reads better now.
+% }
+%
+% \comment{5}{%
+% Lines 59--61, 391--394: It should be stated somewhere that this is really intended for \textbf{engineering purposes}.
+% }
+%
+% \response{
+% Thanks for pointing this out.
+Since Anderson developed metrics for engineering application, we assumed it is implicitly mentioned.
+However, we agree that it should be explicitly clarified.
+We modified the text accordingly.
+% }
+
+\comment{6}{%
+Line 61: This sentence should end with a ``?''.
 }
@@
 \response{
-Response
+We disagree.
+In our view, these sentences were written as indirect questions that do not require the use of a question mark.
+To make this clearer we modified the introduction to the questions slightly by removing the colon and adding the word ``regarding''.
 }
 
-\comment{4}{%
-L. 42-43 I would suggest moving the description of the Anderson (2004) method (i.e., L. 48+) to directly follow this sentence.
+% \comment{7}{%
+% Lines 98--99: It is not clear how the equation for S is applied to values that are a function of period (e.g., the response or Fourier spectrum).
+Do you follow the guidelines provided by Anderson?
+If so, what frequency range/bands are you examining?
+% }
+%
+% \response{
+% In this study we only use broadband results. We added a clarification into the text. Thanks for the comment.
+% }
+
+\comment{8}{%
+Lines 128/132/137/etc.: ``... the three component\textbf{s}...''
 }
@@
 \response{
-Response
+Fixed. Thanks.
 }
 
-\comment{5}{%
-L. 59-61/L. 391-394 It should be stated somewhere that this is really intended for \textbf{engineering purposes}.
+% \comment{9}{%
+% Line 236: Where is the discussion on the consequences of overfitting? The only other place I see overfitting mentioned is line 328.
+% }
+%
+% \response{
+% Added. Thanks.
+% }
+%
+% \comment{10}{%
+% Line 236: Remove ``A matter we discuss later.''
+% }
+%
+% \response{
+% Removed. Thanks.
+% }
+%
+% \comment{11}{%
+% Lines 241--243: It is not helpful to list different algorithm names for making decision trees, without any information about them or how they differ.
+% }
+%
+% \response{
+% We added a brief description about them. Thanks.
+% }
+%
+% \comment{12}{%
+% Lines 257--258: I don't understand this sentence. What is a continuous attribute?
+% }
+%
+% \response{
+% We meant to say numerical continues attribute. Thanks for the comment. Fixed.
+% }
+%
+% \comment{13}{%
+% Lines 259--264: This paragraph contains a lot of jargon.
+% }
+%
+% \response{
+% Thanks for the comment. Additional clarification added.
+% }
+%
+% \comment{14}{%
+% Line 300: Isn't this also true for the C5-C7 combination?
+% }
+%
+% \response{
+% That is right. Added.
+% }
+%
+% \comment{15}{%
+% Line 328: What is ``a strong pruning process''?
+% }
+%
+% \response{
+% We modified it to heavy pruning process and also added its explanation.
+% }
+
+\comment{16}{%
+Line 433: ``We present the result\textbf{s} of...''
 }
@@
 \response{
-Response
+Added. Thanks.
 }
 
-\comment{6}{%
-L. 61 This sentence should end with a "?"
+\comment{17}{%
+Line 435: ``...goal of prioritiz\textbf{ing} and reduc\textbf{ing}...''
 }
@@
 \response{
-Response
+Indeed. We modified them accordingly. Thanks.
 }
 
-\comment{7}{%
-L. 98-99 It is not clear how the equation for S is applied to values that are a function of period (e.g., the response or Fourier spectrum). Do you follow the guidelines provided by Anderson? If
-so, what frequency range/bands are you examining?
+\comment{18}{%
+Line 443: Remove ``Similarly'' or ``similar'' (redundant).
 }
@@
 \response{
-Response
+We eliminated ``Similarly''. Thanks.
 }
 
-\comment{8}{%
-L. 128/132/137/etc. "... the three component\textbf{s} ..."
+\comment{19}{%
+Line 450: ``...remain valid.''
 }
@@
 \response{
-Response
+The extra ``s'' was deleted. Thanks.
 }
 
-\comment{9}{%
-L. 236 Where is the discussion on the consequences of overfitting? The only other place I see overfitting mentioned is L. 328.
-}
-
-\response{
-Response
-}
-
-\comment{10}{%
-L. 236 Remove "A matter we discuss later."
-}
-
-\response{
-Response
-}
-
-\comment{11}{%
-L. 241-243 It is not helpful to list different algorithm names for making decision trees, without any information about them or how they differ.
-}
-
-\response{
-Response
-}
-
-\comment{12}{%
-L. 257-258 I don't understand this sentence. What is a continuous attribute?
-}
-
-\response{
-Response
-}
-
-\comment{13}{%
-L. 259-264 This paragraph contains a lot of jargon.
-}
-
-\response{
-Response
-}
-
-\comment{14}{%
-L. 300 Isn't this also true for the C5-C7 combination?
-}
-
-\response{
-Response
-}
-
-\comment{15}{%
-L. 328 What is "a strong pruning process"?
-}
-
-\response{
-Response
-}
-
-\comment{16}{%
-L. 433 "We present the result\textbf{s} of..."
-}
-
-\response{
-Response
-}
-
-\comment{17}{%
-L. 435 "... goal of prioritiz\textbf{ing} and reduc\textbf{ing}..."
-}
-
-\response{
-Response
-}
-
-\comment{18}{%
-L. 443 Remove "Similarly" or "similar" (redundant)
-}
-
-\response{
-Response
-}
-
-\comment{19}{%
-L. 450 "... remain valid."
-}
-
-\response{
-Response
-}
-
-\comment{20}{%
-L. 455 "... as the decisive \textbf{parameters} ..." ? (This short concluding paragraph would be much stronger without the use of unnecessary phrases, i.e., "This latter point", "One the one hand",
-"on the other hand."
-}
-
-\response{
-Response
-}
+% \comment{20}{%
+% Line 455: ``...as the decisive \textbf{parameters}...'' ?
+(This short concluding paragraph would be much stronger without the use of unnecessary phrases, i.e., ``This latter point'', ``One the one hand'',
+% ``on the other hand.''
+% }
+%
+% \response{
+% \color{red}{Ricardo, please address this comment.}
+% }
 

### 2018-04-06 abstract.tex
went over the simple changes, more attention needed for the remaining ones.

@@
 This leads to their selection often being subjective, either based on intended applications or personal preferences.
-As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could \myremove{focused} \myrevision{focus} on a reduced number of metrics.
-We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques that help label and classify goodness-of-fit results, with the objective \myrevision{of} prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation.
+As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could \change{focus} on a reduced number of metrics.
+We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques that help label and classify goodness-of-fit results, with the objective \change{of} prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation.
 \myrevision{We use a dataset of three physic-based regional scale ground motion simulations based on one earthquake and three velocity models with eleven goodness of fit metrics.} In particular, we study the relationships that exist between \myremove{eleven} \myrevision{these} different metrics.
@@
 We then develop a family of decision trees using a C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four categories.
-These decision trees can be understood as rapid predictors of the quality of a simulation, or as  \myremove{a} data-informed classifiers that can help prioritize validation metrics.
+These decision trees can be understood as rapid predictors of the quality of a simulation, \change{or as data-informed classifiers} that can help prioritize validation metrics.
 Our analysis indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.

### 2018-04-06 conclusions.tex
went over the simple changes, more attention needed for the remaining ones.

@@
 
-We present the result\myrevision{s} of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of \myremove{prioritize} \myrevision{prioritizing} and \myremove{reduce} \myrevision{reducing} the number of metrics, and develop an application independent decision algorithm.
+We present the \change{results} of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of \change{prioritizing} and \change{reducing} the number of metrics, and develop an application independent decision algorithm.
 As a result of the data processing analysis, we propose a simplified algorithm based on a decision tree which uses only two metrics as opposed to the initial eleven available in the dataset.
 In particular, the proposed algorithm uses three (disjunctive decision) steps based on the values obtained for the metrics of the total energy and acceleration response spectra.
-\myremove{w}\myrevision{W}e also propose rules to allow for the new class-based validation criteria to combine results obtained from different components of motion, and demonstrate that the results obtained with the proposed algorithm using two metrics are comparable to those obtained with the score-based validation results used in other recent validation studies.
-\myremove{Similarly} \myrevision{Also}, one could implement similar rules to combine results from a frequency-band analysis, using, for example, the mode or majority validation class.
+We also propose rules to allow for the new class-based validation criteria to combine results obtained from different components of motion, and demonstrate that the results obtained with the proposed algorithm using two metrics are comparable to those obtained with the score-based validation results used in other recent validation studies.
+\change{One} could implement similar rules to combine results from a frequency-band analysis, using, for example, the mode or majority validation class.
 
@@
 In a future follow-up study it would be ideal to refine the tree adding other simulation datasets (i.e., from different earthquakes, regions, models, \myrevision{metrics,} etcetera) in order to arrive to a sufficiently sound and robust decision tree.
-The procedural steps laid out here, nonetheless, remain\myremove{s} valid.
+The procedural steps laid out here, nonetheless, \change{remain} valid.
 We showed that the procedure and background information used for clustering and decision making is stable, and it is likely that the metrics of energy and response spectra (along with peak acceleration and velocity as suggested by the additional trees) will prevail as those among the most decisive ones.
@@
 On the one hand, this contributes to clarifying a standing question in the area of validation, and on the other hand, it signals to simulators where the focus should be in the search for improvements in their models.
-
-% OLD NAEEM VERSION
-% -----------------
-
-% In this study we use two machine learning algorithms to address the long-term question about choosing the appropriate metrics to evaluate the ground motion simulation process independent of application and components.
-Assuming hypothetical stations, we labeled the dataset through constraint \kmeans{} algorithm with subspace analysis.
-Using the labeled data we developed two decision trees algorithm to evaluate the ground motion simulation.
-We generate two models where the first model only uses two metrics to evaluate the stations and the second model which is more accurate and also more complicated than the first model uses 4 metrics.
-According to our analysis, Response Spectra is the most important metric in classifying the stations.
-After response spectra, Total Energy are mostly used.
-We studied different combination of components's classes and developed a model to predict a components independent class.
-We applied the models on 2008 Chino Hills earthquake simulation synthetics and data GOF scores and discussed the results.
-A comprehensive study is needed to address the significance of the results or variations to earthquake magnitude, velocity model, station distance from epicenter, earthquake depth, and so on.
-The proposed model could be used with high accuracy (about 90 \%) in determining the accuracy of simulation independent of application and components, uniformly among physics-based ground motion simulation developers.

### 2018-04-06 dataset.tex
went over the simple changes, more attention needed for the remaining ones.

@@
 The modeling domain covered an area of \adomain{180}{135}{km} that included all the major sedimentary basins and other relevant geologic structures in the greater Los Angeles region.
-Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three component\myrevision{s} of motion (EW, NS, and UD).
+Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three \change{components} of motion (EW, NS, and UD).
 The simulation domain and the stations used for the analysis are shown in Figure \ref{fig:chino-hills}, and a sample of the validation results obtained by \citet{Taborda_2014_BSSA} using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps}.
-This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, averaged for all metrics, and the three component\myrevision{s} of motion.
-\myremove{This type of so-called GOF maps are commonly used in validation studies.}
+This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, averaged for all metrics, and the three \change{components} of motion.
 
@@
 
-Here, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the component of motions.
+Here, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the \change{components of motion}.
 Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be described in the following section was done using the whole dataset of scores.

### 2018-04-06 decision-tree.tex
went over the simple changes, more attention needed for the remaining ones.

@@
 This is done for all the trees obtained using all possible combinations of CF and $S_{\min}$ values.
-Then, we select the tree with the highest level of effectiveness as \myremove{measure} \myrevision{measured} by $F_1$, commensurate to such tree having a low complexity, i.e., a reduced number of metrics and a shallow depth, which is precisely the goal set at the start.
+Then, we select the tree with the highest level of effectiveness as \change{measured} by $F_1$, commensurate to such tree having a low complexity, i.e., a reduced number of metrics and a shallow depth, which is precisely the goal set at the start.
 Note that $F_1$ is obtained based on the testing dataset as opposed to the training dataset.

### 2018-04-06 introduction.tex
went over the simple changes, more attention needed for the remaining ones.

@@
 Out of the various validation methods just described, \citet{Anderson_2004_Proc} is perhaps the most widely used today \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
-\myrevision{ In essence, \citeauthor{Anderson_2004_Proc}'s method assesses the similarity of two signals based on the average score of 10 metrics.
+\myrevision{In essence, \citeauthor{Anderson_2004_Proc}'s method assesses the similarity of two signals based on the average score of 10 metrics.
 These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), their peak values (in acceleration, velocity and displacement), frequency content (in terms of Fourier and response spectra), and their time synchronization (through cross-correlation).}
@@
 
-The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics are \myrevision{relatively} well understood in both engineering and seismology contexts.
+% The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics are \myrevision{relatively} well understood in both engineering and seismology contexts. 
+The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics \change{convey physical meaning to both seismologists and engineers}.
 However, it has been pointed out that some of these metrics are redundant, or that other relevant metrics need to be added.
@@
 In the same spirit, \citet{Maufroy_2015_BSSA} reduced the number of metrics, limiting them to only those with comparable units.
-Unfortunately, none of these alternatives addresses the underlying questions: what are the \myrevision{most important} parameters that ought to be taken into account when validating ground motion simulations \myrevision{for engineering applications without specific target application (e.g., designing tall buildings or power plants)}, and what level of priority should these be given\myremove{.}\myrevision{?}
+Unfortunately, none of these alternatives addresses the underlying questions \change{regarding} what are the \myrevision{most important} parameters that ought to be taken into account when validating ground motion simulations \myrevision{for engineering applications without specific target application (e.g., designing tall buildings or power plants)}, and what level of priority should these \change{parameters} be given.
 

### 2018-04-06 response.tex
updates

@@
 It has been difficult to address this question because the answer varies depending on the application of the simulations (e.g., GMPE development, PSHA, structural analysis response of interest, etc).
-This paper is interesting because it provides a statistical approach to reduce the number of validation metrics and select a few representatives from a bigger set without compromising the final validation results.
+This paper is interesting because it provides a statistical approach to reduce the number of validation metrics and select a few representatives from a bigger set without
+% compromising the final validation results.
 The authors have done a great job providing background on the topic.
@@
 Both Abstract and Introduction need modification to specify the application for which the results and conclusions are true (i.e., the application here is to narrow down the 11 Anderson2004+ parameters to select fewer parameters that give similar results as the original 11 for the simulations they chose in this paper).
-Currently, the Abstract and Introduction give the impression that the conclusions are true for any engineering application or simulation method.
+Currently, the Abstract and Introduction give the impression that the conclusions are true for any engineering application or
+% simulation method.
 Furthermore, there should be a short discussion in the paper on how the users can apply the same method for different applications, for example for structural responses such as drift ratio or responses that are sensitive to high-frequency content of simulations (theoretically, this method should be applicable to any response of interest if they replace the GOF scores used in this study?
@@
 \response{%
-We appreciate the suggestion made by the reviewer, but we disagree.
-There should be nothing relative about the fact that these metrics are well understood by seismologists and engineers, and the strength of Anderson's method relying on this characteristic has been pointed out before by others.
-That being said, we modified the sentence with a choice of words that we have used in two previous papers published in BSSA, which convey the same idea we wanted to transmit.
+We appreciate the suggestion made by the reviewer, and in attention to it we modified the sentence with a choice of words that we have used in two previous papers published in BSSA, which convey the same idea we wanted to transmit.
+That being said, our view is that there should be nothing relative about the fact that these metrics are well understood by seismologists and engineers, and the strength of Anderson's because of this has been pointed out before by others.
 }
 
-% \comment{7}{%
-% The two questions posed in the introduction lines 59--61: I don't think the first one is really the question, all parameters should be considered if anyone has thought of them as important (i.e., has proposed them).
+\comment{7}{%
+The two questions posed in the introduction lines 59--61: I don't think the first one is really the question, all parameters should be considered if anyone has thought of them as important (i.e., has proposed them).
 The second question is the main question and the one being addressed in this paper.
-% }
-%
-% \response{%
-% We modified the text for clarification and also to acknowledge each individual proposed metrics' importance.
-% }
-%
+}
+
+\response{%
+We thank the reviewer for expressing his/her opinion about this point.
+First, we changed slightly the sentences as a whole to emphasize these are indirect questions as opposed to direct questions.
+That said, we do think they both are valid and relevant questions.
+The argument made by the reviewer that any or all parameters should be included if someone has thought of them as important is precisely one of the points we argue against, and the results of this research support that argument.
+The fact that someone thinks of a parameter as relevant does not guarantee its relevance, it is only the reflect of subjective (expert) opinion.
+There may be parameters that are more influential, and these can be so influential to the point of proving others irrelevant.
+This is the advantage of the procedure we follow here, which allows one to evaluate objectively, based on how data relates to the results, how relevant or irrelevant is each one of the metrics involved in reaching the same conclusions.
+This, in turn, allows one to prioritize the metrics, which is the point of the second indirect question posed in this part of the introduction.
+}
+
 % \comment{8}{%
@@
 
+\noindent
 \textbf{Typos:}
+
+\vspace{2em}
 

### 2018-04-06 introduction.tex
updates

@@
 % The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics are \myrevision{relatively} well understood in both engineering and seismology contexts.
+
 The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics \change{convey physical meaning to both seismologists and engineers}.
@@
 In the same spirit, \citet{Maufroy_2015_BSSA} reduced the number of metrics, limiting them to only those with comparable units.
-Unfortunately, none of these alternatives addresses the underlying questions \change{regarding} what are the \myrevision{most important} parameters that ought to be taken into account when validating ground motion simulations \myrevision{for engineering applications without specific target application (e.g., designing tall buildings or power plants)}, and what level of priority should these \change{parameters} be given.
+Unfortunately, none of these alternatives addresses the underlying questions \change{regarding} what are the \change{most important} parameters that ought to be taken into account when validating ground motion simulations \change{for a specific application}, and what level of priority should these \change{parameters} be given.
+
+% Unfortunately, none of these alternatives addresses the underlying questions \change{regarding} what are the \myrevision{most important} parameters that ought to be taken into account when validating ground motion simulations \myrevision{for engineering applications without specific target application (e.g., designing tall buildings or power plants)}, and what level of priority should these \change{parameters} be given.
 

### 2018-04-07 response.tex
updates

@@
 
+\thispagestyle{empty}
+\setcounter{page}{0}
+
+~
+
+\vspace{6em}
+\noindent\today
+
+\vspace{5em}
+\noindent
+Dr.~Thomas Brocher\\
+Associate Editor, BSSA\\
+BSSA Editorial Office\\
+400 Evelyn Ave, Ste 201, Albany, CA 94706
+
+\vspace{4em}
+\noindent
+Dear Dr. Brocher,
+
+\vspace{1em}
+\noindent
+Thank you for coordinating the review of our manuscript (BSSA-D-18-00056) and for your own comments about it.
+Altogether, they have certainly helped improve our submission to the Bulletin.
+We are enclosing a response to both your suggestions and those of the reviewers.
+We were able to satify most, if not all of them.
+To help in this second review cycle, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from your comments and those of the reviewers in green font; and below we provide a response to each one of your comments, followed by our responses to the reviewers.
+In all three cases we reproduce the original comments in italic black font, and our responses in regular blue font.\\
+
+\noindent
+We look forward to hearing back from you.
+
+\vspace{2em}
+\noindent
+Sincerely yours,
+
+\vspace{5em}
+\noindent
+Ricardo Taborda\\
+Department of Civil Engineering, and\\
+Center for Earthquake Research and Information\\
+The University of Memphis
+
+\newpage
 \begin{center}
@@
 	\large
-	Authors' Response to Associate Editor (Dr. Thomas Brocher)
+	Authors' Response to the Associate Editor
 \end{center}
 
-% \noindent
-% We thank Dr.
-Thomas Brocher for his comments and suggestions, which helped us improve the manuscript.
-To help the review process, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from these comments in green font and the removed text in red font.
-In the following, we provide a response to each comment.
+\noindent
+We thank the Associate Editor for his comments and suggestions.
+Below, we provide a response to each comment.
 The original comments are in italic black font, followed by our responses in regular blue font.
-% \vspace{2ex}
-% \newline
-%
-% \introcomment{~}{%
-% I agree with the peer reviewers that this paper addresses a topical subject of great importance to BSSA readers.
+\vspace{2ex}
+\newline
+
+\introcomment{~}{%
+I agree with the peer reviewers that this paper addresses a topical subject of great importance to BSSA readers.
 The science is in the process of transitioning from purely observational based analysis of ground motions to a mixture of observations and simulations of ground motions.
@@
 The peer reviewers have made some very helpful suggestions for improving the paper that I hope the authors will strongly consider.
-% }
-%
-% \introcomment{~}{%
-% In particular, I agree with their comment that the limitations of the study, i.e., the choice of 11 metrics and earthquake simulations for only one earthquake, needs to be included in the abstract and the conclusions.
+}
+
+\introcomment{~}{%
+In particular, I agree with their comment that the limitations of the study, i.e., the choice of 11 metrics and earthquake simulations for only one earthquake, needs to be included in the abstract and the conclusions.
 In addition, I also agree that all the terms and methodology should be well defined for the BSSA readership, so that the paper can be understood by them without resort to frequent online searches.
-% }
-%
-% \introcomment{~}{%
-% I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit.
+}
+
+\introcomment{~}{%
+I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit.
 It clearly shows that only these metrics are change significantly with the goodness of fit (P,F, G, E).
@@
 In any case, more
-% discussion of this figure might help the authors make the case for the main conclusions.
-% }
+discussion of this figure might help the authors make the case for the main conclusions.
+}
 %

### 2018-04-08 response.tex
updated

@@
 \usepackage{simplemargins}
+\usepackage{ifpdf}
+
+\ifpdf
+	\usepackage[pdftex,bookmarks=false]{hyperref}
+\else
+	\usepackage{graphics}
+	\usepackage{graphicx}
+	\usepackage{epsfig}
+	\usepackage[dvipdfm,bookmarks=false]{hyperref}
+\fi
+
+\hypersetup{
+    pdfpagelayout=OneColumn,
+	colorlinks = true,
+	urlcolor   = blue,
+	citecolor  = blue,
+	linkcolor  = blue
+}
 
@@
 BSSA Editorial Office\\
-400 Evelyn Ave, Ste 201, Albany, CA 94706
+400 Evelyn Ave, Ste 201\\
+Albany, CA 94706
 
@@
 
+\response{%
+Thank you.
+We were able to address all the suggestions made by the reviewers.
+In the vast majority of the cases we adopted their suggestions and we only disagreed with a very small portion of them, in which case we presented our arguments in the response to the reviews enclosed here.
+}
+
 \introcomment{~}{%
-In particular, I agree with their comment that the limitations of the study, i.e., the choice of 11 metrics and earthquake simulations for only one earthquake, needs to be included in the abstract and the conclusions.
+In particular, I agree with their comment that the limitations of the study, i.e., the choice of 11 metrics and earthquake simulations for only one earthquake, needs to be included in the abstract and the conclusions.}
+
+\response{
+We modified the abstract and the conclusions accordingly.
+In the case of the abstract we added minor modifications to emphasize the use of 11 metrics, and a whole new sentence explicitly referencing the used of three simulation datasets corresponding to the 2008 Chino Hills earthquake.
+In the case of the conclusions, although we already had a paragraph describing the limitations of the study, we added a small sentence reemphasizing it.
+}
+
+\introcomment{~}{%
 In addition, I also agree that all the terms and methodology should be well defined for the BSSA readership, so that the paper can be understood by them without resort to frequent online searches.
@@
 I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit.
-It clearly shows that only these metrics are change significantly with the goodness of fit (P,F, G, E).
+It clearly shows that only these metrics are change significantly with the goodness of fit (P, F, G, E).
 Plotting these values as a function of P, F, G, and E together on the same figure might allow the reader to see this relationship more clearly and might help provide a heuristic understanding of how C4, C5, C6, and C8 are diagnostic whereas C11 is constant with P, F, G, and E, and thus not diagnostic.
@@
 
-% \introcomment{~}{%
-% This is an interesting article for the readers of the BSSA journal.
+\introcomment{~}{%
+This is an interesting article for the readers of the BSSA.
 The authors propose a new approach based on machine learning techniques to determine the importance of validation metrics for ground motion simulation.
@@
 It has been difficult to address this question because the answer varies depending on the application of the simulations (e.g., GMPE development, PSHA, structural analysis response of interest, etc).
-This paper is interesting because it provides a statistical approach to reduce the number of validation metrics and select a few representatives from a bigger set without
-% compromising the final validation results.
+This paper is interesting because it provides a statistical approach to reduce the number of validation metrics and select a few representatives from a bigger set without compromising the final validation results.
 The authors have done a great job providing background on the topic.
-However, their application is very narrow, they only consider the 11 Anderson2004+ parameters (as they've noted there are other parameters) for one simulation method for one earthquake, and as a result their conclusions are only valid for this particular case.
+However, their application is very narrow, they only consider the 11 Anderson2004+ parameters (as they have noted, there are other parameters) for one simulation method for one earthquake, and as a result their conclusions are only valid for this particular case.
 They have mentioned this limitation in the conclusions, but the abstract and introduction do not explain the limitations and the application-dependency of the proposed approach.
 Both Abstract and Introduction need modification to specify the application for which the results and conclusions are true (i.e., the application here is to narrow down the 11 Anderson2004+ parameters to select fewer parameters that give similar results as the original 11 for the simulations they chose in this paper).
-Currently, the Abstract and Introduction give the impression that the conclusions are true for any engineering application or
-% simulation method.
+Currently, the Abstract and Introduction give the impression that the conclusions are true for any engineering application or simulation method.
 Furthermore, there should be a short discussion in the paper on how the users can apply the same method for different applications, for example for structural responses such as drift ratio or responses that are sensitive to high-frequency content of simulations (theoretically, this method should be applicable to any response of interest if they replace the GOF scores used in this study?
 If true this should be mentioned as it will be of interest to others).
-% }
-%
-% \response{%
-% The reviewer pointed out a very important topic that we believe we did not discuss it well in the manuscript.
-The manuscript presents steps of moving from considerable amount of data with specialties knowledge (background knowledge) to a final product that gives a better understanding about the data.
-According the reviewer suggestion we added a paragraph  to discuss this important point.
-Also we modified abstract an introduction for acknowledging the limitations with the study.
-% }
-%
+}
+
+\response{%
+We thank the reviewer for the positive feedback in this general review.
+We agree that there were some important points that we did not discuss fully in the manuscript and we have corrected them.
+In attention to this general review, we modified the abstract, the introduction, and the conclusions to (i) transparently point out the limitations of our study, and (ii) offer an additional perspective on the potential refinement of it or use with additional datasets or in reference to additional validation metrics.
+}
+
 % \comment{1}{
@@
 % }
-%
-% \comment{2}{%
-% The terms ``common practice'' or ``common standards'' have been mentioned several times.
+
+\comment{2}{%
+The terms ``common practice'' or ``common standards'' have been mentioned several times.
 Because GMSV is such a new field (and I don't think there is consensus, I for one don't agree with some statements but I agree that others may), I think these terms should be removed.
 Just say this is your opinion, or it is based on the references that you have listed.
-% }
-%
-% \response{%
-% Thanks for the comment. We fixed it.
-% }
-%
+}
+
+\response{%
+Thanks for the comment. We limited the use of these terms to a minimum and/or replaced them by sentences providing additional context.
+}
+
 % \comment{3}{%
-% In abstract, ``semi-supervised'' and ``supervised'' learning techniques are mentioned, but in the paper these terms are not explained, what part of the learning techniques is semi-supervised
-% and what does that mean?
+% In abstract, ``semi-supervised'' and ``supervised'' learning techniques are mentioned, but in the paper these terms are not explained, what part of the learning techniques is semi-supervised and what does that mean?
 % }
@@
 % \response{%
+% not present in the abstract, present in the introduction, I will come back to this.
 % \color{red}{Ricardo, please address this comment.}
@@
 % }
-%
-% \comment{13}{%
-% Page 17, lines 391--394: this sentence should be re-worded, what is an ``accurate conclusion''?
+
+\comment{13}{%
+Page 17, lines 391--394: this sentence should be re-worded, what is an ``accurate conclusion''?
 You should be very specific that your conclusion is that instead of using all 11 validation metrics you can select a few in order to get the same categorization as all 11 metrics would have given you.}
-%
+
 % \response{%

### 2018-04-08 abstract.tex
updated

@@
 As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could \change{focus} on a reduced number of metrics.
-We present an analysis of ground motion validation metrics using semi-supervised and supervised learning techniques that help label and classify goodness-of-fit results, with the objective \change{of} prioritizing and narrowing the choice of metrics available and commonly used in ground motion simulation validation.
-\myrevision{We use a dataset of three physic-based regional scale ground motion simulations based on one earthquake and three velocity models with eleven goodness of fit metrics.} In particular, we study the relationships that exist between \myremove{eleven} \myrevision{these} different metrics.
-In the process carried out, these metrics are understood as part of a multi-dimensional space.
+We present an analysis \change{that looks into eleven} ground motion validation metrics using semi-supervised and supervised learning techniques\change{.
+These techniques} help label and classify goodness-of-fit results with the objective \change{of} prioritizing and narrowing the choice of \change{these metrics.
+In particular, we use the validation datasets of a series of physics-based ground motion simulations done for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake using different velocity models.
+We} study the relationships that exist between \change{the eleven metrics, and carry out a process where} these metrics are understood as part of a multi-dimensional space.
 We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address high-dimensional effects.
-This allows us to label a given validation dataset into four categories (poor, fair, good, excellent) following \myrevision{previous studies} \myremove{common practice}.
-We then develop a family of decision trees using a C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four categories.
+This allows us to label a given validation dataset into four categories (poor, fair, good, excellent) following \change{previous studies}.
+We then develop a family of decision trees using the C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four referenced categories.
 These decision trees can be understood as rapid predictors of the quality of a simulation, \change{or as data-informed classifiers} that can help prioritize validation metrics.
-Our analysis indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.
-\myrevision{Adding other simulation datasets (i.e., from different earthquakes, regions, models, metrics, etcetera) will improve the results in terms of developing a sufficiently sound and robust decision tree, however, the procedural steps laid out in this study, nonetheless, remain valid.}
+Our analysis\change{, although limited to the particular datasets used here,} indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.

### 2018-04-08 conclusions.tex
updated

@@
 
-We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulations done for a particular earthquake \myrevision{(one earthquake and three velocity models)} in a particular region.
-In a future follow-up study it would be ideal to refine the tree adding other simulation datasets (i.e., from different earthquakes, regions, models, \myrevision{metrics,} etcetera) in order to arrive to a sufficiently sound and robust decision tree.
-The procedural steps laid out here, nonetheless, \change{remain} valid.
-We showed that the procedure and background information used for clustering and decision making is stable, and it is likely that the metrics of energy and response spectra (along with peak acceleration and velocity as suggested by the additional trees) will prevail as those among the most decisive ones.
+We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulations done for a particular earthquake, in a particular region\change{, and using a particular set of metrics}.
+In a future follow-up study, it would be ideal to refine the tree adding other simulation datasets (i.e., from different earthquakes, regions, models, \change{and using additional metrics)} in order to arrive to a sufficiently sound and robust decision tree.
+The procedural steps laid out here, nonetheless, \change{remain} valid\change{, and there is no reason why not to use it also in other contexts utilizing other metrics (e.g., structural responses metrics such as drift ratio) provided that they are properly normalized.
+In summary, we can say that for the case of the metrics used here,} we showed that the procedure and background information used for clustering and decision making is stable, and it is likely that\change{---despite the limitations just described---}the metrics of energy and response spectra (along with peak acceleration and velocity as suggested by the additional trees) will prevail as those among the most decisive ones.
 

### 2018-04-08 introduction.tex
updated

@@
 We use a constrained \kmeans{} method \citep[e.g.,][]{Macqueen_1967_Proc, Wagstaff_2001_Proc} and conduct a subspace clustering analysis to address high-dimensional effects in order to add labels to the dataset in terms of four categories: poor, fair, good, and excellent.
-These categories \myrevision{are commonly used in previous validation studies} \myremove{adhere to the common practice in validation}.
+These categories \change{were defined by \citet{Anderson_2004_Proc} and have been adopted---despite some differences---by other GOF methods used in verification and validation \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}.}
+% \myrevision{are commonly used in previous validation studies} 
+% \myremove{adhere to the common practice in validation}. 
 \myrevision{In the machine learning linguistic, \kmeans{} clustering approach is considered as an unsupervised learning method.
-However, when we add background knowledge into the process it becomes constrained \kmeans{} method and is considered as a semi-supervised learning method.} The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}.
+However, when we add background knowledge into the process it becomes constrained \kmeans{} method and is considered as a semi-supervised learning method.}
+The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}.
 Upon labeling the dataset, we develop a family of decision trees using a C5.0 algorithm \citep[][]{Quinlan_1993_Book, Quinlan_1996_JAIR}, and select a few to prioritize the metrics involved in the validation process.
@@
 
-In summary, our results indicate that \myrevision{for \citeauthor{Anderson_2004_Proc}'s metrics based the results of 2008 $M_w$ 5.4 Chino Hills earthquake for three velocity models,} among the eleven metrics considered in the analysis, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.
+In summary, \change{although limited to the datasets from specific simulation method and event}, our results indicate that 
+% \myrevision{for \citeauthor{Anderson_2004_Proc}'s metrics based the results of 2008 $M_w$ 5.4 Chino Hills earthquake for three velocity models,} 
+among the eleven metrics considered in the analysis, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.
 We test our prediction model and offer a discussion about its implications and potential use in future validation efforts.

### 2018-04-08 metrics.tex
updated

@@
 He concluded that for a typical distribution of the scores, the GOF could be categorized into four groups: poor (for scores from 0 to 4), fair (4 to 6), good (6 to 8), and excellent (for scores from 8 to 10).
-While this categorization was arguably subjective, over the years it has  \myrevision{ been frequently used and} \myremove{become common practice and has} been adopted or matched by alternative validation methods \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}, making the method itself a reference baseline for various validation studies \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
+While this categorization is arguably subjective, over the years it has \change{been adopted---despite some differences in the ranges---by other GOF methods employed in verification and validation \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}.
+As a result, \citeauthor{Anderson_2004_Proc}'s method has been used as} a reference baseline for various validation studies \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
+
+% \myrevision{ been frequently used and} 
+% \myremove{become common practice and has} 
+% been adopted or matched by alternative validation methods \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}, making the method itself a reference baseline for various validation studies \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
 

### 2018-04-08 results.tex
went over the simple changes, more attention needed for the remaining ones.
updates
updated

@@
 We used the oversampling approach described in the previous section.
-We replicated data from the excellent class \myrevision{through random selection of data from that class until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.
-This oversampling is considered as a $\times$10 factor.} \myremove{randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.} By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}.
+We replicated data from the excellent class
+\myrevision{through random selection of data from that class until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.
+This oversampling is considered as a $\times$10 factor.}
+\myremove{randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.} 
+% By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}. 
+\mychange{According to \citet{Weiss_2003_JAIR}, an oversample rate of $\times$10 is considered to be acceptable in this type of data processing.}
 (Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary.
@@
 
-%\begin{figure*}[t]
-%	\centering
-%	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
-%	\caption{Accuracy of tree predictions in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed with respect to the number of attributes (GOF metrics) as a function of: \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees.
-The rings around some of the dots indicate select trees used as reference in the text, other figures and tables.
-The color version of this figure is available only in the electronic edition.}
-%	\label{fig:nodes-depth}
-%\end{figure*}
-
 \begin{figure*}[t]
 	\centering
-	\includegraphics[width=\textwidth]{figures/pdf/figure-08_mod}
+	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
 	\caption{Accuracy of tree predictions in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed with respect to the number of attributes (GOF metrics) as a function of: \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees.
@@
 All coincide in the use of the total energy (C4) and response spectra (C8) as key metrics.
-T2 adds peak acceleration (C5), and T3 adds peak velocity \myremove{(C5)} \myrevision{(C6)} to the previous metrics.
+T2 adds peak acceleration (C5), and T3 adds peak velocity \change{(C6)} to the previous metrics.
 

### 2018-04-19 response.tex
update

@@
 We modified the abstract and the conclusions accordingly.
-In the case of the abstract we added minor modifications to emphasize the use of 11 metrics, and a whole new sentence explicitly referencing the used of three simulation datasets corresponding to the 2008 Chino Hills earthquake.
+In the case of the abstract we added minor modifications to emphasize the use of 11 metrics, and a whole new sentence explicitly referencing the use of three simulation datasets corresponding to the 2008 Chino Hills earthquake.
 In the case of the conclusions, although we already had a paragraph describing the limitations of the study, we added a small sentence reemphasizing it.
@@
 
-\introcomment{~}{%
-In addition, I also agree that all the terms and methodology should be well defined for the BSSA readership, so that the paper can be understood by them without resort to frequent online searches.
-}
-
-\introcomment{~}{%
-I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit.
+% \introcomment{~}{%
+% In addition, I also agree that all the terms and methodology should be well defined for the BSSA readership, so that the paper can be understood by them without resort to frequent online searches.
+% }
+%
+% \introcomment{~}{%
+% I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit.
 It clearly shows that only these metrics are change significantly with the goodness of fit (P, F, G, E).
@@
 In any case, more
-discussion of this figure might help the authors make the case for the main conclusions.
-}
+% discussion of this figure might help the authors make the case for the main conclusions.
+% }
 %
@@
 \response{%
-We appreciate the suggestion made by the reviewer, and in attention to it we modified the sentence with a choice of words that we have used in two previous papers published in BSSA, which convey the same idea we wanted to transmit.
+We appreciate the suggestion made by the reviewer.
+In attention to it, we modified the sentence with a choice of words that we have used in two previous papers published in BSSA, which convey the same idea we wanted to transmit but in a slightly different way.
 That being said, our view is that there should be nothing relative about the fact that these metrics are well understood by seismologists and engineers, and the strength of Anderson's because of this has been pointed out before by others.
@@
 The argument made by the reviewer that any or all parameters should be included if someone has thought of them as important is precisely one of the points we argue against, and the results of this research support that argument.
-The fact that someone thinks of a parameter as relevant does not guarantee its relevance, it is only the reflect of subjective (expert) opinion.
-There may be parameters that are more influential, and these can be so influential to the point of proving others irrelevant.
+The fact that someone thinks of a parameter as relevant does not guarantee its relevance, it is only the reflect of his/her expert (though still subjective) opinion.
+There may be parameters that are more influential, and these can be as influential as to the point of proving others irrelevant.
 This is the advantage of the procedure we follow here, which allows one to evaluate objectively, based on how data relates to the results, how relevant or irrelevant is each one of the metrics involved in reaching the same conclusions.
@@
 
-% \response{%
-% You are right. It could be misleading. We re-worded the sentence.
-% }
-%
-% \comment{14}{%
-% Page 18, line 405: this is implying that fig 10 is similar to fig2!
+\response{%
+You are right. It could be misleading. We reworded the second half of this paragraph to make things more transparent and to keep them in the context of what we used in our study.
+}
+
+\comment{14}{%
+Page 18, line 405: this is implying that fig 10 is similar to fig2!
 I personally cannot see this in the figures!
@@
 It might be better to regenerate fig 11a but with different color scales that only vary by four choices, then the comparison would be much easier.
-% }
-%
+}
+
 % \response{%
@@
 % }
-%
+
 % \response{%
@@
 
-% \introcomment{~}{%
-% Review of Khoshnevis and Taborda for BSSA
-% }
-%
+\introcomment{~}{%
+Review of Khoshnevis and Taborda for BSSA
+}
+
 % \introcomment{~}{%
@@
 % }
-%
-% \comment{4}{%
-% Lines 42--43: I would suggest moving the description of the Anderson (2004) method (i.e., Line 48+) to directly follow this sentence.}
-%
-% \response{
-% Thanks for the suggestion. We moved the description of the Anderson (2004) method (i.e., Line 48+) to the mentioned location. It reads better now.
-% }
-%
-% \comment{5}{%
-% Lines 59--61, 391--394: It should be stated somewhere that this is really intended for \textbf{engineering purposes}.
-% }
-%
-% \response{
-% Thanks for pointing this out.
-Since Anderson developed metrics for engineering application, we assumed it is implicitly mentioned.
-However, we agree that it should be explicitly clarified.
-We modified the text accordingly.
-% }
+
+\comment{4}{%
+Lines 42--43: I would suggest moving the description of the Anderson (2004) method (i.e., Line 48+) to directly follow this sentence.}
+
+\response{
+Thanks for the suggestion. We moved it and it does read better this way.
+}
+
+\comment{5}{%
+Lines 59--61, 391--394: It should be stated somewhere that this is really intended for \textbf{engineering purposes}.
+}
+
+\response{While it is true that the initial intent of Anderson (2004) was to develop credibility of synthetic seismograms for engineering applications, we do not share the need to make that intent explicit here.
+On the one hand, these metrics have already been used by several authors who did not intend to validate seismograms for any engineering application, but to simply show---by means of them---that their synthetics were realistic in a general sense.
+On the other hand, we do not want to convey the idea that the steps that we followed are exclusive to narrowing validation criteria for engineering applications.
+Others may see our work and these steps as applicable to other ends.
+That being said, we hope the reviewer will be pleased to notice that we modified the Introduction to point out that engineering applications are the one of the motivating factors for the increased attention in verification and validation.
+Please see the first sentence in the introduction (see line 29).}
 
@@
 We disagree.
-In our view, these sentences were written as indirect questions that do not require the use of a question mark.
-To make this clearer we modified the introduction to the questions slightly by removing the colon and adding the word ``regarding''.
+In our view, these sentences are written as indirect questions and do not require the use of a question mark.
+To make this clearer, we modified the beginning of the sentences slightly by removing the colon and adding the word ``regarding''.
 }
 
-% \comment{7}{%
-% Lines 98--99: It is not clear how the equation for S is applied to values that are a function of period (e.g., the response or Fourier spectrum).
+\comment{7}{%
+Lines 98--99: It is not clear how the equation for $S$ is applied to values that are a function of period (e.g., the response or Fourier spectrum).
 Do you follow the guidelines provided by Anderson?
 If so, what frequency range/bands are you examining?
-% }
-%
-% \response{
-% In this study we only use broadband results. We added a clarification into the text. Thanks for the comment.
-% }
+}
+
+\response{
+Yes.
+We follow the guidelines of Anderson (2004), but the question of the reviewer gives way to two separate points of those guidelines: first, that in general the equation for $S$ is applied to all values of frequency (or period) for which is computed and then averaged to produce the GOF scores for C8 and C9; and second, that one can do this process for the broadband signals as well as for different frequency bands, and then average the results.
+The latter of these two points is applicable to not only C8 and C9 but to all metrics.
+In this study we only use broadband results to avoid the adding other parameters to the validation process such as the width of the frequency bands or the characteristics of the filters.
+We added a clarification about how the method guidelines were used to address this comment.
+}
 
@@
 
-% \comment{9}{%
-% Line 236: Where is the discussion on the consequences of overfitting? The only other place I see overfitting mentioned is line 328.
-% }
-%
-% \response{
-% Added. Thanks.
-% }
-%
-% \comment{10}{%
-% Line 236: Remove ``A matter we discuss later.''
-% }
-%
-% \response{
-% Removed. Thanks.
-% }
-%
+\comment{9}{%
+Line 236: Where is the discussion on the consequences of overfitting? The only other place I see overfitting mentioned is line 328.
+}
+
+\response{
+We added it at this point. Thank you.
+}
+
+\comment{10}{%
+Line 236: Remove ``A matter we discuss later.''
+}
+
+\response{
+We removed it, and in its place added an explanation about overfitting in response to the previous comment.
+}
+
 % \comment{11}{%
@@
 % }
-%
-% \comment{12}{%
-% Lines 257--258: I don't understand this sentence. What is a continuous attribute?
-% }
-%
-% \response{
-% We meant to say numerical continues attribute. Thanks for the comment. Fixed.
-% }
-%
+
+\comment{12}{%
+Lines 257--258: I don't understand this sentence. What is a continuous attribute?
+}
+
+\response{
+Following Equation (3) we define $A[a]$ as all the possible values of attribute $A$.
+The use of $[a]$ implies that $A$ is considered a numerically discrete attribute.
+We therefore meant to say that if $A$ continues attribute.
+Thanks for the comment.
+Fixed.
+}
+
 % \comment{13}{%
@@
 % }
-%
-% \comment{14}{%
-% Line 300: Isn't this also true for the C5-C7 combination?
-% }
-%
-% \response{
-% That is right. Added.
-% }
-%
-% \comment{15}{%
-% Line 328: What is ``a strong pruning process''?
-% }
-%
-% \response{
-% We modified it to heavy pruning process and also added its explanation.
-% }
+
+\comment{14}{%
+Line 300: Isn't this also true for the C5-C7 combination?
+}
+
+\response{
+That is right. We modified the sentence to reflect that.
+}
+
+\comment{15}{%
+Line 328: What is ``a strong pruning process''?
+}
+
+\response{
+We modified it to ``heavy pruning process'' and added an explanation earlier in the Decision Trees section.
+}
 

### 2018-04-19 decision-tree.tex
update

@@
 In our case, as we will see later, we apply oversampling to one of our dataset classes.
-This process has consequences, mainly that it increases the likelihood of overfitting.
-\myremove{A matter we discuss later.} \myrevision{Overfitting happens when the prediction algorithm predicts the final results too closely or exactly to the training data, however, it provides non-accurate results for unseen or test data.}
+\change{In general, oversampling increases the possibility of overfitting, which occurs when the training data leads the algorithm to produce a decision tree yields that predicts outcomes too close to or exactly the same as the training data.
+This is not desirable because such a tree would lead to inaccurate predictions for other unseen data.
+Overfitting can be avoided by using conservative oversampling rates \citet{Weiss_2003_JAIR} and/or by applying heavy pruning methods.
+(Pruning, as the word implies, is the process of cutting branches off a tree.
+Heavy pruning methods limit the complexity of the tree by constraining the number of branches and/or depth of a tree.)}
+
+% This process has consequences, mainly that it increases the likelihood of overfitting.
+\myrevision{Overfitting happens when the prediction algorithm predicts the final results too closely or exactly to the training data, however, it provides non-accurate results for unseen or test data.}
 
@@
 Conversely, an extremely low value of entropy would mean most of the data falls within a particular class.
-With this in mind, the information gain $G$ of an attribute $A$ for the dataset $S$, or $G(S,A)$ in equation (\ref{eq:gain}), is the expected reduction in entropy caused by partitioning the dataset according to attribute $A$ \citep{Mitchell_1997_Book}.
-For selection of a threshold for a \myrevision{numerical} continuous attribute refer to \citet{Quinlan_1996_JAIR}.
+With this in mind, the information gain $G$ of an attribute $A$ for the dataset $S$, or $G(S,A)$ in equation (\ref{eq:gain}), is the expected reduction in entropy caused by partitioning the dataset according to \change{the attribute $A$ \citep{Mitchell_1997_Book}, which here we consider to be numerically discrete.
+For selection of a threshold for a numerically continuous attribute, please refer to \citet{Quinlan_1996_JAIR}.}
+
+% attribute $A$ \citep{Mitchell_1997_Book}. For selection of a threshold for a \myrevision{numerical} continuous attribute refer to \citet{Quinlan_1996_JAIR}.
 

### 2018-04-19 introduction.tex
update

@@
 
-Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation.
+Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation\change{, and a growing interest in the use of synthetic seismograms for engineering applications}.
 \myrevision{Validation is the process concerned with determining whether a conceptual model and the results it leads to are accurate representations of the physical system under study.
@@
 Out of the various validation methods just described, \citet{Anderson_2004_Proc} is perhaps the most widely used today \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
-\myrevision{In essence, \citeauthor{Anderson_2004_Proc}'s method assesses the similarity of two signals based on the average score of 10 metrics.
+\change{In essence, this method assesses the similarity of two signals based on the average score of 10 metrics.
 These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), their peak values (in acceleration, velocity and displacement), frequency content (in terms of Fourier and response spectra), and their time synchronization (through cross-correlation).}
-
-We, in particular, have consistently relied on a modified version of this method for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake, and for the evaluation of velocity models in southern California through the validation of a large suite of moderate-magnitude earthquake simulations \citep{Taborda_2013_BSSA, Taborda_2014_BSSA, Taborda_2016_GJI}.
-
-\myremove{In essence, \citeauthor{Anderson_2004_Proc}'s method assesses the similarity of two signals based on the average score of 10 metrics.
+% 
+We, in particular, have consistently relied on a modified version of \change{\citeauthor{Anderson_2004_Proc}'s method} for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake \citep{Taborda_2013_BSSA, Taborda_2014_BSSA}, and for the evaluation of velocity models in southern California through the validation of a large suite of moderate-magnitude earthquake simulations \citep{Taborda_2016_GJI}.
+% 
+% \myremove{In essence, \citeauthor{Anderson_2004_Proc}'s method assesses the similarity of two signals based on the average score of 10 metrics.
 These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), their peak values (in acceleration, velocity and displacement), frequency content (in terms of Fourier and response spectra), and their time synchronization (through cross-correlation).
 }
-
+% 
 % The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics are \myrevision{relatively} well understood in both engineering and seismology contexts.

### 2018-04-19 metrics.tex
update

@@
 where $S$ is the GOF score that results from comparing values $p_1$ and $p_2$ from signals 1 and 2, respectively, for each one of the different metrics in Table \ref{tab:metrics}.
-\citeauthor{Anderson_2004_Proc} also provided guidelines for the process to be applied to the original signals as well as to those resulting from a sequential set of band-pass filters covering the whole frequency range of interest.
+\change{In the case of C8 and C9, where the values $p_1$ and $p_2$ are function of the period ($T$) and frequency ($f$), respectively, $S$ is computed for all values of $T$ and $f$ and averaged to produce a single mean score for metrics C8 and C9.} \citeauthor{Anderson_2004_Proc} also provided guidelines for the process to be applied to the original signals as well as to those resulting from a sequential set of band-pass filters covering the whole frequency range of interest.
 In his method, this frequency-domain decomposition should have pass bands defined following a logarithmic distribution in order to give more weight to the low-frequencies, and the GOF scores of all sub-bands and the broadband be averaged into a single GOF final score.
-\myrevision{However, in this study, in order to reduce the uncertainties that can come from different filtering process \citep[e.g., see ][]{Khoshnevis_2015_Proc} we use broadband results (\feq{0}{4}).
+\change{However, in this study, in order to avoid the additional parameters that would result from this approach (i.e., sub-bands width and filter characteristics), we used only the broadband (\frange{0}{4}) results.}
+
+% reduce the uncertainties that can come from different filtering process \citep[e.g., see ][]{Khoshnevis_2015_Proc} we use broadband results (\feq{0}{4}).
+For metrics which are functions of frequency (or period) like Fourier spectra and response spectra, the final result is the mean value of application of equation(\ref{eq:gof-function}) on all frequencies (or periods).}
+
+% \myrevision{However, in this study, in order to reduce the uncertainties that can come from different filtering process \citep[e.g., see ][]{Khoshnevis_2015_Proc} we use broadband results (\feq{0}{4}).
 For metrics which are functions of frequency (or period) like Fourier spectra and response spectra, the final result is the mean value of application of equation(\ref{eq:gof-function}) on all frequencies (or periods).}

### 2018-04-19 results.tex
update

@@
 
-Some aspects of Fig.~\ref{fig:clusters} are worth discussing.
-Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case.
-In the C5-C9 and C7-C9 combinations in the 3-feature analysis, for instance, the excellent cluster has a limited presence.
+Some aspects of Fig.~\ref{fig:clusters} are worth discussing. Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case. In 
+% the C5-C9 and C7-C9 combinations in 
+\change{the combinations in}
+the 3-feature analysis, for instance, the excellent cluster has a limited presence.
 This is due to the influence of the lower C9 values (see Fig.~\ref{fig:data-box-plot}) on the correlations with C5 and C7.
@@
 % By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}. 
-\mychange{According to \citet{Weiss_2003_JAIR}, an oversample rate of $\times$10 is considered to be acceptable in this type of data processing.}
+\change{According to \citet{Weiss_2003_JAIR}, an oversample rate of $\times$10 is considered to be acceptable in this type of data processing.}
 (Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary.
-Instead, we used a \myremove{strong} \myrevision{heavy} pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.
+Instead, we used a \change{heavy} pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.
 
@@
 Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
-What the results we present here mean, in the context of GOF validation \myrevision{for engineering applications} , is that in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more decisive in \myrevision{narrowing eleven validation metrics into fewer numbers in order to get the same categorization as all eleven metrics}  \myremove{reaching an accurate conclusion about validation quality} than the duration (C11).
+What the results we present here mean \change{is that, in the context of this particular set of 11 metrics, in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more influential in the final result than the duration (C11).
+This, of course, is done under the assumption that one seeks to narrow the selection of metrics without loss of insight about the outcome of the validation process when compared to the outcome obtained with the whole suite of the eleven metrics used here.}
+
+% Original:
+% A low participation does not necessarily mean that a given metric is not important.
+Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
+What the results we present here mean, in the context of GOF validation, is that in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more decisive in reaching an accurate conclusion about validation quality than the duration (C11).
+
+% Naeem suggestion:
+% A low participation does not necessarily mean that a given metric is not important.
+Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
+What the results we present here mean, in the context of GOF validation for engineering applications, is that in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more decisive in narrowing eleven validation metrics into fewer numbers in order to get the same categorization as all eleven metrics than the duration (C11).
 
@@
 
-%\begin{table}[t]
-%	\centering
-%	\caption{Metrics data analysis participation in select trees (in percent)}
-%	\label{tab:participation}
-%	\small
-%	\begin{tabular}{lrrrr}
-%	Metric 	& T1 	& T2	& T3 	& T66 	\\ 
-%	\hline
-%	C1 		& --	& --	& -- 	& 1.5	\\
-%	C2 		& -- 	& --	& -- 	& 0.5	\\
-%	C3 		& -- 	& -- 	& -- 	& 9.6	\\
-%	\rowcolor{light-light-gray}
-%	C4 		& 46.8	& 52.4	& 60.2	& 100.0	\\
-%	\rowcolor{light-light-gray}
-%	C5 		& -- 	& 21.5	& 21.5	& 72.2	\\
-%	\rowcolor{light-light-gray}
-%	C6 		& -- 	& -- 	& 37.0	& 56.3	\\
-%	C7 		& -- 	& -- 	& -- 	& 19.4	\\
-%	\rowcolor{light-light-gray}
-%	C8 		& 100.0	& 100.0	& 100.0	& 100.0	\\
-%	C9 		& -- 	& -- 	& --	& 23.8	\\
-%	C10		& -- 	& -- 	& -- 	& 22.2	\\
-%	C11		& -- 	& -- 	& -- 	& -- 	\\
-%	\hline
-%	\end{tabular}
-%\end{table}
-
-\begin{table}[t]
-	\centering
-	\caption{Metrics data analysis participation in select trees (in percent). \myrevision{Stars represent the trees with same complexity in pruning process, however, not exact tree structures.}}
-	\label{tab:participation}
-	\small
-	\myrevision{
-	\begin{tabular}{lrrrrrrrrrrrr}
-	Metric 	& T1 	& T2	& T3 	& T66  & T1* 	& T2*	& T3* 	& T66*  & T1* 	& T2*	& T3* 	& T66* 	\\ 
-	\hline
-	&\multicolumn{4}{c}{Including C8} & \multicolumn{4}{c}{Without C8}&  \multicolumn{4}{c}{Without C5 and C8}\\
-	\hline
-	C1 		& --	& --	& -- 	& 1.5 & --	& --	& -- 	& --- & --	& --	& -- 	& --	\\
-	C2 		& -- 	& --	& -- 	& 0.5 & --	& --	& -- 	& 2.27 & --	& --	& -- 	& --	\\
-	C3 		& -- 	& -- 	& -- 	& 9.6 & --	& --	& -- 	& 53.22 & --	& 60.35	& 61.19 	& 99.11	\\
-	\rowcolor{light-light-gray}
-	C4 		& 46.8	& 52.4	& 60.2	& 100.0  & 44.35	& 81.77	& 81.77 	& 83.20 & 47.44	& 48.67	& 80.54 	& 54.64	\\
-	\rowcolor{light-light-gray}
-	C5 		& -- 	& 21.5	& 21.5	& 72.2 & 100.0	& 100.0	& 100.0 	& 100.0 & --	& --	& -- 	& --	\\
-	\rowcolor{light-light-gray}
-	C6 		& -- 	& -- 	& 37.0	& 56.3 & 58.65	& 63.51	& 62.77 	& 82.39 & 100.0	& 100	& 100 	& 100\\
-	C7 		& -- 	& -- 	& -- 	& 19.4 & --	& --	& -- 	& 32.37 & --	& --	& -- 	& 43.24	\\
-	\rowcolor{light-light-gray}
-	C8 		& 100.0	& 100.0	& 100.0	& 100.0 & --	& --	& -- 	& --	 & --	& --	& -- 	& --\\
-	C9 		& -- 	& -- 	& --	& 23.8 & --	& --	& -- 	& 48.63 & --	& --	& -- 	& 72.22	\\
-	C10		& -- 	& -- 	& -- 	& 22.2 & --	& --	& -- 	& 24.93 & --	& --	& -- 	& 20.46	\\
-	C11		& -- 	& -- 	& -- 	& -- 	    & --	& --	& -- 	& --- & --	& --	& -- 	& 8.09\\
-	\hline
-	\end{tabular}
-	}
-\end{table}
+\input{tab-partition}

### 2018-04-25 response.tex
updated

@@
 % }
-%
-% \comment{4}{%
-% In introduction, the difference between ``verification'' and ``validation'' should be discussed.
-% }
-%
-% \response{%
-% A discussion about verification and validation is added.
-% }
-%
+
+\comment{4}{%
+In introduction, the difference between ``verification'' and ``validation'' should be discussed.
+}
+
+\response{%
+A brief description of the differences between verification and validation was added to the Introduction.
+}
+
 % \comment{5}{%
@@
 
-% \comment{10}{%
-% Page 14, line 310: you say ``...the latter can essentially be discarded...'' I don't understand this statement, wouldn't this mean that both C1 and C7 are important to consider since they are independent of each other (one gives information that the other one doesn't)?
-% }
-%
-% \response{%
-% In that specific subspace (C1,C7), C1 does not help in clustering process.
+\comment{10}{%
+Page 14, line 310: you say ``...the latter can essentially be discarded...'' I don't understand this statement, wouldn't this mean that both C1 and C7 are important to consider since they are independent of each other (one gives information that the other one doesn't)?
+}
+
+\response{%
+In the specific case of the 2-feature clustering in the C1-C7 subspace, what the results show is that only C7 (in the vertical axis) is playing a role in establishing boundaries between the different classes.
+C1 does not help to do that.
+C1 may be important in other context, but here it does not add any value to the decision making process.
 That is why we call it irrelevant feature, because one can define the clusters based on only defining boundaries on C7.
-C1 can be important, however, it does not add any positive or negative values in decision making process in this subspace analysis.
-% }
-%
+It is normal to have a tendency to interpret this as a correlation problem, in which the interpretation of independence (i.e., lack of correlation between C1 and C7) would indicate that both are relevant.
+But that is not the goal of the clustering process.
+The goal is to identify which metrics play a role in delimiting the classes when correlated to other metrics.
+In other words, the colors and the way boundaries end up forming in these figures, is as important as the shape of the clouds of data-points.
+}
+
 % \comment{11}{%
@@
 
-% \response{%
-% That is correct.
-The statement convey part of the message.
-Although we expect to see similarities between the current results and previous studies, we keep in mind that this study is based on lack of consensus on using GOF metrics for validation purposes.
-Since the previous study is representing a uniform average of all metrics, we do not expect to match all stations.
-Therefore, we prefer to keep it as numerical.
-We modified the text accordingly.
-% }
+\response{%
+In hindsight, we may have allowed ourselves to be carried away by our own optimism.
+We have therefore modified our comments in this section to a more self-critical approach about the results.
+The question is really not whether they look the same or not, but whether they lead to similar---if not equivalent---conclusions.
+It is our interpretation that they do in the sense that looking at the results obtained with either approach one may walk away with the same conclusion about the simulation as a whole and about how well (or not well) did the simulations turned out with respect to specific areas in the simulation domain.
+We modified the text to convey this message, now hopefully more clearly than before.
+}
 

### 2018-04-25 introduction.tex
updated

@@
 
-Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation\change{, and a growing interest in the use of synthetic seismograms for engineering applications}.
-\myrevision{Validation is the process concerned with determining whether a conceptual model and the results it leads to are accurate representations of the physical system under study.
-This entails a comparison of simulation results with experimental data and observations.
-Verification, on the other hand, is concerned with the correctness and accuracy of a solution approach and its proper implementation with respect to the intended conceptual model, which involves a comparison of simulation results with exact results, if available, or with other alternative or approximate results \citep{Taborda_2013_BSSA}}.
+Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation\change{, as well as a growing interest in the use of synthetic seismograms for engineering applications}.
+\change{Validation entails comparisons of simulations with observations, whereas verification involves comparisons of simulations with a exact or alternative solutions \citep[see][and references therein]{Bielak_2010_GJI, Taborda_2013_BSSA}}.
 Various methods or schemes have been proposed to evaluate, through direct quantitative comparisons or indirect statistical analysis, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.

### 2018-04-25 testing.tex
updated

@@
-\section{Verification Test}
+% \section{Verification Test}
+\section{\change{Testing}}
 
-We now use tree T1 on the original simulation results from \citet{Taborda_2014_BSSA}, for different velocity models and components of motion.
+We now \change{test} tree T1 on the original simulation results from \citet{Taborda_2014_BSSA}, for different velocity models and components of motion.
 In this case we no longer aggregate all data from the simulation but now look at individual simulations done with different velocity models, and the three components of motion separately, as it would normally be done during any ground motion simulation validation.
 The results obtained with the T1 validation algorithm are shown in Fig.~\ref{fig:res-gof-maps}.
-These results are in essence very similar to those obtained by \citet{Taborda_2014_BSSA}, though using the T1 algorithm, we now rely only on the two metrics C4 and C8.
-
-A relative drawback from the T1 algorithm is that because its outcome are GOF validation classes (poor, fair, good, excellent) as opposed to GOF validation scores (with values from 0 to 10), then once one obtains the results for the validation process for different components of motion (as shown in the rows of Fig.~\ref{fig:res-gof-maps}) then there is no natural way of taking averages as one would do for scores in a numerical scale.
+\change{In an ideal case, we would expect these results to closely resemble those obtained by \citet{Taborda_2014_BSSA}, if grouped only into four categories.
+Relying only on the two metrics C4 and C8 using T1 is, however, not exactly the same.
+In addition, a}
+% 
+% These results are in essence very similar to those obtained by \citet{Taborda_2014_BSSA}, though using the T1 algorithm, we now rely only on the two metrics C4 and C8.
+% 
+% A relative 
+drawback from the T1 algorithm is that because its outcome are GOF validation classes (poor, fair, good, excellent) as opposed to GOF validation scores (with values from 0 to 10), then once one obtains the results for the validation process for different components of motion (as shown in the rows of Fig.~\ref{fig:res-gof-maps}) then there is no natural way of taking averages as one would do for scores in a numerical scale \change{\citep[as typically done when using][]{Anderson_2004_Proc}}.
 Therefore, in order to combine results from the three components into a single validation classification for each station, we define the following rules:
@@
 Applying them to the validation results shown in Fig.~\ref{fig:res-gof-maps} for the case of the simulation done using the model CVM-S is shown in Fig.~\ref{fig:avg-gof-maps}, which compares the outcome with the result of the average scores obtained using the traditional \citet{Anderson_2004_Proc}-type GOF method.
-This figure exemplifies the difference between the numeric GOF validation-scores scale and the proposed T1 GOF validation classes.
-At the same time, it highlights the similarities between the outcomes.
+This figure exemplifies the \change{differences or similarities} between the numeric GOF validation-scores scale and the proposed T1 GOF validation classes.
+% At the same time, it highlights the similarities between the outcomes.
 
@@
 In our view, the T1 results are simpler to assimilate and equally informative.
-Looking at both plots in Fig.~\ref{fig:avg-gof-maps} it is possible to argue they lead to \myremove{the same} \myrevision{similar} conclusions and present a similar picture of the validation of the simulation, not only globally but at specific locations, with only minor differences---especially if one is to consider the assumed equivalence between scores and classes (0--4: poor; 4--6: fair; 6--8:good; 8--10: excellent).
+Looking at both plots in Fig.~\ref{fig:avg-gof-maps} it is possible to argue they lead to \change{similar conclusions in respect to the overall validation of the simulation, and in respect to some particular areas and specific locations.
+Nonetheless, it must be said that we do not expect to see a 1-to-1 relationship.
+That is the scope of a future work, whereas here our focus was on identifying the relevance of the different metrics.
+This is important because the original approach proposed by \citet{Anderson_2004_Proc} favors a uniform weighing of metrics, and the results shown here indicate this may not be the preferred strategy.}
+
+% conclusions and present a similar picture of the validation of the simulation, not only globally but at specific locations, with only minor differences---especially if one is to consider the assumed equivalence between scores and classes (0--4: poor; 4--6: fair; 6--8:good; 8--10: excellent).
 \myrevision{However, we do not expect to see the same results in both figures.

### 2018-04-25 response.tex
update

@@
 
-% \comment{3}{%
-% Lines 25--41: The frequent use of phrases such as ``Among these methods,''; ``Among the latter,''; ``In this category,''; and ``Within this group,'' make the text difficult to follow.
-% }
-%
+\comment{3}{%
+Lines 25--41: The frequent use of phrases such as ``Among these methods,''; ``Among the latter,''; ``In this category,''; and ``Within this group,'' make the text difficult to follow.
+}
+
 % \response{
@@
 
-% \comment{13}{%
-% Lines 259--264: This paragraph contains a lot of jargon.
-% }
-%
-% \response{
-% Thanks for the comment. Additional clarification added.
-% }
+\comment{13}{%
+Lines 259--264: This paragraph contains a lot of jargon.
+}
+
+\response{
+Indeed. We expanded the paragraph to include an explanation about the options used with the C5.0 algorithm.
+}
 
@@
 
-% \comment{20}{%
-% Line 455: ``...as the decisive \textbf{parameters}...'' ?
-(This short concluding paragraph would be much stronger without the use of unnecessary phrases, i.e., ``This latter point'', ``One the one hand'',
-% ``on the other hand.''
-% }
-%
-% \response{
-% \color{red}{Ricardo, please address this comment.}
-% }
+\comment{20}{%
+Line 455: ``...as the decisive \textbf{parameters}...'' ?
+(This short concluding paragraph would be much stronger without the use of unnecessary phrases, i.e., ``This latter point'', ``One the one hand'', ``on the other hand.''
+}
+
+\response{
+Indeed. We modified the last paragraph of the Conclusions by removing those sentences.
+}
 

### 2018-04-25 conclusions.tex
update

@@
 
-This latter point is perhaps the most important conclusion of our study.
+\change{Identifying the total energy and response spectra metrics as decisive metrics in the comparisons between synthetics and observations is a key contribution to validation of ground motion simulations.
+This contributes to clarifying a standing question in the area of validation, and it provides an indication to simulators about where the focus in the search for improvements in their models.}
+
+% This latter point is perhaps the most important conclusion of our study.
 Identifying the total energy and response spectra metrics as the decisive ones is a key contribution to validation of ground motion simulations.

### 2018-04-25 decision-tree.tex
update

@@
 
-In essence, given a dataset, the process of building the tree consists of recursively identifying the attributes in the training data that are most likely to predict an outcome.
+In \change{general}, given a dataset, the process of building the tree consists of recursively identifying the attributes in the training data that are most likely to predict an outcome.
 The process is recursive because new branches and decision nodes are created based on the remaining data at every branch and level, until the tree reaches a certain depth or when other conditions are met.
@@
 
-In the particular case of the C5.0 algorithm the result of this process depends on several parameters, logical and numerical.
+\change{
+In the particular case of the C5.0 algorithm, in addition to the general steps just described, the result of the process of building a decision tree also depends on other logical and numerical parameters.
+We set the options to: (a) perform feature selection or winnowing, (b) evaluate possible advanced splits of data, \textcolor{red}{(c) discard the use of boosted models,} and (d) use a confidence factor and a threshold in the number of samples that go in th tree leaves.
+The feature selection option is used by the algorithm to choose the most important attribute over others.
+The option of evaluating advanced splits prevents the use of a hard threshold during the classification by considering different probabilities in assigning data to different classes.
+Discarding the use of boosted models prevents the algorithm from...
+\textcolor{red}{[pending for discussion with Naeem]}.
+Last, the options regarding the use of a confidence factor and a minimum number of samples ($S_{\min}$, also known as \textit{minCases}) that can be put in at least two of the branches are set to limit the level of complexity of the tree.
+In essence, the C5.0 algorithm grows a tree with many branches and then goes through a pruning process to satisfy a minimum number of samples in each leaf and minimize the error of misclassifying the data, which is measure by the confidence factor (CF).
+Additional details about these options are available in \citet{Quinlan_1993_Book} and \citet{Kuhn_2017_Manual}.
+For this study, in order to obtain a variety of tree options, we build trees for varying values of CF and $S_{\min}$.
+}
+
+% SUGGESTED BY NAEEM:
+% In the particular case of the C5.0 algorithm the result of this process depends on several parameters, logical and numerical.
 We set the options to do feature selection \myrevision{or winnowing.
 Therefore, the algorithm choose to use the most important attributes and ignore others.
-We also activate the evaluation of possible advanced splits of data.
+% We also activate the evaluation of possible advanced splits of data.
 Sometimes data is very close to the splitting threshold where assigning a hard threshold may incorrectly change the classification, whereas, that small difference could happen because of numerical errors or error in measurement.
 Therefore, advanced split can define a soft threshold to consider different probabilities in assigning data to different classes.
-A decision tree can grow so deep to finally classify all training data correctly.
+% A decision tree can grow so deep to finally classify all training data correctly.
 However, the tree will be very complex and will suffer from overfitting.
 In other words it will have very poor functionality for test data.
-Therefore, decision tree algorithms (here C5.0) first grow the tree up to the most complex tree then based on user defined factors start to prune the grown tree.The pruning process helps to reduce the complexity of the tree and also increase it's generalizability.
+Therefore, decision tree algorithms (here C5.0) first grow the tree up to the most complex tree then based on user defined factors start to prune the grown tree.
+The pruning process helps to reduce the complexity of the tree and also increase it's generalizability.
 Consequently the pruned tree's leaf nodes will have training data which belongs to more than one class which the final class will be the mode of that leaf node.
@@
 Therefore, error estimates for leaf nodes and subtrees are computed assuming that they were used to classify a set of unseen cases with a predicted error rate which is computed based on CF and binomial distribution of number of misclassified events (E) and all training data at that leaf (N).
-Decreasing CF and increasing $S_{\min}$ cause more heavy pruning process \citep[see][]{Kuhn_2017_Manual}.} \myremove{and evaluate possible advanced splits of data, but do not use the boosted models option \citep[see][]{Kuhn_2017_Manual}; and use the confidence factor (CF) and the smallest number of samples that must be placed in at least two of the splits ($S_{\min}$, also known as \textit{minCases}) to fine tune the prediction models.} In particular, we build a suite of trees for varying values of CF and $S_{\min}$.
+Decreasing CF and increasing $S_{\min}$ cause more heavy pruning process \citep[see][]{Kuhn_2017_Manual}.} In particular, we build a suite of trees for varying values of CF and $S_{\min}$.
+
+% ORIGINAL:
+% In the particular case of the C5.0 algorithm the result of this process depends on several parameters, logical and numerical.
+We set the options to do feature selection and evaluate possible advanced splits of data, but do not use the boosted models option \citep[see][]{Kuhn_2017_Manual}; and use the confidence factor (CF) and the smallest number of samples that must be placed in at least two of the splits ($S_{\min}$, also known as \textit{minCases}) to fine tune the prediction models.
+In particular, we build a suite of trees for varying values of CF and $S_{\min}$.
 

### 2018-04-25 introduction.tex
update

@@
 Various methods or schemes have been proposed to evaluate, through direct quantitative comparisons or indirect statistical analysis, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
-Among these methods, there are some that are better suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for validation.
+\change{Some of these methods are better} suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for validation.
 Among the latter, some are better for direct application purposes because they compare the behavior of systems (e.g., buildings or earth structures) at the response level \citep{Burks_BSSA_2014}, while the remaining methods focus on the validation of the overall characteristics of seismograms.

### 2018-04-26 response.tex
close to finishing

@@
 
-\response{
+\response{%
 We modified the abstract and the conclusions accordingly.
@@
 
-% \introcomment{~}{%
-% In addition, I also agree that all the terms and methodology should be well defined for the BSSA readership, so that the paper can be understood by them without resort to frequent online searches.
-% }
-%
-% \introcomment{~}{%
-% I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit.
+\introcomment{~}{%
+In addition, I also agree that all the terms and methodology should be well defined for the BSSA readership, so that the paper can be understood by them without resort to frequent online searches.
+}
+
+\response{%
+\textcolor{red}{[ *** Pending ***]}
+}
+
+\introcomment{~}{%
+I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit.
 It clearly shows that only these metrics are change significantly with the goodness of fit (P, F, G, E).
@@
 In any case, more
-% discussion of this figure might help the authors make the case for the main conclusions.
-% }
-%
-% \response{%
+discussion of this figure might help the authors make the case for the main conclusions.
+}
+
+\response{%
+\textcolor{red}{[ *** Pending ***]}
 % Thanks for the constructive suggestions.
@@
 % We added text to abstract and conclusion to give a heads up to readers about 11 metrics and one earthquake with 3 velocity model  simulations.
-% }
+}
 
@@
 
-% \comment{1}{
-% The terminologies used in the paper are not clear in many cases.
+\comment{1}{%
+The terminologies used in the paper are not clear in many cases.
 For example, what are ``features'' ``attributes'' ``dimensions'' ``trees, nodes and leafs'' ``class''?
@@
 I suggest that the authors streamline the paper and the terminologies to be consistent and clear in the final version.
-% }
-%
-% \response{%
+}
+
+\response{%
+\textcolor{red}{[ *** Pending ***]}
 % We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
@@
 We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.
-% }
+}
 
@@
 
-% \comment{3}{%
-% In abstract, ``semi-supervised'' and ``supervised'' learning techniques are mentioned, but in the paper these terms are not explained, what part of the learning techniques is semi-supervised and what does that mean?
-% }
-%
-% \response{%
+\comment{3}{%
+In abstract, ``semi-supervised'' and ``supervised'' learning techniques are mentioned, but in the paper these terms are not explained, what part of the learning techniques is semi-supervised and what does that mean?
+}
+
+\response{%
+\textcolor{red}{[ *** Pending ***]}
 % We briefly explained these terms \textbf{Data Analysis Method} section. We also added clarification about semi-supervised and supervised learning into \textbf{Introduction} section.
-% }
+}
 
@@
 
-% \comment{5}{%
-% In abstract and introduction, what does a ``direct application'' mean?
+\comment{5}{%
+In abstract and introduction, what does a ``direct application'' mean?
 it's not clear.
 Similarly, in Validation Metrics section: ``direct quantitative comparisons'' and ``indirect statistical analysis''.
-% }
-%
-% \response{%
-% not present in the abstract, present in the introduction, I will come back to this.
-% \color{red}{Ricardo, please address this comment.}
-% }
+}
+
+\response{%
+These terms were not present in the Abstract.
+The term ``direct application'' was present in the Introduction, but it has been removed due to improvements in the text based on other suggestion by this and a second Reviewer.
+The terms ``direct quantitative comparisons'' and ``indirect statistical analysis'' were present in the Introduction and the Validation Metrics sections.
+In our view, the first of these terms makes reference to validation through direct, i.e., signal to signal, quantitative comparisons of metrics; while the second of these terms makes reference to validation by means of statistical analysis showcasing the overall results of comparisons, and thus the use of ``indirect''.
+We understand, however, that this may not be a view shared by any other reader, so we have modified the text to be more explicit about this distinction.
+}
 
@@
 
-% \comment{8}{%
-% You are treating these metrics as ``independent variables'', but as you have also noted many of them are not.
+\comment{8}{%
+You are treating these metrics as ``independent variables'', but as you have also noted many of them are not.
 For example, Arias intensity and energy are correlated.
@@
 If not explain why in the Results section.
-% }
+}
 %
-% \response{%
+\response{%
+\textcolor{red}{[ *** Pending ***]}
 % This is a very interesting comment.
@@
 However, in this study, we only deal with 11 metrics and the results show that it works fairly well.
-% }
+}
 
@@
 
-% \comment{11}{%
-% Page 14, line 323: It's not clear how you ``replicate data'', provide more detail.
-% }
-%
-% \response{%
-% Thanks. More details provided.
-% }
-%
-% \comment{12}{%
-% I don't think ``confusion matrix'' given in tables is explained in the paper!
-% }
-%
-% \response{%
-% In \textbf{Decision Trees} section we provided a definition of confusion matrix for the simplest case of two categories. The idea can be extended for more than two (in our case four) categories.
-% }
+\comment{11}{%
+Page 14, line 323: It's not clear how you ``replicate data'', provide more detail.
+}
+
+\response{%
+We modified this paragraph to provide more details about the oversampling process.
+}
+
+\comment{12}{%
+I don't think ``confusion matrix'' given in tables is explained in the paper!
+}
+
+\response{%
+We are a little surpirsed by this comment.
+The paper does have an explanation about confusion matrices.
+In the original manuscript, this explanation is the section titled \textbf{Decision Trees}, page 13, between lines 280 and 282.
+The explanation includes an illustrative example for the simplest case of two categories and it should not be difficult for any reader of the BSSA to understand this concept is extendable to larger matrics such as those shown in Tables 2 and 3.
+We added a sentence in the explanation to make it explicit that they can be extended to deal with more attributes.
+}
 
@@
 
-% \comment{15}{%
-% Figure 8: I don't know what you mean by y-label ``decision attributes''. Do you mean ``the number of GOF metrics''? why there is no data point for 1 or 11?
-% }
+\comment{15}{%
+Figure 8: I don't know what you mean by y-label ``decision attributes''. Do you mean ``the number of GOF metrics''? why there is no data point for 1 or 11?
+}
 
-% \response{% *** needs attention and modification in text too ***
-% That is right.
+\response{%
+That is right.
 It should be the number of decision attributes.
 We modified the figure accordingly.
-The reason why there is no data for 1 is because there are no trees that would classify the data using a single metric into the four classes because each node is a binary decision, so one needs at least 2 metrics with 3 or more nodes with a minimum depth of 2.
-And the reason why there is no data for 11 metrics is because when using all metrics, C11 (duration) never had a participation---as found by the algorithm.
-The new Table 6 shows that if we forcefully eliminate
-%
-% ...
-%
-% }
+The reason why there is no data for 1 is because there are no trees that would classify the data using a single metric into the four classes because these are binary trees, i.e., each node represents a binary decision, so one needs at least 2 metrics with 3 or more nodes with a minimum depth of 2.
+On the other end of things, the reason why there is no data for 11 metrics is because the analysis of the data resulted in the metric C11 not having any participation in the decision process, as shown in the original Table 6.
+In new version of Table 6 we include cases in which we forcefully removed metric C8, first, and then C5 and C8.
+Once one does that, the participations of the metrics changes and so would the topology of the trees.
+}
 
-% \comment{16}{%
-% The conclusion basically says that response spectrum is the most important validation metric.
+\comment{16}{%
+The conclusion basically says that response spectrum is the most important validation metric.
 This is not surprising to any engineer or modeler.
@@
 Or would it be influenced by removal of Sa as a parameter (this is possible since there are correlations between Sa and other metrics).
-% }
+}
 
-% \response{%
+\response{%
+\textcolor{red}{[ *** Pending ***]}
 % That is a very good suggestion.
@@
 C11 also become somewhat effective in the most complicated tree.
-% }
+}
 
@@
 \introcomment{~}{%
-Review of Khoshnevis and Taborda for BSSA
-}
-
-% \introcomment{~}{%
-% Earthquake ground motion simulations are usually validated against observational data or empirically-derived solutions.
+Earthquake ground motion simulations are usually validated against observational data or empirically-derived solutions.
 However, exactly which parameters to compare (e.g., peak ground velocity, duration, spectral content) in order to assess the goodness-of-fit, is not always clear.
@@
 The paper is generally well written.
+}
+
+\response{%
+We thank the reviewer for the generally positive view of our paper.
+}
+
+\introcomment{~}{%
 As a strong ground motion seismologist (i.e., I assume I'm part of the audience you hope to reach), there is a lot of machine learning-related jargon that makes certain sections difficult to follow.
-It would also be useful to have some discussion regarding which parameters were found to be repetitive, incase future authors wish to look at more than the four recommended parameters.
-% }
-%
-% \response{
+}
+
+\response{%
+\textcolor{red}{[ *** Pending ***]}
 % We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
@@
 We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.
-% \vspace{2ex}
-% \newline
+}
+
+\introcomment{~}{%
+It would also be useful to have some discussion regarding which parameters were found to be repetitive, in case future authors wish to look at more than the four recommended parameters.
+}
+
+\response{%
+\textcolor{red}{[ *** Pending ***]}
 % We also studied the models without response spectra (C8) and without response spectra (C8) and peak ground acceleration (C5).
-% }
+}
 
@@
 
-\response{
+\response{%
 Done. Thanks.
@@
 
-\response{
+\response{%
 Added. Thanks.
@@
 
-% \response{
-% \color{red}{Ricardo, please address this comment.}
-% }
+\response{%
+We modified the second half of the first paragraph in the Introduction to be more direct in the description of the alternative verification and validation metrics.
+}
 
@@
 
-\response{
+\response{%
 Thanks for the suggestion. We moved it and it does read better this way.
@@
 
-\response{While it is true that the initial intent of Anderson (2004) was to develop credibility of synthetic seismograms for engineering applications, we do not share the need to make that intent explicit here.
+\response{%
+While it is true that the initial intent of Anderson (2004) was to develop credibility of synthetic seismograms for engineering applications, we do not share the need to make that intent explicit here.
 On the one hand, these metrics have already been used by several authors who did not intend to validate seismograms for any engineering application, but to simply show---by means of them---that their synthetics were realistic in a general sense.
@@
 
-\response{
+\response{%
 We disagree.
@@
 
-\response{
+\response{%
 Yes.
@@
 
-\response{
+\response{%
 Fixed. Thanks.
@@
 
-\response{
+\response{%
 We added it at this point. Thank you.
@@
 
-\response{
+\response{%
 We removed it, and in its place added an explanation about overfitting in response to the previous comment.
@@
 
-% \comment{11}{%
-% Lines 241--243: It is not helpful to list different algorithm names for making decision trees, without any information about them or how they differ.
-% }
-%
-% \response{
-% We added a brief description about them. Thanks.
-% }
+\comment{11}{%
+Lines 241--243: It is not helpful to list different algorithm names for making decision trees, without any information about them or how they differ.
+}
+
+\response{%
+We disagree about the need to make detailed distinctions about these different algorithms.
+That is why we provided references.
+If the reader is interested in learning the differences, the most we can do without expanding the scope of our study is to point him/her in the right direction, which we already do.
+Nonetheless, in attention to the Reviewer, we modified slightly the descritpion to give some additional context in what regards to the C5.0 algorithm and added a sentence explaining that detailing the alternative algorithms is out of the scope of this study.
+}
 
@@
 
-\response{
+\response{%
 Following Equation (3) we define $A[a]$ as all the possible values of attribute $A$.
@@
 
-\response{
+\response{%
 Indeed. We expanded the paragraph to include an explanation about the options used with the C5.0 algorithm.
@@
 
-\response{
+\response{%
 That is right. We modified the sentence to reflect that.
@@
 
-\response{
+\response{%
 We modified it to ``heavy pruning process'' and added an explanation earlier in the Decision Trees section.
@@
 
-\response{
+\response{%
 Indeed. We modified them accordingly. Thanks.
@@
 
-\response{
+\response{%
 We eliminated ``Similarly''. Thanks.
@@
 
-\response{
+\response{%
 The extra ``s'' was deleted. Thanks.
@@
 
-\response{
+\response{%
 Indeed. We modified the last paragraph of the Conclusions by removing those sentences.

### 2018-04-26 decision-tree.tex
close to finishing

@@
 Upon completing the clustering and resampling processes, the next step is to subdivide the dataset in two, a training dataset (with 70 percent of the samples picked randomly), and a testing dataset (with the remaining 30 percent).
-We first use the training dataset to build a large suite of potential decision trees, and then use the testing dataset to evaluate and pick the best possible decision tree(s) among them.
+We first use the training dataset to build a large suite of potential decision trees, and then use the testing dataset to evaluate and pick the best possible decision tree(s).
 There are several algorithms for building decision trees \citep[e.g., ID3, C4.5, C5.0, CART; see][]{Quinlan_1986_ML, Quinlan_1993_Book, Quinlan_1996_JAIR, Breiman_1984_Book}.
-\myrevision{C5.0 is considered as the developed version of ID3 and C4.5, where it does not have any limitations in handling missing data, numerical attributes, and developing boosted models.
-Internal algorithm of CART is different than other methods, however, discussing technical details of these algorithms are beyond the scope of this study.} Here we use the C5.0 implementation \citep{Kuhn_2017_Manual}, which is an extension of the C4.5 algorithm developed by \citet{Quinlan_1993_Book, Quinlan_1996_JAIR}.
+\change{Here, we use the C5.0 algorithm as implemented by \citet{Kuhn_2017_Manual}.
+This algorithm is the latest update to the original ID3 and subsequent C4.5 algorithms \citep{Quinlan_1993_Book, Quinlan_1996_JAIR}.
+C5.0 is superior to its predecessors because it reduces the limitations for handling numerical attributes, missing data, and it has additional features such as the development of boosted models.
+Discussing the differences between these algorithms is out of the scope of this study.}
+
+% \myrevision{C5.0 is considered as the developed version of ID3 and C4.5, where it does not have any limitations in handling missing data, numerical attributes, and developing boosted models.
+Internal algorithm of CART is different than other methods, however, discussing technical details of these algorithms are beyond the scope of this study.}
+
+% Here we use the C5.0 implementation \citep{Kuhn_2017_Manual}, which is an extension of the C4.5 algorithm developed by \citet{Quinlan_1993_Book, Quinlan_1996_JAIR}.
 
@@
 where TN is the number of true-negative samples.
+\change{These confusion matrices can be larger depending on the number of attributes.
+In our case, the confusion matrices are size 4$\times$4, in order to compare the actual number of samples with the predicted ones for the four given categories (poor, fair, good and excellent).}
 

### 2018-04-26 introduction.tex
close to finishing

@@
 Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation\change{, as well as a growing interest in the use of synthetic seismograms for engineering applications}.
-\change{Validation entails comparisons of simulations with observations, whereas verification involves comparisons of simulations with a exact or alternative solutions \citep[see][and references therein]{Bielak_2010_GJI, Taborda_2013_BSSA}}.
-Various methods or schemes have been proposed to evaluate, through direct quantitative comparisons or indirect statistical analysis, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
+\change{Validation entails the comparison of simulations with observations, whereas verification involves the comparison of simulations with a exact or alternative solutions \citep[see][and references therein]{Bielak_2010_GJI, Taborda_2013_BSSA}}.
+Various methods or schemes have been proposed to evaluate, through direct \change{signal-to-signal} quantitative comparisons or \change{overall} statistical analysis, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
 \change{Some of these methods are better} suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for validation.
-Among the latter, some are better for direct application purposes because they compare the behavior of systems (e.g., buildings or earth structures) at the response level \citep{Burks_BSSA_2014}, while the remaining methods focus on the validation of the overall characteristics of seismograms.
+\change{\citet{Burks_BSSA_2014} and \citet{Rezaeian_2015_BSSA}, for instance, are used for validation in the context of engineering applications, i.e., at the level of the dynamic response of buildings.
+\citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL}, on the other hand, are used for validation in the context of the overall characteristics of seismograms when comparing synthetics with data in both the time and frequency domains.
+These methods are preferred for validating high-frequency or broadband simulations, where applications may be undefined and for which the expectation of matching seismograms at the waveform level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}), and because they offer simple schemes to quantify the goodness-of-fit (GOF) of simulation results.}
+
+% Among the latter, some are better for direct application purposes because they compare the behavior of systems (e.g., buildings or earth structures) at the response level \citep{Burks_BSSA_2014}, while the remaining methods focus on the validation of the overall characteristics of seismograms.
 In this category, the comparisons are done both in time and frequency.

### 2018-04-26 metrics.tex
close to finishing

@@
 
-There are various methods and algorithms available to evaluate the similarity between two or more seismograms, both through direct quantitative comparisons or indirect statistical analysis \citep[e.g.,][]{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
+There are various methods and algorithms available to evaluate the similarity between two or more seismograms, both through direct \chage{signal-to-signal} quantitative comparisons or \change{overall} statistical analysis \citep[e.g.,][]{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
 In earthquake ground motion simulation, they are primarily used for verification with respect to benchmark or analytical solutions, and for validation with respect to data (i.e., ground motion records).

### 2018-04-26 results.tex
close to finishing

@@
 As it can be seen in this figure, the number of samples in the excellent class is significantly less than those in the poor, fair, and good classes.
-Therefore, before moving on with the decision tree analysis, it is necessary to resample the subset of the excellent class.
-We used the oversampling approach described in the previous section.
-We replicated data from the excellent class
-\myrevision{through random selection of data from that class until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.
+Therefore, before moving on with the decision tree analysis, it is necessary to resample the subset of the excellent class\change{, for which we use the oversampling approach described in the previous section.
+Oversampling is nothing else but a replication of data samples.
+This process is done randomly, i.e., through a random selection of data samples from the original set to be duplicated until one increases the number of total samples in the oversampled set to a desired target number.
+In the case of the excellent class, we applied oversampling until reaching a total of 760 samples, as indicated with the dashed line in Fig.~\ref{fig:count-classes}.
+Since the original size was 76 samples, that means we applied an oversampling ratio of 10$\times$.
+According to \citet{Weiss_2003_JAIR}, an oversample rate of 10$\times$ is considered to be acceptable in this type of data processing.} (Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary.
+Instead, we used a \change{heavy} pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.
+
+% We used the oversampling approach described in the previous section. We replicated data from the excellent class 
+% \myrevision{through random selection of data from that class until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.
 This oversampling is considered as a $\times$10 factor.}
-\myremove{randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.} 
+% \myremove{randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.} 
 % By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}.
-\change{According to \citet{Weiss_2003_JAIR}, an oversample rate of $\times$10 is considered to be acceptable in this type of data processing.}
-(Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary.
-Instead, we used a \change{heavy} pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.
 
@@
 For each one of these unique trees, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}), and extracted the total number of nodes in the trees and their depth.
-Figure \ref{fig:nodes-depth} shows the results of $F_1$ for all the trees and its distribution in terms of the number of attributes used in the trees as a function of the number of nodes and the depth of each tree.
+Figure \ref{fig:nodes-depth} shows the results of $F_1$ for all the trees and its distribution in terms of the number of \change{decision} attributes \change{or GOF metrics} used in the trees as a function of the number of nodes and the depth of each tree.
 Recall that we are interested in finding a sequence of decisions (represented by disjunctive decision nodes in a tree) that can lead to good GOF predictions (i.e, high values of $F_1$) using a reduced number of attributes (GOF metrics).

### 2018-04-26 response.tex
making progress

@@
 \textcolor{red}{[ *** Pending ***]}
-}
-
-\introcomment{~}{%
-I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit.
-It clearly shows that only these metrics are change significantly with the goodness of fit (P, F, G, E).
-Plotting these values as a function of P, F, G, and E together on the same figure might allow the reader to see this relationship more clearly and might help provide a heuristic understanding of how C4, C5, C6, and C8 are diagnostic whereas C11 is constant with P, F, G, and E, and thus not diagnostic.
-In fact, in my opinion, Figure 6 is a simple and easy way of determining which metrics are diagnostic that might even eliminate the need for the decision trees.
-In any case, more
-discussion of this figure might help the authors make the case for the main conclusions.
-}
-
-\response{%
-\textcolor{red}{[ *** Pending ***]}
 % Thanks for the constructive suggestions.
@@
 % \newline
-% We also modified Figure 6 in order to easily track the variation of different metrics for final classes.
-We do agree with you that Figure 6 clearly gives a good understanding about the metrics, however, we add decision trees, because they can convert the qualitative judgment to quantitative conditional form to be able to exactly decide on the final class based on their numerical GOF scores.
-% \vspace{2ex}
-% \newline
 % We added text to abstract and conclusion to give a heads up to readers about 11 metrics and one earthquake with 3 velocity model  simulations.
+}
+
+\introcomment{~}{%
+I found Figure 6 very helpful in giving me a very simple if heuristic understanding why C4, C5, C6, and C8 are the most diagnostic of the goodness of fit.
+It clearly shows that only these metrics are change significantly with the goodness of fit (P, F, G, E).
+Plotting these values as a function of P, F, G, and E together on the same figure might allow the reader to see this relationship more clearly and might help provide a heuristic understanding of how C4, C5, C6, and C8 are diagnostic whereas C11 is constant with P, F, G, and E, and thus not diagnostic.
+In fact, in my opinion, Figure 6 is a simple and easy way of determining which metrics are diagnostic that might even eliminate the need for the decision trees.
+In any case, more discussion of this figure might help the authors make the case for the main conclusions.
+}
+
+\response{%
+Thank you.
+This is very true.
+Figure 6 certainly conveys a clear message about how the metrics vary for each validation classes.
+The decision tree analysis, nonetheless, is still needed in order to convert this qualitative insight into a set of quantitative conditionals to decide the final class based on the GOF scores.
+In attention to this comment, we modified Fig.~6 to highlight the boundaries of the categories defined by Anderson (2004) and expanded the text in the Results section to provide a better explanation of the meaning of the results shown in this figure.
 }

### 2018-04-26 dataset.tex
making progress

@@
     \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component\myrevision{s} of motion (NS, EW, UD).
-In each case, the median is indicated by a notch in the box, the boxes represents the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), and the line shows the range of the data, with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots.
+In each case, \change{the boxes represent the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), the medians are indicated by a notch in the boxes, and the vertical lines show} the range of the data, with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots.
 The color version of this figure is available only in the electronic edition.}

### 2018-04-26 metrics.tex
making progress

--- 2018-04-19 metrics.tex
@@
 
-There are various methods and algorithms available to evaluate the similarity between two or more seismograms, both through direct quantitative comparisons or indirect statistical analysis \citep[e.g.,][]{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
+There are various methods and algorithms available to evaluate the similarity between two or more seismograms, both through direct \change{signal-to-signal} quantitative comparisons or \change{overall} statistical analysis \citep[e.g.,][]{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
 In earthquake ground motion simulation, they are primarily used for verification with respect to benchmark or analytical solutions, and for validation with respect to data (i.e., ground motion records).

### 2018-04-26 results.tex
making progress

@@
 Separately, we prepared similar plots to look at the influence of the velocity models and components of motions on the results of the clustering process and, as observed before, they had no significant differences with the aggregate of all the samples in the dataset.
+\change{Figure 6 is particularly important because it highlights the outcome of the clustering process and provides a preview into the decision tree analysis results.
+Notice that metrics C5 through C8 consistently fall within the poor, fair, good, and excellent classification and that metrics C3 and C4 also show a progressive variation in sync with the validation classes.
+This means that these metrics are likely the best predictors for the outcome of the validation process, as we will see upon performing the decision tree analysis.
+On the other hand, metrics C1 and C2, and C9 through C11 are almost invariant, therefore less or not decisive in the validation process.
+C11, in particular, shows a broader (larger boxes) distribution, which indicates that it is less effective in predicting the final validation class.}
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{figures/pdf/figure-06_mod}
+	\includegraphics[width=\textwidth]{figures/pdf/figure-06}
 	\caption{Statistical distribution of the dataset as partitioned into the four validation categories (poor, fair, good, excellent) after the clustering analysis.
 The distributions are shown in the form of box-plots for each metric (C1 through C11).
-In each case, the median is indicated by a notch in the box of the central quartiles, and the lines represent the interquartile range, with outliers shown as scattered dots.
-The color version of this figure is available only in the electronic edition.
-\myrevision{Dashed lines represent the boundaries of poor, fair, good, and excellent categories according to  \citeauthor{Anderson_2004_Proc}'s method.}}
+In each case, \change{the medians are indicated by a notch in the boxes of the central quartiles, and the vertical} lines represent the interquartile range, with outliers shown as scattered dots.
+\change{The dashed lines and background shadows represent the boundaries of the poor, fair, good, and excellent categories as defined by \citet{Anderson_2004_Proc}.} The color version of this figure is available only in the electronic edition.}
 	\label{fig:boxed-clusters}
@@
 \input{tab-partition}
+
+\input{tab-partition-extra}

### 2018-04-27 response.tex
updated

@@
 \response{%
-\textcolor{red}{[ *** Pending ***]}
-% This is a very interesting comment.
-You are totally right.
-There are fairly high correlation between Arias intensity and energy integral.
-However, decision tree algorithm is robust enough to realize this similarity and pick the best one in each decision making process.
-As we remove the peak ground acceleration and response spectra from the prediction model, the algorithm looks for other best parameters.
-In case of hundreds of metrics it would be a reasonable choice to do feature selection analysis and choose one metric out of several correlated metrics.
-However, in this study, we only deal with 11 metrics and the results show that it works fairly well.
+This is a very valid point.
+However, an important aspect to keep in mind is that the clustering process and the subsequent construction of the decision is not the same as a correlation analysis, and it is therefore not compromised by having highly correlated metrics.
+On the contrary, the process naturally selects the metrics with the best chance of predicting the outcome.
+In the case of two metrics being highly correlated, the process would still analyze which of the two is a better predictor.
+In some cases it may decide to keep both, and in some cases it may decide that one is enough.
+In attention to this and other related comments we decided to include an additional table (Table 7) showing how the level of participation of the different metrics would change if one removes some of the initially most relevant metrics.
+This should help the readers better understand the results of our work.
 }
@@
 \response{%
-We are a little surpirsed by this comment.
-The paper does have an explanation about confusion matrices.
-In the original manuscript, this explanation is the section titled \textbf{Decision Trees}, page 13, between lines 280 and 282.
-The explanation includes an illustrative example for the simplest case of two categories and it should not be difficult for any reader of the BSSA to understand this concept is extendable to larger matrics such as those shown in Tables 2 and 3.
+The explanation about confusion matrices in the original manuscript was in the section titled \textbf{Decision Trees}, page 13, between lines 280 and 282.
+The explanation included an illustrative example for the simplest case of two categories.
+We assumed it would be clear that this concept is extendable to larger matrics such as those shown in Tables 2 and 3.
 We added a sentence in the explanation to make it explicit that they can be extended to deal with more attributes.
@@
 \response{%
-\textcolor{red}{[ *** Pending ***]}
-% That is a very good suggestion.
-We excluded C8(response spectra) and C5(peak acceleration) from analysis and reproduce similar trees with same complexities in terms of pruning process with previous four trees.
-As we expected, most important, metrics are still C4 and C6.
-C11 also become somewhat effective in the most complicated tree.
+We also mention in the conclusion that the next metrics in line are the energy (C4), the peak acceleration (C5) and peak velocity (C6).
+In order to emphasize this point, as described in the response to comment \#8, we also included a new table (Table 7) showing the participation of the different metrics in the absence of the response spectra (C8) and the peak acceleration (C5).
+The results confirm that in such cases the scond most relevant metrics are still the energy (C4) and the peak ground velocity (C6) as mentioned before.
+This additional analysis also confirms the very low or null participation of the duration (C11) in other trees.
 }
@@
 \response{%
-\textcolor{red}{[ *** Pending ***]}
-% We also studied the models without response spectra (C8) and without response spectra (C8) and peak ground acceleration (C5).
+We addressed this by considering additional trees in the cases of removing metrics C5 and C8 (see Table 7).
+This was also done in response to the comments \#8 and \#16 from Reviewer 1.
+As we explain in the response to those comments, the procedure used here is different from a simpler correlation analisis in which a metric can be directly related by other if ``repetitive''.
+Here, the algorithm searches for the potential of each metric to serve as a predictor.
+In some cases, one may need two well correlated metrics to aid in that decision process, while not in others.
+By including the results shown in the new Table 7, instead of trying to understand these correlations, we can see what metrics become relevant in the absence of others.
+The results indicate that in addition to the energy and the response spectra, the next in line are the peak response in acceleration, velocity, and the Arias intensity.
+They also confirm the lack of relevance of the duration (C11), the time integrals (C1) and (C2), and the cross correlation (C10).
 }
@@
 \response{%
-We disagree.
 In our view, these sentences are written as indirect questions and do not require the use of a question mark.

### 2018-04-27 dataset.tex
updated

@@
     \includegraphics[width=0.48\textwidth]{figures/pdf/figure-03}
-    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and component\myrevision{s} of motion (NS, EW, UD).
+    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and \change{components of motion} (NS, EW, UD).
 In each case, \change{the boxes represent the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), the medians are indicated by a notch in the boxes, and the vertical lines show} the range of the data, with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots.

### 2018-04-27 method.tex
updated

@@
 
-The following sections explain the data processing analysis we put in place to cluster the dataset and obtain \myremove{two} \myrevision{three} alternative decision trees.
+The following sections explain the data processing analysis we put in place to cluster the dataset and obtain \change{three} alternative decision trees.
 

### 2018-04-27 results.tex
switching to office computer
updated

@@
 (Note that the only cases in which this is a direct view into the clustering results are the 2-feature analysis cases.)
+
+\begin{figure*}[ht!]
+	\centering
+	\includegraphics[width=\textwidth]{figures/pdf/figure-05}
+	\caption{A sample of the results from the multi-dimensional clustering analysis showing the clusters from a two-dimensional perspective into the relationships between different GOF metrics.
+In each case, the labels near the upper-left corner indicate the features considered in the subspace analysis, and the features associated to the horizontal ($x$) and vertical ($y$) axes in the plot.
+The top, middle, and bottom rows correspond to 2-, 3-, and 4-feature subspace analyses.
+In each case, the poor, fair, good, and excellent clusters are indicated with circle, triangle, square, and cross symbols, respectively.
+The empty circles indicate the location of the artificial cannot-link stations we introduced as background knowledge to the clustering process.
+The color version of this figure is available only in the electronic edition.}
+	\label{fig:clusters}
+\end{figure*}
 
@@
 
-\begin{figure*}[ht!]
-	\centering
-	\includegraphics[width=\textwidth]{figures/pdf/figure-05}
-	\caption{A sample of the results from the multi-dimensional clustering analysis showing the clusters from a two-dimensional perspective into the relationships between different GOF metrics.
-In each case, the labels near the upper-left corner indicate the features considered in the subspace analysis, and the features associated to the horizontal ($x$) and vertical ($y$) axes in the plot.
-The top, middle, and bottom rows correspond to 2-, 3-, and 4-feature subspace analyses.
-In each case, the poor, fair, good, and excellent clusters are indicated with circle, triangle, square, and cross symbols, respectively.
-The empty circles indicate the location of the artificial cannot-link stations we introduced as background knowledge to the clustering process.
-The color version of this figure is available only in the electronic edition.}
-	\label{fig:clusters}
-\end{figure*}
-
 Figure \ref{fig:boxed-clusters} shows the statistical distribution (box-plots) of the samples once the dataset is partitioned into the four GOF validation classes.
@@
 Separately, we prepared similar plots to look at the influence of the velocity models and components of motions on the results of the clustering process and, as observed before, they had no significant differences with the aggregate of all the samples in the dataset.
-\change{Figure 6 is particularly important because it highlights the outcome of the clustering process and provides a preview into the decision tree analysis results.
+\change{Figure \ref{fig:boxed-clusters} is particularly important because it highlights the outcome of the clustering process and provides a preview into the decision tree analysis results.
 Notice that metrics C5 through C8 consistently fall within the poor, fair, good, and excellent classification and that metrics C3 and C4 also show a progressive variation in sync with the validation classes.
@@
 Therefore, in reality, from the 20,000 combinations for which we ran the algorithm, only 66 unique trees were found.
+
+For each one of these unique trees, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}), and extracted the total number of nodes in the trees and their depth.
+Figure \ref{fig:nodes-depth} shows the results of $F_1$ for all the trees and its distribution in terms of the number of \change{decision} attributes \change{or GOF metrics} used in the trees as a function of the number of nodes and the depth of each tree.
+Recall that we are interested in finding a sequence of decisions (represented by disjunctive decision nodes in a tree) that can lead to good GOF predictions (i.e, high values of $F_1$) using a reduced number of attributes (GOF metrics).
+In general, all the trees obtained with the C5.0 algorithm are good in terms of the effectiveness factor ($F_1$ values close to 1).
+Then our choice comes down to using a reduced number of metrics.
+Having several trees with 2, 3, and 4 metrics (as opposed to 11), the following factor in the decision is choosing trees with algorithms using a small number of steps to reach the prediction.
+This is given by a combination between the depth of the tree and the number of nodes in the tree (i.e., trees with low complexity).
 
@@
 
-For each one of these unique trees, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}), and extracted the total number of nodes in the trees and their depth.
-Figure \ref{fig:nodes-depth} shows the results of $F_1$ for all the trees and its distribution in terms of the number of \change{decision} attributes \change{or GOF metrics} used in the trees as a function of the number of nodes and the depth of each tree.
-Recall that we are interested in finding a sequence of decisions (represented by disjunctive decision nodes in a tree) that can lead to good GOF predictions (i.e, high values of $F_1$) using a reduced number of attributes (GOF metrics).
-In general, all the trees obtained with the C5.0 algorithm are good in terms of the effectiveness factor ($F_1$ values close to 1).
-Then our choice comes down to using a reduced number of metrics.
-Having several trees with 2, 3, and 4 metrics (as opposed to 11), the following factor in the decision is choosing trees with algorithms using a small number of steps to reach the prediction.
-This is given by a combination between the depth of the tree and the number of nodes in the tree (i.e., trees with low complexity).
-
 Based on these criteria, we selected three candidate trees: T1, T2, and T3.
@@
 T3 is the most complex of them.
-More complex trees tend to be deeper, have more nodes, and employ more attributes.
-\myrevision{complexity of trees is controlled by pruning process.} This leads to higher $F_1$ values, but they may not necessarily be more practical.
+More complex trees tend to be deeper, have more nodes, and employ more attributes\change{, all of which depends---in part---on the pruning process.
+More complex trees have} higher $F_1$ values, but they may not necessarily be more practical.
 In total T1, T2, and T3 use 2, 3 and 4 attributes respectively.
@@
 T2 adds peak acceleration (C5), and T3 adds peak velocity \change{(C6)} to the previous metrics.
+
+At the bottom of each tree in Fig.~\ref{fig:trees} are the histograms of the data that land on each leaf node.
+(Note that the count of samples here is done based on the training dataset.) The decision tree assigns the final validation category based on the dominant class in each leaf.
+As such, the samples in the second leaf-node from the right in T3 are categorized as good despite a significant portion of them being excellent.
+This is a natural trade-off embedded in the use of decision trees.
+
+% , which actually results in more effective trees.
 
@@
 \end{figure*}
+
+The actual effectiveness (i.e., the average $F_1$ values shown in Fig.~\ref{fig:nodes-depth}), however, is measured based on the testing dataset.
+Recall that $F_1$ depends on the number of true predictions and false predictions measured by the confusion matrix and the precision and recall factors.
+Tables \ref{tab:t1:confusion} and \ref{tab:t3:confusion} show the confusion matrices for T1 and T3, respectively; and Tables \ref{tab:t1:f1} and \ref{tab:t3:f1} show the corresponding results for $P$, $R$, and $F_1$.
+As it can be seen in the Tables \ref{tab:t1:confusion} and \ref{tab:t3:confusion}, both trees lead to strong diagonal confusion matrices, meaning that the classification into poor, fair, good and excellent is well defined.
+The differences between both matrices are actually minor and small with respect to their diagonal values, and the $F_1$ values in Tables \ref{tab:t1:f1} and \ref{tab:t3:f1} are all near to or above 0.9.
+The values obtained for tree T2 are consistent with \change{these observations}.
+
+Another aspect of interest is the level of participation of each metric in the analysis of data, as the GOF values are run through \change{the nodes of the decision trees (i.e., as the tree is traversed)}.
+The results of such level of participation are listed in Table \ref{tab:participation} for trees T1, T2, T3, and T66.
+(The tree T66 is the most complex of them all as inferred from the number of metrics involved and the number of nodes and levels seen in Fig.~\ref{fig:nodes-depth}.) The percentages in this table represent the amount of data that is seen by a decision node associated to any particular metric.
+The nodes are not unique, and metrics are often used in different nodes in a given tree.
+Therefore, these percentage accumulate for each \change{metric as the tree is traversed}.
+A percentage of 100 means that a given metric has the opportunity to see all the data samples through one or multiple nodes, a low percentage means only a few data samples are evaluated by that metric, and a null percentage means the metric plays no role in the decision tree (it is not present in any node).
 
@@
 
-At the bottom of each tree in Fig.~\ref{fig:trees} are the histograms of the data that land on each leaf node.
-(Note that the count of samples here is done based on the training dataset.) The decision tree assigns the final validation category based on the dominant class in each leaf.
-As such, the samples in the second leaf-node from the right in T3 are categorized as good despite a significant portion of them being excellent.
-This is a natural trade-off embedded in the use of \myrevision{pruned} decision trees which actually results in more effective trees.
-
-The actual effectiveness (i.e., the average $F_1$ values shown in Fig.~\ref{fig:nodes-depth}), however, is measured based on the testing dataset.
-Recall that $F_1$ depends on the number of true predictions and false predictions measured by the confusion matrix and the precision and recall factors.
-Tables \ref{tab:t1:confusion} and \ref{tab:t3:confusion} show the confusion matrices for T1 and T3, respectively; and Tables \ref{tab:t1:f1} and \ref{tab:t3:f1} show the corresponding results for $P$, $R$, and $F_1$.
-As it can be seen in the Tables \ref{tab:t1:confusion} and \ref{tab:t3:confusion}, both trees lead to strong diagonal confusion matrices, meaning that the classification into poor, fair, good and excellent is well defined.
-The differences between both matrices are actually minor and small with respect to their diagonal values, and the $F_1$ values in Tables \ref{tab:t1:f1} and \ref{tab:t3:f1} are all near to or above 0.9.
-The values obtained for tree T2 are consistent with this.
-
-Another aspect of interest is the level of participation of each metric in the analysis of data as GOF values are run through the decision nodes in the trees.
-The results of such level of participation are listed in Table \ref{tab:participation} for trees T1, T2, T3, and T66.
-(The tree T66 is the most complex of them all as inferred from the number of metrics involved and the number of nodes and levels as seen in Fig.~\ref{fig:nodes-depth}.) The percentages in this table represent the amount of data that is seen by a decision node associated by any particular metric.
-The nodes are not unique and metrics are often used in different nodes in a given tree, thus these percentage are cumulative for each individual metric.
-A percentage of 100 means a given metric has the opportunity to see all data samples through a single nodes or multiple ones.
-A low percentage means only a few data samples are evaluated by that metric, and a null percentage means the metric plays no role in the decision tree (it is not present in any node).
-
-Notice that in all trees \myrevision{,using all metrics,} the metrics C4 (total energy) and C8 (response spectra) are consistently the most relevant ones in the decision algorithms.
+\change{The results in Table 6 highlight the fact that, for all trees,} the metrics C4 (total energy) and C8 (response spectra) are consistently the most relevant ones in the decision algorithms.
 They are followed by C5 (peak acceleration) and C6 (peak velocity).
-On the other end, notice that C11 (strong phase duration) plays no role whatsoever in any of the trees, and that C1 (Arias integral), C2 (energy integral), and C3 (Arias intensity) have only small to marginal participations.
+On the other end, C11 (strong phase duration) plays no role whatsoever in any of the trees, and that C1 (Arias integral), C2 (energy integral), and C3 (Arias intensity) have only small to marginal participations.
 The remaining metrics---C7 (peak displacement), C9 (Fourier spectrum), and C10 (Cross correlation)---have a somewhat significant participation, but only in the more complex trees.
 
-\myrevision{It is obvious that C8(response spectra) and C5(peak acceleration) are the most important metrics in developing the prediction models.
+\change{An additional aspect worth highlighting here is the fact that the clustering and decision tree processes address the problem of correlation between the metrics in a natural way.
+For instance, one would expect C3 and C5, or C3 and C4 to be related.
+Considering them all without distinction could lead to double weighing their influence, and assuming a 1-to-1 relationship may overstate their level of correlation.
+Instead, the process used here determines which of them has a better chance to predict the outcome in the presence of the other, and preserves that metric as part of the decision tree.}
+
+\change{One can put this to test by removing one or more metrics and see how the participation of the metrics rearranges.
+Table \ref{tab:participation-extra} shows these participations for trees with similar topologies (i) when removing C8, and (ii) when removing C5 and C8.
+In the first case, C5 takes the place of the most determinant metric, followed by C4 and C6, depending on the tree.
+In the second case, C6 becomes the most dominant, followed by C3 and C4, which vary in their level of participation depending on the tree.
+These additional alternatives emphasizes the C4 and C6, and show that other metrics such as C5, C7 and C9 are also relevant.
+They also reaffirm the low or null levels of contribution of metrics C1, C2, C10 and C11.
+All this is consistent with what we observed in Fig.~\ref{fig:boxed-clusters}.}
+
+% \myrevision{It is obvious that C8(response spectra) and C5(peak acceleration) are the most important metrics in developing the prediction models.
 Therefore, in order to understand other parameters importance we developed two sets of trees according to complexity of the original trees (the structure of generated trees are not presented here), while once dropping C8 (response spectra) and once dropping C8 (response spectra) and C5(peak acceleration) from the prediction attributes.
@@
 
-A low participation does not necessarily mean that a given metric is not important.
+\change{That being said, a} low participation does not necessarily mean that a given metric is not relevant at all.
 Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
@@
 In the end, the final selection of a preferred tree comes down to reducing complexity in the analysis without compromising the interpretation of results.
-We favor tree T1 because it is based only in three decision steps, and two GOF metrics: C4 (total energy), and C8 (response spectra).
+We favor tree T1 because it is based only on three decision steps, and two GOF metrics: C4 (total energy), and C8 (response spectra).
 

### 2018-04-27 testing.tex
updated

@@
 
-\myrevision{It is worth mentioning that, in this study, we presented an integration of scientists knowledge and field data using machine learning algorithms to develop an acceptable and robust prediction model.
+% \myrevision{It is worth mentioning that, in this study, we presented an integration of scientists knowledge and field data using machine learning algorithms to develop an acceptable and robust prediction model.
 This method can be used in any field where there are some field data as well as expertise knowledge about the data.

### 2018-04-27 decision-tree.tex
updated
almost there

@@
 Building decision trees is a supervised learning process used to approximate a target function as a sequence of disjunctive conditions designed to measure the effectiveness of a set of attributes to classify data and predict outcomes.
-The theory behind decision trees is well established \citep[e.g.,][]{Quinlan_1986_ML, Mitchell_1997_Book}.
-\myrevision{A decision tree can be used to classify a case by starting at the root of the tree and moving through it until the leaf is encountered.
+The theory behind decision trees is well established \citep[e.g.,][]{Quinlan_1986_ML, Quinlan_1993_Book, Mitchell_1997_Book}.
+% \myrevision{A decision tree can be used to classify a case by starting at the root of the tree and moving through it until the leaf is encountered.
 Each tree is a structure including a root, decision nodes and leaf nodes.
-Decision nodes specifies some test to carried out on attributes and leaf nodes indicate the final class \citep{Quinlan_1993_Book}.} Decision trees are, however, non-unique.
+Decision nodes specifies some test to carried out on attributes and leaf nodes indicate the final class \citep{Quinlan_1993_Book}.}
+Decision trees are, however, non-unique.
 They depend on the dataset, the user parameters, and the algorithm employed to build them \citep{Murthy_1998_DMKD}.
@@
 Therefore, the algorithm choose to use the most important attributes and ignore others.
-% We also activate the evaluation of possible advanced splits of data.
+We also activate the evaluation of possible advanced splits of data.
 Sometimes data is very close to the splitting threshold where assigning a hard threshold may incorrectly change the classification, whereas, that small difference could happen because of numerical errors or error in measurement.
 Therefore, advanced split can define a soft threshold to consider different probabilities in assigning data to different classes.
-% A decision tree can grow so deep to finally classify all training data correctly.
+A decision tree can grow so deep to finally classify all training data correctly.
 However, the tree will be very complex and will suffer from overfitting.

### 2018-04-27 introduction.tex
almost there

@@
 To that end, we study the relationships that exist between eleven different metrics---those proposed by \citet{Anderson_2004_Proc}, plus the strong motion duration---with the objective of offering a prioritized reduced number of metrics that can help predict validation results.
-We treat these metrics as independent variables defining a multi-dimensional space, and analyze them using machine learning techniques.
-We use a constrained \kmeans{} method \citep[e.g.,][]{Macqueen_1967_Proc, Wagstaff_2001_Proc} and conduct a subspace clustering analysis to address high-dimensional effects in order to add labels to the dataset in terms of four categories: poor, fair, good, and excellent.
-These categories \change{were defined by \citet{Anderson_2004_Proc} and have been adopted---despite some differences---by other GOF methods used in verification and validation \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}.}
-% \myrevision{are commonly used in previous validation studies} 
-% \myremove{adhere to the common practice in validation}. 
+We \change{initially} treat these metrics as independent variables defining a multi-dimensional space, and analyze them using machine learning techniques.
+\change{In particular, we used semi-supervised and supervised methods.
+In machine learning, whether a method is considered unsupervised, semi-supervised or supervised, depends on the level of constraints applied to the process or prior level of classification given to the data.
+Here, we first use a semi-supervised method called \textit{constrained k-means} in order to conduct a subspace clustering analysis to label data samples in a given validation dataset \citep[e.g.,][]{Macqueen_1967_Proc, Wagstaff_2001_Proc}.
+This dataset corresponds to the validation results from a series of simulations done for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake \citep{Taborda_2014_BSSA}.
+Here, the clustering done by the constrained $k$-means method allows us to label the data using for validation categories: poor, fair, good, and excellent.
+These categories were defined by \citet{Anderson_2004_Proc} and have been adopted---despite some differences---by other GOF methods used in verification and validation \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}.
+Upon labeling the data samples in our dataset into these four categories, we use a supervised learning method as implemented in the C5.0 algorithm to obtain a family of decision trees \citep[][]{Quinlan_1993_Book, Quinlan_1996_JAIR}.
+We then select a few of these trees to gain insight about and} prioritize the metrics involved in the validation process.
+The decision trees narrow the number of metrics offering a prediction algorithm into the aforementioned validation categories.
+
+% We use a constrained \kmeans{} method \citep[e.g.,][]{Macqueen_1967_Proc, Wagstaff_2001_Proc} and conduct a subspace clustering analysis to address high-dimensional effects in order to add labels to the dataset in terms of four categories: poor, fair, good, and excellent.
+These categories \change{were defined by \citet{Anderson_2004_Proc} and have been adopted---despite some differences---by other GOF methods used in verification and validation \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}.} \myrevision{are commonly used in previous validation studies} \myremove{adhere to the common practice in validation}.
 \myrevision{In the machine learning linguistic, \kmeans{} clustering approach is considered as an unsupervised learning method.
-However, when we add background knowledge into the process it becomes constrained \kmeans{} method and is considered as a semi-supervised learning method.}
-The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}.
+However, when we add background knowledge into the process it becomes constrained \kmeans{} method and is considered as a semi-supervised learning method.} The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}.
 Upon labeling the dataset, we develop a family of decision trees using a C5.0 algorithm \citep[][]{Quinlan_1993_Book, Quinlan_1996_JAIR}, and select a few to prioritize the metrics involved in the validation process.

### 2018-04-27 decision-tree.tex
update

@@
 First, regarding data classes we refer to whether the classes of data are properly distinguishable or not, and to whether the data attributes contribute to those classes or not.
-In our case, data classes are handled by means of the clustering and labeling process.
+In our case, data classes are handled by means of the clustering process.
 As we described in the previous section, at the end of the clustering process, all the data-points (or samples) in our dataset get labeled with one out of four outcome classes (or clusters) defined in terms of the validation categories (poor, fair, good, excellent).
@@
 
-\change{
-In the particular case of the C5.0 algorithm, in addition to the general steps just described, the result of the process of building a decision tree also depends on other logical and numerical parameters.
-We set the options to: (a) perform feature selection or winnowing, (b) evaluate possible advanced splits of data, \textcolor{red}{(c) discard the use of boosted models,} and (d) use a confidence factor and a threshold in the number of samples that go in th tree leaves.
-The feature selection option is used by the algorithm to choose the most important attribute over others.
-The option of evaluating advanced splits prevents the use of a hard threshold during the classification by considering different probabilities in assigning data to different classes.
-Discarding the use of boosted models prevents the algorithm from...
-\textcolor{red}{[pending for discussion with Naeem]}.
-Last, the options regarding the use of a confidence factor and a minimum number of samples ($S_{\min}$, also known as \textit{minCases}) that can be put in at least two of the branches are set to limit the level of complexity of the tree.
-In essence, the C5.0 algorithm grows a tree with many branches and then goes through a pruning process to satisfy a minimum number of samples in each leaf and minimize the error of misclassifying the data, which is measure by the confidence factor (CF).
+\change{In the particular case of the C5.0 algorithm, in addition to the general steps just described, the result of the process of building a decision tree also depends on other logical and numerical parameters.
+We set the options to: (a) perform feature selection or winnowing, (b) evaluate possible advanced splits of data, (c) use a confidence factor, and (d) use a threshold for the number of samples that go in the tree leaves.
+The feature selection option is used by the algorithm to choose the most important attribute over others; the option of evaluating advanced splits prevents the use of a hard threshold during the classification by considering different probabilities in assigning data to different classes; \colortext{red}{the confidence factor (CF) is used to estimate the errors in which one would incur should a given branch be removed, and then used to optimize the pruning process}; and the threshold for the number of samples in a leaf is set to a minimum number $S_{\min}$ (also known as \textit{minCases}) to limit the level of complexity of the tree.
 Additional details about these options are available in \citet{Quinlan_1993_Book} and \citet{Kuhn_2017_Manual}.
-For this study, in order to obtain a variety of tree options, we build trees for varying values of CF and $S_{\min}$.
-}
+For this study, in order to obtain a variety of tree options, we build trees for varying values of CF and $S_{\min}$.}
 

### 2018-04-27 metrics.tex
update

@@
 In his method, this frequency-domain decomposition should have pass bands defined following a logarithmic distribution in order to give more weight to the low-frequencies, and the GOF scores of all sub-bands and the broadband be averaged into a single GOF final score.
-\change{However, in this study, in order to avoid the additional parameters that would result from this approach (i.e., sub-bands width and filter characteristics), we used only the broadband (\frange{0}{4}) results.}
+\change{However, in this study, in order to avoid the additional parameters that would result from this approach \citep[i.e., sub-bands width and filter characteristics][]{Khoshnevis_2015_Proc}, we used only the broadband (\frange{0}{4}) results.}
 

### 2018-04-27 decision-tree.tex
update

@@
 Conversely, an extremely low value of entropy would mean most of the data falls within a particular class.
-With this in mind, the information gain $G$ of an attribute $A$ for the dataset $S$, or $G(S,A)$ in equation (\ref{eq:gain}), is the expected reduction in entropy caused by partitioning the dataset according to \change{the attribute $A$ \citep{Mitchell_1997_Book}, which here we consider to be numerically discrete.
-For selection of a threshold for a numerically continuous attribute, please refer to \citet{Quinlan_1996_JAIR}.}
+With this in mind, the information gain $G$ of an attribute $A$ for the dataset $S$, or $G(S,A)$ in equation (\ref{eq:gain}), is the expected reduction in entropy caused by partitioning the dataset according to \change{the attribute $A$ \citep{Mitchell_1997_Book}, which here we consider to be discrete.
+For selection of a threshold for a continuous attribute, please refer to \citet{Quinlan_1996_JAIR}.}
 
@@
 We set the options to: (a) perform feature selection or winnowing, (b) evaluate possible advanced splits of data, (c) use a confidence factor, and (d) use a threshold for the number of samples that go in the tree leaves.
-The feature selection option is used by the algorithm to choose the most important attribute over others; the option of evaluating advanced splits prevents the use of a hard threshold during the classification by considering different probabilities in assigning data to different classes; \colortext{red}{the confidence factor (CF) is used to estimate the errors in which one would incur should a given branch be removed, and then used to optimize the pruning process}; and the threshold for the number of samples in a leaf is set to a minimum number $S_{\min}$ (also known as \textit{minCases}) to limit the level of complexity of the tree.
+The feature selection option is used by the algorithm to choose the most important attribute over others; the option of evaluating advanced splits prevents the use of a hard threshold during the classification by considering different probabilities in assigning data to different classes; \textcolor{red}{the confidence factor (CF) is used to estimate the errors in which one would incur should a given branch be removed, and then used to optimize the pruning process}; and the threshold for the number of samples in a leaf is set to a minimum number $S_{\min}$ (also known as \textit{minCases}) to limit the level of complexity of the tree.
 Additional details about these options are available in \citet{Quinlan_1993_Book} and \citet{Kuhn_2017_Manual}.
@@
 This is done for all the trees obtained using all possible combinations of CF and $S_{\min}$ values.
-Then, we select the tree with the highest level of effectiveness as \change{measured} by $F_1$, commensurate to such tree having a low complexity, i.e., a reduced number of metrics and a shallow depth, which is precisely the goal set at the start.
+Then, we select trees with high levels of effectiveness as \change{measured} by $F_1$, commensurate to such trees having low complexities, i.e., reduced number of metrics and shallow depths, which is precisely the goal set at the start.
 Note that $F_1$ is obtained based on the testing dataset as opposed to the training dataset.

### 2018-04-28 response.tex
almost there
update

@@
 \response{%
-\textcolor{red}{[ *** Pending ***]}
-% We briefly explained these terms \textbf{Data Analysis Method} section. We also added clarification about semi-supervised and supervised learning into \textbf{Introduction} section.
+Thank you for bringing this up.
+We now provide a better approach to understanding these terms in the introduction and also included small changes to highlight the differences in the Data Analysis Method section.
 }
@@
 We modified the figure accordingly.
-The reason why there is no data for 1 is because there are no trees that would classify the data using a single metric into the four classes because these are binary trees, i.e., each node represents a binary decision, so one needs at least 2 metrics with 3 or more nodes with a minimum depth of 2.
-On the other end of things, the reason why there is no data for 11 metrics is because the analysis of the data resulted in the metric C11 not having any participation in the decision process, as shown in the original Table 6.
-In new version of Table 6 we include cases in which we forcefully removed metric C8, first, and then C5 and C8.
-Once one does that, the participations of the metrics changes and so would the topology of the trees.
+The reason why there is no data for 1 is because there are no trees that would classify the data using a single metric.
+On the other end, the reason why there is no data for 11 metrics is because the analysis of the data resulted in the metric C11 not having any participation in the decision process, as shown in Table 6.
+We also tried other trees for which we removed metrics C5 and C8.
+In such cases the participations of the metrics changes and so do the topology of the trees.
+Some of them include the metric C11.
 }

### 2018-04-28 decision-tree.tex
update

@@
 We set the options to: (a) perform feature selection or winnowing, (b) evaluate possible advanced splits of data, (c) use a confidence factor, and (d) use a threshold for the number of samples that go in the tree leaves.
-The feature selection option is used by the algorithm to choose the most important attribute over others; the option of evaluating advanced splits prevents the use of a hard threshold during the classification by considering different probabilities in assigning data to different classes; \textcolor{red}{the confidence factor (CF) is used to estimate the errors in which one would incur should a given branch be removed, and then used to optimize the pruning process}; and the threshold for the number of samples in a leaf is set to a minimum number $S_{\min}$ (also known as \textit{minCases}) to limit the level of complexity of the tree.
+The feature selection option is used by the algorithm to choose the most important attribute over others; the option of evaluating advanced splits prevents the use of a hard threshold during the classification by considering different probabilities in assigning data to different classes; the confidence factor (CF) is used to control the level of pruning; and the threshold for the number of samples in a leaf is set to a minimum number $S_{\min}$ (also known as \textit{minCases}) to limit the level of complexity of the tree.
 Additional details about these options are available in \citet{Quinlan_1993_Book} and \citet{Kuhn_2017_Manual}.

### 2018-04-28 response.tex
ricardo finished his review

@@
 We are enclosing a response to both your suggestions and those of the reviewers.
-We were able to satify most, if not all of them.
+We were able to satisfy most, if not all of them.
 To help in this second review cycle, in the annotated manuscript version of the paper, we have highlighted the most relevant changes that resulted from your comments and those of the reviewers in green font; and below we provide a response to each one of your comments, followed by our responses to the reviewers.
@@
 \response{%
-\textcolor{red}{[ *** Pending ***]}
-% Thanks for the constructive suggestions.
-We agree with you and peer reviewers that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
-The confusion happened because of integrating three different topics into a method to carry out this research.
-One technical term in one field is different in another field and that cause a confusion.
-For example cluster, class, and label all are referring to the same concept, however, it depends on the application.
-We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.
-% \vspace{2ex}
-% \newline
-% We added text to abstract and conclusion to give a heads up to readers about 11 metrics and one earthquake with 3 velocity model  simulations.
+Thank you for this suggestion.
+We revised the \textbf{Introduction}, \textbf{Study Dataset}, and \textbf{Data Analysis} sections to carefully introduce and explain the different terms and their equivalencies, when necessary.
+We think the introduced modifications should solve this issue and provide a long-lasting reference for the BSSA readership.
 }
@@
 \response{%
-\textcolor{red}{[ *** Pending ***]}
-% We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
-The confusion happened because of integrating three different topics into a method to carry out this research.
-One technical term in one field is different in another field and that cause a confusion.
-For example cluster, class, and label all are referring to the same concept, however, it depends on the application.
-We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.
+We revised the \textbf{Introduction}, \textbf{Study Dataset}, and \textbf{Data Analysis} sections to carefully introduce and explain the different terms and their equivalencies.
 }
@@
 \response{%
-These terms were not present in the Abstract.
-The term ``direct application'' was present in the Introduction, but it has been removed due to improvements in the text based on other suggestion by this and a second Reviewer.
+The term ``direct application'' was present in the Introduction, but it has been removed due to improvements in the text based on other suggestion by this and a Reviewer \#2.
 The terms ``direct quantitative comparisons'' and ``indirect statistical analysis'' were present in the Introduction and the Validation Metrics sections.
@@
 The explanation included an illustrative example for the simplest case of two categories.
-We assumed it would be clear that this concept is extendable to larger matrics such as those shown in Tables 2 and 3.
+We assumed it would be clear that this concept is extendable to larger matrices such as those shown in Tables 2 and 3.
 We added a sentence in the explanation to make it explicit that they can be extended to deal with more attributes.
@@
 We modified the figure accordingly.
-The reason why there is no data for 1 is because there are no trees that would classify the data using a single metric.
-On the other end, the reason why there is no data for 11 metrics is because the analysis of the data resulted in the metric C11 not having any participation in the decision process, as shown in Table 6.
+The reason why there is no data for 1 is because the analysis did not yield any tree that would classify the data using a single metric.
+On the other end, the reason there are no trees with the 11 metrics is because the analysis of the data resulted in the metric C11 not having any participation in the decision process, as shown in Table 6.
 We also tried other trees for which we removed metrics C5 and C8.
 In such cases the participations of the metrics changes and so do the topology of the trees.
-Some of them include the metric C11.
+Some of the trees resulting from that analysis did include the metric C11, but with a very small participation, as shown un the new Table 7.
 }
@@
 \response{%
-We also mention in the conclusion that the next metrics in line are the energy (C4), the peak acceleration (C5) and peak velocity (C6).
-In order to emphasize this point, as described in the response to comment \#8, we also included a new table (Table 7) showing the participation of the different metrics in the absence of the response spectra (C8) and the peak acceleration (C5).
-The results confirm that in such cases the scond most relevant metrics are still the energy (C4) and the peak ground velocity (C6) as mentioned before.
-This additional analysis also confirms the very low or null participation of the duration (C11) in other trees.
+These are very good points.
+Yes, the general consensus about the importance of the response spectra (C8) may not be surprising, but it is still very useful to confirm such importance through a data-informed process.
+Most important, as pointed out by the reviewer, is to answer the question about what other parameters are relevant.
+We think our work contributes precisely to that.
+We mention in the conclusion that the next metrics in line are the energy (C4), the peak acceleration (C5) and the peak velocity (C6).
+In order to further emphasize this point, as described in the response to comment \#8, we also included a new table (Table 7) showing the participation of the different metrics in the absence of the response spectra (C8) and the peak acceleration (C5).
+The results confirm that in such cases the next most relevant metrics are still the energy (C4) and the peak ground velocity (C6), as mentioned before.
+This additional analysis also reafirms the initial conclusion about the very small (and in some cases null) participation of the duration (C11) in most of the decision trees built in the data-processing analysis carried out here.
 }
@@
 \response{%
-\textcolor{red}{[ *** Pending ***]}
-% We agree with you that the technical terms are not commonly used in seismology and strong motion earthquake engineering.
-The confusion happened because of integrating three different topics into a method to carry out this research.
-One technical term in one field is different in another field and that cause a confusion.
-For example cluster, class, and label all are referring to the same concept, however, it depends on the application.
-We added a paragraph in the \textbf{Data Analysis Method} section to elaborate this confusion.
+Thank you for this suggestion.
+We revised the \textbf{Introduction}, \textbf{Study Dataset}, and \textbf{Data Analysis} sections to carefully introduce and explain the different terms and their equivalencies, when necessary.
 }
@@
 This was also done in response to the comments \#8 and \#16 from Reviewer 1.
-As we explain in the response to those comments, the procedure used here is different from a simpler correlation analisis in which a metric can be directly related by other if ``repetitive''.
+As we explain in the response to those comments, the procedure used here is different from a simpler correlation analysis in which a metric can be directly related by other if ``repetitive''.
 Here, the algorithm searches for the potential of each metric to serve as a predictor.
@@
 \response{%
-We disagree about the need to make detailed distinctions about these different algorithms.
-That is why we provided references.
-If the reader is interested in learning the differences, the most we can do without expanding the scope of our study is to point him/her in the right direction, which we already do.
-Nonetheless, in attention to the Reviewer, we modified slightly the descritpion to give some additional context in what regards to the C5.0 algorithm and added a sentence explaining that detailing the alternative algorithms is out of the scope of this study.
+Although it could certainly be good to have a comprehensive discussion about different methods, that would take much space, and it is not within the scope of this study.
+Instead, we provide adequate references to suggest further reading for the interested audience.
+That being said, in attention to the Reviewer's comment, we modified slightly the descritpion to give some additional context in what regards to the C5.0 algorithm and added a sentence explaining that detailing the alternative algorithms is out of the scope of this study.
 }
@@
 Following Equation (3) we define $A[a]$ as all the possible values of attribute $A$.
-The use of $[a]$ implies that $A$ is considered a numerically discrete attribute.
-We therefore meant to say that if $A$ continues attribute.
-Thanks for the comment.
-Fixed.
+The use of $[a]$ instead of $(a)$ implies that $A$ is considered to be discrete.
+We therefore meant to say that if $A$ was continuous, as in $A(a)$, then the reader should refer to Quinlan (1996).
+We improved the text to make this clearer.
 }

### 2018-04-28 abstract.tex
ricardo finished his review

@@
 %
-It has become common practice to validate ground motion simulations based on a variety of time and frequency metrics scaled to quantify the level of agreement between synthetics and data, or with respect to reference solutions.
+It has become common practice to validate ground motion simulations based on a variety of time and frequency metrics scaled to quantify the level of agreement between synthetics and data or other reference solutions.
 There is, however, no agreement about the importance or weight that it ought to be given to each metric.
@@
 As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could \change{focus} on a reduced number of metrics.
-We present an analysis \change{that looks into eleven} ground motion validation metrics using semi-supervised and supervised learning techniques\change{.
+We present an analysis \change{that looks into eleven} ground motion validation metrics using semi-supervised and supervised machine learning techniques\change{.
 These techniques} help label and classify goodness-of-fit results with the objective \change{of} prioritizing and narrowing the choice of \change{these metrics.
-In particular, we use the validation datasets of a series of physics-based ground motion simulations done for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake using different velocity models.
+In particular, we use a validation dataset of a series of physics-based ground motion simulations done for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake.
 We} study the relationships that exist between \change{the eleven metrics, and carry out a process where} these metrics are understood as part of a multi-dimensional space.
-We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address high-dimensional effects.
-This allows us to label a given validation dataset into four categories (poor, fair, good, excellent) following \change{previous studies}.
+We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address the implicit high-dimensional effects.
+This allows us to label the data in out dataset into four validation categories (poor, fair, good, excellent) following \change{previous studies}.
 We then develop a family of decision trees using the C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four referenced categories.
 These decision trees can be understood as rapid predictors of the quality of a simulation, \change{or as data-informed classifiers} that can help prioritize validation metrics.
-Our analysis\change{, although limited to the particular datasets used here,} indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.
+Our analysis\change{, although limited to the particular dataset used here,} indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of acceleration and velocity.

### 2018-04-28 dataset.tex
ricardo finished his review

@@
 
-We select the simulation results and validation analysis done by \citet{Taborda_2014_BSSA} as our reference dataset, hereafter the dataset.
+We select the simulation results and validation analysis done by \citet{Taborda_2014_BSSA} as our \change{\textit{validation dataset}}.
 In that study, the authors carried out deterministic simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake using a finite-element approach.
@@
 Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three \change{components} of motion (EW, NS, and UD).
-The simulation domain and the stations used for the analysis are shown in Figure \ref{fig:chino-hills}, and a sample of the validation results obtained by \citet{Taborda_2014_BSSA} using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps}.
-This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, averaged for all metrics, and the three \change{components} of motion.
+The simulation domain and the stations used in that study are shown in Figure \ref{fig:chino-hills}, and a sample of the validation results obtained by the authors using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps}.
+This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, averaged for all metrics and the three \change{components} of motion.
+
+\begin{figure*}[ht!]
+    \centering
+    \includegraphics[width=\textwidth]{figures/pdf/figure-01}
+    \caption{\textbf{(a)} Region of interest used by \citet{Taborda_2014_BSSA} for the simulations of the 2008 $M_W ~ 5.4$ Chino Hills earthquake, including the epicenter, focal mechanism, and major quaternary faults.
+In the background, the main roads and topography are shown for visual reference.
+\textbf{(b)} Distribution of the 336 stations (gray dots) used by \citet{Taborda_2014_BSSA} for the validation analysis of their simulations within the modeling domain shown in part (a).
+Roads, city names, and the hillshade topography are also shown here in the background for reference.
+The color version of this figure is available only in the electronic edition.}
+    \label{fig:chino-hills}
+\end{figure*}
 
@@
 \end{figure*}
-% 
+
+Here, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the \change{components of motion}.
+\change{We refer to those scores as the \textit{data samples} in our dataset.} Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be described in the following section was done using \change{all the data samples (i.e., GOF scores) in the validation dataset}.
+The motivation behind this choice was that the dataset, as a collection of GOF values, was independent of the simulations and serves here as a generic set of data samples for the purpose of identifying the correlations that exist between the different metrics in Table \ref{tab:metrics}.
+As such, given the simulations for each velocity model (3), the components of motion (3), and the number of stations used in the validation (336), gave us a large enough \change{dataset, with} 3,024 \change{data samples}.
+Figure \ref{fig:data-box-plot} illustrates the idea of a lack of dependence on the simulations by comparing the statistical distribution of the GOF scores of the simulations organized by velocity models and components.
+It is clear that although there are some differences between them, these are negligible.
+In other words, the statistical distribution of the \change{data samples} for each metric is about the same independently of model or component.
+Consequently, we felt it was alright to combine all the \change{data samples into a single dataset}.
+
 \begin{figure}[ht!]
@@
 \end{figure}
-
-Here, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the \change{components of motion}.
-Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be described in the following section was done using the whole dataset of scores.
-The motivation behind this choice was that the dataset, as a sample of GOF values, was independent of the simulation and serves here as a generic set for the purpose of identifying the correlations that exist between the different metrics in Table \ref{tab:metrics}.
-As such, given the simulations for each velocity model (3), the motion components (3), and the number of stations used in the validation (336) gave us a large enough sample of 3,024 GOF scores.
-Figure \ref{fig:data-box-plot} illustrates this idea by comparing the statistical distribution of the broadband GOF scores of the simulations classified by velocity models and components.
-It is clear that although there are differences between them, these are negligible; in other words, the statistical distribution of results for each metric is about the same independently of model or component.
-Consequently, we combine all the data into a single dataset.

### 2018-04-28 decision-tree.tex
ricardo finished his review

@@
 
-First, regarding data classes we refer to whether the classes of data are properly distinguishable or not, and to whether the data attributes contribute to those classes or not.
+First, by data classes we refer to whether the classes of data are properly distinguishable or not, and to whether the data attributes contribute to those classes or not.
 In our case, data classes are handled by means of the clustering process.
-As we described in the previous section, at the end of the clustering process, all the data-points (or samples) in our dataset get labeled with one out of four outcome classes (or clusters) defined in terms of the validation categories (poor, fair, good, excellent).
+As we described in the previous section, at the end of the clustering process, all the data-samples in our dataset are labeled \change{into one out the four validation categories.
+In decision trees, however, we refer to these categories as \textit{classes}.
+We are therefore interested in working with a labeled dataset or a properly \textit{classified} dataset using the validation classes P, F, G and E, which is something that has been guaranteed by the prior clustering process.}
 
-Second, attributes equivalence refers to whether the data attributes are comparable with each other on equivalent terms or not.
+% outcome classes (or clusters) defined in terms of the validation categories (poor, fair, good, excellent).
+
+Second, by attributes equivalence, we refer to whether the data attributes are comparable with each other on equivalent terms or not.
 There are cases, either because of the method or because of the data, that the data attributes need to be standardized.
@@
 Normalization is also necessary when data attributes come in different unit scales or value ranges \citep[e.g.,][]{Wu_2010_JH}.
-In our case, this is not necessary because (a) all the data attributes---defined by the 11 GOF metrics---are on the same dimensionless numerical scale (varying from 0 to 10), and (b) because the algorithm we use is not susceptible to the attributes scale.
+In our case, \change{the concept of \textit{attributes} in the context of the decision tree analysis is the same as the concept of \textit{features} in clustering or GOF metrics in validation.
+Since all our metrics are dimensionless and defined within the same numerical scale (varying from 0 to 10), normalization is not necessary and the process of building a decision tree} is not susceptible to the attributes scales.
+
+% In our case, this is not necessary because (a) all the data attributes---defined by the 11 GOF metrics---are on the same dimensionless numerical scale (varying from 0 to 10), and (b) because the algorithm we use is not susceptible to the attributes scale.
 
@@
 In our case, as we will see later, we apply oversampling to one of our dataset classes.
-\change{In general, oversampling increases the possibility of overfitting, which occurs when the training data leads the algorithm to produce a decision tree yields that predicts outcomes too close to or exactly the same as the training data.
+
+\change{In general, oversampling increases the possibility of overfitting, which occurs when the training data leads the algorithm to produce a decision tree that predicts outcomes too close to or exactly the same as the training data.
 This is not desirable because such a tree would lead to inaccurate predictions for other unseen data.
-Overfitting can be avoided by using conservative oversampling rates \citet{Weiss_2003_JAIR} and/or by applying heavy pruning methods.
+Overfitting can be avoided by using conservative oversampling rates \citep{Weiss_2003_JAIR} and/or by applying heavy pruning methods.
 (Pruning, as the word implies, is the process of cutting branches off a tree.
@@
 
-Upon completing the clustering and resampling processes, the next step is to subdivide the dataset in two, a training dataset (with 70 percent of the samples picked randomly), and a testing dataset (with the remaining 30 percent).
+Upon completing the clustering and resampling processes, the next step is to subdivide the dataset in two, a training dataset (with 70 percent of the samples, picked randomly), and a testing dataset (with the remaining 30 percent).
 We first use the training dataset to build a large suite of potential decision trees, and then use the testing dataset to evaluate and pick the best possible decision tree(s).
@@
 %
-where $A[a]$ is the set of all possible values of attribute $A$, $S_v$ is the subset of $S$ for which attribute $A$ has value $v$, and $E(\cdot)$ is the entropy function given by
+where $A[a]$ is the \change{discrete} set of all possible values of attribute $A$, $S_v$ is the subset of $S$ for which attribute $A$ has value $v$, and $E(\cdot)$ is the entropy function given by
 %
@@
 Conversely, an extremely low value of entropy would mean most of the data falls within a particular class.
-With this in mind, the information gain $G$ of an attribute $A$ for the dataset $S$, or $G(S,A)$ in equation (\ref{eq:gain}), is the expected reduction in entropy caused by partitioning the dataset according to \change{the attribute $A$ \citep{Mitchell_1997_Book}, which here we consider to be discrete.
-For selection of a threshold for a continuous attribute, please refer to \citet{Quinlan_1996_JAIR}.}
+With this in mind, the information gain $G$ of an attribute $A$ for the dataset $S$, or $G(S,A)$ in equation (\ref{eq:gain}), is the expected reduction in entropy caused by partitioning the dataset according to \change{the attribute $A$ \citep{Mitchell_1997_Book}, which we consider to be discrete.
+For the selection of a threshold for a continuous attribute $A(a)$, please refer to \citet{Quinlan_1996_JAIR}.}
 
@@
 We are interested in selecting a tree that is highly effective, but with a low level of complexity.
-In other words, we seek a tree with a good ratio of accuracy for predicting the classification outcome of the data, while using a reduced number of attributes (GOF metrics) in only a few number of steps (as given by a small number of decision nodes or a tree with shallow depth).
+In other words, we seek a tree with a good ratio of accuracy for predicting the classification outcome of the data, while using a reduced number of attributes (GOF metrics) in only a few number of steps (as given by a small number of decision nodes, or a tree with shallow depth).
 The number of metrics, the number of nodes, and the depth of a tree are directly seen from the topology of the tree, and are often proportional (shallow trees tend to use less nodes, and thus less metrics).
@@
 Here, TP, FP, and FN are the number of samples in the testing dataset considered to be true-positives, false-positives, and false-negatives, respectively.
-These values are typically ordered in what is known as a confusion matrix.
-For the simplest case of two categories, a confusion matrix takes the form
+
+These values are typically ordered in a confusion matrix. For the simplest case of two categories, a confusion matrix takes the form
 %
@@
 where TN is the number of true-negative samples.
-\change{These confusion matrices can be larger depending on the number of attributes.
-In our case, the confusion matrices are size 4$\times$4, in order to compare the actual number of samples with the predicted ones for the four given categories (poor, fair, good and excellent).}
+\change{Confusion matrices can be larger depending on the number of classes.
+In our case, the confusion matrices are size 4$\times$4, in order to compare the actual number of data samples classified as P, F, G and E with the number of samples predicted by the decision trees for each one of the same validation classes.}
 
@@
 Note that $F_1$ is obtained based on the testing dataset as opposed to the training dataset.
-The results as applied to our dataset are presented next.
+
+% The results as applied to our dataset are presented next.

### 2018-04-28 introduction.tex
ricardo finished his review

@@
 Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation\change{, as well as a growing interest in the use of synthetic seismograms for engineering applications}.
-\change{Validation entails the comparison of simulations with observations, whereas verification involves the comparison of simulations with a exact or alternative solutions \citep[see][and references therein]{Bielak_2010_GJI, Taborda_2013_BSSA}}.
-Various methods or schemes have been proposed to evaluate, through direct \change{signal-to-signal} quantitative comparisons or \change{overall} statistical analysis, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
+\change{Validation entails the comparison of simulations with observations, whereas verification involves the comparison of simulations with exact or alternative solutions \citep[see][and references therein]{Bielak_2010_GJI, Taborda_2013_BSSA}}.
+Various methods or schemes have been proposed to evaluate, through direct \change{signal-to-signal} quantitative comparisons or \change{overall} statistical analyses, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
 \change{Some of these methods are better} suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for validation.
 \change{\citet{Burks_BSSA_2014} and \citet{Rezaeian_2015_BSSA}, for instance, are used for validation in the context of engineering applications, i.e., at the level of the dynamic response of buildings.
-\citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL}, on the other hand, are used for validation in the context of the overall characteristics of seismograms when comparing synthetics with data in both the time and frequency domains.
-These methods are preferred for validating high-frequency or broadband simulations, where applications may be undefined and for which the expectation of matching seismograms at the waveform level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}), and because they offer simple schemes to quantify the goodness-of-fit (GOF) of simulation results.}
+\citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL}, on the other hand, are used for validation in the context of the overall characteristics of seismograms when comparing synthetics with data, in both time and frequency.
+These methods are preferred for validating high-frequency or broadband simulations, where applications may be undefined or the expectation of matching seismograms at the waveforms level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}), and because they offer simple schemes to quantify the goodness-of-fit (GOF) of the simulation results.}
 
@@
 
-In summary, \change{although limited to the datasets from specific simulation method and event}, our results indicate that 
+In summary, \change{although specific to the chosen dataset (from the given set of simulations for a single event)}, our results indicate that 
 % \myrevision{for \citeauthor{Anderson_2004_Proc}'s metrics based the results of 2008 $M_w$ 5.4 Chino Hills earthquake for three velocity models,} 
-among the eleven metrics considered in the analysis, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.
+among the eleven metrics considered in the analysis, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of acceleration and velocity.
 We test our prediction model and offer a discussion about its implications and potential use in future validation efforts.

### 2018-04-28 method.tex
ricardo finished his review

--- 2018-02-05 method.tex
@@
 
-We are interested in proposing an algorithm with a reduced and prioritized number of validation metrics based on previously acquired validation data (i.e., our dataset).
-A common method to do this is to identify rules with disjunctive characteristics in the form of a decision tree that leads to an outcome representative of the overall analysis.
-In our case, we define such outcome in terms of four categories representative of the quality of the validation: poor, fair, good, excellent.
-The development of such a decision tree requires a proper classification of the data, which can be done through a clustering process.
-The following sections explain the data processing analysis we put in place to cluster the dataset and obtain two alternative decision trees.
+We are interested in proposing an algorithm with a reduced and prioritized number of validation metrics based on previously acquired \change{collection of validation data samples} (i.e., our dataset).
+A common method to do this is to identify rules with disjunctive characteristics, in the form of decision trees, that lead to outcomes representative of the overall validation analysis.
+In our case, we define such outcome in terms of four \change{\textit{validation categories} or \textit{classes}} representative of the quality of the validation, namely: poor (P), fair (F), good (G), and excellent (E).
+The development of such decision trees requires a proper classification of the data, which can be done through a clustering process.
+\change{This this implies applying the \textit{labels} P, F, G and E to the data samples in our dataset.} The following sections explain the data processing analysis we put in place to \change{label the data samples in our} dataset and subsequently obtain \change{three} alternative decision trees.
 
@@
 \input{decision-tree}
+
+% SOME MATERIAL PROVIDED BY NAEEM
+% \myrevision{In this study we integrate three different topics (i.e., ground motion simulation, semi-supervised, and supervised learning), therefore, the same concept may have different technical terms in different part of the paper based on their application.
+In machine learning linguistic, attributes and features, mostly are used interchangeably.
+In this study, we use features for clustering and attributes for classification purposes.
+The reason for choosing these terms is keeping consistency with the references that we use to develop the methods.
+Therefore, in this study feature, attribute, and metric all refer to the eleven GOF scores.
+Consequently, dimension refers to the number of features in each constrained \kmeans{} clustering analysis.
+For example, 3-Dimensional subspace analysis is application of constrained \kmeans{} on three out of eleven features.
+Each observation with eleven metrics is categorized in one of Poor(p), Fair(F), Good(g), and Excellent (E) categories.
+The process of assigning these categories to each observation observation is called clustering or labeling.
+The resulted dataset is called the labeled dataset.
+When we use these labeled dataset in classification, the labels or clusters are called classes.
+Therefore, label, class, and cluster all refer to four categories of GOF level.
+}

### 2018-04-28 metrics.tex
ricardo finished his review

--- 2018-04-26 metrics.tex
@@
 
-There are various methods and algorithms available to evaluate the similarity between two or more seismograms, both through direct \chage{signal-to-signal} quantitative comparisons or \change{overall} statistical analysis \citep[e.g.,][]{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
-In earthquake ground motion simulation, they are primarily used for verification with respect to benchmark or analytical solutions, and for validation with respect to data (i.e., ground motion records).
+There are various methods and algorithms available to evaluate the similarity between two or more seismograms, both through direct \change{signal-to-signal} quantitative comparisons or \change{overall} statistical analyses \citep[e.g.,][]{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
+In earthquake ground motion simulation, they are used primarily for verification with respect to benchmark or analytical solutions, and for validation with respect to data (i.e., ground motion records).
 Here, we focus on the list of metrics proposed by \citet{Anderson_2004_Proc}, with an additional metric for duration, as suggested by others \citep{Olsen_2010_SRL, Maufroy_2015_BSSA} and as implemented in \citet{Taborda_2013_BSSA}.
+These metrics are listed in Table \ref{tab:metrics}.
 
@@
 
-These metrics are listed in Table \ref{tab:metrics}.
 Following \citet{Anderson_2004_Proc}, when applied to a pair of signals, each one of these metrics yields a GOF score ranging from 0 to 10, where a value of 10 corresponds to two signals having identical characteristics.
@@
 where $S$ is the GOF score that results from comparing values $p_1$ and $p_2$ from signals 1 and 2, respectively, for each one of the different metrics in Table \ref{tab:metrics}.
-\change{In the case of C8 and C9, where the values $p_1$ and $p_2$ are function of the period ($T$) and frequency ($f$), respectively, $S$ is computed for all values of $T$ and $f$ and averaged to produce a single mean score for metrics C8 and C9.} \citeauthor{Anderson_2004_Proc} also provided guidelines for the process to be applied to the original signals as well as to those resulting from a sequential set of band-pass filters covering the whole frequency range of interest.
-In his method, this frequency-domain decomposition should have pass bands defined following a logarithmic distribution in order to give more weight to the low-frequencies, and the GOF scores of all sub-bands and the broadband be averaged into a single GOF final score.
-\change{However, in this study, in order to avoid the additional parameters that would result from this approach (i.e., sub-bands width and filter characteristics), we used only the broadband (\frange{0}{4}) results.}
+\change{In the case of C8 and C9, where the values $p_1$ and $p_2$ are function of the period ($T$) and frequency ($f$), respectively, $S$ is computed for all values of $T$ and $f$ and averaged to produce a single mean score.} \citeauthor{Anderson_2004_Proc} also provided guidelines for the process to be applied to the original signals as well as to those resulting from a sequential set of band-pass filters covering the whole frequency range of interest.
+In his method, this frequency-domain decomposition should have pass bands defined following a logarithmic distribution in order to give more weight to the low frequencies, and the GOF scores of all sub-bands and the broadband be averaged into a single GOF final score.
+\change{However, in this study, in order to avoid the additional parameters that would result from this approach \citep[i.e., sub-bands width and filter characteristics][]{Khoshnevis_2015_Proc}, we use only the broadband (\frange{0}{4}) results.}
 
@@
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{figures/pdf/figure-01}
-    \caption{\textbf{(a)} Region of interest used by \citet{Taborda_2014_BSSA} for the simulations of the 2008 $M_W ~ 5.4$ Chino Hills earthquake, including the epicenter, focal mechanism, and major quaternary faults.
-In the background, the main roads and topography are shown for visual reference.
-\textbf{(b)} Distribution of the 336 stations (gray dots) used by \citet{Taborda_2014_BSSA} for the validation analysis of their simulations within the modeling domain shown in part (a).
-Roads, city names, and the hillshade topography are also shown here in the background for reference.
-The color version of this figure is available only in the electronic edition.}
-    \label{fig:chino-hills}
-\end{figure*}
-
-\citet{Anderson_2004_Proc} calibrated the results of GOF scores comparing horizontal components of recorded seismograms and other simulations using the first ten metrics (C1 through C10) in Table \ref{tab:metrics}.
-He concluded that for a typical distribution of the scores, the GOF could be categorized into four groups: poor (for scores from 0 to 4), fair (4 to 6), good (6 to 8), and excellent (for scores from 8 to 10).
-While this categorization is arguably subjective, over the years it has \change{been adopted---despite some differences in the ranges---by other GOF methods employed in verification and validation \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}.
+One important aspect of \citet{Anderson_2004_Proc} is that the results of the GOF scores were calibrated by comparing horizontal components of recorded seismograms and other simulations using the first ten metrics (C1 through C10) in Table \ref{tab:metrics} in order to find out how the scores were distributed throughout the scoring scale.
+\citeauthor{Anderson_2004_Proc} concluded that for a typical distribution of scores, the GOF values could be classified into four validation categories: poor (for scores from 0 to 4), fair (4 to 6), good (6 to 8), and excellent (for scores from 8 to 10).
+While these categories are arguably subjective, over the years they have \change{been adopted---despite some differences in the ranges---by other GOF methods employed in verification and validation \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}.
 As a result, \citeauthor{Anderson_2004_Proc}'s method has been used as} a reference baseline for various validation studies \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.

### 2018-04-28 results.tex
ricardo finished his review

@@
 
-As explained in the methodology, the first step is to carry out a clustering process on the dataset in order to properly label the samples according to the classes to be used in the decision tree algorithm.
+As explained in the methodology, the first step is to carry out a clustering process on the dataset in order to properly label the data samples according to the classes to be used in the decision tree algorithm.
 Figure \ref{fig:clusters} presents a sample of the results obtained after applying the constrained \kmeans{} method, including subspacing.

### 2018-04-28 testing.tex
ricardo finished his review

@@
 % A relative 
-drawback from the T1 algorithm is that because its outcome are GOF validation classes (poor, fair, good, excellent) as opposed to GOF validation scores (with values from 0 to 10), then once one obtains the results for the validation process for different components of motion (as shown in the rows of Fig.~\ref{fig:res-gof-maps}) then there is no natural way of taking averages as one would do for scores in a numerical scale \change{\citep[as typically done when using][]{Anderson_2004_Proc}}.
+drawback from the T1 algorithm is that because its outcome are GOF validation classes (poor, fair, good, excellent) as opposed to GOF validation scores (with values from 0 to 10), once one obtains the results for the validation process for different components of motion (as shown in the rows of Fig.~\ref{fig:res-gof-maps}), there is no natural way of taking averages as one would do for scores in a numerical scale \change{\citep[i.e., as typically done when using][]{Anderson_2004_Proc}}.
 Therefore, in order to combine results from the three components into a single validation classification for each station, we define the following rules:

### 2018-04-30 data-resources.tex
finished naeem.s revision

--- 2018-02-10 data-resources.tex
@@
 
-Calculations, data processing, and some initial figures were done using R, the language and environment for statistical computing and graphics (\url{https://www.r-project.org}, last accessed February 2018).
-Additional calculations and figures were done using MATLAB B (\url{http://www.mathworks.com}, last accessed February 2018).
-Map figures were prepared using the Generic Mapping Tools, GMT (\url{http://gmt.soest.hawaii.edu/}, last accessed February 2018).
-Additional editing of figures was done using Adobe Illustrator (\url{http://www.adobe.com/Illustrator‎}, last accessed February 2018).
+Calculations, data processing, and some initial figures were done using R, the language and environment for statistical computing and graphics (\url{https://www.r-project.org}, last accessed April 2018), and the C5.0 package developed by \citet{Kuhn_2017_Manual}.
+Additional calculations and figures were done using MATLAB (\url{http://www.mathworks.com}, last accessed April 2018).
+Map figures were prepared using the Generic Mapping Tools, GMT (\url{http://gmt.soest.hawaii.edu/}, last accessed April 2018).
+Additional editing of figures was done using Adobe Illustrator (\url{http://www.adobe.com/Illustrator‎}, last accessed April 2018).
 The validation dataset used here was readily available to the authors and can be provided without restrictions upon request.

### 2018-04-30 dataset.tex
finished naeem.s revision

@@
 Here, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the \change{components of motion}.
-\change{We refer to those scores as the \textit{data samples} in our dataset.} Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be described in the following section was done using \change{all the data samples (i.e., GOF scores) in the validation dataset}.
+\change{We refer to the set of 11 scores associated with a pair of signals as one of the \textit{data samples} that compose our dataset.} Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be described in the following section was done using \change{all the data samples in the validation dataset}.
 The motivation behind this choice was that the dataset, as a collection of GOF values, was independent of the simulations and serves here as a generic set of data samples for the purpose of identifying the correlations that exist between the different metrics in Table \ref{tab:metrics}.
@@
 It is clear that although there are some differences between them, these are negligible.
-In other words, the statistical distribution of the \change{data samples} for each metric is about the same independently of model or component.
+In other words, the statistical distribution of the data for each metric is about the same independently of model or component.
 Consequently, we felt it was alright to combine all the \change{data samples into a single dataset}.

### 2018-04-30 decision-tree.tex
finished naeem.s revision

@@
 In our case, data classes are handled by means of the clustering process.
-As we described in the previous section, at the end of the clustering process, all the data-samples in our dataset are labeled \change{into one out the four validation categories.
+As we described in the previous section, at the end of the clustering process, all the data samples in our dataset are labeled \change{into one out the four validation categories.
 In decision trees, however, we refer to these categories as \textit{classes}.
@@
 This is not desirable because such a tree would lead to inaccurate predictions for other unseen data.
-Overfitting can be avoided by using conservative oversampling rates \citep{Weiss_2003_JAIR} and/or by applying heavy pruning methods.
+Overfitting can be avoided
+% by using conservative oversampling rates \citep{Weiss_2003_JAIR} and/or
+by applying heavy pruning methods.
 (Pruning, as the word implies, is the process of cutting branches off a tree.
@@
 %
-where $P$ and $R$ are the precision and recall factors, respectively, and $\beta$ is a weighting factor between the two.
-$P$ measures the fraction of data samples that are correctly classified within a category, whereas $R$ measures the fraction of the data classified within that category regardless of whether it was correctly classified or not.
-Mathematically, they are defined as
+where $P$ and $R$ are the precision and recall factors, respectively, and $\beta$ is a weighting factor between the two. $P$ and $R$ are defined as
+% $P$ measures the fraction of data samples that are correctly classified \change{in that class, whereas $R$ measures the fraction that are correctly recalled.}
+%
+% \textcolor{red}{of the data classified within that category regardless of whether it was correctly classified or not.}
+%
+% Mathematically, they are defined as
 %

### 2018-04-30 method.tex
finished naeem.s revision

@@
 The development of such decision trees requires a proper classification of the data, which can be done through a clustering process.
-\change{This this implies applying the \textit{labels} P, F, G and E to the data samples in our dataset.} The following sections explain the data processing analysis we put in place to \change{label the data samples in our} dataset and subsequently obtain \change{three} alternative decision trees.
+\change{This implies applying the \textit{labels} P, F, G and E to the data samples in our dataset.} The following sections explain the data processing analysis we put in place to \change{label the data samples in our} dataset and subsequently obtain \change{three} alternative decision trees.
 

### 2018-04-30 results.tex
finished naeem.s revision

@@
 \change{Figure \ref{fig:boxed-clusters} is particularly important because it highlights the outcome of the clustering process and provides a preview into the decision tree analysis results.
-Notice that metrics C5 through C8 consistently fall within the poor, fair, good, and excellent classification and that metrics C3 and C4 also show a progressive variation in sync with the validation classes.
+Notice that metrics C5 through C8 consistently fall within \citeauthor{Anderson_2004_Proc}'s poor, fair, good, and excellent classifications, and that metrics C3 and C4 also show a progressive variation in sync with these categories.
 This means that these metrics are likely the best predictors for the outcome of the validation process, as we will see upon performing the decision tree analysis.
@@
 Since the original size was 76 samples, that means we applied an oversampling ratio of 10$\times$.
-According to \citet{Weiss_2003_JAIR}, an oversample rate of 10$\times$ is considered to be acceptable in this type of data processing.} (Arguably, we could have as well undersampled the fair class, but we deemed that unnecessary.
-Instead, we used a \change{heavy} pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.
+According to \citet{Weiss_2003_JAIR}, an oversample rate of 10$\times$ is considered to be acceptable in this type of data processing.} (Arguably, we could have undersampled the fair class as well but we deemed that unnecessary.
+Besides, we used a \change{heavy} pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.
 
@@
 Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
-What the results we present here mean \change{is that, in the context of this particular set of 11 metrics, in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more influential in the final result than the duration (C11).
+What the results we present here mean \change{is that, in the context of this particular set of 11 metrics, in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C4) are much more influential in the final result than the duration (C11).
 This, of course, is done under the assumption that one seeks to narrow the selection of metrics without loss of insight about the outcome of the validation process when compared to the outcome obtained with the whole suite of the eleven metrics used here.}

### 2018-04-30 testing.tex
finished naeem.s revision

@@
 The results obtained with the T1 validation algorithm are shown in Fig.~\ref{fig:res-gof-maps}.
-\change{In an ideal case, we would expect these results to closely resemble those obtained by \citet{Taborda_2014_BSSA}, if grouped only into four categories.
+\change{We should note that they are not supposed to be the same as those obtained by \citet{Taborda_2014_BSSA} because T1 no longer gives the same weight to all metrics but relies only on the two metrics C4 and C8.
+A
+%
+% In an ideal case, we would expect these results to closely resemble those obtained by \citet{Taborda_2014_BSSA}, if grouped only into four categories.
 Relying only on the two metrics C4 and C8 using T1 is, however, not exactly the same.

### 2018-04-30 introduction.tex
finished naeem.s revision
finalized revision, after cleaning

@@
 
-% Among the latter, some are better for direct application purposes because they compare the behavior of systems (e.g., buildings or earth structures) at the response level \citep{Burks_BSSA_2014}, while the remaining methods focus on the validation of the overall characteristics of seismograms.
-In this category, the comparisons are done both in time and frequency.
-This type of validation is better suited for high-frequency or broadband simulations, where applications may be undefined and for which the expectation of matching seismograms at the waveform level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}).
-Within this group, methods such as those proposed by \citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL} are preferred because they offer simple schemes to quantify the goodness-of-fit (GOF) of synthetics with respect to data in a broad sense.
-
 Out of the various validation methods just described, \citet{Anderson_2004_Proc} is perhaps the most widely used today \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
 \change{In essence, this method assesses the similarity of two signals based on the average score of 10 metrics.
-These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), their peak values (in acceleration, velocity and displacement), frequency content (in terms of Fourier and response spectra), and their time synchronization (through cross-correlation).}
-% 
-We, in particular, have consistently relied on a modified version of \change{\citeauthor{Anderson_2004_Proc}'s method} for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake \citep{Taborda_2013_BSSA, Taborda_2014_BSSA}, and for the evaluation of velocity models in southern California through the validation of a large suite of moderate-magnitude earthquake simulations \citep{Taborda_2016_GJI}.
-% 
-% \myremove{In essence, \citeauthor{Anderson_2004_Proc}'s method assesses the similarity of two signals based on the average score of 10 metrics.
-These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), their peak values (in acceleration, velocity and displacement), frequency content (in terms of Fourier and response spectra), and their time synchronization (through cross-correlation).
-}
-% 
-% The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics are \myrevision{relatively} well understood in both engineering and seismology contexts.
+These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), their peak values (in acceleration, velocity and displacement), frequency content (in terms of Fourier and response spectra), and their time synchronization (through cross-correlation).} We, in particular, have consistently relied on a modified version of \change{\citeauthor{Anderson_2004_Proc}'s method} for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake \citep{Taborda_2013_BSSA, Taborda_2014_BSSA}, and for the evaluation of velocity models in southern California through the validation of a large suite of moderate-magnitude earthquake simulations \citep{Taborda_2016_GJI}.
 
@@
 
-% Unfortunately, none of these alternatives addresses the underlying questions \change{regarding} what are the \myrevision{most important} parameters that ought to be taken into account when validating ground motion simulations \myrevision{for engineering applications without specific target application (e.g., designing tall buildings or power plants)}, and what level of priority should these \change{parameters} be given.
-
 Lack of consensus about how to answer these questions makes the choice of validation methods and the selection of the comparison metrics a subjective one, mostly based on personal preferences and---arguably---on the application intended for the simulation and/or the validation itself.
@@
 We \change{initially} treat these metrics as independent variables defining a multi-dimensional space, and analyze them using machine learning techniques.
-\change{In particular, we used semi-supervised and supervised methods.
+\change{In particular, we use semi-supervised and supervised methods.
 In machine learning, whether a method is considered unsupervised, semi-supervised or supervised, depends on the level of constraints applied to the process or prior level of classification given to the data.
@@
 
-% We use a constrained \kmeans{} method \citep[e.g.,][]{Macqueen_1967_Proc, Wagstaff_2001_Proc} and conduct a subspace clustering analysis to address high-dimensional effects in order to add labels to the dataset in terms of four categories: poor, fair, good, and excellent.
-These categories \change{were defined by \citet{Anderson_2004_Proc} and have been adopted---despite some differences---by other GOF methods used in verification and validation \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}.} \myrevision{are commonly used in previous validation studies} \myremove{adhere to the common practice in validation}.
-\myrevision{In the machine learning linguistic, \kmeans{} clustering approach is considered as an unsupervised learning method.
-However, when we add background knowledge into the process it becomes constrained \kmeans{} method and is considered as a semi-supervised learning method.} The dataset used corresponds to a series of simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake, and their corresponding validation results \citep{Taborda_2014_BSSA}.
-Upon labeling the dataset, we develop a family of decision trees using a C5.0 algorithm \citep[][]{Quinlan_1993_Book, Quinlan_1996_JAIR}, and select a few to prioritize the metrics involved in the validation process.
-The decision trees narrow the number of metrics offering a prediction algorithm into the aforementioned validation categories.
-\myrevision{Decision trees are considered as supervised learning algorithms.}
-
-In summary, \change{although specific to the chosen dataset (from the given set of simulations for a single event)}, our results indicate that 
-% \myrevision{for \citeauthor{Anderson_2004_Proc}'s metrics based the results of 2008 $M_w$ 5.4 Chino Hills earthquake for three velocity models,} 
-among the eleven metrics considered in the analysis, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of acceleration and velocity.
+In summary, \change{although specific to the chosen dataset (from the given set of simulations for a single event)}, our results indicate that among the eleven metrics considered in the analysis, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of acceleration and velocity.
 We test our prediction model and offer a discussion about its implications and potential use in future validation efforts.

### 2018-04-30 testing.tex
finalized revision, before cleaning
finalized revision, after cleaning

@@
-% \section{Verification Test}
 \section{\change{Testing}}
@@
 \change{We should note that they are not supposed to be the same as those obtained by \citet{Taborda_2014_BSSA} because T1 no longer gives the same weight to all metrics but relies only on the two metrics C4 and C8.
-A
-%
-% In an ideal case, we would expect these results to closely resemble those obtained by \citet{Taborda_2014_BSSA}, if grouped only into four categories.
-Relying only on the two metrics C4 and C8 using T1 is, however, not exactly the same.
-In addition, a}
-%
-% These results are in essence very similar to those obtained by \citet{Taborda_2014_BSSA}, though using the T1 algorithm, we now rely only on the two metrics C4 and C8.
-%
-% A relative
-drawback from the T1 algorithm is that because its outcome are GOF validation classes (poor, fair, good, excellent) as opposed to GOF validation scores (with values from 0 to 10), once one obtains the results for the validation process for different components of motion (as shown in the rows of Fig.~\ref{fig:res-gof-maps}), there is no natural way of taking averages as one would do for scores in a numerical scale \change{\citep[i.e., as typically done when using][]{Anderson_2004_Proc}}.
+A} drawback from the T1 algorithm is that because its outcome are GOF validation classes (poor, fair, good, excellent) as opposed to GOF validation scores (with values from 0 to 10), once one obtains the results for the validation process for different components of motion (as shown in the rows of Fig.~\ref{fig:res-gof-maps}), there is no natural way of taking averages as one would do for scores in a numerical scale \change{\citep[i.e., as typically done when using][]{Anderson_2004_Proc}}.
 Therefore, in order to combine results from the three components into a single validation classification for each station, we define the following rules:
@@
 This figure exemplifies the \change{differences or similarities} between the numeric GOF validation-scores scale and the proposed T1 GOF validation classes.
-% At the same time, it highlights the similarities between the outcomes.
 
@@
 
-% conclusions and present a similar picture of the validation of the simulation, not only globally but at specific locations, with only minor differences---especially if one is to consider the assumed equivalence between scores and classes (0--4: poor; 4--6: fair; 6--8:good; 8--10: excellent).
-\myrevision{However, we do not expect to see the same results in both figures.
-\citet{Taborda_2014_BSSA} results is developed based on assigning uniform weight on all metrics, however, this study research question is addressing the importance of different metrics and implying that assigning uniform weights might not provide accurate results.}
-
 \begin{figure*}[th!]
@@
 \end{figure*}
-
-% \myrevision{It is worth mentioning that, in this study, we presented an integration of scientists knowledge and field data using machine learning algorithms to develop an acceptable and robust prediction model.
-This method can be used in any field where there are some field data as well as expertise knowledge about the data.
-Validation metrics are note limited to \citeauthor{Anderson_2004_Proc}'s metrics.
-Therefore, adding more metrics from different field can be used through the proposed method to develop prediction models.
-}

### 2018-05-14 abstract.tex
finished naeem.s revision
last minute changes

@@
 In particular, we use a validation dataset of a series of physics-based ground motion simulations done for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake.
-We} study the relationships that exist between \change{the eleven metrics, and carry out a process where} these metrics are understood as part of a multi-dimensional space.
+We} study the relationships that exist between \change{eleven metrics, and carry out a process where} these metrics are understood as part of a multi-dimensional space.
 We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address the implicit high-dimensional effects.
-This allows us to label the data in out dataset into four validation categories (poor, fair, good, excellent) following \change{previous studies}.
+This allows us to label the data in our dataset into four validation categories (poor, fair, good, excellent) following \change{previous studies}.
 We then develop a family of decision trees using the C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four referenced categories.

### 2018-05-20 response-one.tex
completed first pass on second review by associate editor

--- 2018-04-28 response.tex
@@
 In such cases the participations of the metrics changes and so do the topology of the trees.
-Some of the trees resulting from that analysis did include the metric C11, but with a very small participation, as shown un the new Table 7.
+Some of the trees resulting from that analysis did include the metric C11, but with a very small participation, as shown in the new Table 7.
 }
@@
 \response{%
-Yes.
-We follow the guidelines of Anderson (2004), but the question of the reviewer gives way to two separate points of those guidelines: first, that in general the equation for $S$ is applied to all values of frequency (or period) for which is computed and then averaged to produce the GOF scores for C8 and C9; and second, that one can do this process for the broadband signals as well as for different frequency bands, and then average the results.
+Yes, we follow the guidelines of Anderson (2004), but the question of the reviewer gives way to two separate points of those guidelines: first, that in general the equation for $S$ is applied to all values of frequency (or period) for which is computed and then averaged to produce the GOF scores for C8 and C9; and second, that one can do this process for the broadband signals as well as for different frequency bands, and then average the results.
 The latter of these two points is applicable to not only C8 and C9 but to all metrics.

### 2018-05-20 response-two.tex
completed first pass on second review by associate editor

% \setlength{\paperwidth}{210mm}
% \setlength{\paperheight}{297mm}% fixed.

\documentclass{article}

\usepackage[review]{myreviewpckg}

\usepackage{textcomp}
\usepackage{simplemargins}
\usepackage{ifpdf}

\ifpdf
	\usepackage[pdftex,bookmarks=false]{hyperref}
\else
	\usepackage{graphics}
	\usepackage{graphicx}
	\usepackage{epsfig}
	\usepackage[dvipdfm,bookmarks=false]{hyperref}
\fi

\hypersetup{
    pdfpagelayout=OneColumn,
	colorlinks = true,
	urlcolor   = blue,
	citecolor  = blue,
	linkcolor  = blue
}

\clubpenalty=10000  % Orphan - First of paragraph left behind
\widowpenalty=10000 % Widow  - Last of paragraph sent ahead

\setallmargins{1in}

\begin{document}

% *****************************************************************
% *********************** Associate Editor ************************
% *****************************************************************

\thispagestyle{empty}
\setcounter{page}{0}

~

\vspace{6em}
\noindent\today

\vspace{5em}
\noindent
Dr.~Thomas Brocher\\
Associate Editor, BSSA\\
BSSA Editorial Office\\
400 Evelyn Ave, Ste 201\\
Albany, CA 94706

\vspace{4em}
\noindent
Dear Dr. Brocher,

\vspace{1em}
\noindent
Thank you again for reviewing our manuscript (BSSA-D-18-00056R1). We are enclosing a response to your comments and suggestions, which we hope you will find satisfactory. To help with the final review, we have assumed all previous changes as accepted, and produced a new annotated manuscript highlighting the latest modifications for your convenience.
\\

\noindent
We look forward to hearing back from you.

\vspace{2em}
\noindent
Sincerely yours,

\vspace{5em}
\noindent
Ricardo Taborda\\
Department of Civil Engineering, and\\
Center for Earthquake Research and Information\\
The University of Memphis

\newpage
\begin{center}
	\bf
	\large
	Authors' Response
\end{center}

\noindent
As done before, below, we provide a response to each comment. The original comments are in italic black font, followed by our responses in regular blue font.
\vspace{2ex}
\newline

\introcomment{~}{%
I appreciate the author's thoughtful response to the reviewers suggestions and to the revisions that they made to their manuscript, which have substantially improved the paper. I think the paper is close to being acceptable for publication, but I have a few minor suggestions for the authors to consider.
}

\response{%
Thank you. We think we have addressed all your suggestions, and hope you will find the new version of our submission acceptable for publication.
}

\comment{1}{%
Figure 2. In the boxes showing the counts of the GOF factors, consider indicating the ranges of Poor, Fair, Good, and Excellent to allow the reader to quickly see that most GOF factors are in the Fair range. Would it also be useful to indicate the average GOF score for each model? This would allow the reader interested in velocity models to more easily which model produced the highest average GOF.
}

\response{%
We modified Figure 2 by including the ranges for poor, fair, good and excellent, as suggested, and modified the figure's caption accordingly. We, however, did not include the average score as it may be misleading to the readers. The reason for that is that the values for the particular case of the 2008 Chino Hills earthquake at high frequencies would suggest that the velocity model CVM-H+GTL is better than CVM-S or CVM-H, but we know that is not generally the case. In a subsequent study (Taborda et al.,~2016), which is cited in the paper, we showed that over a larger collection of events (30), and for the lower frequencies at which the models can be more properly evaluated, the velocity model CVM-S (in its version 4.26.M01) is the one that is most consistent in yielding better simulation results when these are compared to data. We had cited that paper in the Introduction, and recognizing it may be of interest at this point as well, we now modified the end of the first paragraph in the \textbf{Study Dataset} section to reflect that.
}

\comment{2}{%
Line 146. I assume that all 11 of Anderson's metrics were used for the GOFs shown in Figure 2, but this is not explicitly stated. It should be explicitly stated which of Anderson's metrics were used in Figure 2. It may also be worthwhile to comment that the general fit was Fair, and to compare the average GOF score for the 3 velocity models. This information all may be in Taborda and Bielak (2014) but this paper should repeat them so that the readers don't have to look up the older paper.
}

\response{%
That is correct. We now make explicit mention to the fact that those results were obtained using the eleven metrics described earlier and shown in Table 1. Following the explanation in our response to the previous comment, we modified the end of the first paragraph in the \textbf{Study Dataset} section to make reference to the evaluation of the velocity models done later on by Taborda et al.~(2016). We think that, on what regards to comparing velocity models, that reference is better than Taborda and Bielak (2014).
}

\comment{3}{%
Line 165. ``Alright'' is not quite the right word here. I think you mean either acceptable or reasonable. Did you consider other options, such as using only one component for one velocity model? I assume that you wanted as large a dataset as possible, but you should state the reason you followed the approach of using all data.
}

\response{%
Certainly. We fixed the language and added a sentence to explain the reason behind our choice, which was indeed to use a dataset as large as possible. Thank you.
}

\comment{4}{%
Figure 5. Consider adding lines to each plot that would indicate a 1:1 relation between the x and y axes. These lines would highlight datasets that lie on and off this line, as well as help the reader better see the scatter in the data.
}

\response{%
We appreciate this comment. However, we are reluctant to add the 1:1 relationship lines. The reason being that these plots should not be interpreted as typical correlation plots of scattered data are usually interpreted. If we add the line we would be erroneously leading the readers to think these are as any other $x$-$y$ plots correlating showing the relationship of two metrics as seen through the data, and here one also needs to understand that the symbols (or colors) and how well defined are the transitions from one to another are actually equally, and perhaps more important than whether the cloud of data points follow a 1:1 relationship or not. To make this point clearer we added a comment in the text reflecting this point... \textcolor{red}{Adding the comment in the text of the paper is pending on Naeem agreeing with this response}.
}

\comment{5}{%
Line 280. Suggest ``scope of the paper'' rather than ``scope of the study''.
}

\response{%
Done. Thanks for the sugestion.
}

\comment{6}{%
Tables 6 and 7. To make it easier for the reader to easily understand your results, I suggest adding the full title of each metric to the tables from Table 1.
}

\response{%
Thank you for the suggestion. We adopted it and modified Tables 6 and 7 accordingly.
}

\comment{7}{%
Lines 437-443 and 450-456. Throughout the text, but especially in these paragraphs, I suggest making it easier for the reader to understand which metrics matter and which don't. For this reason I would state the full name of the metric, followed by its code. For example, I would write total energy (C4) instead of just C4 or C4 (total energy). The full name of the metric is easily understood by any reader, the code requires the reader to constantly refer to Table 1.
}

\response{%
Thank you for the suggestion. We thought that doing it all throughout the paper would likely be too repetitive and at some points unnecessary. That being said, we completely agree that in these paragraphs of the \textbf{Results} section using the full names would help carry over the main observations we draw from the analysis. We have therefore modified this part of the manuscript in attention to this comment. We also review the format to be consistent, i.e., first the code, then the name within parentheses.
}

\comment{8}{%
Line 479. Suggest ``of the algorithm'' rather than ``from the algorithm''.
}

\response{%
Fixed. Thanks.
}

\comment{9}{%
Lines 497-499. One of the differences between the two results are that the Tree T1 Combination appears to yield a higher percentage of poor and fair scores than the 11-metric score. But I'm not sure because it's actually very hard to compare the counts in Figure 11. A more direct comparison could be made if the counts in the numerical All GOF Scores were combined into Poor, Fair, Good, and Excellent, and plotted over the Tree T1 Combination count in the lower right hand side. These boxes could be shown as dashed or dotted lines and would make this comparison much more quantitative.
}

\response{%
We followed this advise and modified Figure 11 accordingly. We also modified the text in the \textbf{Testing} section.
}

\comment{10}{%
Line 533. Suggest ``to focus'' instead of ``the focus''.
}

\response{%
Done. Thanks.
}

\comment{11}{%
Lastly, it is customary to acknowledge and/or to thank, the journal reviewers for their helpful suggestions.
}

\response{%
Absolutely! We totally forgot about that during the previous revision. Our apologies for that. We have now added our thanks to the reviewers and to you as well.
}

\end{document}

### 2018-05-20 abstract.tex
completed first pass on second review by associate editor

@@
 This leads to their selection often being subjective, either based on intended applications or personal preferences.
-As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could \change{focus} on a reduced number of metrics.
-We present an analysis \change{that looks into eleven} ground motion validation metrics using semi-supervised and supervised machine learning techniques\change{.
-These techniques} help label and classify goodness-of-fit results with the objective \change{of} prioritizing and narrowing the choice of \change{these metrics.
+As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could focus on a reduced number of metrics.
+We present an analysis that looks into eleven ground motion validation metrics using semi-supervised and supervised machine learning techniques.
+These techniques help label and classify goodness-of-fit results with the objective of prioritizing and narrowing the choice of these metrics.
 In particular, we use a validation dataset of a series of physics-based ground motion simulations done for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake.
-We} study the relationships that exist between \change{eleven metrics, and carry out a process where} these metrics are understood as part of a multi-dimensional space.
+We study the relationships that exist between eleven metrics, and carry out a process where these metrics are understood as part of a multi-dimensional space.
 We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address the implicit high-dimensional effects.
-This allows us to label the data in our dataset into four validation categories (poor, fair, good, excellent) following \change{previous studies}.
+This allows us to label the data in our dataset into four validation categories (poor, fair, good, excellent) following previous studies.
 We then develop a family of decision trees using the C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four referenced categories.
-These decision trees can be understood as rapid predictors of the quality of a simulation, \change{or as data-informed classifiers} that can help prioritize validation metrics.
-Our analysis\change{, although limited to the particular dataset used here,} indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of acceleration and velocity.
+These decision trees can be understood as rapid predictors of the quality of a simulation, or as data-informed classifiers that can help prioritize validation metrics.
+Our analysis, although limited to the particular dataset used here, indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of acceleration and velocity.

### 2018-05-20 conclusions.tex
finalized revision, after cleaning
completed first pass on second review by associate editor

--- 2018-04-06 conclusions.tex
@@
 
-We present the \change{results} of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of \change{prioritizing} and \change{reducing} the number of metrics, and develop an application independent decision algorithm.
+We present the results of a machine learning analysis using semi-supervised and supervised methods on a large dataset comparing synthetics and data based on multiple goodness-of-fit metrics used in ground motion validation with the goal of prioritizing and reducing the number of metrics, and develop an application independent decision algorithm.
 As a result of the data processing analysis, we propose a simplified algorithm based on a decision tree which uses only two metrics as opposed to the initial eleven available in the dataset.
@@
 We also propose rules to allow for the new class-based validation criteria to combine results obtained from different components of motion, and demonstrate that the results obtained with the proposed algorithm using two metrics are comparable to those obtained with the score-based validation results used in other recent validation studies.
-\change{One} could implement similar rules to combine results from a frequency-band analysis, using, for example, the mode or majority validation class.
+One could implement similar rules to combine results from a frequency-band analysis, using, for example, the mode or majority validation class.
 
-We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulations done for a particular earthquake \myrevision{(one earthquake and three velocity models)} in a particular region.
-In a future follow-up study it would be ideal to refine the tree adding other simulation datasets (i.e., from different earthquakes, regions, models, \myrevision{metrics,} etcetera) in order to arrive to a sufficiently sound and robust decision tree.
-The procedural steps laid out here, nonetheless, \change{remain} valid.
-We showed that the procedure and background information used for clustering and decision making is stable, and it is likely that the metrics of energy and response spectra (along with peak acceleration and velocity as suggested by the additional trees) will prevail as those among the most decisive ones.
+We recognize, however, that the proposed decision tree algorithm may not be a definitive one because of a potential bias on the fact that the dataset used here, although large enough from a statistical point of view, came from simulations done for a particular earthquake, in a particular region, and using a particular set of metrics.
+In a future follow-up study, it would be ideal to refine the tree adding other simulation datasets (i.e., from different earthquakes, regions, models, and using additional metrics) in order to arrive to a sufficiently sound and robust decision tree.
+The procedural steps laid out here, nonetheless, remain valid, and there is no reason why not to use it also in other contexts utilizing other metrics (e.g., structural responses metrics such as drift ratio) provided that they are properly normalized.
+In summary, we can say that for the case of the metrics used here, we showed that the procedure and background information used for clustering and decision making is stable, and it is likely that---despite the limitations just described---the metrics of energy and response spectra (along with peak acceleration and velocity as suggested by the additional trees) will prevail as those among the most decisive ones.
 
-This latter point is perhaps the most important conclusion of our study.
-Identifying the total energy and response spectra metrics as the decisive ones is a key contribution to validation of ground motion simulations.
-On the one hand, this contributes to clarifying a standing question in the area of validation, and on the other hand, it signals to simulators where the focus should be in the search for improvements in their models.
+Identifying the total energy and response spectra metrics as decisive metrics in the comparisons between synthetics and observations is a key contribution to validation of ground motion simulations.
+This contributes to clarifying a standing question in the area of validation, and it provides an indication to simulators about where \change{to focus} in the search for improvements in their models.

### 2018-05-20 dataset.tex
completed first pass on second review by associate editor

@@
 
-We select the simulation results and validation analysis done by \citet{Taborda_2014_BSSA} as our \change{\textit{validation dataset}}.
+We select the simulation results and validation analysis done by \citet{Taborda_2014_BSSA} as our \textit{validation dataset}.
 In that study, the authors carried out deterministic simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake using a finite-element approach.
@@
 The modeling domain covered an area of \adomain{180}{135}{km} that included all the major sedimentary basins and other relevant geologic structures in the greater Los Angeles region.
-Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three \change{components} of motion (EW, NS, and UD).
+Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three components of motion (EW, NS, and UD).
 The simulation domain and the stations used in that study are shown in Figure \ref{fig:chino-hills}, and a sample of the validation results obtained by the authors using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps}.
-This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, averaged for all metrics and the three \change{components} of motion.
+This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, \change{for the three velocity models used by \citet{Taborda_2014_BSSA}.
+In each case, the results are the average of the GOF results for the three components of motion and the scores for the eleven metrics from Table \ref{tab:metrics}.
+A subsequent study by \citet{Taborda_2016_GJI} using multiple events found that the model CVM-S \citep[version 4.26.M01, see][]{Small_2017_SRL} to be the model that most consistently led to better simulation results.
+Here, however, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the components of motion}.
 
@@
     \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between simulations for three different velocity models and data recorded at ground motion monitoring stations.
-The color version of this figure is available only in the electronic edition.}
+\change{The distribution of the results at the bottom show the count of stations on the GOF scale defined by \citet{Anderson_2004_Proc} and the ranges of the validation categories: poor, fair, good and excellent.} The color version of this figure is available only in the electronic edition.}
     \label{fig:ref-gof-maps}
@@
 
-Here, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the \change{components of motion}.
-\change{We refer to the set of 11 scores associated with a pair of signals as one of the \textit{data samples} that compose our dataset.} Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be described in the following section was done using \change{all the data samples in the validation dataset}.
+We refer to the set of 11 scores associated with a pair of signals \change{from the work done by \citet{Taborda_2014_BSSA}} as one of the \textit{data samples} that compose our dataset.
+Although at times we will make distinctions between the velocity models and the components of motion for visualization purposes, the clustering analysis to be described in the following section was done using all the data samples in the validation dataset.
 The motivation behind this choice was that the dataset, as a collection of GOF values, was independent of the simulations and serves here as a generic set of data samples for the purpose of identifying the correlations that exist between the different metrics in Table \ref{tab:metrics}.
-As such, given the simulations for each velocity model (3), the components of motion (3), and the number of stations used in the validation (336), gave us a large enough \change{dataset, with} 3,024 \change{data samples}.
+As such, given the simulations for each velocity model (3), the components of motion (3), and the number of stations used in the validation (336), gave us a large enough dataset, with 3,024 data samples.
 Figure \ref{fig:data-box-plot} illustrates the idea of a lack of dependence on the simulations by comparing the statistical distribution of the GOF scores of the simulations organized by velocity models and components.
@@
 In other words, the statistical distribution of the data for each metric is about the same independently of model or component.
-Consequently, we felt it was alright to combine all the \change{data samples into a single dataset}.
+Consequently, \change{and in order to use a dataset as large as possible, we thought it acceptable} to combine all the data samples into a single dataset.
 
@@
     \includegraphics[width=0.48\textwidth]{figures/pdf/figure-03}
-    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and \change{components of motion} (NS, EW, UD).
-In each case, \change{the boxes represent the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), the medians are indicated by a notch in the boxes, and the vertical lines show} the range of the data, with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots.
+    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and components of motion (NS, EW, UD).
+In each case, the boxes represent the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), the medians are indicated by a notch in the boxes, and the vertical lines show the range of the data, with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots.
 The color version of this figure is available only in the electronic edition.}

### 2018-05-20 decision-tree.tex
finalized revision, after cleaning
completed first pass on second review by associate editor

--- 2018-04-28 decision-tree.tex
@@
 The theory behind decision trees is well established \citep[e.g.,][]{Quinlan_1986_ML, Quinlan_1993_Book, Mitchell_1997_Book}.
-% \myrevision{A decision tree can be used to classify a case by starting at the root of the tree and moving through it until the leaf is encountered.
-Each tree is a structure including a root, decision nodes and leaf nodes.
-Decision nodes specifies some test to carried out on attributes and leaf nodes indicate the final class \citep{Quinlan_1993_Book}.}
 Decision trees are, however, non-unique.
@@
 In our case, data classes are handled by means of the clustering process.
-As we described in the previous section, at the end of the clustering process, all the data-samples in our dataset are labeled \change{into one out the four validation categories.
+As we described in the previous section, at the end of the clustering process, all the data samples in our dataset are labeled into one out the four validation categories.
 In decision trees, however, we refer to these categories as \textit{classes}.
-We are therefore interested in working with a labeled dataset or a properly \textit{classified} dataset using the validation classes P, F, G and E, which is something that has been guaranteed by the prior clustering process.}
-
-% outcome classes (or clusters) defined in terms of the validation categories (poor, fair, good, excellent).
+We are therefore interested in working with a labeled dataset or a properly \textit{classified} dataset using the validation classes P, F, G and E, which is something that has been guaranteed by the prior clustering process.
 
@@
 Normalization is also necessary when data attributes come in different unit scales or value ranges \citep[e.g.,][]{Wu_2010_JH}.
-In our case, \change{the concept of \textit{attributes} in the context of the decision tree analysis is the same as the concept of \textit{features} in clustering or GOF metrics in validation.
-Since all our metrics are dimensionless and defined within the same numerical scale (varying from 0 to 10), normalization is not necessary and the process of building a decision tree} is not susceptible to the attributes scales.
-
-% In our case, this is not necessary because (a) all the data attributes---defined by the 11 GOF metrics---are on the same dimensionless numerical scale (varying from 0 to 10), and (b) because the algorithm we use is not susceptible to the attributes scale.
+In our case, the concept of \textit{attributes} in the context of the decision tree analysis is the same as the concept of \textit{features} in clustering or GOF metrics in validation.
+Since all our metrics are dimensionless and defined within the same numerical scale (varying from 0 to 10), normalization is not necessary and the process of building a decision tree is not susceptible to the attributes scales.
 
@@
 
-\change{In general, oversampling increases the possibility of overfitting, which occurs when the training data leads the algorithm to produce a decision tree that predicts outcomes too close to or exactly the same as the training data.
+In general, oversampling increases the possibility of overfitting, which occurs when the training data leads the algorithm to produce a decision tree that predicts outcomes too close to or exactly the same as the training data.
 This is not desirable because such a tree would lead to inaccurate predictions for other unseen data.
-Overfitting can be avoided by using conservative oversampling rates \citep{Weiss_2003_JAIR} and/or by applying heavy pruning methods.
+Overfitting can be avoided by applying heavy pruning methods.
 (Pruning, as the word implies, is the process of cutting branches off a tree.
-Heavy pruning methods limit the complexity of the tree by constraining the number of branches and/or depth of a tree.)}
-
-% This process has consequences, mainly that it increases the likelihood of overfitting.
-\myrevision{Overfitting happens when the prediction algorithm predicts the final results too closely or exactly to the training data, however, it provides non-accurate results for unseen or test data.}
+Heavy pruning methods limit the complexity of the tree by constraining the number of branches and/or depth of a tree.)
 
@@
 There are several algorithms for building decision trees \citep[e.g., ID3, C4.5, C5.0, CART; see][]{Quinlan_1986_ML, Quinlan_1993_Book, Quinlan_1996_JAIR, Breiman_1984_Book}.
-\change{Here, we use the C5.0 algorithm as implemented by \citet{Kuhn_2017_Manual}.
+Here, we use the C5.0 algorithm as implemented by \citet{Kuhn_2017_Manual}.
 This algorithm is the latest update to the original ID3 and subsequent C4.5 algorithms \citep{Quinlan_1993_Book, Quinlan_1996_JAIR}.
 C5.0 is superior to its predecessors because it reduces the limitations for handling numerical attributes, missing data, and it has additional features such as the development of boosted models.
-Discussing the differences between these algorithms is out of the scope of this study.}
+Discussing the differences between these algorithms is out of the scope of \change{the paper}.
 
-% \myrevision{C5.0 is considered as the developed version of ID3 and C4.5, where it does not have any limitations in handling missing data, numerical attributes, and developing boosted models.
-Internal algorithm of CART is different than other methods, however, discussing technical details of these algorithms are beyond the scope of this study.}
-
-% Here we use the C5.0 implementation \citep{Kuhn_2017_Manual}, which is an extension of the C4.5 algorithm developed by \citet{Quinlan_1993_Book, Quinlan_1996_JAIR}.
-
-In \change{general}, given a dataset, the process of building the tree consists of recursively identifying the attributes in the training data that are most likely to predict an outcome.
+In general, given a dataset, the process of building the tree consists of recursively identifying the attributes in the training data that are most likely to predict an outcome.
 The process is recursive because new branches and decision nodes are created based on the remaining data at every branch and level, until the tree reaches a certain depth or when other conditions are met.
@@
 %
-where $A[a]$ is the \change{discrete} set of all possible values of attribute $A$, $S_v$ is the subset of $S$ for which attribute $A$ has value $v$, and $E(\cdot)$ is the entropy function given by
+where $A[a]$ is the discrete set of all possible values of attribute $A$, $S_v$ is the subset of $S$ for which attribute $A$ has value $v$, and $E(\cdot)$ is the entropy function given by
 %
@@
 Conversely, an extremely low value of entropy would mean most of the data falls within a particular class.
-With this in mind, the information gain $G$ of an attribute $A$ for the dataset $S$, or $G(S,A)$ in equation (\ref{eq:gain}), is the expected reduction in entropy caused by partitioning the dataset according to \change{the attribute $A$ \citep{Mitchell_1997_Book}, which we consider to be discrete.
-For the selection of a threshold for a continuous attribute $A(a)$, please refer to \citet{Quinlan_1996_JAIR}.}
+With this in mind, the information gain $G$ of an attribute $A$ for the dataset $S$, or $G(S,A)$ in equation (\ref{eq:gain}), is the expected reduction in entropy caused by partitioning the dataset according to the attribute $A$ \citep{Mitchell_1997_Book}, which we consider to be discrete.
+For the selection of a threshold for a continuous attribute $A(a)$, please refer to \citet{Quinlan_1996_JAIR}.
 
-% attribute $A$ \citep{Mitchell_1997_Book}. For selection of a threshold for a \myrevision{numerical} continuous attribute refer to \citet{Quinlan_1996_JAIR}.
-
-\change{In the particular case of the C5.0 algorithm, in addition to the general steps just described, the result of the process of building a decision tree also depends on other logical and numerical parameters.
+In the particular case of the C5.0 algorithm, in addition to the general steps just described, the result of the process of building a decision tree also depends on other logical and numerical parameters.
 We set the options to: (a) perform feature selection or winnowing, (b) evaluate possible advanced splits of data, (c) use a confidence factor, and (d) use a threshold for the number of samples that go in the tree leaves.
@@
 Additional details about these options are available in \citet{Quinlan_1993_Book} and \citet{Kuhn_2017_Manual}.
-For this study, in order to obtain a variety of tree options, we build trees for varying values of CF and $S_{\min}$.}
-
-% SUGGESTED BY NAEEM:
-% In the particular case of the C5.0 algorithm the result of this process depends on several parameters, logical and numerical.
-We set the options to do feature selection \myrevision{or winnowing.
-Therefore, the algorithm choose to use the most important attributes and ignore others.
-We also activate the evaluation of possible advanced splits of data.
-Sometimes data is very close to the splitting threshold where assigning a hard threshold may incorrectly change the classification, whereas, that small difference could happen because of numerical errors or error in measurement.
-Therefore, advanced split can define a soft threshold to consider different probabilities in assigning data to different classes.
-A decision tree can grow so deep to finally classify all training data correctly.
-However, the tree will be very complex and will suffer from overfitting.
-In other words it will have very poor functionality for test data.
-Therefore, decision tree algorithms (here C5.0) first grow the tree up to the most complex tree then based on user defined factors start to prune the grown tree.
-The pruning process helps to reduce the complexity of the tree and also increase it's generalizability.
-Consequently the pruned tree's leaf nodes will have training data which belongs to more than one class which the final class will be the mode of that leaf node.
-In this study we use two parameters to control the pruning process including confidence or certainty factor (CF) and the smallest number of samples that must be put in at least two of the splits ($S_{\min}$, also known as \textit{minCases}).
-C5.0 algorithm uses error-based pruning method (see \citet{Quinlan_1993_Book} for more details).
-The objective is to minimize the observed error rate.
-Therefore, error estimates for leaf nodes and subtrees are computed assuming that they were used to classify a set of unseen cases with a predicted error rate which is computed based on CF and binomial distribution of number of misclassified events (E) and all training data at that leaf (N).
-Decreasing CF and increasing $S_{\min}$ cause more heavy pruning process \citep[see][]{Kuhn_2017_Manual}.} In particular, we build a suite of trees for varying values of CF and $S_{\min}$.
-
-% ORIGINAL:
-% In the particular case of the C5.0 algorithm the result of this process depends on several parameters, logical and numerical.
-We set the options to do feature selection and evaluate possible advanced splits of data, but do not use the boosted models option \citep[see][]{Kuhn_2017_Manual}; and use the confidence factor (CF) and the smallest number of samples that must be placed in at least two of the splits ($S_{\min}$, also known as \textit{minCases}) to fine tune the prediction models.
-In particular, we build a suite of trees for varying values of CF and $S_{\min}$.
+For this study, in order to obtain a variety of tree options, we build trees for varying values of CF and $S_{\min}$.
 
@@
 %
-where $P$ and $R$ are the precision and recall factors, respectively, and $\beta$ is a weighting factor between the two.
-$P$ measures the fraction of data samples that are correctly classified within a category, whereas $R$ measures the fraction of the data classified within that category regardless of whether it was correctly classified or not.
-Mathematically, they are defined as
+where $P$ and $R$ are the precision and recall factors, respectively, and $\beta$ is a weighting factor between the two. $P$ and $R$ are defined as
 %
@@
 	R = \frac{ \mathrm{TP} }{ \mathrm{TP} + \mathrm{FN} }
-	\, .
+	\, ,
 \end{equation}
 %
-Here, TP, FP, and FN are the number of samples in the testing dataset considered to be true-positives, false-positives, and false-negatives, respectively.
+respectively. Here, TP, FP, and FN are the number of samples in the testing dataset considered to be true-positives, false-positives, and false-negatives, respectively.
 
@@
 where TN is the number of true-negative samples.
-\change{Confusion matrices can be larger depending on the number of classes.
-In our case, the confusion matrices are size 4$\times$4, in order to compare the actual number of data samples classified as P, F, G and E with the number of samples predicted by the decision trees for each one of the same validation classes.}
+Confusion matrices can be larger depending on the number of classes.
+In our case, the confusion matrices are size 4$\times$4, in order to compare the actual number of data samples classified as P, F, G and E with the number of samples predicted by the decision trees for each one of the same validation classes.
 
@@
 This is done for all the trees obtained using all possible combinations of CF and $S_{\min}$ values.
-Then, we select trees with high levels of effectiveness as \change{measured} by $F_1$, commensurate to such trees having low complexities, i.e., reduced number of metrics and shallow depths, which is precisely the goal set at the start.
+Then, we select trees with high levels of effectiveness as measured by $F_1$, commensurate to such trees having low complexities, i.e., reduced number of metrics and shallow depths, which is precisely the goal set at the start.
 Note that $F_1$ is obtained based on the testing dataset as opposed to the training dataset.
-
-% The results as applied to our dataset are presented next.

### 2018-05-20 introduction.tex
completed first pass on second review by associate editor

@@
 
-Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation\change{, as well as a growing interest in the use of synthetic seismograms for engineering applications}.
-\change{Validation entails the comparison of simulations with observations, whereas verification involves the comparison of simulations with exact or alternative solutions \citep[see][and references therein]{Bielak_2010_GJI, Taborda_2013_BSSA}}.
-Various methods or schemes have been proposed to evaluate, through direct \change{signal-to-signal} quantitative comparisons or \change{overall} statistical analyses, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
-\change{Some of these methods are better} suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for validation.
-\change{\citet{Burks_BSSA_2014} and \citet{Rezaeian_2015_BSSA}, for instance, are used for validation in the context of engineering applications, i.e., at the level of the dynamic response of buildings.
+Verification and validation of ground motion synthetics have received increasing attention in recent years due to advances in deterministic and non-deterministic physics-based earthquake ground motion simulation, as well as a growing interest in the use of synthetic seismograms for engineering applications.
+Validation entails the comparison of simulations with observations, whereas verification involves the comparison of simulations with exact or alternative solutions \citep[see][and references therein]{Bielak_2010_GJI, Taborda_2013_BSSA}.
+Various methods or schemes have been proposed to evaluate, through direct signal-to-signal quantitative comparisons or overall statistical analyses, the similarity between simulation synthetics and recorded data, or with respect to other reference solutions \citep{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
+Some of these methods are better suited for verification purposes because they compare the signals at the waveforms level \citep{Kristekova_2006_BSSA, Kristekova_2009_GJI}, whereas others are better suited for validation.
+\citet{Burks_BSSA_2014} and \citet{Rezaeian_2015_BSSA}, for instance, are used for validation in the context of engineering applications, i.e., at the level of the dynamic response of buildings.
 \citet{Anderson_2004_Proc} and \citet{Olsen_2010_SRL}, on the other hand, are used for validation in the context of the overall characteristics of seismograms when comparing synthetics with data, in both time and frequency.
-These methods are preferred for validating high-frequency or broadband simulations, where applications may be undefined or the expectation of matching seismograms at the waveforms level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}), and because they offer simple schemes to quantify the goodness-of-fit (GOF) of the simulation results.}
+These methods are preferred for validating high-frequency or broadband simulations, where applications may be undefined or the expectation of matching seismograms at the waveforms level is less relevant (unless dealing with low-frequencies, \fmaxleq{1}), and because they offer simple schemes to quantify the goodness-of-fit (GOF) of the simulation results.
 
 Out of the various validation methods just described, \citet{Anderson_2004_Proc} is perhaps the most widely used today \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
-\change{In essence, this method assesses the similarity of two signals based on the average score of 10 metrics.
-These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), their peak values (in acceleration, velocity and displacement), frequency content (in terms of Fourier and response spectra), and their time synchronization (through cross-correlation).} We, in particular, have consistently relied on a modified version of \change{\citeauthor{Anderson_2004_Proc}'s method} for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake \citep{Taborda_2013_BSSA, Taborda_2014_BSSA}, and for the evaluation of velocity models in southern California through the validation of a large suite of moderate-magnitude earthquake simulations \citep{Taborda_2016_GJI}.
+In essence, this method assesses the similarity of two signals based on the average score of 10 metrics.
+These metrics measure the cumulative strength of the signals (in terms of Arias intensity and absolute energy), the comparative evolution of the signals in time (through normalized Arias and energy integrals), their peak values (in acceleration, velocity and displacement), frequency content (in terms of Fourier and response spectra), and their time synchronization (through cross-correlation).
+We, in particular, have consistently relied on a modified version of \citeauthor{Anderson_2004_Proc}'s method for the validation of a series of simulations of the 2008 $M_w$ 5.4 Chino Hills earthquake \citep{Taborda_2013_BSSA, Taborda_2014_BSSA}, and for the evaluation of velocity models in southern California through the validation of a large suite of moderate-magnitude earthquake simulations \citep{Taborda_2016_GJI}.
 
-The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics \change{convey physical meaning to both seismologists and engineers}.
+The strength of the method proposed by \citet{Anderson_2004_Proc} is that its metrics convey physical meaning to both seismologists and engineers.
 However, it has been pointed out that some of these metrics are redundant, or that other relevant metrics need to be added.
@@
 In the same spirit, \citet{Maufroy_2015_BSSA} reduced the number of metrics, limiting them to only those with comparable units.
-Unfortunately, none of these alternatives addresses the underlying questions \change{regarding} what are the \change{most important} parameters that ought to be taken into account when validating ground motion simulations \change{for a specific application}, and what level of priority should these \change{parameters} be given.
+Unfortunately, none of these alternatives addresses the underlying questions regarding what are the most important parameters that ought to be taken into account when validating ground motion simulations for a specific application, and what level of priority should these parameters be given.
 
@@
 To that end, we study the relationships that exist between eleven different metrics---those proposed by \citet{Anderson_2004_Proc}, plus the strong motion duration---with the objective of offering a prioritized reduced number of metrics that can help predict validation results.
-We \change{initially} treat these metrics as independent variables defining a multi-dimensional space, and analyze them using machine learning techniques.
-\change{In particular, we use semi-supervised and supervised methods.
+We initially treat these metrics as independent variables defining a multi-dimensional space, and analyze them using machine learning techniques.
+In particular, we use semi-supervised and supervised methods.
 In machine learning, whether a method is considered unsupervised, semi-supervised or supervised, depends on the level of constraints applied to the process or prior level of classification given to the data.
@@
 Upon labeling the data samples in our dataset into these four categories, we use a supervised learning method as implemented in the C5.0 algorithm to obtain a family of decision trees \citep[][]{Quinlan_1993_Book, Quinlan_1996_JAIR}.
-We then select a few of these trees to gain insight about and} prioritize the metrics involved in the validation process.
+We then select a few of these trees to gain insight about and prioritize the metrics involved in the validation process.
 The decision trees narrow the number of metrics offering a prediction algorithm into the aforementioned validation categories.
 
-In summary, \change{although specific to the chosen dataset (from the given set of simulations for a single event)}, our results indicate that among the eleven metrics considered in the analysis, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of acceleration and velocity.
+In summary, although specific to the chosen dataset (from the given set of simulations for a single event), our results indicate that among the eleven metrics considered in the analysis, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of acceleration and velocity.
 We test our prediction model and offer a discussion about its implications and potential use in future validation efforts.

### 2018-05-20 method.tex
finalized revision, after cleaning
completed first pass on second review by associate editor

@@
 
-We are interested in proposing an algorithm with a reduced and prioritized number of validation metrics based on previously acquired \change{collection of validation data samples} (i.e., our dataset).
+We are interested in proposing an algorithm with a reduced and prioritized number of validation metrics based on previously acquired collection of validation data samples (i.e., our dataset).
 A common method to do this is to identify rules with disjunctive characteristics, in the form of decision trees, that lead to outcomes representative of the overall validation analysis.
-In our case, we define such outcome in terms of four \change{\textit{validation categories} or \textit{classes}} representative of the quality of the validation, namely: poor (P), fair (F), good (G), and excellent (E).
+In our case, we define such outcome in terms of four \textit{validation categories} or \textit{classes} representative of the quality of the validation, namely: poor (P), fair (F), good (G), and excellent (E).
 The development of such decision trees requires a proper classification of the data, which can be done through a clustering process.
-\change{This implies applying the \textit{labels} P, F, G and E to the data samples in our dataset.} The following sections explain the data processing analysis we put in place to \change{label the data samples in our} dataset and subsequently obtain \change{three} alternative decision trees.
+This implies applying the \textit{labels} P, F, G and E to the data samples in our dataset.
+The following sections explain the data processing analysis we put in place to label the data samples in our dataset and subsequently obtain three alternative decision trees.
 
@@
 \input{decision-tree}
-
-% SOME MATERIAL PROVIDED BY NAEEM
-% \myrevision{In this study we integrate three different topics (i.e., ground motion simulation, semi-supervised, and supervised learning), therefore, the same concept may have different technical terms in different part of the paper based on their application.
-In machine learning linguistic, attributes and features, mostly are used interchangeably.
-In this study, we use features for clustering and attributes for classification purposes.
-The reason for choosing these terms is keeping consistency with the references that we use to develop the methods.
-Therefore, in this study feature, attribute, and metric all refer to the eleven GOF scores.
-Consequently, dimension refers to the number of features in each constrained \kmeans{} clustering analysis.
-For example, 3-Dimensional subspace analysis is application of constrained \kmeans{} on three out of eleven features.
-Each observation with eleven metrics is categorized in one of Poor(p), Fair(F), Good(g), and Excellent (E) categories.
-The process of assigning these categories to each observation observation is called clustering or labeling.
-The resulted dataset is called the labeled dataset.
-When we use these labeled dataset in classification, the labels or clusters are called classes.
-Therefore, label, class, and cluster all refer to four categories of GOF level.
-}

### 2018-05-20 metrics.tex
finalized revision, after cleaning
completed first pass on second review by associate editor

@@
 
-There are various methods and algorithms available to evaluate the similarity between two or more seismograms, both through direct \change{signal-to-signal} quantitative comparisons or \change{overall} statistical analyses \citep[e.g.,][]{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
+There are various methods and algorithms available to evaluate the similarity between two or more seismograms, both through direct signal-to-signal quantitative comparisons or overall statistical analyses \citep[e.g.,][]{Anderson_2004_Proc, Kristekova_2006_BSSA, Kristekova_2009_GJI, Olsen_2010_SRL, Burks_BSSA_2014, Rezaeian_2015_BSSA}.
 In earthquake ground motion simulation, they are used primarily for verification with respect to benchmark or analytical solutions, and for validation with respect to data (i.e., ground motion records).
@@
 where $S$ is the GOF score that results from comparing values $p_1$ and $p_2$ from signals 1 and 2, respectively, for each one of the different metrics in Table \ref{tab:metrics}.
-\change{In the case of C8 and C9, where the values $p_1$ and $p_2$ are function of the period ($T$) and frequency ($f$), respectively, $S$ is computed for all values of $T$ and $f$ and averaged to produce a single mean score.} \citeauthor{Anderson_2004_Proc} also provided guidelines for the process to be applied to the original signals as well as to those resulting from a sequential set of band-pass filters covering the whole frequency range of interest.
+In the case of C8 and C9, where the values $p_1$ and $p_2$ are function of the period ($T$) and frequency ($f$), respectively, $S$ is computed for all values of $T$ and $f$ and averaged to produce a single mean score.
+\citeauthor{Anderson_2004_Proc} also provided guidelines for the process to be applied to the original signals as well as to those resulting from a sequential set of band-pass filters covering the whole frequency range of interest.
 In his method, this frequency-domain decomposition should have pass bands defined following a logarithmic distribution in order to give more weight to the low frequencies, and the GOF scores of all sub-bands and the broadband be averaged into a single GOF final score.
-\change{However, in this study, in order to avoid the additional parameters that would result from this approach \citep[i.e., sub-bands width and filter characteristics][]{Khoshnevis_2015_Proc}, we use only the broadband (\frange{0}{4}) results.}
-
-% reduce the uncertainties that can come from different filtering process \citep[e.g., see ][]{Khoshnevis_2015_Proc} we use broadband results (\feq{0}{4}).
-For metrics which are functions of frequency (or period) like Fourier spectra and response spectra, the final result is the mean value of application of equation(\ref{eq:gof-function}) on all frequencies (or periods).}
-
-% \myrevision{However, in this study, in order to reduce the uncertainties that can come from different filtering process \citep[e.g., see ][]{Khoshnevis_2015_Proc} we use broadband results (\feq{0}{4}).
-For metrics which are functions of frequency (or period) like Fourier spectra and response spectra, the final result is the mean value of application of equation(\ref{eq:gof-function}) on all frequencies (or periods).}
+However, in this study, in order to avoid the additional parameters that would result from this approach \citep[i.e., sub-bands width and filter characteristics][]{Khoshnevis_2015_Proc}, we use only the broadband (\frange{0}{4}) results.
 
@@
 \citeauthor{Anderson_2004_Proc} concluded that for a typical distribution of scores, the GOF values could be classified into four validation categories: poor (for scores from 0 to 4), fair (4 to 6), good (6 to 8), and excellent (for scores from 8 to 10).
-While these categories are arguably subjective, over the years they have \change{been adopted---despite some differences in the ranges---by other GOF methods employed in verification and validation \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}.
-As a result, \citeauthor{Anderson_2004_Proc}'s method has been used as} a reference baseline for various validation studies \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
-
-% \myrevision{ been frequently used and} 
-% \myremove{become common practice and has} 
-% been adopted or matched by alternative validation methods \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}, making the method itself a reference baseline for various validation studies \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
+While these categories are arguably subjective, over the years they have been adopted---despite some differences in the ranges---by other GOF methods employed in verification and validation \citep[e.g.,][]{Kristekova_2009_GJI, Olsen_2010_SRL}.
+As a result, \citeauthor{Anderson_2004_Proc}'s method has been used as a reference baseline for various validation studies \citep[e.g.,][]{Chaljub_2010_BSSA, Bielak_2010_GJI, Guidotti_2011_SRL, Maufroy_2015_BSSA}.
 

### 2018-05-20 results.tex
finalized revision, after cleaning
completed first pass on second review by associate editor

@@
 
-Some aspects of Fig.~\ref{fig:clusters} are worth discussing. Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case. In
-% the C5-C9 and C7-C9 combinations in
-\change{the combinations in}
-the 3-feature analysis, for instance, the excellent cluster has a limited presence.
+Some aspects of Fig.~\ref{fig:clusters} are worth discussing.
+Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case.
+In the combinations in the 3-feature analysis, for instance, the excellent cluster has a limited presence.
 This is due to the influence of the lower C9 values (see Fig.~\ref{fig:data-box-plot}) on the correlations with C5 and C7.
@@
 Separately, we prepared similar plots to look at the influence of the velocity models and components of motions on the results of the clustering process and, as observed before, they had no significant differences with the aggregate of all the samples in the dataset.
-\change{Figure \ref{fig:boxed-clusters} is particularly important because it highlights the outcome of the clustering process and provides a preview into the decision tree analysis results.
+Figure \ref{fig:boxed-clusters} is particularly important because it highlights the outcome of the clustering process and provides a preview into the decision tree analysis results.
 Notice that metrics C5 through C8 consistently fall within \citeauthor{Anderson_2004_Proc}'s poor, fair, good, and excellent classifications, and that metrics C3 and C4 also show a progressive variation in sync with these categories.
@@
 On the other hand, metrics C1 and C2, and C9 through C11 are almost invariant, therefore less or not decisive in the validation process.
-C11, in particular, shows a broader (larger boxes) distribution, which indicates that it is less effective in predicting the final validation class.}
+C11, in particular, shows a broader (larger boxes) distribution, which indicates that it is less effective in predicting the final validation class.
 
@@
 The distributions are shown in the form of box-plots for each metric (C1 through C11).
-In each case, \change{the medians are indicated by a notch in the boxes of the central quartiles, and the vertical} lines represent the interquartile range, with outliers shown as scattered dots.
-\change{The dashed lines and background shadows represent the boundaries of the poor, fair, good, and excellent categories as defined by \citet{Anderson_2004_Proc}.} The color version of this figure is available only in the electronic edition.}
+In each case, the medians are indicated by a notch in the boxes of the central quartiles, and the vertical lines represent the interquartile range, with outliers shown as scattered dots.
+The dashed lines and background shadows represent the boundaries of the poor, fair, good, and excellent categories as defined by \citet{Anderson_2004_Proc}.
+The color version of this figure is available only in the electronic edition.}
 	\label{fig:boxed-clusters}
@@
 As it can be seen in this figure, the number of samples in the excellent class is significantly less than those in the poor, fair, and good classes.
-Therefore, before moving on with the decision tree analysis, it is necessary to resample the subset of the excellent class\change{, for which we use the oversampling approach described in the previous section.
+Therefore, before moving on with the decision tree analysis, it is necessary to resample the subset of the excellent class, for which we use the oversampling approach described in the previous section.
 Oversampling is nothing else but a replication of data samples.
@@
 Since the original size was 76 samples, that means we applied an oversampling ratio of 10$\times$.
-According to \citet{Weiss_2003_JAIR}, an oversample rate of 10$\times$ is considered to be acceptable in this type of data processing.} (Arguably, we could have undersampled the fair class as well but we deemed that unnecessary.
-Besides, we used a \change{heavy} pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.
-
-% We used the oversampling approach described in the previous section. We replicated data from the excellent class
-% \myrevision{through random selection of data from that class until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.
-This oversampling is considered as a $\times$10 factor.}
-% \myremove{randomly by a $\times$10 factor until elevating the number of samples in this class to 760 as indicated in Fig.~\ref{fig:count-classes} with the dashed line.}
-% By common standards, an oversample rate of $\times$10 is considered acceptable \citep{Weiss_2003_JAIR}.
+According to \citet{Weiss_2003_JAIR}, an oversample rate of 10$\times$ is considered to be acceptable in this type of data processing.
+(Arguably, we could have undersampled the fair class as well but we deemed that unnecessary.
+Besides, we used a heavy pruning process to prevent overfitting.) Once this process was completed we went on with the decision tree analysis.
 
@@
 For each one of these unique trees, we computed the effectiveness factor $F_1$ from equation (\ref{eq:f}), and extracted the total number of nodes in the trees and their depth.
-Figure \ref{fig:nodes-depth} shows the results of $F_1$ for all the trees and its distribution in terms of the number of \change{decision} attributes \change{or GOF metrics} used in the trees as a function of the number of nodes and the depth of each tree.
+Figure \ref{fig:nodes-depth} shows the results of $F_1$ for all the trees and its distribution in terms of the number of decision attributes or GOF metrics used in the trees as a function of the number of nodes and the depth of each tree.
 Recall that we are interested in finding a sequence of decisions (represented by disjunctive decision nodes in a tree) that can lead to good GOF predictions (i.e, high values of $F_1$) using a reduced number of attributes (GOF metrics).
@@
 T3 is the most complex of them.
-More complex trees tend to be deeper, have more nodes, and employ more attributes\change{, all of which depends---in part---on the pruning process.
-More complex trees have} higher $F_1$ values, but they may not necessarily be more practical.
+More complex trees tend to be deeper, have more nodes, and employ more attributes, all of which depends---in part---on the pruning process.
+More complex trees have higher $F_1$ values, but they may not necessarily be more practical.
 In total T1, T2, and T3 use 2, 3 and 4 attributes respectively.
-All coincide in the use of the total energy (C4) and response spectra (C8) as key metrics.
-T2 adds peak acceleration (C5), and T3 adds peak velocity \change{(C6)} to the previous metrics.
+All coincide in the use of the C4 (total energy) and C8 (response spectra) as key metrics.
+T2 adds C5 (peak acceleration), and T3 adds C6 (peak velocity) to the previous metrics.
 
@@
 This is a natural trade-off embedded in the use of decision trees.
-
-% , which actually results in more effective trees.
 
@@
 The differences between both matrices are actually minor and small with respect to their diagonal values, and the $F_1$ values in Tables \ref{tab:t1:f1} and \ref{tab:t3:f1} are all near to or above 0.9.
-The values obtained for tree T2 are consistent with \change{these observations}.
+The values obtained for tree T2 are consistent with these observations.
 
-Another aspect of interest is the level of participation of each metric in the analysis of data, as the GOF values are run through \change{the nodes of the decision trees (i.e., as the tree is traversed)}.
+Another aspect of interest is the level of participation of each metric in the analysis of data, as the GOF values are run through the nodes of the decision trees (i.e., as the tree is traversed).
 The results of such level of participation are listed in Table \ref{tab:participation} for trees T1, T2, T3, and T66.
@@
 The nodes are not unique, and metrics are often used in different nodes in a given tree.
-Therefore, these percentage accumulate for each \change{metric as the tree is traversed}.
+Therefore, these percentage accumulate for each metric as the tree is traversed.
 A percentage of 100 means that a given metric has the opportunity to see all the data samples through one or multiple nodes, a low percentage means only a few data samples are evaluated by that metric, and a null percentage means the metric plays no role in the decision tree (it is not present in any node).
@@
 
-\change{The results in Table 6 highlight the fact that, for all trees,} the metrics C4 (total energy) and C8 (response spectra) are consistently the most relevant ones in the decision algorithms.
+The results in Table 6 highlight the fact that, for all trees, the metrics \change{C4 (total energy) and C8 (response spectra) are consistently the most relevant ones in the decision algorithms.
 They are followed by C5 (peak acceleration) and C6 (peak velocity).
 On the other end, C11 (strong phase duration) plays no role whatsoever in any of the trees, and that C1 (Arias integral), C2 (energy integral), and C3 (Arias intensity) have only small to marginal participations.
-The remaining metrics---C7 (peak displacement), C9 (Fourier spectrum), and C10 (Cross correlation)---have a somewhat significant participation, but only in the more complex trees.
+The remaining metrics---C7 (peak displacement), C9 (Fourier spectrum), and C10 (cross correlation)}---have a somewhat significant participation, but only in the more complex trees.
 
-\change{An additional aspect worth highlighting here is the fact that the clustering and decision tree processes address the problem of correlation between the metrics in a natural way.
-For instance, one would expect C3 and C5, or C3 and C4 to be related.
+An additional aspect worth highlighting here is the fact that the clustering and decision tree processes address the problem of correlation between the metrics in a natural way.
+For instance, one would expect \change{C3 (Arias intensity) and C5 (peak acceleration), or C3 and C4 (total energy)} to be related.
 Considering them all without distinction could lead to double weighing their influence, and assuming a 1-to-1 relationship may overstate their level of correlation.
-Instead, the process used here determines which of them has a better chance to predict the outcome in the presence of the other, and preserves that metric as part of the decision tree.}
+Instead, the process used here determines which of them has a better chance to predict the outcome in the presence of the other, and preserves that metric as part of the decision tree.
 
-\change{One can put this to test by removing one or more metrics and see how the participation of the metrics rearranges.
-Table \ref{tab:participation-extra} shows these participations for trees with similar topologies (i) when removing C8, and (ii) when removing C5 and C8.
-In the first case, C5 takes the place of the most determinant metric, followed by C4 and C6, depending on the tree.
-In the second case, C6 becomes the most dominant, followed by C3 and C4, which vary in their level of participation depending on the tree.
-These additional alternatives emphasizes the C4 and C6, and show that other metrics such as C5, C7 and C9 are also relevant.
-They also reaffirm the low or null levels of contribution of metrics C1, C2, C10 and C11.
-All this is consistent with what we observed in Fig.~\ref{fig:boxed-clusters}.}
+One can put this to test by removing one or more metrics and see how the participation of the metrics rearranges.
+Table \ref{tab:participation-extra} shows these participations for trees with similar topologies (i) when removing \change{C8 (response spectrum), and (ii) when removing C5 (peak acceleration) and C8.
+In the first case, C5 takes the place of the most determinant metric, followed by C4 (total energy) and C6 (peak velocity), depending on the tree.
+In the second case, C6 becomes the most dominant, followed by C3 (Arias intensity) and C4, which vary in their level of participation depending on the tree.
+These additional alternatives emphasizes the C4 and C6, and show that other metrics such as C5, C7 (peak displacement) and C9 (Fourier spectrum) are also relevant.
+They also reaffirm the low or null levels of contribution of metrics C1 (Arias integral), C2 (energy integral), C10 (cross correlation) and C11 (strong phase duration)}.
+All this is consistent with what we observed in Fig.~\ref{fig:boxed-clusters}.
 
-% \myrevision{It is obvious that C8(response spectra) and C5(peak acceleration) are the most important metrics in developing the prediction models.
-Therefore, in order to understand other parameters importance we developed two sets of trees according to complexity of the original trees (the structure of generated trees are not presented here), while once dropping C8 (response spectra) and once dropping C8 (response spectra) and C5(peak acceleration) from the prediction attributes.
-According to Table \ref{tab:participation}, these trees also emphasize the importance of C4 and C6.
-However, in the absence of C5(peak acceleration) and C8 (response spectra),  C3 (Arias intensity) and C9 (Fourier spectrum) become more effective in developing prediction models.
-Also C11(strong phase duration) becomes somewhat important in the most complicated tree.}
-
-\change{That being said, a} low participation does not necessarily mean that a given metric is not relevant at all.
+That being said, a low participation does not necessarily mean that a given metric is not relevant at all.
 Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
-What the results we present here mean \change{is that, in the context of this particular set of 11 metrics, in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C4) are much more influential in the final result than the duration (C11).
-This, of course, is done under the assumption that one seeks to narrow the selection of metrics without loss of insight about the outcome of the validation process when compared to the outcome obtained with the whole suite of the eleven metrics used here.}
-
-% Original:
-% A low participation does not necessarily mean that a given metric is not important.
-Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
-What the results we present here mean, in the context of GOF validation, is that in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more decisive in reaching an accurate conclusion about validation quality than the duration (C11).
-
-% Naeem suggestion:
-% A low participation does not necessarily mean that a given metric is not important.
-Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
-What the results we present here mean, in the context of GOF validation for engineering applications, is that in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as energy (C8) are much more decisive in narrowing eleven validation metrics into fewer numbers in order to get the same categorization as all eleven metrics than the duration (C11).
+What the results we present here mean is that, in the context of this particular set of 11 metrics, in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as \change{C4 (energy) are much more influential in the final result than C11 (strong phase duration)}.
+This, of course, is done under the assumption that one seeks to narrow the selection of metrics without loss of insight about the outcome of the validation process when compared to the outcome obtained with the whole suite of the eleven metrics used here.
 
 In the end, the final selection of a preferred tree comes down to reducing complexity in the analysis without compromising the interpretation of results.
-We favor tree T1 because it is based only on three decision steps, and two GOF metrics: C4 (total energy), and C8 (response spectra).
+We favor tree T1 because it is based only on three decision steps, and two GOF metrics: \change{C4 (total energy), and C8 (response spectra)}.
 

### 2018-05-20 testing.tex
completed first pass on second review by associate editor

@@
-\section{\change{Testing}}
+\section{Testing}
 
-We now \change{test} tree T1 on the original simulation results from \citet{Taborda_2014_BSSA}, for different velocity models and components of motion.
+We now test tree T1 on the original simulation results from \citet{Taborda_2014_BSSA}, for different velocity models and components of motion.
 In this case we no longer aggregate all data from the simulation but now look at individual simulations done with different velocity models, and the three components of motion separately, as it would normally be done during any ground motion simulation validation.
 The results obtained with the T1 validation algorithm are shown in Fig.~\ref{fig:res-gof-maps}.
-\change{We should note that they are not supposed to be the same as those obtained by \citet{Taborda_2014_BSSA} because T1 no longer gives the same weight to all metrics but relies only on the two metrics C4 and C8.
-A} drawback from the T1 algorithm is that because its outcome are GOF validation classes (poor, fair, good, excellent) as opposed to GOF validation scores (with values from 0 to 10), once one obtains the results for the validation process for different components of motion (as shown in the rows of Fig.~\ref{fig:res-gof-maps}), there is no natural way of taking averages as one would do for scores in a numerical scale \change{\citep[i.e., as typically done when using][]{Anderson_2004_Proc}}.
+We should note that they are not supposed to be the same as those obtained by \citet{Taborda_2014_BSSA} because T1 no longer gives the same weight to all metrics but relies only on the two metrics C4 and C8.
+\change{A drawback of the T1 algorithm} is that because its outcome are GOF validation classes (poor, fair, good, excellent) as opposed to GOF validation scores (with values from 0 to 10), once one obtains the results for the validation process for different components of motion (as shown in the rows of Fig.~\ref{fig:res-gof-maps}), there is no natural way of taking averages as one would do for scores in a numerical scale \citep[i.e., as typically done when using][]{Anderson_2004_Proc}.
 Therefore, in order to combine results from the three components into a single validation classification for each station, we define the following rules:
@@
 Applying them to the validation results shown in Fig.~\ref{fig:res-gof-maps} for the case of the simulation done using the model CVM-S is shown in Fig.~\ref{fig:avg-gof-maps}, which compares the outcome with the result of the average scores obtained using the traditional \citet{Anderson_2004_Proc}-type GOF method.
-This figure exemplifies the \change{differences or similarities} between the numeric GOF validation-scores scale and the proposed T1 GOF validation classes.
+This figure exemplifies the differences or similarities between the numeric GOF validation-scores scale and the proposed T1 GOF validation classes.
 
@@
 In our view, the T1 results are simpler to assimilate and equally informative.
-Looking at both plots in Fig.~\ref{fig:avg-gof-maps} it is possible to argue they lead to \change{similar conclusions in respect to the overall validation of the simulation, and in respect to some particular areas and specific locations.
-Nonetheless, it must be said that we do not expect to see a 1-to-1 relationship.
-That is the scope of a future work, whereas here our focus was on identifying the relevance of the different metrics.
-This is important because the original approach proposed by \citet{Anderson_2004_Proc} favors a uniform weighing of metrics, and the results shown here indicate this may not be the preferred strategy.}
+Looking at both plots in Fig.~\ref{fig:avg-gof-maps} it is possible to argue they lead to similar conclusions in respect to the overall validation of the simulation, and in respect to some particular areas and specific locations.
+Nonetheless, it must be said that we do not expect to see a 1-to-1 relationship\change{, as evidenced by the differences in the histograms when compared in equal terms (see dashed-line counts on the right-hand side corresponding to the count of the original results from the left-hand side histogram drawn next to the counts of the results from the T1 algorithm).
+Finding the best possible match} is the scope of a future work, whereas here our focus was on identifying the relevance of the different metrics.
+This is important because the original approach proposed by \citet{Anderson_2004_Proc} favors a uniform weighing of metrics, and the results shown here indicate this may not be the preferred strategy.
 
@@
 The contours are drawn for illustration purposes only.
-The bottom plots show histograms with the count of stations for each GOF score interval or validation class.}
+The bottom plots show histograms with the count of stations for each GOF score interval or validation class.
+\change{On the right-hand side, the results of both methods are compared by superimposing the results of the counts from using \citeauthor{Anderson_2004_Proc}'s method (dashed-line empty bars) next to those obtained using the T1 algorithm (filled bars).} The color version of this figure is available only in the electronic edition.}
 	\label{fig:avg-gof-maps}

### 2018-05-21 results.tex
updates

@@
 Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
-What the results we present here mean is that, in the context of this particular set of 11 metrics, in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as \change{C4 (energy) are much more influential in the final result than C11 (strong phase duration)}.
+What the results we present here mean is that, in the context of this particular set of 11 metrics, in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as \change{C4 (total energy) are much more influential in the final result than C11 (strong phase duration)}.
 This, of course, is done under the assumption that one seeks to narrow the selection of metrics without loss of insight about the outcome of the validation process when compared to the outcome obtained with the whole suite of the eleven metrics used here.

### 2018-05-22 response-two.tex
updates
completed second review

@@
 The reason being that these plots should not be interpreted as typical correlation plots of scattered data are usually interpreted.
-If we add the line we would be erroneously leading the readers to think these are as any other $x$-$y$ plots correlating showing the relationship of two metrics as seen through the data, and here one also needs to understand that the symbols (or colors) and how well defined are the transitions from one to another are actually equally, and perhaps more important than whether the cloud of data points follow a 1:1 relationship or not.
-To make this point clearer we added a comment in the text reflecting this point...
-\textcolor{red}{Adding the comment in the text of the paper is pending on Naeem agreeing with this response}.
+If we add the line we would be erroneously leading the readers to think these are as any other $x$-$y$ plots showing the relationship of two metrics as seen through the data.
+Here, one also needs to understand that the symbols (or colors) and how well defined are the transitions from one to another are actually equally, and perhaps more important than whether the cloud of data points follow a 1:1 relationship or not.
+We will therefore appreciate it if we can leave the figure as is, and instead, we added a short sentence to make this point clearer to the reader on page 15, lines 347--349.
 }

### 2018-05-22 results.tex
completed second review

--- 2018-05-20 results.tex
@@
 Some aspects of Fig.~\ref{fig:clusters} are worth discussing.
-Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case.
+\change{First of all, one must keep in mind that these are not typical correlation plots.
+How well defined the clusters are from each other or with respect to the metrics are important aspects to consider.} Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case.
 In the combinations in the 3-feature analysis, for instance, the excellent cluster has a limited presence.
@@
 In total T1, T2, and T3 use 2, 3 and 4 attributes respectively.
-All coincide in the use of the C4 (total energy) and C8 (response spectra) as key metrics.
-T2 adds C5 (peak acceleration), and T3 adds C6 (peak velocity) to the previous metrics.
+All coincide in the use of \change{the total energy (C4) and response spectra (C8) as key metrics.
+T2 adds the peak acceleration (C5), and T3 adds the peak velocity (C6)} to the previous metrics.
 
@@
 
-The results in Table 6 highlight the fact that, for all trees, the metrics \change{C4 (total energy) and C8 (response spectra) are consistently the most relevant ones in the decision algorithms.
-They are followed by C5 (peak acceleration) and C6 (peak velocity).
-On the other end, C11 (strong phase duration) plays no role whatsoever in any of the trees, and that C1 (Arias integral), C2 (energy integral), and C3 (Arias intensity) have only small to marginal participations.
-The remaining metrics---C7 (peak displacement), C9 (Fourier spectrum), and C10 (cross correlation)}---have a somewhat significant participation, but only in the more complex trees.
+The results in Table 6 highlight the fact that, for all trees, the metrics \change{for total energy (C4) and response spectra (C8) are consistently the most relevant ones in the decision algorithms.
+They are followed by the peak acceleration (C5) and peak velocity (C6).
+On the other end, the strong phase duration (C11) plays no role whatsoever in any of the trees, and the Arias integral (C1), energy integral (C2), and Arias intensity (C3) have only small to marginal participations.
+The remaining metrics---peak displacement (C7), Fourier spectrum (C9), and cross correlation (C10)}---have a somewhat significant participation, but only in the more complex trees.
 
 An additional aspect worth highlighting here is the fact that the clustering and decision tree processes address the problem of correlation between the metrics in a natural way.
-For instance, one would expect \change{C3 (Arias intensity) and C5 (peak acceleration), or C3 and C4 (total energy)} to be related.
+For instance, one would expect \change{the Arias intensity (C3) and peak acceleration (C5), or the Arias intensity (C3) and total energy (C4)} to be related.
 Considering them all without distinction could lead to double weighing their influence, and assuming a 1-to-1 relationship may overstate their level of correlation.
@@
 One can put this to test by removing one or more metrics and see how the participation of the metrics rearranges.
-Table \ref{tab:participation-extra} shows these participations for trees with similar topologies (i) when removing \change{C8 (response spectrum), and (ii) when removing C5 (peak acceleration) and C8.
-In the first case, C5 takes the place of the most determinant metric, followed by C4 (total energy) and C6 (peak velocity), depending on the tree.
-In the second case, C6 becomes the most dominant, followed by C3 (Arias intensity) and C4, which vary in their level of participation depending on the tree.
-These additional alternatives emphasizes the C4 and C6, and show that other metrics such as C5, C7 (peak displacement) and C9 (Fourier spectrum) are also relevant.
-They also reaffirm the low or null levels of contribution of metrics C1 (Arias integral), C2 (energy integral), C10 (cross correlation) and C11 (strong phase duration)}.
+Table \ref{tab:participation-extra} shows these participations for trees with similar topologies (i) when removing \change{the response spectrum (C8), and (ii) when removing both the peak acceleration (C5) and the response spectrum (C8).
+In the first case, the peak acceleration (C5) takes the place of the most determinant metric, followed by the total energy (C4) and peak velocity (C6), depending on the tree.
+In the second case, the peak velocity (C6) becomes the most dominant, followed by the Arias intensity (C3) and the total energy (C4), which vary in their level of participation depending on the tree.
+These additional alternatives emphasize the role of the total energy (C4) and the peak velocity (C6), and show that other metrics such as the peak acceleration (C5), peak displacement (C7) and Fourier spectrum (C9) are also relevant.
+They also reaffirm the low or null levels of contribution of the metrics associated with the shape of the Arias integral (C1), the energy integral (C2), and the cross correlation (C10) and strong phase duration (C11)}.
 All this is consistent with what we observed in Fig.~\ref{fig:boxed-clusters}.
@@
 Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
-What the results we present here mean is that, in the context of this particular set of 11 metrics, in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as \change{C4 (energy) are much more influential in the final result than C11 (strong phase duration)}.
+What the results we present here mean is that, in the context of this particular set of 11 metrics, in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as \change{the total energy (C4) are much more influential in the final result than the strong phase duration (C11)}.
 This, of course, is done under the assumption that one seeks to narrow the selection of metrics without loss of insight about the outcome of the validation process when compared to the outcome obtained with the whole suite of the eleven metrics used here.
@@
 In the end, the final selection of a preferred tree comes down to reducing complexity in the analysis without compromising the interpretation of results.
-We favor tree T1 because it is based only on three decision steps, and two GOF metrics: \change{C4 (total energy), and C8 (response spectra)}.
+We favor tree T1 because it is based only on three decision steps, and two GOF metrics: \change{the total energy (C4) and response spectra (C8)}.
 

### 2018-05-23 conclusions.tex
changes before final submission

@@
 Identifying the total energy and response spectra metrics as decisive metrics in the comparisons between synthetics and observations is a key contribution to validation of ground motion simulations.
-This contributes to clarifying a standing question in the area of validation, and it provides an indication to simulators about where \change{to focus} in the search for improvements in their models.
+This contributes to clarifying a standing question in the area of validation, and it provides an indication to simulators about where to focus in the search for improvements in their models.

### 2018-05-23 data-resources.tex
changes before final submission

@@
 
-Calculations, data processing, and some initial figures were done using R, the language and environment for statistical computing and graphics (\url{https://www.r-project.org}, last accessed April 2018), and the C5.0 package developed by \citet{Kuhn_2017_Manual}.
+Calculations, data processing, and some initial figures were done using R, the language and environment for statistical computing and graphics (\url{https://www.r-project.org}, last accessed May 2018), and the C5.0 package (\url{https://CRAN.R-project.org/package=C50}, last accessed May 2018).
 Additional calculations and figures were done using MATLAB (\url{http://www.mathworks.com}, last accessed April 2018).
-Map figures were prepared using the Generic Mapping Tools, GMT (\url{http://gmt.soest.hawaii.edu/}, last accessed April 2018).
-Additional editing of figures was done using Adobe Illustrator (\url{http://www.adobe.com/Illustrator‎}, last accessed April 2018).
+Map figures were prepared using the Generic Mapping Tools, GMT (\url{http://gmt.soest.hawaii.edu/}, last accessed May 2018).
+Additional editing of figures was done using Adobe Illustrator (\url{http://www.adobe.com/Illustrator‎}, last accessed May 2018).
 The validation dataset used here was readily available to the authors and can be provided without restrictions upon request.

### 2018-05-23 dataset.tex
changes before final submission

@@
 The simulation domain and the stations used in that study are shown in Figure \ref{fig:chino-hills}, and a sample of the validation results obtained by the authors using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps}.
-This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, \change{for the three velocity models used by \citet{Taborda_2014_BSSA}.
+This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, for the three velocity models used by \citet{Taborda_2014_BSSA}.
 In each case, the results are the average of the GOF results for the three components of motion and the scores for the eleven metrics from Table \ref{tab:metrics}.
 A subsequent study by \citet{Taborda_2016_GJI} using multiple events found that the model CVM-S \citep[version 4.26.M01, see][]{Small_2017_SRL} to be the model that most consistently led to better simulation results.
-Here, however, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the components of motion}.
+Here, however, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the components of motion.
 
@@
     \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between simulations for three different velocity models and data recorded at ground motion monitoring stations.
-\change{The distribution of the results at the bottom show the count of stations on the GOF scale defined by \citet{Anderson_2004_Proc} and the ranges of the validation categories: poor, fair, good and excellent.} The color version of this figure is available only in the electronic edition.}
+The distribution of the results at the bottom show the count of stations on the GOF scale defined by \citet{Anderson_2004_Proc} and the ranges of the validation categories: poor, fair, good and excellent.
+The color version of this figure is available only in the electronic edition.}
     \label{fig:ref-gof-maps}
@@
 
-We refer to the set of 11 scores associated with a pair of signals \change{from the work done by \citet{Taborda_2014_BSSA}} as one of the \textit{data samples} that compose our dataset.
+We refer to the set of 11 scores associated with a pair of signals from the work done by \citet{Taborda_2014_BSSA} as one of the \textit{data samples} that compose our dataset.
 Although at times we will make distinctions between the velocity models and the components of motion for visualization purposes, the clustering analysis to be described in the following section was done using all the data samples in the validation dataset.
@@
 In other words, the statistical distribution of the data for each metric is about the same independently of model or component.
-Consequently, \change{and in order to use a dataset as large as possible, we thought it acceptable} to combine all the data samples into a single dataset.
+Consequently, and in order to use a dataset as large as possible, we thought it acceptable to combine all the data samples into a single dataset.
 

### 2018-05-23 results.tex
changes before final submission

@@
 Some aspects of Fig.~\ref{fig:clusters} are worth discussing.
-\change{First of all, one must keep in mind that these are not typical correlation plots.
-How well defined the clusters are from each other or with respect to the metrics are important aspects to consider.} Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case.
+First of all, one must keep in mind that these are not typical correlation plots.
+How well defined the clusters are from each other or with respect to the metrics are important aspects to consider.
+Although in most plots all four clusters are clearly distinguishable, there are others in which that is not the case.
 In the combinations in the 3-feature analysis, for instance, the excellent cluster has a limited presence.
@@
 	\includegraphics[width=\textwidth]{figures/pdf/figure-08}
-	\caption{Accuracy of tree predictions in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by colored dots distributed with respect to the number of attributes (GOF metrics) as a function of: \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees.
+	\caption{Accuracy of tree predictions in terms of the factor ($F_1$) from equation \ref{eq:f} as indicated by dots distributed with respect to the number of attributes (GOF metrics) as a function of: \textbf{(a)} the number of nodes, and \textbf{(b)} the depth of the trees.
 The rings around some of the dots indicate select trees used as reference in the text, other figures and tables.
@@
 In total T1, T2, and T3 use 2, 3 and 4 attributes respectively.
-All coincide in the use of \change{the total energy (C4) and response spectra (C8) as key metrics.
-T2 adds the peak acceleration (C5), and T3 adds the peak velocity (C6)} to the previous metrics.
+All coincide in the use of the total energy (C4) and response spectra (C8) as key metrics.
+T2 adds the peak acceleration (C5), and T3 adds the peak velocity (C6) to the previous metrics.
 
@@
 
-The results in Table 6 highlight the fact that, for all trees, the metrics \change{for total energy (C4) and response spectra (C8) are consistently the most relevant ones in the decision algorithms.
+The results in Table \ref{tab:participation} highlight the fact that, for all trees, the metrics for total energy (C4) and response spectra (C8) are consistently the most relevant ones in the decision algorithms.
 They are followed by the peak acceleration (C5) and peak velocity (C6).
 On the other end, the strong phase duration (C11) plays no role whatsoever in any of the trees, and the Arias integral (C1), energy integral (C2), and Arias intensity (C3) have only small to marginal participations.
-The remaining metrics---peak displacement (C7), Fourier spectrum (C9), and cross correlation (C10)}---have a somewhat significant participation, but only in the more complex trees.
+The remaining metrics---peak displacement (C7), Fourier spectrum (C9), and cross correlation (C10)---have a somewhat significant participation, but only in the more complex trees.
 
 An additional aspect worth highlighting here is the fact that the clustering and decision tree processes address the problem of correlation between the metrics in a natural way.
-For instance, one would expect \change{the Arias intensity (C3) and peak acceleration (C5), or the Arias intensity (C3) and total energy (C4)} to be related.
+For instance, one would expect the Arias intensity (C3) and peak acceleration (C5), or the Arias intensity (C3) and total energy (C4) to be related.
 Considering them all without distinction could lead to double weighing their influence, and assuming a 1-to-1 relationship may overstate their level of correlation.
@@
 One can put this to test by removing one or more metrics and see how the participation of the metrics rearranges.
-Table \ref{tab:participation-extra} shows these participations for trees with similar topologies (i) when removing \change{the response spectrum (C8), and (ii) when removing both the peak acceleration (C5) and the response spectrum (C8).
+Table \ref{tab:participation-extra} shows these participations for trees with similar topologies (i) when removing the response spectrum (C8), and (ii) when removing both the peak acceleration (C5) and the response spectrum (C8).
 In the first case, the peak acceleration (C5) takes the place of the most determinant metric, followed by the total energy (C4) and peak velocity (C6), depending on the tree.
@@
 These additional alternatives emphasize the role of the total energy (C4) and the peak velocity (C6), and show that other metrics such as the peak acceleration (C5), peak displacement (C7) and Fourier spectrum (C9) are also relevant.
-They also reaffirm the low or null levels of contribution of the metrics associated with the shape of the Arias integral (C1), the energy integral (C2), and the cross correlation (C10) and strong phase duration (C11)}.
+They also reaffirm the low or null levels of contribution of the metrics associated with the shape of the Arias integral (C1), the energy integral (C2), and the cross correlation (C10) and strong phase duration (C11).
 All this is consistent with what we observed in Fig.~\ref{fig:boxed-clusters}.
@@
 Arguably, the strong phase duration is highly regarded as an important parameter for strong motion records in engineering.
-What the results we present here mean is that, in the context of this particular set of 11 metrics, in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as \change{the total energy (C4) are much more influential in the final result than the strong phase duration (C11)}.
+What the results we present here mean is that, in the context of this particular set of 11 metrics, in order to classify whether a simulation result is poor, fair, good, or excellent, other metrics such as the total energy (C4) are much more influential in the final result than the strong phase duration (C11).
 This, of course, is done under the assumption that one seeks to narrow the selection of metrics without loss of insight about the outcome of the validation process when compared to the outcome obtained with the whole suite of the eleven metrics used here.
@@
 In the end, the final selection of a preferred tree comes down to reducing complexity in the analysis without compromising the interpretation of results.
-We favor tree T1 because it is based only on three decision steps, and two GOF metrics: \change{the total energy (C4) and response spectra (C8)}.
+We favor tree T1 because it is based only on three decision steps, and two GOF metrics: the total energy (C4) and response spectra (C8).
 

### 2018-05-23 testing.tex
changes before final submission

@@
 We should note that they are not supposed to be the same as those obtained by \citet{Taborda_2014_BSSA} because T1 no longer gives the same weight to all metrics but relies only on the two metrics C4 and C8.
-\change{A drawback of the T1 algorithm} is that because its outcome are GOF validation classes (poor, fair, good, excellent) as opposed to GOF validation scores (with values from 0 to 10), once one obtains the results for the validation process for different components of motion (as shown in the rows of Fig.~\ref{fig:res-gof-maps}), there is no natural way of taking averages as one would do for scores in a numerical scale \citep[i.e., as typically done when using][]{Anderson_2004_Proc}.
+A drawback of the T1 algorithm is that because its outcome are GOF validation classes (poor, fair, good, excellent) as opposed to GOF validation scores (with values from 0 to 10), once one obtains the results for the validation process for different components of motion (as shown in the rows of Fig.~\ref{fig:res-gof-maps}), there is no natural way of taking averages as one would do for scores in a numerical scale \citep[i.e., as typically done when using][]{Anderson_2004_Proc}.
 Therefore, in order to combine results from the three components into a single validation classification for each station, we define the following rules:
@@
 Looking at both plots in Fig.~\ref{fig:avg-gof-maps} it is possible to argue they lead to similar conclusions in respect to the overall validation of the simulation, and in respect to some particular areas and specific locations.
-Nonetheless, it must be said that we do not expect to see a 1-to-1 relationship\change{, as evidenced by the differences in the histograms when compared in equal terms (see dashed-line counts on the right-hand side corresponding to the count of the original results from the left-hand side histogram drawn next to the counts of the results from the T1 algorithm).
-Finding the best possible match} is the scope of a future work, whereas here our focus was on identifying the relevance of the different metrics.
+Nonetheless, it must be said that we do not expect to see a 1-to-1 relationship, as evidenced by the differences in the histograms when compared in equal terms (see dashed-line counts on the right-hand side corresponding to the count of the original results from the left-hand side histogram drawn next to the counts of the results from the T1 algorithm).
+Finding the best possible match is the scope of a future work, whereas here our focus was on identifying the relevance of the different metrics.
 This is important because the original approach proposed by \citet{Anderson_2004_Proc} favors a uniform weighing of metrics, and the results shown here indicate this may not be the preferred strategy.
@@
 The bottom plots show histograms with the count of stations for each GOF score interval or validation class.
-\change{On the right-hand side, the results of both methods are compared by superimposing the results of the counts from using \citeauthor{Anderson_2004_Proc}'s method (dashed-line empty bars) next to those obtained using the T1 algorithm (filled bars).} The color version of this figure is available only in the electronic edition.}
+On the right-hand side, the results of both methods are compared by superimposing the results of the counts from using \citeauthor{Anderson_2004_Proc}'s method (dashed-line empty bars) next to those obtained using the T1 algorithm (filled bars).
+The color version of this figure is available only in the electronic edition.}
 	\label{fig:avg-gof-maps}

### 2018-05-24 dataset.tex
new changes to final submission

@@
     \includegraphics[width=\textwidth]{figures/pdf/figure-02}
-    \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between simulations for three different velocity models and data recorded at ground motion monitoring stations.
+    \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between data recorded at ground motion monitoring stations and simulations for three different velocity models: \textbf{(a)} CVM-S, \textbf{(b)} CVM-H, and \textbf{(c)} CVM-H+GTL.
 The distribution of the results at the bottom show the count of stations on the GOF scale defined by \citet{Anderson_2004_Proc} and the ranges of the validation categories: poor, fair, good and excellent.
@@
     \includegraphics[width=0.48\textwidth]{figures/pdf/figure-03}
-    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and components of motion (NS, EW, UD).
+    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), and components of motion (NS, EW, UD), for the simulations done with velocity models: \textbf{(a)} CVM-S, \textbf{(b)} CVM-H, and \textbf{(c)} CVM-H+GTL.
 In each case, the boxes represent the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), the medians are indicated by a notch in the boxes, and the vertical lines show the range of the data, with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots.

### 2018-05-24 results.tex
new changes to final submission

@@
 	\includegraphics[width=\textwidth]{figures/pdf/figure-06}
-	\caption{Statistical distribution of the dataset as partitioned into the four validation categories (poor, fair, good, excellent) after the clustering analysis.
+	\caption{Statistical distribution of the dataset after the clustering analysis, as partitioned into the four validation categories: \textbf{(a)} poor, \textbf{(b)} fair, \textbf{(c)} good, and \textbf{(d)} excellent.
 The distributions are shown in the form of box-plots for each metric (C1 through C11).

### 2018-05-24 testing.tex
new changes to final submission

@@
 	\includegraphics[width=\textwidth]{figures/pdf/figure-11}
-	\caption{Comparison of GOF validation scores obtained using an 11-metric \citet{Anderson_2004_Proc}-type GOF scoring (left) and the T1 combination GOF validation classification (right).
-The top plots show maps with a distribution of the scores or classes outcome at each station.
-The contours are drawn for illustration purposes only.
-The bottom plots show histograms with the count of stations for each GOF score interval or validation class.
+	\caption{Comparison of GOF validation scores obtained using \textbf{(a)} an 11-metric \citet{Anderson_2004_Proc}-type GOF scoring, and \textbf{(b)} the T1 combination GOF validation classification.
+In both cases, the top plots show maps with a distribution of the scores or classes outcome at each station, where the contours are drawn for illustration purposes only; while the bottom plots show histograms with the count of stations for each GOF score interval or validation class.
 On the right-hand side, the results of both methods are compared by superimposing the results of the counts from using \citeauthor{Anderson_2004_Proc}'s method (dashed-line empty bars) next to those obtained using the T1 algorithm (filled bars).
